---
title: "UPF3A and UPF3B Differential Gene Expression Analysis in Mouse L-cells"
author: "Urwah Nawaz"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_float: true
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  autodep = TRUE,
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

```{r load-libraries}
library(tximport)
library(edgeR)
library(limma)
library(EnsDb.Mmusculus.v79)
library(org.Mm.eg.db)
library(tidyverse)
library(magrittr)
library(scales)
library(reshape2)
library(gridExtra)
library(pheatmap)
source(here::here("code/plots.R"))
source(here::here("code/functions.R"))
```

# Overview

This notebook covers RNA-seq quality control and differential gene expression (DEG) analysis of L-cell lines with variable *Upf3a* and *Upf3b* expression (Control, *Upf3a* KD, *Upf3a* OE, *Upf3b* KD, *Upf3
dKD*, *Upf3b* KD + *Upf3a* OE). Only QC and volcano plots are retained here.



# Data import and filtering

```{r import-counts}
salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Salmon")
md.files = here::here("data/LTK_Sample Metafile_V3.txt")

txdf <- transcripts(EnsDb.Mmusculus.v79, return.type = "DataFrame")
tx2gene <- as.data.frame(txdf[, c("tx_id", "gene_id")])

salmon <- list.files(salmon.files, pattern = "transcripts$", full.names = TRUE)
all_files <- file.path(salmon, "quant.sf")
names(all_files) <- paste(c(212:229))

txi_genes <- tximport(
  all_files,
  type = "salmon",
  txOut = FALSE,
  countsFromAbundance = "scaledTPM",
  tx2gene = tx2gene,
  ignoreTxVersion = TRUE,
  ignoreAfterBar = TRUE
)


md = read.table(md.files, header= TRUE) %>%  mutate(files = file.path(salmon, "quant.sf"))
md$Group = factor(md$Group)
md$Sample = as.character(md$Sample)

group_cols <- hcl.colors(n = length(unique(md$Group)), palette = "Zissou 1") %>%
  setNames(unique(md$Group))
```

Genes were removed if they were not detectable (TPM < 0.5 in more than 3 samples) or lacked annotation. This retained 10,957 genes for downstream analysis.

```{r filter-genes}
keep <- (rowSums(txi_genes$abundance >= 0.5) >= 3)

txi_genes_filtered <- txi_genes$counts[keep, ] %>%
  as.data.frame() %>%
  rownames_to_column("geneID") %>%
  mutate(geneName = mapIds(org.Mm.eg.db, keys = .$geneID,
                            column = "SYMBOL", keytype = "ENSEMBL",
                            multiVals = "first")) %>%
  drop_na(geneName) %>%
  column_to_rownames("geneID") %>%
  dplyr::select(-geneName)
```

# Quality control

## Expression distributions before and after filtering

```{r qc-density, fig.height=4, fig.width=9, fig.cap="Expression density plots before and after low-abundance gene filtering.** logCPM distributions are shown for all samples before (left) and after (right) removing genes with TPM < 0.5 in fewer than 3 samples. Filtering removes the peak of low/undetectable expression while preserving the expressed gene distribution across samples."}

exp_den <- txi_genes$counts %>%
  cpm(log = TRUE) %>%
  reshape2::melt() %>%
  dplyr::filter(is.finite(value)) %>%
  ggplot(aes(x = value, color = as.character(Var2))) +
  geom_density() +
  ggtitle("Before filtering") +
  labs(x = "logCPM", y = "Proportion of genes") +
  theme_bw() +
  theme(legend.position = "none")

exp_den_filt <- txi_genes$counts %>%
  cpm(log = TRUE) %>%
  magrittr::extract(keep, ) %>%
  reshape2::melt() %>%
  dplyr::filter(is.finite(value)) %>%
  ggplot(aes(x = value, color = as.character(Var2))) +
  geom_density() +
  ggtitle("After filtering") +
  labs(x = "logCPM", y = "Proportion of genes") +
  theme_bw() +
  theme(legend.position = "none")

grid.arrange(exp_den, exp_den_filt, ncol = 2)

```

## Principal component analysis

```{r qc-pca, fig.height=4.5, fig.width=6, fig.cap="Principal component analysis of logCPM expression values across all L-cell lines.** Each point represents one sample, coloured by cell line and shaped by *Upf3b* knockdown status. *Upf3a* KD and *Upf3b* KD samples cluster away from Controls along PC1/PC2 but do not cluster close to one another, while *Upf3a* OE samples overlap with Controls, consistent with minimal impact of UPF3A overexpression on the transcriptome."}
pca <- log2(txi_genes$abundance[keep, ] + 10^-6) %>%
  t() %>%
  prcomp(scale = TRUE)

pcaVars <- percent_format(0.1)(summary(pca)$importance["Proportion of Variance", ])

gene_pca <- pca$x[, 1:2] %>%
  as.data.frame() %>%
  cbind(md) %>%
  as_tibble() %>%
  ggplot(aes(PC1, PC2, shape = Condition_UPF3B, color = Group)) +
  geom_point(size = 6) +
  theme_bw() +
  scale_shape_manual(values = c(0, 16)) +
  labs(shape = "UPF3B status", colour = "Cell line") +
  scale_color_manual(
    values = c("#FCC98A", "#6E74B4", "#E12F8F", "#EA2A5F", "#FD9675", "#2F124B"),
    labels = c(
      "UPF3A_KD_UPF3B_KD" = "Double KD",
      "UPF3A_KD" = "UPF3A KD",
      "UPF3B_KD" = "UPF3B KD",
      "UPF3A_OE" = "UPF3A OE",
      "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD"
    )
  ) +
  labs(
    x = paste0("PC1 (", pcaVars[["PC1"]], ")"),
    y = paste0("PC2 (", pcaVars[["PC2"]], ")")
  )

gene_pca
```

The principal component analysis was performed using the logCPM values from each sample. Each group appears to be clustering together. UPF3A over-expression is clustering with Controls, whereas the UPF3B KD samples are clustering closely with UPF3A overexpression in UPF3B KD files. 


## Per-sample expression and library size

```{r qc-boxplot, fig.height=4, fig.cap = "Distribution of gene expression (log TPM) per sample** after filtering, used to identify outlier samples prior to differential expression testing."}
txi_genes$abundance[keep, ] %>%
  as.data.frame() %>%
  gather(sample, value) %>%
  ggplot(aes(x = sample, y = log(value), fill = sample)) +
  geom_boxplot() +
  ylab("log(TPM)") +
  xlab("Sample") +
  theme_bw() +
  theme(legend.position = "none")
```


# Differential gene expression using limma 

## Model description 

Five pairwise comparisons (each condition vs. Control) were tested using `limma`/`voom`.

```{r}
y <- DGEList(txi_genes_filtered)


design <- model.matrix(~0+Group, data = md) %>% 
    set_colnames(gsub(pattern = "Group", replacement ="", x = colnames(.)))

contrasts <- makeContrasts(levels = colnames(design), 
                           UPF3B_vs_Control = UPF3B_KD - Control, 
                           UPF3A_vs_Control = UPF3A_KD - Control, 
                           UPF3A_OE_vs_Control = UPF3A_OE - Control, 
                           DoubleKD_vs_Control = UPF3A_KD_UPF3B_KD - Control, 
                           UPF3A_OE_UPF3B_KD_vs_Control = UPF3A_OE_UPF3B_KD - Control)
y <- calcNormFactors(y)
v <- voom(y, design)

fit = lmFit(v, design) %>% 
    contrasts.fit(contrasts) %>% 
    eBayes()
```


## Relative log expression (RLE) after normalisation

```{r r qc-rle, fig.height=4, fig.cap="Relative log expression (RLE) of voom-normalised data** per sample, faceted by cell line. RLE values centred on zero indicate successful normalisation, with no evidence of systematic sample-level bias within each group."}
v$E %>% 
  as.data.frame() %>% 
  pivot_longer(
    cols = everything(), names_to = "Sample", values_to = "logCPM"
  ) %>% 
  left_join(md) %>% 
  group_by(Sample) %>% 
  mutate(RLE = logCPM - median(logCPM))%>% 
  ggplot(aes(Sample, RLE, fill = Group)) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_wrap(~Group, scales = "free_x") +
  scale_fill_manual(values = group_cols) +
  labs(
    x = "Sample", fill = "Group"
  )
```

## Extract per-contrast results

```{r de-results}
res <- colnames(contrasts) %>%
  lapply(function(x) {
    topTable(fit, coef = x, number = Inf) %>%
      mutate(contrast = as.character(x)) %>%
      as.data.frame() %>%
      rownames_to_column("EnsID") %>%
      mutate(
        gene = mapIds(org.Mm.eg.db, keys = EnsID, column = "SYMBOL",
                       keytype = "ENSEMBL", multiVals = "first")
      )
  })
names(res) <- colnames(contrasts)

results <- decideTests(fit, p.value = 0.05, adjust.method = "fdr",
                        method = "global", lfc = log2(1.5))

limma_results <- write_fit(fit, results = results, adjust = "fdr") %>%
  rownames_to_column("ensembl_gene_id") %>%
  mutate(
    gene = mapIds(org.Mm.eg.db, keys = ensembl_gene_id, column = "SYMBOL",
                   keytype = "ENSEMBL", multiVals = "first")
  ) %>%
  drop_na(gene)
```

```{r}
library(cowplot)
library(readr)
res <- colnames(contrasts) %>%
        lapply(function(x){

            # For each contrast, we need to extract out a ranked list of genes for
            # fgsea, ranked by t-statistic for DE in that particular comparison.
            results <- topTable(fit, coef = x, number = Inf) %>%
              mutate(contrast = as.character(x)) %>%
             as.data.frame() %>%
  rownames_to_column("EnsID") %>%
  mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           genename = mapIds(org.Mm.eg.db, keys=EnsID,  column="GENENAME",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"),
           type =  mapIds(org.Mm.eg.db, keys=EnsID,  column="GENETYPE",keytype="ENSEMBL", multiVals="first"))  })


names(res) = colnames(contrasts)

DEGenes = lapply(res, function(x)
                 {
  x %>% dplyr::filter(adj.P.Val < 0.05 & abs(logFC) > log2(1.5))
                   })


names(DEGenes) = colnames(contrasts)
```

# Volcano plots

Genes were considered differentially expressed (DEGs) at FDR-adjusted p < 0.05 and |log2 fold-change| > log2(1.5).

## Combined volcano plot across conditions

```{r volcano-combined, fig.height=6, fig.width=12}
DEColours <- c("Downregulated" = "#53A08E", "Upregulated" = "#690020", "Not Sig" = "#CECCCA")

df <- res %>%
  do.call("rbind", .) %>%
  mutate(Result = ifelse(adj.P.Val < 0.05 & logFC > log2(1.5), "Upregulated",
                   ifelse(adj.P.Val < 0.05 & logFC < -log2(1.5), "Downregulated", "Not Sig"))) %>%
  dplyr::filter(contrast != "UPF3A_OE_vs_Control") %>%
  mutate(contrast = factor(contrast, levels = c(
    "UPF3B_vs_Control", "UPF3A_vs_Control",
    "DoubleKD_vs_Control", "UPF3A_OE_UPF3B_KD_vs_Control"
  )))

contrast_labeller <- as_labeller(c(
  "UPF3B_vs_Control" = "italic(Upf3b)~'KD DEGs'",
  "UPF3A_vs_Control" = "italic(Upf3a)~'KD DEGs'",
  "DoubleKD_vs_Control" = "italic(Upf3)~'dKD DEGs'",
  "UPF3A_OE_UPF3B_KD_vs_Control" = "italic(Upf3a)~'OE'~italic(Upf3b)~'KD DEGs'"
), label_parsed)

deg_volcano_plot <- df %>%
  ggplot(aes(x = logFC, y = -log10(adj.P.Val), colour = Result)) +
  geom_point(alpha = 0.75, size = 3) +
  scale_colour_manual(values = DEColours, breaks = c("Upregulated", "Downregulated"), drop = TRUE) +
  xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value"))) +
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), linetype = "dashed") +
  ylim(0, 10) +
  facet_grid(~contrast, scales = "free_x", labeller = contrast_labeller) +
  theme_bw() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    strip.text = element_text(size = 12)
  )

deg_volcano_plot 

```


# Data export


```{r}
#require(openxlsx)
#save(DEGenes,res, file=here::here("output/DEG-list.Rda"))
#save(limma_results, file=here::here("output/DEG-limma-results.Rda"))
#save(results, fit,y,v,contrasts, design, contrasts,  file = here::here("output/limma-matrices.Rda"))
#write.xlsx(DEGenes,file = here::here("output/DEG-results.xlsx"))

```
