---
title: "DEG-analysis"
author: "unawaz1996"
date: "2023-03-14"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	message = FALSE
)
```

```{r load}
source("code/plots.R") 
source("code/functions.R")
source("code/libraries.R")
library(biomaRt)
library(nVennR)
library(eulerr)
library(ggsci)
```


# Introduction 

# Count data 

Gene-level count data was retrieved from salmon files by `tximport` into a DGEList object. 
During this process, genes were removed if 

* They were not considered as detectable (TPM < 0.5 in > 3 samples)
* They did not contain any annotation 

These filtering steps returned gene-level counts for 10,957 genes. 

```{r}
# Quality control of genes 
salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Salmon")
md.files = "data/LTK_Sample Metafile_V3.txt"

# Loading data
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id")])

salmon = list.files(salmon.files, pattern = "transcripts$", full.names = TRUE)
all_files = file.path(salmon, "quant.sf")
names(all_files) <- paste(c(212:229))

txi_genes <- tximport(all_files, type="salmon", txOut=FALSE,
                      countsFromAbundance="scaledTPM", tx2gene = tx2gene, ignoreTxVersion = TRUE, ignoreAfterBar = TRUE)


md = read.table(md.files, header= TRUE) %>%  mutate(files = file.path(salmon, "quant.sf"))
md$Group = factor(md$Group)
md$Sample = as.character(md$Sample)
```


```{r}
group_cols <- hcl.colors(
  n = length(unique(md$Group)), 
  palette = "Zissou 1"
  ) %>%
  setNames(unique(md$Group))
```



```{r fig.height = 4, fig.cap="Expression density plots for all samples before and after filtering, showing logCPM values."}
keep = (rowSums(txi_genes$abundance >= .5 ) >= 3)

exp_den = txi_genes$counts %>% cpm(log=TRUE) %>%
    reshape2::melt() %>%
    dplyr::filter(is.finite(value)) %>% 
    ggplot(aes(x=value, color = as.character(Var2))) +
    geom_density() +
    ggtitle("Before filtering") +
    labs(x = "logCPM", y = "Proportion of Genes") + 
    theme_bw() +
    theme(legend.position = "none")

exp_den_filt = txi_genes$counts %>%
    cpm(log=TRUE) %>%
    magrittr::extract(keep,) %>% 
    reshape2::melt() %>%
    dplyr::filter(is.finite(value)) %>% 
    ggplot(aes(x=value, color = as.character(Var2))) +
    geom_density() +
    ggtitle("After filtering") +
    labs(x = "logCPM", y = "Proportion of Genes") + theme_bw() +
    theme(legend.position = "none")

grid.arrange(exp_den,exp_den_filt, ncol=2 )

```


```{r fig.height= 4}
txi_genes_filtered = txi_genes$counts[keep,] %>%  
    as.data.frame() %>% 
    rownames_to_column("geneID") %>%
    mutate(geneName = mapIds(org.Mm.eg.db, keys=.$geneID,
                             column="SYMBOL",keytype="ENSEMBL",
                         multiVals="first")) %>%
    drop_na(geneName) %>% 
    column_to_rownames("geneID") %>% dplyr::select(-geneName)


```

# Exploratory Analysis and quality control

- PCA 

```{r fig.height=4.5}
pca = log2(txi_genes$abundance[keep,]  + 10^-6 ) %>%
    t() %>% 
    prcomp(scale = TRUE)

pcaVars <- percent_format(0.1)(summary(pca)$importance["Proportion of Variance",]) 

gene_pca = pca$x[,1:2] %>% 
    as.data.frame()  %>%
    cbind(md) %>%
    as.tibble() %>%
    ggplot(aes(PC1, PC2, shape = Condition_UPF3B, label = Condition_UPF3B, color=Group)) +
    geom_point(size = 11.5) +
    theme_bw() + scale_shape_manual(values = c(0,16)) + labs(shape="UPF3B status", colour= "Samples") +
    theme(axis.title.y = element_text(size = 12), axis.title.x = element_text(size = 12), 
          axis.text = element_text(size = 10)) +
    xlab(paste0("Principal Component 1 (28.1%)")) + ylab("Principal Component 2 (16.6%)") +
    scale_color_manual(values = c("#FCC98A","#6E74B4", "#E12F8F","#EA2A5F", "#FD9675", "#2F124B"),
                      labels = c("UPF3A_KD_UPF3B_KD" = "Double KD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  labs(x = paste0("PC1 (", pcaVars[["PC1"]], ")"),
    y = paste0("PC2 (", pcaVars[["PC2"]], ")"))

#ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/Gene_PCA.svg", plot = gene_pca, 
#       units = c("in"), width = 6.65, height = 5.60)

gene_pca
```

The principal component analysis was performed using the logCPM values from each sample. Each group appears to be clustering together. UPF3A over-expression is clustering with Controls, whereas the UPF3B KD samples are clustering closely with UPF3A overexpression in UPF3B KD files. 

- Boxplots of gene expression and library size 

```{r fig.height =4}
txi_genes$abundance[keep,] %>% 
    as.data.frame() %>% 
    gather(sample,value) %>% 
    ggplot(aes(x=sample, y = log(value), fill = sample)) + geom_boxplot() +
    ylab("Log(TPM)") + xlab("Samples") +
    theme_bw() + theme(legend.position = "none") + scale_fill_manual(values = hydrangea[1:18])
```

- Library size
```{r fig.height = 3}
# Library size 

countsSum <- colSums(txi_genes$counts) # library size
countsSum %>% reshape2::melt() %>% rownames_to_column("Sample") %>% 
    ggplot(aes(x=Sample, y = value)) + geom_bar(stat = "identity") + ylab("Library size") +
    theme_bw()

#max(countsSum) / min(countsSum)

```


 - Heatmap of NMD substrates
 
```{r}
genes_of_interest = c("Upf3b", "Upf3a", "Ire1", 
                      "Smad5", "Smad7", "Ttyh1", "Gas5", 
                      "Snord22", "Upf2", "Gadd45g")
heatmap = txi_genes$abundance  %>%
  as.data.frame() %>%
  rownames_to_column("gene_id") %>% 
  mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%
  dplyr::select(-c("gene_id")) %>%
  dplyr::filter(gene %in% genes_of_interest) %>% 
  column_to_rownames("gene")
  

log(heatmap + 0.05) %>% 
  as.data.frame() %>%
  dplyr::select("212", "218", "224",
               # "215", "221", "227", 
                #"213", "219", "225",
                #"216", "222", "228", 
                #"214", "220", "226", 
                "217", "223", "229") %>% 
  pheatmap(
           cluster_cols = FALSE) 
 
```


# Differential gene expression using limma 

## Model description 

In this analysis, we are interested in seeing how the knockdowns behave differently to controls. Given the quality and clustering of the data, we will perform five pairwise comparisons using `limma` where all conditions (KD/OE) will be compared to controls. 

```{r}
y <- DGEList(txi_genes_filtered)


design <- model.matrix(~0+Group, data = md) %>% 
    set_colnames(gsub(pattern = "Group", replacement ="", x = colnames(.)))

contrasts <- makeContrasts(levels = colnames(design), 
                           UPF3B_vs_Control = UPF3B_KD - Control, 
                           UPF3A_vs_Control = UPF3A_KD - Control, 
                           UPF3A_OE_vs_Control = UPF3A_OE - Control, 
                           DoubleKD_vs_Control = UPF3A_KD_UPF3B_KD - Control, 
                           UPF3A_OE_UPF3B_KD_vs_Control = UPF3A_OE_UPF3B_KD - Control)
```


```{r}
y <- calcNormFactors(y)
v <- voom(y, design)

fit = lmFit(v, design) %>% 
    contrasts.fit(contrasts) %>% 
    eBayes()
```

```{r}
v$E %>% 
  as.data.frame() %>% 
  pivot_longer(
    cols = everything(), names_to = "Sample", values_to = "logCPM"
  ) %>% 
  left_join(md) %>% 
  group_by(Sample) %>% 
  mutate(RLE = logCPM - median(logCPM))%>% 
  ggplot(aes(Sample, RLE, fill = Group)) +
  geom_boxplot(alpha = 0.9) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_wrap(~Group, scales = "free_x") +
  scale_fill_manual(values = group_cols) +
  labs(
    x = "Sample", fill = "Group"
  )
```


## Results

```{r}
library(cowplot)
library(readr)
res <- colnames(contrasts) %>%
        lapply(function(x){

            # For each contrast, we need to extract out a ranked list of genes for
            # fgsea, ranked by t-statistic for DE in that particular comparison.
            results <- topTable(fit, coef = x, number = Inf) %>%
              mutate(contrast = as.character(x)) %>%
             as.data.frame() %>%
  rownames_to_column("EnsID") %>%
  mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           genename = mapIds(org.Mm.eg.db, keys=EnsID,  column="GENENAME",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"),
           type =  mapIds(org.Mm.eg.db, keys=EnsID,  column="GENETYPE",keytype="ENSEMBL", multiVals="first"))  })


names(res) = colnames(contrasts)

DEGenes = lapply(res, function(x)
                 {
  x %>% dplyr::filter(adj.P.Val < 0.05 & abs(logFC) > log2(1.5))
                   })


names(DEGenes) = colnames(contrasts)
```



<!-- # ```{r} -->
<!-- # # DEGenes$UPF3B_vs_Control %>% -->
<!-- # #   with(., -->
<!-- # #    table(sign.lfc=sign(logFC))) %>% -->
<!-- # #   as.data.frame() -->
<!-- #  -->
<!-- #  -->
<!-- # summary_stats= lapply(DEGenes, function(x){ -->
<!-- # x %>% -->
<!-- #   with(., -->
<!-- #    table(sign.lfc=sign(logFC))) %>% -->
<!-- #   as.data.frame() -->
<!-- # }) -->
<!-- #  -->
<!-- # results_summary = do.call("rbind", summary_stats) %>% -->
<!-- #   rownames_to_column("comparison") %>% -->
<!-- #   mutate_at(.vars = "comparison", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") -->
<!-- #  -->
<!-- #  -->
<!-- # comparisons = c("UPF3B_vs_Control","UPF3A_vs_Control"        , "UPF3A_OE_vs_Control", "DoubleKD_vs_Control"         , "UPF3A_OE_UPF3B_KD_vs_Control") -->
<!-- #  -->
<!-- # ratios = data.frame() -->
<!-- # for (c in comparisons[-3]){ -->
<!-- #   temp = results_summary %>% -->
<!-- #     dplyr::filter(comparison == c) -->
<!-- #   ratio= temp$Freq[2]/temp$Freq[1] -->
<!-- #   ratio = data.frame(Comparisons = c, -->
<!-- #            ratio = ratio) -->
<!-- #   ratios = rbind(ratios, ratio) -->
<!-- # } -->
<!-- #  -->
<!-- # ratio_plot = ratios  %>% -->
<!-- #   ggplot(aes(x =Comparisons, y = ratio)) + -->
<!-- #   geom_bar(stat = "identity",   colour = "white", width = .75, fill = "#e64055") + -->
<!-- #     theme_classic() + -->
<!-- #     theme(axis.text.x = element_text(angle = 70,  hjust=1, face = "bold"), -->
<!-- #           legend.position = "none", -->
<!-- #           legend.title= element_blank(), axis.title.y = element_text(size = 12), -->
<!-- #           axis.text.y = element_text(size = 14, face = "bold")) + -->
<!-- #     scale_fill_manual(values = c("#2F124B","#EA2A5F","#6E74B4","#E12F8F", "#FD9675")) + -->
<!-- #     ylab("ratio up/down") + xlab("") + scale_x_discrete(labels=c("DoubleKD_vs_Control" = "Double KD", "UPF3A_vs_Control" = "UPF3A KD", -->
<!-- #                                                             "UPF3B_vs_Control" = "UPF3B KD", -->
<!-- #                                                             "UPF3A_OE_UPF3B_KD_vs_Control" = "UPF3A OE UPF3B KD")) -->
<!-- #  -->
<!-- #  -->
<!-- #  -->
<!-- # #ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/DEG_ratios.svg", plot = ratio_plot,  units = c("in"), height = 4, width = 4) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # summary_DEG = summary(res) %>% reshape2::melt() %>% -->
<!-- #     dplyr::filter(Var1 != "NotSig") -->
<!-- #  -->
<!-- #  -->
<!-- # summary_DEG$Var1 = factor(summary_DEG$Var1, levels = c("Up", "Down")) -->
<!-- # summary_DEG$Var2 = factor(summary_DEG$Var2, levels = c("UPF3B_vs_Control", -->
<!-- #                                                        "UPF3A_vs_Control", -->
<!-- #                                                        "DoubleKD_vs_Control", -->
<!-- #                                                        "UPF3A_OE_vs_Control", -->
<!-- #                                                        "UPF3A_OE_UPF3B_KD_vs_Control")) -->
<!-- #  -->
<!-- # summary_DEG %<>% -->
<!-- #   mutate(Var2= factor(summary_DEG$Var2, levels = c("UPF3A_OE_UPF3B_KD_vs_Control", -->
<!-- #                                                "UPF3A_OE_vs_Control", -->
<!-- #                                                "DoubleKD_vs_Control", -->
<!-- #                                                "UPF3A_vs_Control", -->
<!-- #                                                "UPF3B_vs_Control"))) -->
<!-- # up = summary_DEG %>% -->
<!-- #   dplyr::filter(Var1 == "Up") -->
<!-- # down = summary_DEG %>% -->
<!-- #   dplyr::filter(Var1=="Down") -->
<!-- #  -->
<!-- #  -->
<!-- # summary_dot_plot = summary_DEG %>% -->
<!-- #   mutate(X2= factor(summary_DEG$Var2, levels = c("UPF3A_OE_UPF3B_KD_vs_Control", -->
<!-- #                                                "UPF3A_OE_vs_Control", -->
<!-- #                                                "DoubleKD_vs_Control", -->
<!-- #                                                "UPF3A_vs_Control", -->
<!-- #                                                "UPF3B_vs_Control"))) %>% -->
<!-- # ggplot() + geom_segment(data = down, aes(x=value, y=Var2, -->
<!-- #                                        yend=up$Var2, xend= up$value), -->
<!-- #                                        color = "black", -->
<!-- #               size = 0.7) + -->
<!-- #  -->
<!-- #   geom_point(aes(x = value, y = Var2, color = Var1, size = value)) + -->
<!-- #   theme_bw() + -->
<!-- #         theme( -->
<!-- #               strip.background =element_rect(fill="#AA9A9C", color = "#E1DFDB"), -->
<!-- #               panel.border = element_rect(color = "#E1DFDB", fill = NA, size = 1)) + -->
<!-- #   scale_color_manual(values= c("#C32137", "#4657A6")) + xlim(0,1000) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # ggarrange(summary_plot_poster, ratio_plot, ncol=2) -->
<!-- # ``` -->

<!-- ## Volcano plots  -->

<!-- After performing pairwise differential gene expression analysis, we can see that majority of all significantly expressed genes in all contrasts are up-regulated. Overall, we find that a total of 395 genes are differentially expressed between UPF3B KD and controls. 388 genes are differentially expressed between UPF3A KD vs Controls. -->

<!-- We now want to explore the genes from each of these analyses and identify similarities and differences of expression. -->


<!-- ```{r} -->
<!-- DEColours <- c("Downregulated" = "#53A08E","Upregulated" = "#690020",  "Not Sig" = "#CECCCA") -->
<!-- upf3b_KD = res$UPF3B_vs_Control %>%  -->
<!--     mutate(Result = ifelse(adj.P.Val < 0.05 & logFC > log2(1.5), "Upregulated",         -->
<!--                ifelse(adj.P.Val < 0.05 & abs(logFC) > log2(1.5), "Downregulated", "Not Sig"))) %>% -->
<!--     ggplot(aes(x= logFC, -->
<!--                y = -log10(adj.P.Val), -->
<!--                colour = as.factor(as.character(Result )))) + -->
<!--     geom_point(alpha = 0.75, size = 5) + -->
<!--     scale_colour_manual(values = DEColours) + theme_bw() + -->
<!--   theme(legend.position = "none",  -->
<!--         axis.text = element_text(family = "serif", size =14, color = "black")) +  -->
<!--   geom_vline(xintercept = c(-log2(1.5),log2(1.5)), linetype = "dashed") + xlab(bquote(log["2"](FC))) + -->
<!--   ylab(bquote(log["10"]("adjusted P-value"))) + ylim(0, 10) -->

<!-- ggsave(plot = upf3b_KD, filename = here::here("output/Thesis_plots/DEGs/Upf3b_KD_volcano.png"), -->
<!--        dpi =1200, height = 6, width =5.5) -->


<!-- upf3a_KD = res$UPF3A_vs_Control %>%  -->
<!--     mutate(Result = ifelse(adj.P.Val < 0.05 & logFC > log2(1.5), "Upregulated",         -->
<!--                ifelse(adj.P.Val < 0.05 & abs(logFC) > log2(1.5), "Downregulated", "Not Sig"))) %>% -->
<!--     ggplot(aes(x= logFC, -->
<!--                y = -log10(adj.P.Val), -->
<!--                colour = as.factor(as.character(Result )))) + -->
<!--     geom_point(alpha = 0.75, size = 5) + -->
<!--     scale_colour_manual(values = DEColours) + theme_bw() + -->
<!--   theme(legend.position = "none",  -->
<!--         axis.text = element_text(family = "serif", size =14, color = "black")) +  -->
<!--   geom_vline(xintercept = c(-log2(1.5),log2(1.5)), linetype = "dashed") + xlab(bquote(log["2"](FC))) + -->
<!--   ylab(bquote(log["10"]("adjusted P-value"))) + ylim(0, 10) -->

<!-- ggsave(plot = upf3a_KD, filename = here::here("output/Thesis_plots/DEGs/Upf3a_KD_volcano.png"), -->
<!--        dpi =1200, height = 6, width =5.5) -->

<!-- upf3_dKD = res$DoubleKD_vs_Control %>%  -->
<!--     mutate(Result = ifelse(adj.P.Val < 0.05 & logFC > log2(1.5), "Upregulated",         -->
<!--                ifelse(adj.P.Val < 0.05 & abs(logFC) > log2(1.5), "Downregulated", "Not Sig"))) %>% -->
<!--     ggplot(aes(x= logFC, -->
<!--                y = -log10(adj.P.Val), -->
<!--                colour = as.factor(as.character(Result )))) + -->
<!--     geom_point(alpha = 0.75, size = 5) + -->
<!--     scale_colour_manual(values = DEColours) + theme_bw() + -->
<!--   theme(legend.position = "none",  -->
<!--         axis.text = element_text(family = "serif", size =14, color = "black")) +  -->
<!--   geom_vline(xintercept = c(-log2(1.5),log2(1.5)), linetype = "dashed") + xlab(bquote(log["2"](FC))) + -->
<!--   ylab(bquote(log["10"]("adjusted P-value"))) + ylim(0, 10) -->

<!-- ggsave(plot = upf3_dKD, filename = here::here("output/Thesis_plots/DEGs/Upf3_dKD_volcano.png"), -->
<!--        dpi =1200, height = 6, width =5.5) -->


<!-- upf3oe_KD = res$UPF3A_OE_UPF3B_KD_vs_Control %>%  -->
<!--     mutate(Result = ifelse(adj.P.Val < 0.05 & logFC > log2(1.5), "Upregulated",         -->
<!--                ifelse(adj.P.Val < 0.05 & abs(logFC) > log2(1.5), "Downregulated", "Not Sig"))) %>% with(table(.$Result)) -->
<!--     ggplot(aes(x= logFC, -->
<!--                y = -log10(adj.P.Val), -->
<!--                colour = as.factor(as.character(Result )))) + -->
<!--     geom_point(alpha = 0.75, size = 5) + -->
<!--     scale_colour_manual(values = DEColours) + theme_bw() + -->
<!--   theme(legend.position = "none",  -->
<!--         axis.text = element_text(family = "serif", size =14, color = "black")) +  -->
<!--   geom_vline(xintercept = c(-log2(1.5),log2(1.5)), linetype = "dashed") + xlab(bquote(log["2"](FC))) + -->
<!--   ylab(bquote(log["10"]("adjusted P-value"))) + ylim(0, 10) -->

<!-- ggsave(plot = upf3oe_KD , filename = here::here("output/Thesis_plots/DEGs/Upf3aOE_Upf3bKD_volcano.png"), -->
<!--        dpi =1200, height = 6, width =5.5) -->


<!-- ``` -->


<!-- ## Overlaps between all the DEGs  -->

<!-- ```{r fig.cap="Upset plot showing overlapping DEGs from each comparison"} -->
<!-- library(ComplexUpset) -->
<!-- library(ggplot2) -->

<!-- x=list("UPF3B KD" = rownames(DEGenes$UPF3B_vs_Control),  -->
<!--                        "UPF3A KD" = rownames(DEGenes$UPF3A_vs_Control), -->
<!--                     "UPF3 dKD" = rownames(DEGenes$DoubleKD_vs_Control), -->
<!--                     "UPF3A OE" = rownames(DEGenes$UPF3A_OE_vs_Control),  -->
<!--                   "UPF3A OE, UPF3B KD" = rownames(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control)) -->


<!-- ggvenn(x[c(1,5)]) -->

<!-- UpSetR::upset(fromList(x)) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- for_upf3b = colnames(contrasts)[-c(1,3)] -->

<!-- overlap_upf3b = list() -->
<!-- for (c in for_upf3b ){ -->
<!--   overlap = DEGenes$UPF3B_vs_Control %>%  -->
<!--     inner_join(DEGenes[[c]] , by="EnsID") %>%  -->
<!--       mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant"))  -->
<!--   total = table(overlap$Label) %>% as.data.frame()  -->
<!--   overlap_upf3b[[paste0(c, "_vs_Upf3b_KD")]] = total -->

<!-- } -->

<!-- for_upf3a = colnames(contrasts)[c(4:5)] -->
<!-- overlap_upf3a = list() -->
<!-- for (c in for_upf3a ){ -->
<!--   overlap = DEGenes$UPF3A_vs_Control %>%  -->
<!--     inner_join(DEGenes[[c]], by="EnsID") %>%  -->
<!--       mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant"))  -->
<!--   total = table(overlap$Label) %>% as.data.frame()  -->
<!--   overlap_upf3a[[paste0(c, "_vs_Upf3A_KD")]] = total -->

<!-- } -->


<!-- for_dkDupf3 = colnames(contrasts)[c(5)] -->
<!-- overlap_dkDupf3 = list() -->
<!-- for (c in for_dkDupf3 ){ -->
<!--   overlap = DEGenes$DoubleKD_vs_Control %>%  -->
<!--     inner_join(DEGenes[[c]], by="EnsID") %>%  -->
<!--       mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant"))  -->
<!--   total = table(overlap$Label) %>% as.data.frame()  -->
<!--   overlap_dkDupf3 [[paste0(c, "_vs_Upf3_dKD")]] = total -->

<!-- } -->



<!-- overlap_upf3b =do.call("rbind",overlap_upf3b) %>%  -->
<!--   rownames_to_column("Comparisons") %>%  -->
<!--    mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")  -->


<!-- overlap_upf3a =do.call("rbind",overlap_upf3a) %>%  -->
<!--   rownames_to_column("Comparisons") %>%  -->
<!--    mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")  -->


<!-- overlap_dkDupf3  =do.call("rbind",overlap_dkDupf3 ) %>%  -->
<!--   rownames_to_column("Comparisons") %>%  -->
<!--    mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")  -->


<!-- all_comparisons = rbind(overlap_upf3b,overlap_upf3a, overlap_dkDupf3) %>%  -->
<!--   mutate(Var1 = factor(Var1, levels = c("Discordant", "Concordance" )),  -->
<!--          Comparisons = factor(Comparisons, levels = c("UPF3A_OE_UPF3B_KD_vs_Control_vs_Upf3_dKD",  -->
<!--                                                       "UPF3A_OE_UPF3B_KD_vs_Control_vs_Upf3A_KD", -->
<!--                                                        "DoubleKD_vs_Control_vs_Upf3A_KD",  -->
<!--                                                       "UPF3A_OE_UPF3B_KD_vs_Control_vs_Upf3b_KD", -->
<!--                                                         "DoubleKD_vs_Control_vs_Upf3b_KD", -->
<!--                                                       "UPF3A_vs_Control_vs_Upf3b_KD" -->
<!--                                                       ) )) %>% -->
<!--   ggplot(aes(x = Comparisons, y = Freq, fill = Var1)) +  -->
<!--   geom_bar(stat= "identity", width= 0.65) + coord_flip() + scale_fill_manual(values = c( "#BAB8B5","#EB635E")) + -->
<!--   theme_bw() + -->
<!--      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=10, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(y= "Number of Overlapping transcripts", -->
<!--                   fill = "Expression",  -->
<!--                   x= "") -->
<!-- all_comparisons  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps.svg", plot = all_comparisons,  -->
<!--        units = c("in"), height = 7, width = 9.26) -->
<!-- ``` -->

<!-- Significance of overlap and lists  -->
<!-- ```{r} -->

<!-- library(GeneOverlap) -->

<!-- enrichment_DEGs_P = list() -->
<!-- enrichment_DEGs_OR = list() -->


<!-- test.list = list() -->
<!-- test.list$UPF3B_vs_Control = DEGenes$UPF3B_vs_Control$EnsID -->
<!-- test.list$UPF3A_vs_Control = DEGenes$UPF3A_vs_Control$EnsID -->
<!-- test.list$DoubleKD_vs_Control = DEGenes$DoubleKD_vs_Control$EnsID -->
<!-- test.list$UPF3A_OE_UPF3B_KD_vs_Control = DEGenes$UPF3A_OE_UPF3B_KD_vs_Control$EnsID -->
<!-- test.list$BG = res$UPF3B_vs_Control$EnsID -->

<!-- for_upf3b = colnames(contrasts)[-c(1,3)] -->

<!-- for (c in for_upf3b ){ -->

<!--   go.obj <- newGeneOverlap(test.list$UPF3B_vs_Control,  -->
<!--                          test.list[[c]], -->
<!--                          genome.size=length(test.list$BG)) -->
<!--   go.obj <- testGeneOverlap(go.obj) -->
<!--   enrichment_DEGs_P[[paste0("Upf3b_KD", "_", c)]] = go.obj@pval -->
<!--   enrichment_DEGs_OR[[paste0("Upf3b_KD", "_", c)]] = go.obj@odds.ratio  -->
<!-- } -->

<!-- for_upf3a = colnames(contrasts)[c(4:5)] -->

<!-- for (c in for_upf3a ){ -->
<!--     go.obj <- newGeneOverlap(test.list$UPF3A_vs_Control,  -->
<!--                          test.list[[c]], -->
<!--                          genome.size=length(test.list$BG)) -->
<!--   go.obj <- testGeneOverlap(go.obj) -->
<!--   enrichment_DEGs_P[[paste0("Upf3a_KD", "_", c)]] = go.obj@pval -->
<!--   enrichment_DEGs_OR[[paste0("Upf3a_KD", "_", c)]] = go.obj@odds.ratio  -->
<!-- } -->


<!-- for_dkDupf3 = colnames(contrasts)[c(5)] -->
<!-- for (c in for_dkDupf3 ){ -->
<!--       go.obj <- newGeneOverlap(test.list$DoubleKD_vs_Control,  -->
<!--                          test.list[[c]], -->
<!--                          genome.size=length(test.list$BG)) -->
<!--   go.obj <- testGeneOverlap(go.obj) -->
<!--   enrichment_DEGs_P[[paste0("Upf3_dKD", "_", c)]] = go.obj@pval -->
<!--   enrichment_DEGs_OR[[paste0("Upf3_dKD", "_", c)]] = go.obj@odds.ratio  -->

<!-- } -->

<!-- odds_ratio = enrichment_DEGs_OR %>% do.call(rbind,.) %>%  -->
<!--   set_colnames("OR") %>%  -->
<!--   cbind(enrichment_DEGs_P %>% do.call(rbind,.) %>% set_colnames("pvalue")) %>%  -->
<!--   as.data.frame() %>% -->
<!--   rownames_to_column("comparisons") %>% -->
<!--   mutate(comparsions = factor(comparisons, levels = c("Upf3b_KD_UPF3A_vs_Control",  -->
<!--                                                       "Upf3b_KD_DoubleKD_vs_Control",  -->
<!--                                                       "Upf3b_KD_UPF3A_OE_UPF3B_KD_vs_Control", -->
<!--                                                       "Upf3a_KD_DoubleKD_vs_Control",  -->
<!--                                                       "Upf3a_KD_UPF3A_OE_UPF3B_KD_vs_Control",  -->
<!--                                                       "Upf3_dKD_UPF3A_OE_UPF3B_KD_vs_Control"))) %>% -->
<!--   ggplot(aes(x=OR, y = comparisons)) + geom_bar(stat = "identity", fill = "#edafb8") + -->
<!--   theme_bw() + geom_vline(xintercept = 1) +  -->
<!--    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=12, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(x="Odds ratio") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps_enrichment.svg", plot = odds_ratio,  -->
<!--        units = c("in"), height = 7, width = 9.26) -->

<!-- ``` -->



<!-- ### Upf3a and Upf3b  -->


<!-- ```{r} -->
<!-- cor_upf3 =  DEGenes$UPF3B_vs_Control %>%  -->

<!--   inner_join(DEGenes$UPF3A_vs_Control,  by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.75, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3a KD (LFC)") + xlab("Upf3b KD (LFC)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#EB635E", "#BAB8B5")) +   -->
<!--   theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 13, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3_paralog_cor.png", plot = cor_upf3 ,  -->
<!--        units = c("in"), height = 4.65, width = 4, dpi = 1200) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- x = list("Upf3b KD" = DEGenes$UPF3B_vs_Control$EnsID,  -->
<!--          "Upf3a KD" = DEGenes$UPF3A_vs_Control$EnsID,  -->
<!--          "Upf3 dKD"= DEGenes$DoubleKD_vs_Control$EnsID) -->

<!-- ggvenn(x, show_percentage = FALSE,  -->
<!--         fill_color = c("#1c174a", "#ef476f", "#7175b3")) -->


<!-- ``` -->


<!-- ```{r} -->
<!-- DEGenes$UPF3B_vs_Control %>%  -->

<!--   inner_join(DEGenes$UPF3A_vs_Control,  by = "EnsID") %>%  -->
<!--   dplyr::filter(logFC.x > 0 & logFC.y >0) %>% dim() -->

<!-- ``` -->

<!-- ```{r} -->
<!-- cor_genes = DEGenes$UPF3B_vs_Control %>%  -->

<!--   inner_join(DEGenes$UPF3A_vs_Control,  by = "EnsID") -->
<!-- ``` -->

<!-- ```{r fig.cap="Cumulative frequency of unique genes"} -->
<!-- ks.test(upf3b$logFC, upf3a$logFC) -->

<!-- cumulative_freq =  DEGenes$UPF3B_vs_Control[rownames(DEGenes$UPF3B_vs_Control) %in% rownames(DEGenes$UPF3A_vs_Control),] %>% -->
<!--   mutate(Condition = "Upf3b genes") %>%  -->
<!--   rbind(DEGenes$UPF3A_vs_Control[rownames(DEGenes$UPF3A_vs_Control) %in% rownames(DEGenes$UPF3B_vs_Control),] %>%  -->
<!--   mutate(Condition = "Upf3a genes")) %>%  -->
<!--   ggplot(aes(x=logFC, color = Condition)) + stat_ecdf(geom = "step", size=1) + -->
<!--   scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() + -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") + -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") + -->
<!--   ggtitle("148 overlapping genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") +  -->
<!--       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         ))  +  -->
<!--    annotate("point", x = 0.819, y = 0, colour = "#2F124B") + -->
<!--   annotate("point", x = 0.8886537, y = 0, colour = "#EA2A5F") + -->
<!--   annotate("text", x=-1.5, y = 0.85, label = "D = 0.126,  p-value = 0.029", fontface =7) -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3_cumulative.svg", plot =cumulative_freq  ,  -->
<!--        units = c("in"), height = 5.65, width = 5) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- cor_upf3 =  DEGenes$UPF3B_vs_Control %>%  -->

<!--   inner_join(DEGenes$DoubleKD_vs_Control,  by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.75, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3 dKD (LFC)") + xlab("Upf3b KD (LFC)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#EB635E", "#BAB8B5")) +   -->
<!--   theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 13, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b_Upf3dKD_cor.png", plot = cor_upf3 ,  -->
<!--        units = c("in"), height = 4.65, width = 4, dpi = 1200) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- cor_upf3 =  DEGenes$UPF3A_vs_Control %>%  -->

<!--   inner_join(DEGenes$DoubleKD_vs_Control,  by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.75, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3 dKD (LFC)") + xlab("Upf3a KD (LFC)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#EB635E", "#BAB8B5")) +   -->
<!--   theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 13, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a_Upf3dKD_cor.png", plot = cor_upf3 ,  -->
<!--        units = c("in"), height = 4.65, width = 4, dpi = 1200) -->
<!-- ``` -->



<!-- ### Upf3a OE Upf3b KD -->

<!-- ```{r} -->
<!-- Upf3a_oR_cor = DEGenes$UPF3B_vs_Control %>%  -->
<!--   #rownames_to_column("EnsID") %>%  -->

<!--   inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control, #%>%  -->
<!--               # rownames_to_column("EnsID"), -->
<!--              by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordant", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.75, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3a OE- Upf3b KD (LFC)") + xlab("Upf3b KD (LFC)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#EB635E", "#BAB8B5")) +   -->
<!--   theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 13, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->

<!-- Upf3a_oR_cor -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a_OE_Upf3b_KD_cor.png", plot = Upf3a_oR_cor ,  -->
<!--        units = c("in"), height = 4.65, width = 4, dpi = 1200) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- ## QC plots - deciding log2FC thresholds  -->
<!-- results <- decideTests(fit, p.value = 0.05, adjust.method = "fdr", method= "global") -->
<!-- limma_results = write_fit(fit, results= results, adjust = "fdr") %>%  -->
<!--     rownames_to_column("ensembl_gene_id") %>%  -->
<!--     mutate(gene = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),  -->
<!--            genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first"),  -->
<!--            entrez = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"),  -->
<!--            type =  mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENETYPE",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez) -->


<!-- gene_expression_density = data.frame() -->
<!-- comparisons =c("UPF3B_vs_Control", "UPF3A_vs_Control",  "UPF3A_OE_vs_Control",  -->
<!--                "DoubleKD_vs_Control", "UPF3A_OE_UPF3B_KD_vs_Control")  -->

<!-- for (c in comparisons) { -->
<!--   sig_genes = limma_results %>% as.data.frame() %>%  -->
<!--     dplyr::select(contains("Coef"), contains ("p.value.adj"), contains(c)) %>% -->
<!--   dplyr::filter(p.value.adj.UPF3B_vs_Control < 0.05) %>% -->
<!--   dplyr::select(contains("Coef")) %>% -->
<!--   dplyr::select(contains("UPF3B_vs_Control")) %>% reshape2::melt() -->
<!--   gene_expression_density = rbind(gene_expression_density, sig_genes ) -->

<!-- } -->





<!-- #ggplot(aes(x=Coef.UPF3B_vs_Control)) +  -->
<!--  # geom_histogram(aes(y=..density..), colour="black", fill="white") +   -->
<!--  # geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept =c(log2(1.5), -log2(1.5)),color="red", linetype = "dashed") +  -->
<!-- #  theme_bw() + ylab("Density") -->



<!-- ``` -->


<!-- ### Upf2 levels -->

<!-- ```{r} -->
<!-- upf2_exp = res %>% -->
<!--   do.call(rbind, .) %>%  -->
<!--    rownames_to_column("comparison") %>%  -->
<!--   mutate_at(.vars = "comparison", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% -->
<!--   dplyr::filter(EnsID== "ENSMUSG00000043241") %>%  -->

<!--   mutate(Significance = c("Not significant")) %>% -->
<!--   mutate(comparison = factor(comparison, levels = c("UPF3B_vs_Control", -->
<!--                                                 "UPF3A_vs_Control", -->
<!--                                                 "DoubleKD_vs_Control",  -->
<!--                                                 "UPF3A_OE_vs_Control", -->
<!--                                                 "UPF3A_OE_UPF3B_KD_vs_Control"))) %>% -->
<!--   ggplot(aes(x=logFC, y=comparison, shape=Significance, color= Significance)) + -->
<!--     geom_point(stat="identity", size =8) + -->
<!--  xlab(bquote(log["2"](FC))) + theme_bw() + -->
<!--   theme(axis.text.y = element_text(size=10),  -->
<!--         axis.text.x = element_text(size=10)) +  -->
<!--    scale_color_manual(values=c("#adc178", "#FD9675")) + -->
<!--   geom_vline(xintercept = 0, color="black", linetype = "dashed") + -->
<!--   scale_y_discrete(labels=c("DoubleKD_vs_Control" = "Upf3 dKD",  -->
<!--                             "UPF3A_vs_Control" = "Upf3a KD",  -->
<!--                             "UPF3B_vs_Control" = "Upf3b KD",  -->
<!--                             "UPF3A_OE_vs_Control" = "Upf3a OE", -->
<!--                             "UPF3A_OE_UPF3B_KD_vs_Control" = "Upf3a OE Upf3b KD"),  -->
<!--                    limits =rev) + -->
<!--   ggtitle("Upf2 expression") + ylab("") +  -->
<!--     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20), -->
<!--               legend.position = "top", -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         ))  -->


<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf2.svg", plot = upf2_exp,  -->
<!--        units = c("in"), height = 6.65, width = 8) -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf2_exp.png", plot =upf2_exp ,  -->
<!--        units = c("in"), height = 5, width = 6, dpi = 1200) -->
<!-- ``` -->






```{r}
results <- decideTests(fit, p.value = 0.05, adjust.method = "fdr", method= "global", 
                       lfc = log2(1.5))
limma_results = write_fit(fit, results= results, adjust = "fdr") %>% 
    rownames_to_column("ensembl_gene_id") %>% 
    mutate(gene = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"), 
           genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first"), 
           entrez = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"), 
           type =  mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENETYPE",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


#save(limma_results, file="/home/neuro/Documents/NMD_analysis/Analysis/Results/DEGs/DEG-limma-results.Rda")
#save(results, fit,y,v,contrasts, design, contrasts,  file = "/home/neuro/Documents/NMD_analysis/Analysis/Results/DEGs/limma-matrices.Rda")

```



```{r}
comparisons = c("UPF3B_vs_Control","UPF3A_vs_Control" , "UPF3A_OE_vs_Control", "DoubleKD_vs_Control"         , "UPF3A_OE_UPF3B_KD_vs_Control")

ratios = data.frame()
for (c in comparisons){
  temp = summary(results) %>% 
    reshape2::melt() %>% 
    dplyr::filter(Var2 == c, Var1 != "NotSig") 
  ratio= temp$value[2]/temp$value[1]
  ratio = data.frame(Comparisons = c, 
           ratio = ratio)
  ratios = rbind(ratios, ratio)
  }

ratios$Comparisons = factor(ratios$Comparisons, levels = c("UPF3B_vs_Control", 
                                                           "UPF3A_vs_Control", 
                                                           "DoubleKD_vs_Control", 
                                                           "UPF3A_OE_vs_Control", 
                                                           "UPF3A_OE_UPF3B_KD_vs_Control"))

ratios_fig =ratios  %>%
  ggplot(aes(x =Comparisons, y = ratio, fill = Comparisons)) + 
  geom_bar(stat = "identity",   colour = "white", width = .75) + 
    theme_classic() + 
    theme(axis.text.x = element_text(angle = 70,  hjust=1, face = "bold"), 
          legend.position = "none", 
          legend.title= element_blank(), axis.title.y = element_text(size = 12), 
          axis.text.y = element_text(size = 14, face = "bold")) + 
    scale_fill_manual(values = c("#2F124B","#EA2A5F","#6E74B4","#E12F8F", "#FD9675")) +
    ylab("ratio up/down") + xlab("") + scale_x_discrete(labels=c("DoubleKD_vs_Control" = "Double KD", "UPF3A_vs_Control" = "UPF3A KD", 
                                                            "UPF3B_vs_Control" = "UPF3B KD", "UPF3A_OE_vs_Control" = "UPF3A OE", 
                                                            "UPF3A_OE_UPF3B_KD_vs_Control" = "UPF3A OE UPF3B KD"))



# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/ratio_DEG.svg", plot = ratios_fig,        units = c("in"), height = 3.30, width = 9.30)

```


```{r}
summary_DEG = summary(results) %>% reshape2::melt() %>%
    dplyr::filter(Var1 != "NotSig")


summary_DEG$Var1 = factor(summary_DEG$Var1, levels = c("Up", "Down"))
summary_DEG$Var2 = factor(summary_DEG$Var2, levels = c("UPF3B_vs_Control",
                                                       "UPF3A_vs_Control",
                                                       "DoubleKD_vs_Control",
                                                       "UPF3A_OE_vs_Control",
                                                       "UPF3A_OE_UPF3B_KD_vs_Control"))

summary_DEG %<>%
  mutate(Var2= factor(summary_DEG$Var2, levels = c("UPF3A_OE_UPF3B_KD_vs_Control",
                                               "UPF3A_OE_vs_Control",
                                               "DoubleKD_vs_Control",
                                               "UPF3A_vs_Control",
                                               "UPF3B_vs_Control")))
up = summary_DEG %>%
  dplyr::filter(Var1 == "Up")
down = summary_DEG %>%
  dplyr::filter(Var1=="Down")


summary_dot_plot = summary_DEG %>%
  mutate(X2= factor(summary_DEG$Var2, levels = c("UPF3A_OE_UPF3B_KD_vs_Control",
                                               "UPF3A_OE_vs_Control",
                                               "DoubleKD_vs_Control",
                                               "UPF3A_vs_Control",
                                               "UPF3B_vs_Control"))) %>%
ggplot() + geom_segment(data = down, aes(x=value, y=Var2,
                                       yend=up$Var2, xend= up$value),
                                       color = "black",
              size = 0.7) +

  geom_point(aes(x = value, y = Var2, color = Var1, size = value)) +
  theme_bw() +
        theme(
              strip.background =element_rect(fill="#AA9A9C", color = "#E1DFDB"),
              panel.border = element_rect(color = "#E1DFDB", fill = NA, size = 1)) +
  scale_color_manual(values= c("#C32137", "#4657A6")) + xlim(0,1000)

```


```{r}
ggarrange(summary_dot_plot, ratios_fig, ncol=2)
```


After performing pairwise differential gene expression analysis, we can see that majority of all significantly expressed genes in all contrasts are up-regulated. Overall, we find that a total of 395 genes are differentially expressed between UPF3B KD and controls. 388 genes are differentially expressed between UPF3A KD vs Controls.

We now want to explore the genes from each of these analyses and identify similarities and differences of expression. 

Distribution of Log fold changes 


```{r}
only_coef = list()
only_coef$
  Upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::select(contains("UPF3B_vs_Control"))
only_coef$Upf3a = DEGenes$UPF3A_vs_Control %>% 
  dplyr::select(contains("UPF3A_vs_Control"))
only_coef$upf3d = DEGenes$DoubleKD_vs_Control %>% 
  dplyr::select(contains("DoubleKD_vs_Control"))
```

## Upf3b KD DEGs 

```{r fig.cap="Volcano plot showing differentially expressed genes when Upf3b is knockdown in L-cells"}
DEColours <- c("-1" = "#2F124B","1" = "#EA2A5F",  "0" = "#D3D5E3")
UPF3B_volc = limma_results %>%
    ggplot(aes(x = Coef.UPF3B_vs_Control, 
               y = -log10(p.value.adj.UPF3B_vs_Control), 
               colour = as.factor(as.character(Res.UPF3B_vs_Control)))) +
    geom_point(alpha = 0.8, size =4) +
    scale_colour_manual(values = DEColours) + theme_classic() + 
    theme(legend.position = "none") +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("UPF3A KD vs. Control") + 
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), size = 0.5, lty = "dashed", color = "grey60") + xlim(-8.5, 8.5) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20)
        )


UPF3B_volc_int = UPF3B_volc  + geom_point_interactive(aes(tooltip = gene, data_id = gene), 
    size = 1, hover_nearest = TRUE)
girafe(ggobj = UPF3B_volc)


```

```{r}
DEGenes$UPF3B_vs_Control %>% 
   # dplyr::select(EnsID,gene,entrez, contains("UPF3B_vs")) %>% 
   # dplyr::select(-contains("Res")) %>% 
 #   mutate(genename = mapIds(org.Mm.eg.db, keys=EnsID,  column="GENENAME",keytype="ENSEMBL", multiVals="first"))
    DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), caption = "Genes significantly expressed in Upf3b KD relative to Controls", style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


## Upf3a KD DEGs 

```{r fig.cap="Volcano plot showing differentially expressed genes when Upf3a is knockdown in L-cells"}
DEColours <- c("-1" = "#2F124B","1" = "#EA2A5F",  "0" = "#D3D5E3")
UPF3A_volc = limma_results %>%
    ggplot(aes(x = Coef.UPF3A_vs_Control, 
               y = -log10(p.value.adj.UPF3A_vs_Control), 
               colour = as.factor(as.character(Res.UPF3A_vs_Control)))) +
    geom_point(alpha = 0.8, size =0.8) +
    scale_colour_manual(values = DEColours) + theme_classic() + 
    theme(legend.position = "none") +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("UPF3A KD vs. Control") + 
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), size = 0.5, lty = "dashed", color = "grey60") + xlim(-8.5, 8.5) 


UPF3A_volc_int = UPF3A_volc  + geom_point_interactive(aes(tooltip = gene, data_id = gene), 
    size = 1, hover_nearest = TRUE)
girafe(ggobj = UPF3A_volc)


```

```{r}
DEGenes$UPF3A_vs_Control %>% 
    #dplyr::select(ensembl_gene_id,gene,entrez, contains("UPF3A_vs")) %>% 
    #dplyr::select(-contains("Res")) %>% 
    #mutate(genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first")) %>%    
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), caption = "Genes significantly expressed in Upf3a KD relative to Controls", style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


## Upf3 dKD DEGs


```{r}
dKds_volc = limma_results %>%
    ggplot(aes(x = Coef.DoubleKD_vs_Control, 
               y = -log10(p.value.adj.DoubleKD_vs_Control), 
               colour = as.factor(as.character(Res.DoubleKD_vs_Control)))) +
    geom_point(alpha = 0.8, size =1.5) +
    scale_colour_manual(values = DEColours) + theme_classic() + 
    theme(legend.position = "none", 
          axis.title.y = element_text(size = 12)) +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    #geom_hline(yintercept = -log10(limma_results %>% 
     #                                  dplyr::filter(Res.UPF3B_vs_Control != 0) %>% 
      #                                 use_series(p.value.adj.UPF3B_vs_Control) %>% max), 
       #        color = "black") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("UPF3B KD vs. Control") + 
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), size = 0.5, lty = "dashed", color = "grey60") + xlim(-8.5, 8.5) 

```

```{r}
dKds_volc_int = dKds_volc  + geom_point_interactive(aes(tooltip = gene, data_id = gene), 
    size = 1, hover_nearest = TRUE)
girafe(ggobj = dKds_volc_int)

```


```{r}
DEGenes$DoubleKD_vs_Control %>% 
  #  dplyr::select(ensembl_gene_id,gene,entrez, contains("UPF3A_vs")) %>% 
   # dplyr::select(-contains("Res")) %>% 
    #mutate(genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first")) %>%     
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), caption = "Genes significantly expressed in Upf3 dKD relative to Controls", style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

## Upf3a overexpression DEGs

```{r}
upf3a_oe_volc = limma_results %>%
    ggplot(aes(x = Coef.UPF3A_OE_vs_Control, 
               y = -log10(p.value.adj.UPF3A_OE_vs_Control), 
               colour = as.factor(as.character(Res.UPF3A_OE_vs_Control)))) +
    geom_point(alpha = 0.8, size =1.5) +
    scale_colour_manual(values = DEColours) + theme_classic() + 
    theme(legend.position = "none", 
          axis.title.y = element_text(size = 12)) +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    #geom_hline(yintercept = -log10(limma_results %>% 
     #                                  dplyr::filter(Res.UPF3B_vs_Control != 0) %>% 
      #                                 use_series(p.value.adj.UPF3B_vs_Control) %>% max), 
       #        color = "black") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("UPF3B KD vs. Control") + 
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), size = 0.5, lty = "dashed", color = "grey60") + xlim(-8.5, 8.5) 

```

```{r}
upf3a_oe_volc_int = upf3a_oe_volc  + geom_point_interactive(aes(tooltip = gene, data_id = gene), 
    size = 1, hover_nearest = TRUE)
girafe(ggobj = upf3a_oe_volc)

```


```{r}
DEGenes$UPF3A_OE_vs_Control %>% 
    #dplyr::select(ensembl_gene_id,gene,entrez, contains("UPF3A_vs")) %>% 
    #dplyr::select(-contains("Res")) %>% 
    #mutate(genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first")) %>%     
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), caption = "Genes significantly expressed in Upf3a overexpression relative to Controls", style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```

## Upf3a overexpression Upf3b KD

```{r}
upf3a_oe_kd_volc = limma_results %>%
    ggplot(aes(x = Coef.UPF3A_OE_UPF3B_KD_vs_Control, 
               y = -log10(p.value.adj.UPF3A_OE_UPF3B_KD_vs_Control), 
               colour = as.factor(as.character(Res.UPF3A_OE_UPF3B_KD_vs_Control)))) +
    geom_point(alpha = 0.8, size =1.5) +
    scale_colour_manual(values = DEColours) + theme_classic() + 
    theme(legend.position = "none", 
          axis.title.y = element_text(size = 12)) +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    #geom_hline(yintercept = -log10(limma_results %>% 
     #                                  dplyr::filter(Res.UPF3B_vs_Control != 0) %>% 
      #                                 use_series(p.value.adj.UPF3B_vs_Control) %>% max), 
       #        color = "black") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("UPF3B KD vs. Control") + 
  geom_vline(xintercept = c(-log2(1.5), log2(1.5)), size = 0.5, lty = "dashed", color = "grey60") + xlim(-8.5, 8.5) 


```

```{r}
upf3a_oe_kd_volc_int = upf3a_oe_kd_volc  + geom_point_interactive(aes(tooltip = gene, data_id = gene), 
    size = 1, hover_nearest = TRUE)
girafe(ggobj = upf3a_oe_kd_volc_int)

```

```{r}
DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
   # dplyr::select(ensembl_gene_id,gene,entrez, contains("UPF3A_vs")) %>% 
    #dplyr::select(-contains("Res")) %>% 
    #mutate(genename = mapIds(org.Mm.eg.db, keys=ensembl_gene_id,  column="GENENAME",keytype="ENSEMBL", multiVals="first")) %>%     
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), caption = "Genes significantly expressed in Upf3a OE in Upf3b KD relative to Controls", style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


## NMD substrates 

```{r}

rg <- limma_results %>% 
  dplyr::select(contains("Coef."), gene) %>% 
  dplyr::filter(gene %in% genes_of_interest) %>% 
  column_to_rownames("gene") %>% max(abs(.))

limma_results %>% 
  dplyr::select(contains("Coef."), gene) %>% 
  dplyr::filter(gene %in% genes_of_interest) %>% 
  column_to_rownames("gene") %>%
  pheatmap(,  
        scale="none", 
        cluster_cols = FALSE,
        breaks = seq(-rg, rg, length.out = 100),
        clustering_distance_rows = "euclidean", 
        border_color = "white", 
        treeheight_row = 0,
        fontsize = 6,
        color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100),
        labels_row = row.names(.), 
        cellwidth = 8,
        cellheight = 8)


limma_results %>% 
  dplyr::select(contains("Coef."), gene) %>% 
  dplyr::filter(gene %in% genes_of_interest) %>% 
  reshape2::melt() %>% 
  ggplot(aes(x=gene, y=value)) +
  geom_bar(stat= "identity") + 
  facet_grid(~variable, scales = "free")

```

- UPF2 levels 

```{r}
upf2_exp = res %>%
  do.call(rbind, .) %>% 
   rownames_to_column("comparison") %>% 
  mutate_at(.vars = "comparison", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>%
  dplyr::filter(EnsID== "ENSMUSG00000043241") %>% 

  mutate(Significance = c("Not significant")) %>%
  mutate(comparison = factor(comparison, levels = c("UPF3B_vs_Control",
                                                "UPF3A_vs_Control",
                                                "DoubleKD_vs_Control", 
                                                "UPF3A_OE_vs_Control",
                                                "UPF3A_OE_UPF3B_KD_vs_Control"))) %>%
  ggplot(aes(x=logFC, y=comparison, shape=Significance, color= Significance)) +
    geom_point(stat="identity", size =8) +
 xlab(bquote(log["2"](FC))) + theme_bw() +
  theme(axis.text.y = element_text(size=10), 
        axis.text.x = element_text(size=10)) + 
   scale_color_manual(values=c("#adc178", "#FD9675")) +
  geom_vline(xintercept = 0, color="black", linetype = "dashed") +
  scale_y_discrete(labels=c("DoubleKD_vs_Control" = "Upf3 dKD", 
                            "UPF3A_vs_Control" = "Upf3a KD", 
                            "UPF3B_vs_Control" = "Upf3b KD", 
                            "UPF3A_OE_vs_Control" = "Upf3a OE",
                            "UPF3A_OE_UPF3B_KD_vs_Control" = "Upf3a OE Upf3b KD"), 
                   limits =rev) +
  ggtitle("Upf2 expression") + ylab("") + 
    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
              legend.position = "top",
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) 

upf2_exp

#ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/UPF2_exp.svg", plot = upf2, 
 #      units = c("in"), height = 4.24, width = 7.65)

```


## Overlapping DEGs 

Next, we wanted to see how differentially expressed genes within each comparison compared. 

```{r fig.cap="Upset plot showing overlapping DEGs from each comparison"}
library(ComplexUpset)
library(ggplot2)

x=list("UPF3B KD" = DEGenes$UPF3B_vs_Control$EnsID[DEGenes$UPF3B_vs_Control$logFC > log2(1.5)], 
                       "UPF3A KD" = DEGenes$UPF3A_vs_Control$EnsID[DEGenes$UPF3A_vs_Control$logFC> log2(1.5)],
                    "UPF3 dKD" = DEGenes$DoubleKD_vs_Control$EnsID[DEGenes$DoubleKD_vs_Control$logFC > log2(1.5)],
                    "UPF3A OE" = DEGenes$UPF3A_OE_vs_Control$EnsID, 
                  "UPF3A OE, UPF3B KD" = DEGenes$UPF3A_OE_UPF3B_KD_vs_Control$EnsID)


ggvenn(x[c(1:3)])

UpSetR::upset(fromList(x))

# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/upset_overlap.svg", plot = plot, 
#        units = c("in"), height = 5.89, width = 8.26)
```


```{r fig.cap="Total number of overlaps between two sets. The DEGs from Upf3a KD are used as a hallmark of disruption to the canonical NMD pathway in L-cells"}
overlaps = data.frame("Comparison" = c("UPF3B KD vs UPF3A KD", 
                                       "UPF3B KD vs UPF3 dKD", 
                                       "UPF3B KD vs UPF3A OE", 
                                       "UPF3B KD vd UPF3A OE, UPF3B KD", 
                                       "UPF3A KD vs UPF3 dKD", 
                                       "UPF3A KD vs UPF3A OE", 
                                       "UPF3A KD vs UPF3A OE, UPF3B KD", 
                                       "UPF3 dKD vs UPF3A OE", 
                                       "UPF3 dKD vs UPF3A OE, UPF3B KD", 
                                       "UPF3A OE vs UPF3A OE, UPF3B KD"), 
                      value = c(144, 316, 12, 256, 
                                302, 9, 119, 
                                13, 261, 8), 
                      sig = c(1,2,3,4, 
                              0,0,0,
                              0,0,0))

overlaps$Comparison = factor(overlaps$Comparison, levels = c("UPF3B KD vs UPF3A KD", 
                                       "UPF3B KD vs UPF3 dKD", 
                                       "UPF3B KD vs UPF3A OE", 
                                       "UPF3B KD vd UPF3A OE, UPF3B KD", 
                                       "UPF3A KD vs UPF3 dKD", 
                                       "UPF3A KD vs UPF3A OE", 
                                       "UPF3A KD vs UPF3A OE, UPF3B KD", 
                                       "UPF3 dKD vs UPF3A OE", 
                                       "UPF3 dKD vs UPF3A OE, UPF3B KD", 
                                       "UPF3A OE vs UPF3A OE, UPF3B KD"))


overlaps_fig = overlaps %>% 
  ggplot(aes(x=Comparison, y=value, fill = as.character(sig))) + geom_bar(stat="identity", color = "black") +
  theme_bw() + theme(axis.text.x = element_text(angle=90), 
                     legend.position = "none") 

overlaps_fig
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/overlaps.svg", plot = overlaps_fig, 
       units = c("in"), height = 5.89, width = 8.26)
```


<!-- ## Using Upf3b KD DEGs as a hallmark of impaired NMD  -->

<!-- Given that previous studies have confirmed UPF3B KD results in a decrease of efficieny of NMD resulting in an increase of upregulated genes, we are interested in seeing how the significantly expressed up-regulated genes behave in other conditions compared to controls. -->


<!-- In order to do this, we will create a heatmap visualisaiton.  -->


<!-- ```{r fig.height=30} -->

<!-- rg <- limma_results %>%  -->
<!--      dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% max(abs(.)); -->

<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 8 -->
<!--     ) -->

<!-- ``` -->

<!-- ```{r fig.height=30} -->

<!-- rg <- limma_results %>%  -->
<!--      dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% max(abs(.)); -->

<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% -->
<!--   as.matrix() %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 8 -->
<!--     ) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- library(ComplexHeatmap) -->
<!-- library(circlize) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- heatmap = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control !=0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% -->
<!--   dplyr::select("UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control",  -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control", -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control", -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control",  -->
<!--                 "UPF3B KD" ="Coef.UPF3B_vs_Control") -->

<!-- split = sample(letters[1:7], 418, replace = TRUE) -->
<!-- split = factor(split, levels = letters[1:7]) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- Heatmap(heatmap, split = 7) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- col_fun1 = colorRamp2(c(-5, 0, 5), c("#2F124B", "#F9F9F9","#EA2A5F")) -->
<!-- color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100) -->

<!-- #circos.par(gap.after = c(2, 2, 10)) -->
<!-- circos.heatmap(heatmap ,col = col_fun1, cell_width = 1, track.height = 0.2, cluster = TRUE) -->
<!-- circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) { -->
<!--     if(CELL_META$sector.numeric.index == 7) { # the last sector -->
<!--         cn = colnames(heatmap) -->
<!--         n = length(cn) -->
<!--         circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"),  -->
<!--             1:n - 0.5, cn,  -->
<!--             cex = 0.5, adj = c(0, 0.5,0), facing = "inside") -->
<!--     } -->
<!-- }, bg.border = NA) -->

<!-- circos.clear() -->
<!-- ``` -->


<!-- From the heatmap above, we can notice a few trends that upregulated genes in UPF3B KD have in other conditions.  -->

<!-- 1) Inverted expression in UPF3A OE genes -->

<!-- 2) Inverted expression in UPF3A KD  -->


<!-- These genes can be explored here:  -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--     dplyr::select(gene, contains("Coef"))  %>% -->
<!--    # column_to_rownames("gene") %>%   -->
<!--     DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->



<!-- For ease of interpretation, I have decided to perform a hierarchical clustering to divide the genes based on their expression. For this, it would be great to use all significant genes in UPF3B KD -->

<!-- ```{r} -->
<!-- # Create a matrix -->
<!-- genes_for_clust  = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") -->


<!-- hclust_matrix <- genes_for_clust%>%  -->
<!--   as.matrix() -->

<!-- # assign rownames -->




<!-- hclust_matrix <- hclust_matrix %>%  -->
<!--   # transpose the matrix so genes are as columns -->
<!--   t() %>%  -->
<!--   # apply scalling to each column of the matrix (genes) -->
<!--   scale() %>%  -->
<!--   # transpose back so genes are as rows again -->
<!--   t() -->

<!-- gene_dist <- dist(hclust_matrix) -->

<!-- gene_hclust <- hclust(gene_dist, method = "complete") -->

<!-- # The default `plot()` function can be used to produce a simple dendrogram -->
<!-- plot(gene_hclust, labels = FALSE) -->
<!-- abline(h = 3, col = "brown", lwd = 2) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- gene_cluster <- cutree(gene_hclust, k = 7) %>%  -->
<!--   # turn the named vector into a tibble -->
<!--   enframe() %>%  -->
<!--   as.data.frame() -->

<!-- colnames(gene_cluster) = c("gene", "cluster") -->
<!-- ``` -->

<!-- Cluster 1 -->

<!-- ```{r fig.width=10} -->

<!-- rg <- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% max(abs(.)); -->


<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 1) %>%  -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control",  -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control",  -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control",  -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control",  -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>%  -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--     pheatmap::pheatmap(, -->
<!--        # scale="row",  -->
<!--         cluster_cols = TRUE,  -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10,  -->
<!--         cellwidth = 10 -->
<!--     ) -->
<!-- ``` -->

<!-- Cluster 2 -->

<!-- ```{r fig.width=20, fig.height=10} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 2) %>%  -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control",  -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control",  -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control",  -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control",  -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>%  -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row",  -->
<!--         cluster_cols = TRUE,  -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10,  -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->

<!-- Cluster 3 -->

<!-- ```{r fig.width=12} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 3) %>%  -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control",  -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control",  -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control",  -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control",  -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>%  -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row",  -->
<!--         cluster_cols = TRUE,  -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10,  -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->


<!-- Cluster 4 -->

<!-- ```{r fig.width=12} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 4) %>%  -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control",  -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control",  -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control",  -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control",  -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>%  -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row",  -->
<!--         cluster_cols = TRUE,  -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10,  -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->


<!-- Cluster 5 -->

<!-- ```{r fig.width=12} -->
<!-- limma_results %>% -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 5) %>% -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control", -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control", -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control", -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control", -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>% -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row", -->
<!--         cluster_cols = TRUE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean", -->
<!--         border_color = "white", -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10, -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->


<!-- Cluster 6 -->

<!-- ```{r fig.width=12} -->
<!-- limma_results %>% -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 6) %>% -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control", -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control", -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control", -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control", -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>% -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row", -->
<!--         cluster_cols = TRUE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean", -->
<!--         border_color = "white", -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10, -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->


<!-- Cluster 7 -->


<!-- ```{r fig.width=12} -->
<!-- limma_results %>% -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--    left_join(gene_cluster, by="gene") %>% -->
<!--   dplyr::filter(cluster == 7) %>% -->
<!--   column_to_rownames("gene") %>% -->
<!--   dplyr::select(-cluster) %>% -->
<!--   dplyr::rename("UPF3B KD" = "Coef.UPF3B_vs_Control", -->
<!--                 "UPF3A KD" = "Coef.UPF3A_vs_Control", -->
<!--                 "UPF3A OE" = "Coef.UPF3A_OE_vs_Control", -->
<!--                 "UPF3 dKD" = "Coef.DoubleKD_vs_Control", -->
<!--                 "UPF3A OE UPF3B KD" = "Coef.UPF3A_OE_UPF3B_KD_vs_Control") %>% -->
<!--   dplyr::select("UPF3B KD", "UPF3A KD", "UPF3 dKD", "UPF3A OE", "UPF3A OE UPF3B KD") %>% -->
<!--   t() %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row", -->
<!--         cluster_cols = TRUE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean", -->
<!--         border_color = "white", -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10, -->
<!--         cellwidth = 10 -->
<!--     ) -->

<!-- ``` -->



<!-- We are now also interested in seeing how other DEGs in Upf3b KD DEGs overlap with significant genes from other conditions. -->


<!-- #### Comparison with UPF3A KD significant genes  -->


<!-- ```{r} -->
<!-- #venn(list("UPF3B KD" = DEGenes$UPF3B_vs_Control$ensembl_gene_id,  -->
<!--  #                      "UPF3A KD" = DEGenes$UPF3A_vs_Control$ensembl_gene_id)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- x = list( "UPF3B KD" = rownames(DEGenes$UPF3B_vs_Control),  -->
<!--     "UPF3A KD" = rownames(DEGenes$UPF3A_vs_Control)) -->

<!-- venn_upf3 = ggvenn(x[c(1,2)], fill_color = c("#2F124B", "#801D2E"), -->
<!--        set_name_color = "black", text_color = "white", text_size = 6,  -->
<!--        stroke_size = 0.5, fill_alpha = 0.8,  -->
<!--        stroke_color = "white")  -->

<!-- venn_upf3 -->
<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/UPF3_venn.svg", plot = venn_upf3 ,  -->
<!--        units = c("in"), height = 6, width = 5) -->

<!-- ``` -->


<!-- When we compare all significant genes in Upf3b KD with Upf3a KD, we can see that there is an overlap of 144 genes.  -->

<!-- ```{r fig.height =4, fig.width=5, fig.cap="*Gene expression distribution of Upf3a and Upf3b KD DEGs. 144 genes overlap between the two comparisons and show positive correlaiton in the direction of gene expression*"} -->

<!-- cor_upf3 =  DEGenes$UPF3B_vs_Control %>%  -->

<!--   inner_join(DEGenes$UPF3A_vs_Control, by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.6, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3a KD (log2FoldChange)") + xlab("Upf3b KD (log2FoldChange)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#F15A2B", "#2E368F")) +   -->
<!--   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3_paralog_cor.svg", plot = cor_upf3 ,  -->
<!--        units = c("in"), height = 4.65, width = 4) -->


<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control !=0) %>%  -->
<!--     mutate(Label = ifelse(Res.UPF3B_vs_Control ==1 & Res.UPF3A_vs_Control ==1 |  -->
<!--                             Res.UPF3B_vs_Control ==-1 & Res.UPF3A_vs_Control ==-1, "Concordance", "Discordant")) %>% -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_vs_Control",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=3, alpha =0.6, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "#EF3829", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3a KD (log2FoldChange)") + xlab("Upf3b KD (log2FoldChange)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") +  -->
<!--   scale_color_manual(values = c("#F15A2B", "#2E368F"))  -->
<!-- ``` -->



<!-- When we compare the direction of gene expression in these 144 genes, we can see that there is a positive correlation (r = 0.56).  -->

<!-- We are now interested in seeing which genes are present in each of the quadrants:  -->

<!-- 1) Quadrant 1:  Defined as x < 0, y > 0 where is significant genes belogning to UPF3A KD group show upregulation and genes belonging to UPF3B KD group show downregulation. A total of 14 genes belong to this group: -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == -1 & Res.UPF3A_vs_Control == 1) %>%   -->
<!--     dplyr::select(ensembl_gene_id,gene,genename,type,  contains(c("UPF3B_vs", "UPF3A_vs"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                    caption ="Overlapping genes between Upf3b KD and Upf3a KD relative to controls in Quadrant 1 (Upregulated in UPF3A KD, downregulated in UPF3B KD)", -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- 2) Quadrant 2: Defined as x > 0, y > 0 where significant genes from both conditions show upregulation. 114 genes belong to this group.  -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1 & Res.UPF3A_vs_Control ==1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,genename,type, contains(c("UPF3B_vs", "UPF3A_vs"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--     DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                   caption ="Overlapping genes between UPF3B KD and UPF3A KD relative to controls in Quadrant 2 (Upregulated in both conditions)", -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- 3)  Quadrant 3 is defined as x < 0, y < 0 where genes in both groups show downregulation.  -->
<!-- 12 genes are in this quadrant: -->


<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == -1 & Res.UPF3A_vs_Control == -1) %>% -->
<!--     dplyr::select(ensembl_gene_id,genename,type, contains(c("UPF3B_vs", "UPF3A_vs"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--     DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                   caption ="Overlapping genes between UPF3B KD and UPF3A KD relative to controls in Quadrant 3 (Downregulated in both conditions)", -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- 4) Quadrant 4 is defined as x > 0, y < 0, where significant genes belonging to UPF3B KD group are upregulated and genes belonging to UPF3A KD group are downregulated. Only 4 genes belong in this group.  -->


<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1 & Res.UPF3A_vs_Control ==-1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,genename,type, contains(c("UPF3B_vs", "UPF3A_vs"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                    caption ="Overlapping genes between UPF3B KD and UPF3A KD relative to controls in Quadrant 4 (Upregulated in UPF3B KD, downregulated in UPF3A KD)", -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->


<!-- We are also interested in seeing how the 144 overlapping genes between Upf3a KD and Upf3b KD behave in other conditions. To do this, we will use a heatmap visualisation again. -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control !=0) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene, contains(c("Coef."))) %>% -->
<!--  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                caption ="Overlapping genes between UPF3B KD and UPF3A KD (padj < 0.05; absolute log2FoldChange > log2(1.5)) in other comparisons)", -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->




<!-- ```{r fig.height =20, out.extra='angle=90'} -->

<!-- rg <- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control !=0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% max(abs(.)); -->

<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control !=0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% -->
<!--      pheatmap::pheatmap(, -->
<!--        # scale="row",  -->
<!--         cluster_cols = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         cutree_rows = 2, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.), -->
<!--         cellheight = 10,  -->
<!--         cellwidth = 12 -->
<!--     ) -->


<!-- ``` -->


<!-- ```{r} -->
<!-- heatmap = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control !=0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") -->

<!-- split = sample(letters[1:5], 144, replace = TRUE) -->
<!-- split = factor(split, levels = letters[1:5]) -->

<!-- col_fun1 = colorRamp2(c(-6, 0, 6), c("#2F124B", "#F9F9F9","#EA2A5F")) -->
<!-- color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100) -->

<!-- circos.par(gap.after = c(2, 2, 2, 2, 10)) -->
<!-- circos.heatmap(heatmap ,col = col_fun1, split = split, rownames.side = "outside") -->
<!-- circos.track(track.index = get.current.track.index(), panel.fun = function(x, y) { -->
<!--     if(CELL_META$sector.numeric.index == 5) { # the last sector -->
<!--         cn = colnames(heatmap) -->
<!--         n = length(cn) -->
<!--         circos.text(rep(CELL_META$cell.xlim[2], n) + convert_x(1, "mm"),  -->
<!--             1:n - 0.5, cn,  -->
<!--             cex = 0.5, adj = c(0, 0.5,0), facing = "inside") -->
<!--     } -->
<!-- }, bg.border = NA) -->


<!-- ``` -->


<!-- We also want to see the genes that are unique to UPF3A KD and unique to UPF3B KD and how their expression is -->

<!-- ```{r} -->
<!-- upf3b_uniq = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control ==0) %>%  -->
<!--     mutate(Label = ifelse( -->
<!--       Coef.UPF3B_vs_Control > 0 & Coef.UPF3A_vs_Control > 0 | -->
<!--       Coef.UPF3B_vs_Control < 0 & Coef.UPF3A_vs_Control < 0,  -->
<!--       "Concordance", "Discordant"  )) %>% -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_vs_Control",  -->
<!--                  size=3, alpha =0.6, color = "Label",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "#EF3829", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("UPF3A KD (log2FoldChange)") + xlab("UPF3B KD (log2FoldChange)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") +  -->
<!--   scale_color_manual(values = c("#F15A2B", "#2E368F")) + theme(legend.position = "none")  -->


<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/UPF3B_corr.svg", plot = upf3b_uniq,  -->
<!--        units = c("in"), height = 6, width = 6) -->

<!-- upf3b_uniq -->
<!-- ``` -->

<!-- ```{r} -->
<!-- upf3a_uniq = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 0 & Res.UPF3A_vs_Control !=0) %>%  -->
<!--    mutate(Label = ifelse( -->
<!--       Coef.UPF3B_vs_Control > 0 & Coef.UPF3A_vs_Control > 0 | -->
<!--       Coef.UPF3B_vs_Control < 0 & Coef.UPF3A_vs_Control < 0,  -->
<!--       "Concordance", "Discordant"  )) %>%  -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_vs_Control",  -->
<!--                size=3, alpha =0.6, color = "Label",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "#EF3829", -->
<!--                                                  fill = "lightgray")) +  -->

<!--     theme_bw() + ylab("UPF3A KD (log2FoldChange)") + xlab("UPF3B KD (log2FoldChange)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + -->
<!--    scale_color_manual(values = c("#F15A2B", "#2E368F")) + theme(legend.position = "none")  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/UPF3A_corr.svg", plot = upf3a_uniq,  -->
<!--        units = c("in"), height = 6, width = 6) -->

<!-- upf3a_uniq -->
<!-- ``` -->


<!-- ### Checking the magnitude of impact from upf3b and upf3a specific genes  -->

<!-- ```{r fig.cap="Cumulative frequency of unique genes"} -->
<!-- ks.test(DEGenes$UPF3B_vs_Control$Coef.UPF3B_vs_Control,  -->
<!--         DEGenes$UPF3B_vs_Control$Coef.UPF3A_vs_Control) -->

<!-- cumulative_freq = limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_vs_Control ==0) %>%  -->
<!--   dplyr::select("log2FoldChange"= "Coef.UPF3B_vs_Control") %>%  -->
<!--   mutate(Condition = "Upf3b Unique Genes") %>%  -->
<!--   rbind(limma_results %>%  -->
<!--           dplyr::filter(Res.UPF3B_vs_Control == 0 & Res.UPF3A_vs_Control !=0) %>%  -->
<!--           dplyr::select("log2FoldChange"= "Coef.UPF3A_vs_Control") %>%  -->
<!--           mutate(Condition = "Upf3a unique genes")) %>%  -->
<!--   ggplot(aes(x=log2FoldChange, color = Condition)) + stat_ecdf(geom = "point", size=1.5) + -->
<!--   scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() + -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") + -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") + -->
<!--   ggtitle("Specific genes logFC") + theme(legend.position = "none") -->

<!-- cumulative_freq -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/upf3_cum_freq.svg", plot = cumulative_freq, -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->

<!-- ``` -->

<!-- ```{r fig.cap="Cumulative frequency of all DEGs"} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0) %>%  -->
<!--   dplyr::select("log2FoldChange"= "Coef.UPF3B_vs_Control") %>%  -->
<!--   mutate(Condition = "Upf3b Unique Genes") %>%  -->
<!--   rbind(limma_results %>%  -->
<!--           dplyr::filter(Res.UPF3A_vs_Control !=0) %>%  -->
<!--           dplyr::select("log2FoldChange"= "Coef.UPF3A_vs_Control") %>%  -->
<!--           mutate(Condition = "Upf3a unique genes")) %>%  -->
<!--   ggplot(aes(x=log2FoldChange, color = Condition)) + stat_ecdf(geom = "point", size=0.8) + -->
<!--   scale_color_manual(values = c("#2F124B", "#EA2A5F")) + theme_bw() + -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") + -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") + -->
<!--   ggtitle("Specific genes logFC") -->

<!-- ``` -->



<!-- #### UPF3B vs Double KD  -->


<!-- ```{r} -->
<!-- ggvenn(x[c(1,3)]) -->
<!-- ``` -->

<!-- A total of 316 genes overlap betweeen the Double KDs and UPF3B KD. When we look at the correlation, we see that there is a strong positive correlation.  -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.DoubleKD_vs_Control !=0) %>%  -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.DoubleKD_vs_Control",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  -->
<!-- ``` -->


<!-- 1) Quadrant 1:  Defined as x < 0, y > 0 where is significant genes belonging to Double KD group show upregulation and genes belonging to UPF3B KD group show downregulation. A total of 2 genes belong to this group: -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == -1 & Res.DoubleKD_vs_Control == 1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,entrez, contains(c("UPF3B_vs", "Double"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->
<!-- 2) Quadrant 2: Defined as x > 0, y > 0 where significant genes from both conditions show upregulation. 255 genes belong to this group.  -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1 & Res.DoubleKD_vs_Control ==1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,entrez, contains(c("UPF3B_vs", "DoubleKD"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--    DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->


<!-- 3)  Quadrant 3 is defined as x < 0, y < 0 where genes in both groups show downregulation.  -->
<!-- 58 genes are in this quadrant: -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == -1 & Res.DoubleKD_vs_Control == -1) %>% -->
<!--     dplyr::select(ensembl_gene_id,gene,entrez, contains(c("UPF3B_vs", "DoubleKD"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->

<!-- ### UPF3B vs UPF3A OE, UPF3B KD  -->


<!-- ```{r} -->
<!-- ggvenn(x[c(1,5)]) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- upf3a_oe_upf3bkd= limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 & Res.UPF3A_OE_UPF3B_KD_vs_Control !=0) %>%  -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_OE_UPF3B_KD_vs_Control",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/Upf3a_oe_Upf3bKD_corr.svg", plot = upf3a_oe_upf3bkd,  -->
<!--        units = c("in"), height = 6, width = 6) -->

<!-- ``` -->


<!-- When we compare UPF3B KD with UPF3A OE in UPF3B KD cell lines, we can see there is a strong positive correlation. This correlation is very likely induced by the UPF3B KD cell line itself, and not the impact of the UPF3A OE.  For interest, here are the gene in Quadrant 2 and Quadrant 3 -->


<!-- 1) Quadrant 2: Defined as x > 0, y > 0 where significant genes from both conditions show upregulation. 183 genes belong to this group.  -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == 1 & Res.UPF3A_OE_UPF3B_KD_vs_Control ==1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,entrez, contains(c("UPF3B_vs", "DoubleKD"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--    DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->


<!-- 2)  Quadrant 3 is defined as x < 0, y < 0 where genes in both groups show downregulation.  -->
<!-- 73 genes are in this quadrant: -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control == -1 & Res.UPF3A_OE_UPF3B_KD_vs_Control== -1) %>%  -->
<!--     dplyr::select(ensembl_gene_id,gene,entrez, contains(c("UPF3B_vs", "DoubleKD"))) %>%  -->
<!--     dplyr::arrange(Res.UPF3B_vs_Control) %>% -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->


<!-- Following up from this, it would be good to see how the genes that were specific to the UPF3B KD (162 genes) and UPF3A OE, UPF3B KD (85) behave relative to the other condition. -->


<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control != 0 &  Res.UPF3A_OE_UPF3B_KD_vs_Control == 0 ) %>%  -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_OE_UPF3B_KD_vs_Control",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3B_vs_Control ==0 &  Res.UPF3A_OE_UPF3B_KD_vs_Control != 0 ) %>%  -->
<!--     ggscatter(., x= "Coef.UPF3B_vs_Control", y="Coef.UPF3A_OE_UPF3B_KD_vs_Control",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  -->
<!-- ``` -->

<!-- Overall the above plots show that there is a strong positive correlation. -->













<!-- ```{r fig.height=30, fig.cap ="Heatmap of log2FCof comparisons based on genes significant in UPF3A OE compared to controls"} -->

<!-- rg <- limma_results %>%  -->
<!--      dplyr::filter(Res.UPF3A_OE_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% max(abs(.)); -->

<!-- limma_results %>%  -->
<!--     dplyr::filter(Res.UPF3A_OE_vs_Control != 0) %>% -->
<!--     dplyr::select(contains("Coef"), gene)  %>% -->
<!--     column_to_rownames("gene") %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 8 -->
<!--     ) -->

<!-- ``` -->


<!-- ## Upf3a only transcripts  -->

<!-- ```{r} -->
<!-- x = list("Upf3a DEGs" = DEGenes$UPF3A_vs_Control$EnsID,  -->
<!--          "Upf3b DEGs" = DEGenes$UPF3B_vs_Control$EnsID) -->

<!-- ggvenn(x) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- overlap_upf3 = calculate.overlap(x) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- non_intersect = overlap_upf3$a1[!overlap_upf3$a1 %in% overlap_upf3$a3 ] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- freq_upf3a = res$UPF3B_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) %>%  -->
<!--   rbind(res$UPF3A_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) ) %>%  -->
<!--     ggplot(aes(x=logFC, color = contrast)) + stat_ecdf(geom = "step", size=1) + -->
<!--   scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() + -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") + -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") + -->
<!--   ggtitle("259 Upf3a KD specific genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") +  -->
<!--       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         ))  +  -->
<!--    annotate("point", x = 0.175, y = 0, colour = "#2F124B") + -->
<!--   annotate("point", x =  0.798, y = 0, colour = "#EA2A5F") + -->
<!--   annotate("text", x=-1.5, y = 0.85, label = "D = 0.676,  p-value < 2.2e-16", fontface =7) -->


<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a_specific_cumulative.svg", plot =freq_upf3a  ,  -->
<!--        units = c("in"), height = 5.65, width = 5) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- ks.test(res$UPF3B_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) %>%  -->
<!--     with(., logFC), res$UPF3A_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) %>% with(., logFC)) -->


<!-- median(res$UPF3B_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) %>%  -->
<!--     with(., logFC)) -->

<!-- median(res$UPF3A_vs_Control %>%  -->
<!--   dplyr::filter(EnsID %in% non_intersect) %>%  -->
<!--   dplyr::select(logFC, contrast) %>% with(., logFC)) -->
<!-- ``` -->


<!-- ```{r} -->
<!--  DEGenes$DoubleKD_vs_Control %>%  -->

<!--   inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control,  by = "EnsID") %>%  -->
<!--   mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 |  -->
<!--                             logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>%  -->
<!--   ggscatter(., x= "logFC.x", y="logFC.y",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               size=5, alpha =0.6, color = "Label", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw() + ylab("Upf3a KD (log2FoldChange)") + xlab("Upf3b KD (log2FoldChange)") + -->
<!--   geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!--   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!--   scale_color_manual(values = c("#F15A2B", "#2E368F")) +   -->
<!--   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!--             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!--         axis.title = element_text(size=15, family = "serif"),  -->
<!--          legend.box.background = element_rect(color = "black"), -->
<!--           legend.text = element_text(family = "serif"),  -->
<!--           legend.title = element_text( family = "serif"), -->
<!--            plot.title = element_text(family = "serif", size =20),          -->
<!--         panel.grid.major.x = element_blank(), -->
<!--         panel.grid.major.y = element_line( size=.1 ), -->
<!--               legend.position = "top", -->
<!--         strip.text.y = element_text( -->
<!--         size = 10, face = "bold.italic", family = "serif" -->
<!--         ), -->
<!--          strip.text.x = element_text( -->
<!--         size = 10, family = "serif" -->
<!--         )) + labs(color = "Expression") -->

<!-- ``` -->

# Data export
```{r}
#require(openxlsx)
#save(DEGenes,res, file=here::here("output/DEG-list.Rda"))
#save(limma_results, file=here::here("output/DEG-limma-results.Rda"))
#save(results, fit,y,v,contrasts, design, contrasts,  file = here::here("output/limma-matrices.Rda"))
#write.xlsx(DEGenes,file = here::here("output/DEG-results.xlsx"))

```
