---
title: "Differential transcript expression analysis"
author: "Urwah Nawaz"
date: "2023-03-26"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: inline
---

# Introduction


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```

```{r libraries}
library(magrittr)
library(fishpond)
library(tximeta)
library(GeneOverlap)
library(ggside)
library(tidyquant)
library(stargazer)
library(ggfortify)
library(glue)
library(edgeR)
library(cowplot)
library(broom)
library(glmpca)
library(naniar)
library(dplyr)
library(bigPint)
source(here::here("code/libraries.R"))
source(here::here("code/functions.R"))
```

## Set-up 

```{r tx-setup}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])

gdf = genes(EnsDb.Mmusculus.v79, return.type="DataFrame") %>% as.data.frame()
gene_stats = as.data.frame(gdf[,c("gene_id", "gene_biotype")])

tx2gene =full_join(tx2gene,gdf, by = "gene_id") %>%
  dplyr::select(-c("gene_biotype", "gene_seq_start",
                   "gene_seq_end",   "seq_name",
                   "seq_strand", "seq_coord_system",         "symbol",     "entrezid"))

```

```{r load-res}
load(here::here("output/limma-matrices.Rda"))
load(here::here("output/DEG-limma-results.Rda"))
load(here::here("output/DEG-list.Rda"))

```



```{r data-setup}
md.files = "data/LTK_Sample Metafile_V3.txt"
salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Salmon")
salmon = list.files(salmon.files, pattern = "transcripts$", full.names = TRUE)
names(salmon) <- paste(c(212:229))
md = read.table(here::here(md.files), header= TRUE) %>%  mutate(files = file.path(salmon, "quant.sf"))
comparisons = unique(md$Group)[-1]
colnames(md)[1] = c("names")
md$names = as.character(md$names)
md$Group <- factor(md$Group, 
                            levels=c("Control","UPF3B_KD", "UPF3A_KD", "UPF3A_OE", 
                                     "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
```

```{r}
group_cols <- hcl.colors(
  n = length(unique(md$Group)), 
  palette = "Zissou 1"
  ) %>%
  setNames(unique(md$Group))
```



Transcript-level counts were retrieved using the `catchSalmon()` function from the edgeR package. 
Once retrieved, the counts underwent rigourous QC. 

# Removal of outlier sample 

According to the quality control analysis, it was revealed that one of the samples was not clustering with its condition group. Investigation into GC content, transcript length and library size revealed no contribution of these factors. To ensure we are getting results that are not skewed we will remove the sample from the transcript level analysis. 


```{r include=FALSE}
catch <- catchSalmon(salmon)
colnames(catch$counts) =  paste(c(212:229))

cts.scaled = catch$counts/catch$annotation$Overdispersion

#md_filt = md %>% dplyr::filter(names != "223")


dge <- DGEList(counts = cts.scaled, 
               genes=catch$annotation, 
               samples = md,
               group = md$Group)

keep <- filterByExpr(dge)
dge.scaled.filtr <- dge[keep, , keep.lib.sizes = FALSE]
dge.scaled.filtr <- calcNormFactors(dge.scaled.filtr)

dim(dge.scaled.filtr$counts)

```

```{r fig.height = 7, fig.width=10, fig.cap="*Principal component analysis of transcript level data. PCA was performed on log2 transformed CPM after filtering lowly expressed genes. The PCA shows that samples are clustering close to their conditions based on PC1, however one of the samples of the UPF3A OE in UPF3B KD cell line (sample 223), seems to deviate from its condition group and the rest of the data, so needs to be further investigated*"}
 pca = cpm(dge.scaled.filtr, log = TRUE) %>% 
  t() %>% 
    prcomp(scale = TRUE) %>% 
  tidy() %>% 
  dplyr::rename(names = row) %>% 
  left_join(dge.scaled.filtr$samples, by = "names") %>%
   dplyr::filter(PC %in% 1:3) %>% 
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>% 
  ggplot(
    aes(PC1, PC2, colour = Group, size = lib.size/1e6)
  ) +
  geom_point() +
  geom_text_repel(aes(label = names), show.legend = FALSE) +
  scale_colour_manual(values = group_cols, 
                      labels = c("UPF3A_KD_UPF3B_KD" = "UPF3 dKD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  #scale_size_continuous(limits = c(5, 15), breaks = seq(5, 15, by = 5)) +
  theme_bw() + ggtitle("Transcript Level")



pca
  
```


```{r fig.height=5, fig.width=5}
 plotMDS(dge.scaled.filtr,col=as.numeric(dge.scaled.filtr$samples$group))
```



# Analysis

```{r include=FALSE}
catch <- catchSalmon(salmon[-12])
colnames(catch$counts) =  paste(c(212:222, 224:229))

cts.scaled = catch$counts/catch$annotation$Overdispersion

md_filt = md %>% dplyr::filter(names != "223")


dge <- DGEList(counts = cts.scaled, 
               genes=catch$annotation, 
               samples = md_filt,
               group = md_filt$Group)

keep <- filterByExpr(dge)
dge.scaled.filtr <- dge[keep, , keep.lib.sizes = FALSE]
dge.scaled.filtr <- calcNormFactors(dge.scaled.filtr)


cpm(dge.scaled.filtr, log = TRUE) %>% 
  t() %>% 
    prcomp(scale = TRUE) %>% 
  tidy() %>% 
  dplyr::rename(names = row) %>% 
  left_join(dge.scaled.filtr$samples, by = "names") %>%
   dplyr::filter(PC %in% 1:3) %>% 
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>% 
  ggplot(
    aes(PC1, PC2, colour = Group, size = lib.size/1e6)
  ) +
  geom_point() +
  geom_text_repel(aes(label = names), show.legend = FALSE) +
  scale_colour_manual(values = group_cols, 
                      labels = c("UPF3A_KD_UPF3B_KD" = "UPF3 dKD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  #scale_size_continuous(limits = c(5, 15), breaks = seq(5, 15, by = 5)) +
  theme_bw() + ggtitle("Transcript Level")
  

```

```{r}
plotMDS(dge.scaled.filtr,col=as.numeric(dge.scaled.filtr$samples$group))

```

## Checking distribution of NIFs in all data 
```{r}
NMD_features = read.table(here::here("data/nif_output/Mus_musculus__nifs.tsv"), sep = "\t", skip =1)

colnames(NMD_features)[1:4] = c("EnsID", "NIF_dEJ", "NIF_uORF", "NIF_long-3'utr")
NMD_features = NMD_features[,1:4]
tx_NMD = NMD_features %>% 
  mutate_at(.vars = "EnsID", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  dplyr::rename("tx_id" = "EnsID") %>% 
  right_join(dge.scaled.filtr$counts %>% 
               as.data.frame() %>%
               rownames_to_column("tx_id") %>%
               mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = ""),
             by="tx_id") 

dEJ = tx_NMD %>% 
  dplyr::filter(NIF_dEJ == "True") %>% 
  dplyr::select(tx_id)


uORF = tx_NMD %>% 
  dplyr::filter(NIF_uORF == "True") %>% 
  dplyr::select(tx_id)
UTR = tx_NMD %>% 
  dplyr::filter(`NIF_long-3'utr` == "True") %>% 
  dplyr::select(tx_id)


x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = tx_NMD$tx_id)

UpSetR::upset(fromList(x), main.bar.color = c("#585A58", "#6d597a", "#eaac8b", "#fcb9b2", 
                                              "#585A58", "#585A58", "#585A58", "#585A58"), 
              sets.bar.color = c("#585A58", "#6d597a", "#eaac8b", "#fcb9b2"),
                  mainbar.y.label = "Overlap of Transcripts with NMD inducing features",
              sets.x.label = "Number of transcripts",
                text.scale = c(1.5,1.6,1.6, 1.6, 1.6), 
              point.size = 4.3)


```


## Differential expression analysis using edgeR

```{r fig.height=6, fig.width=6, fig.cap= "*Biological coeficcient of variation plot against the average abundance of each transcript. The plot shows the square-root estimates of the common, trended and tagwise NB dispersions.*"}
design <-  model.matrix(~0+Group, data = md_filt) %>% 
    set_colnames(gsub(pattern = "Group", replacement ="", x = colnames(.)))

dge.scaled.filtr <- estimateDisp(dge.scaled.filtr,design)

dge.scaled.filtr$common.dispersion

plotBCV(dge.scaled.filtr)
```

```{r fig.height=6, fig.width=6, fig.cap="*Quasi-likelihood dispersion aganist gene abundance. Estimates are shown for raw, tended and squeezed dispersions*"}
fit <- glmQLFit(dge.scaled.filtr,design)
plotQLDisp(fit)
```

```{r}
summary(fit$df.prior)
```


```{r}
contrasts <- makeContrasts(levels = colnames(design), 
                           UPF3B_KD_Control = UPF3B_KD - Control, 
                           UPF3A_KD_Control = UPF3A_KD - Control, 
                           UPF3A_OE_Control = UPF3A_OE - Control, 
                           UPF3A_KD_UPF3B_KD_Control = UPF3A_KD_UPF3B_KD - Control, 
                           UPF3A_OE_UPF3B_KD_Control = UPF3A_OE_UPF3B_KD - Control)


dgeLRT_obj = list()
res_dte = list()
metrics = list()
for (i in 1:ncol(contrasts)){
  temp = glmQLFTest(fit, contrast= contrasts[,i])
  tt = topTags(temp,n = Inf)
  tt %<>% as.data.frame() %>% rownames_to_column("tx_id") %>%
    left_join(decideTests(temp) %>%
                  as.data.frame() %>%
                  set_colnames(c("Results")) %>%
                  rownames_to_column("tx_id"))
  res_dte[[paste0(colnames(contrasts)[i])]] = tt
  dgeLRT_obj[[paste0(colnames(contrasts)[i])]] = temp
  
  setDT(tt)[]
  tt %<>% dplyr::rename("ID" = "tx_id")
  tt <- as.data.frame(tt)
  metrics[[paste0(colnames(contrasts)[i])]] = tt
}
```

### Checking normalization and QC

```{r}
scaled_matrix <- as.data.frame(t(apply(as.matrix(dge.scaled.filtr$counts), 1, scale)))
#scaled_matrix <- as.character(rownames(dge.scaled.filtr$counts))
colLength = length(scaled_matrix)
scaled_matrix<- scaled_matrix[,c(colLength, 1:(colLength-1))]
colnames(scaled_matrix) <- colnames(dge.scaled.filtr$counts)
nID <- which(is.nan(scaled_matrix[,2]))
scaled_matrix[nID,2:length(scaled_matrix)] <- 0
```

```{r}
scaled_matrix %<>% rownames_to_column("ID")

counts = log(dge.scaled.filtr$counts + 0.05)
counts  %<>% as.data.frame() %>%rownames_to_column("ID")
colnames(counts) =c("ID", "Control.1", "UPF3AKD.1", "UPF3AOE.1", "UPF3BKD.1", "UPF3AKDUPF3BKD.1", "UPF3AOEUPF3BKD.1",
                    "Control.2", "UPF3AKD.2", "UPF3AOE.2", "UPF3BKD.2", "UPF3AKDUPF3BKD.2",
                    "Control.3", "UPF3AKD.3", "UPF3AOE.3", "UPF3BKD.3", "UPF3AKDUPF3BKD.3", "UPF3AOEUPF3BKD.2")

names(metrics) = c("UPF3BKD_Control", "UPF3AKD_Control","UPF3AOE_Control", "UPF3AKDUPF3BKD_Control", "UPF3AOEUPF3BKD_Control")
ret = plotSM(counts[c(1:2, 5,8,11,13,16)], dataMetrics = metrics[1], saveFile = FALSE)
ret 
```

```{r}
upf3a_oe = plotSM(counts[c(1:2, 4,8,10,13,15)], dataMetrics = metrics[3], saveFile = FALSE)
upf3a_oe
```

### Distribution of results 
```{r}
plot_hist = function(df, comparison, legend = FALSE){
  ggplot(data= df, aes(x=PValue)) + geom_histogram(fill = "grey", color="black") + theme_bw() + xlab("Pvalue") + ylab("Frequency")
}
plot_fc = function(df){
  df %>% 
    dplyr::filter(FDR < 0.05) %>%  
    ggplot(aes(x=logFC)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white") +  
    geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept =c(log2(1.5), -log2(1.5)),color="red", linetype = "dashed")
}
```

### Fold change distributions 

```{r}
a_f = plot_fc(res_dte$UPF3B_KD_Control) 
b_f = plot_fc(res_dte$UPF3A_KD_Control)
c_f = plot_fc(res_dte$UPF3A_KD_UPF3B_KD_Control)
d_f = plot_fc(res_dte$UPF3A_OE_Control)
e_f = plot_fc(res_dte$UPF3A_OE_UPF3B_KD_Control)
```


```{r fig.height=10, fig.cap = "Distribution of p-values and log2Fold changes across conditions after running edgeR"}
a = plot_hist(res_dte$UPF3B_KD_Control) + ggtitle("(a) UPF3B KD vs Control")
b = plot_hist(res_dte$UPF3A_KD_Control) + ggtitle("(b) UPF3A KD vs Control")
c = plot_hist(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("(c) UPF3 dKD vs Control")
d =  plot_hist(res_dte$UPF3A_OE_Control) + ggtitle("(d) UPF3A OE vs Control")
e =  plot_hist(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("(e) UPF3A OE UPF3B KD vs Control")

#grid.arrange(a,b,c,d,e, ncol = 3, nrow=2)


grid.arrange(a, a_f,
             b,b_f, 
             c, c_f, 
             d, d_f, 
             e,e_f, ncol = 2, nrow=5)
```

```{r}
ma_plot = function(data){
  DEColours <- c("Downregulated" = "#14213d","Upregulated" = "#ff6700",  "Not Sig" = "#e6e8e6")
  temp = data %>% 
    mutate(diffExpressed = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",        
               ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))
  plot = temp %>% 
    ggplot(aes(y= logFC,
               x = logCPM,
               colour = as.factor(as.character(diffExpressed )))) +
    geom_point(alpha = 0.65, size = 3) +
    scale_colour_manual(values = DEColours) + theme_bw()+
    theme(legend.position = "top",
          axis.title.y = element_text(size = 12)) + geom_hline(yintercept = 0, 
                                                               linetype = "dashed", color = "grey") +
    # geom_ysidedensity(aes(x=after_stat(density),  
    #                          color = nif_label, fill = nif_label),  alpha=0.2, size = 1) +
    # scale_fill_manual(values = DEColours) + scale_size_manual(values = c("NIF+" = 3, 
    #                                                                      "NIF-" =2,
    #                                                                      "Not Sig"= 2)) +
    theme(
        axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
         strip.text.x = element_text(
        size = 12, face = "bold.italic", family = "serif"
        )) + labs(color = "Expression", 
                  fill = "",
                  size = "") + scale_alpha(guide = 'none')
  return(plot)
    
}


```


```{r fig.width=10, fig.height=10, fig.cap="MA plot of comprisons of KD/OE with Controls. Pink highlights significantly upregulated transcripts whereas blue highlights significantly downregulated transcripts"}
## Ma plot 

ma_1 = ma_plot(res_dte$UPF3B_KD_Control) + ggtitle("UPF3B KD vs Control")
ma_2 = ma_plot(res_dte$UPF3A_KD_Control) + ggtitle("UPF3A KD vs Control")
ma_3 = ma_plot(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("UPF3 dKD vs Control")
ma_4 = ma_plot(res_dte$UPF3A_OE_Control) + ggtitle("UPF3A OE vs Control")
ma_5 = ma_plot(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("UPF3A OE in UPF3B KD")



ggsave(plot=ma_1, filename = here::here("output/Thesis_plots/DETs/Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_2, filename = here::here("output/Thesis_plots/DETs/Upf3a_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_3, filename = here::here("output/Thesis_plots/DETs/Upf3_dKD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_4, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_5, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE_Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
```

`


### Concordance and discordance of results 


```{r}
names(res_dte)
for_upf3b = names(res_dte)[-c(1)]

overlap_upf3b = list()
for (c in for_upf3b ){
  overlap = res_dte$UPF3B_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>% 
    inner_join(res_dte[[c]] %>% 
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_upf3b[[paste0(c, "_vs_Upf3b_KD")]] = total
  
}

for_upf3a = names(res_dte)[c(3:5)]
overlap_upf3a = list()
for (c in for_upf3a ){
  overlap =res_dte$UPF3A_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>%  
    inner_join(res_dte[[c]] %>%
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_upf3a[[paste0(c, "_vs_Upf3A_KD")]] = total
  
}


for_dkDupf3 = names(res_dte)[c(3,5)]
overlap_dkDupf3 = list()
for (c in for_dkDupf3 ){
  overlap = res_dte$UPF3A_KD_UPF3B_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>%  
    inner_join(res_dte[[c]] %>%
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_dkDupf3 [[paste0(c, "_vs_Upf3_dKD")]] = total
  
}



overlap_upf3b =do.call("rbind",overlap_upf3b) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 
  

overlap_upf3a =do.call("rbind",overlap_upf3a) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 


overlap_dkDupf3  =do.call("rbind",overlap_dkDupf3 ) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 


all_comparisons = rbind(overlap_upf3b,overlap_upf3a, overlap_dkDupf3) %>% 
  mutate(Var1 = factor(Var1, levels = c("Discordant", "Concordance" )), 
         Comparisons = factor(Comparisons, levels =
                                c("UPF3A_OE_UPF3B_KD_Control_vs_Upf3_dKD",
                                  "UPF3A_OE_Control_vs_Upf3_dKD",
                                  "UPF3A_OE_UPF3B_KD_Control_vs_Upf3A_KD",
                                  "UPF3A_OE_Control_vs_Upf3A_KD",
                                  "UPF3A_KD_UPF3B_KD_Control_vs_Upf3A_KD",      
                                  "UPF3A_OE_UPF3B_KD_Control_vs_Upf3b_KD",
                                  "UPF3A_OE_Control_vs_Upf3b_KD",
                                  "UPF3A_KD_UPF3B_KD_Control_vs_Upf3b_KD",
                                  "UPF3A_KD_Control_vs_Upf3b_KD"
                                                      ) )) %>%
  ggplot(aes(x = Comparisons, y = Freq, fill = Var1)) + 
  geom_bar(stat= "identity", alpha = 0.75, width= 0.65) + coord_flip() + scale_fill_manual(values = c( "#2E368F","#F15A2B")) +
  theme_bw() +
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=10, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(y= "Number of Overlapping transcripts",
                  fill = "Expression", 
                  x= "")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps-transcript.svg", plot = all_comparisons, 
       units = c("in"), height = 7, width = 9.26)
```

Significance of overlap and lists 
```{r}

library(GeneOverlap)

enrichment_DEGs_P = list()
enrichment_DEGs_OR = list()


test.list = list()
test.list$UPF3B_KD_Control = res_dte$UPF3B_KD_Control$tx_id[res_dte$UPF3B_KD_Contro$FDR < 0.05]
test.list$UPF3A_KD_Control =  res_dte$UPF3A_KD_Control$tx_id[res_dte$UPF3A_KD_Contro$FDR < 0.05]
test.list$UPF3A_KD_UPF3B_KD_Control =  res_dte$UPF3A_KD_UPF3B_KD_Control$tx_id[res_dte$UPF3A_KD_UPF3B_KD_Control$FDR < 0.05]
test.list$UPF3A_OE_UPF3B_KD_Control = res_dte$UPF3A_OE_UPF3B_KD_Control$tx_id[res_dte$UPF3A_OE_UPF3B_KD_Control$FDR < 0.05]
#test.list$UPF3A_OE_Control = res_dte$UPF3A_OE_Control$tx_id[res_dte$UPF3A_OE_Control$FDR < 0.05]
test.list$BG = res_dte$UPF3B_KD_Control$tx_id

for_upf3b = names(res_dte)[-c(1)]

for (c in for_upf3b ){
  
  go.obj <- newGeneOverlap(test.list$UPF3B_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3b_KD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3b_KD", "_", c)]] = go.obj@odds.ratio 
}
  
for_upf3a = names(res_dte)[c(3:5)]

for (c in for_upf3a ){
    go.obj <- newGeneOverlap(test.list$UPF3A_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3a_KD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3a_KD", "_", c)]] = go.obj@odds.ratio 
}


for_dkDupf3 = names(res_dte)[c(4:5)]
for (c in for_dkDupf3 ){
      go.obj <- newGeneOverlap(test.list$UPF3A_KD_UPF3B_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3_dKD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3_dKD", "_", c)]] = go.obj@odds.ratio 
  
}
    
odds_ratio = enrichment_DEGs_OR %>% do.call(rbind,.) %>% 
  set_colnames("OR") %>% 
  cbind(enrichment_DEGs_P %>% do.call(rbind,.) %>% set_colnames("pvalue")) %>% 
  as.data.frame() %>%
  rownames_to_column("comparisons") %>%
  mutate(comparsions = factor(comparisons, levels = c("Upf3b_KD_UPF3A_vs_Control", 
                                                      "Upf3b_KD_DoubleKD_vs_Control", 
                                                      "Upf3b_KD_UPF3A_OE_UPF3B_KD_vs_Control",
                                                      "Upf3a_KD_DoubleKD_vs_Control", 
                                                      "Upf3a_KD_UPF3A_OE_UPF3B_KD_vs_Control", 
                                                      "Upf3_dKD_UPF3A_OE_UPF3B_KD_vs_Control"))) %>%
  ggplot(aes(x=OR, y = comparisons)) + geom_bar(stat = "identity", fill = "#edafb8") +
  theme_bw() + geom_vline(xintercept = 1) + 
   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=12, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(x="Odds ratio")



#ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps_enrichment.svg", #plot = odds_ratio, 
 #      units = c("in"), height = 7, width = 9.26)

```

```{r}

res_sig_dte = lapply(res_dte, function(x){
  x %<>% 
    dplyr::filter(FDR < 0.05 ) %>% 
    as.data.frame()
})


upf3_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3_paralogs_cor.png", plot = upf3_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3d_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3dKD_Upf3b_cor.png", plot = upf3d_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3bKD_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a OE DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOE_Upf3b_cor.png", plot = upf3a_Upf3bKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3bKD_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a OE Upf3b KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOEKD_Upf3b_cor.png", plot = upf3a_Upf3bKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a OE DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3aOE_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a OE Upf3b KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3aOEKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}
upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOE_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}
upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE Upf3b KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOEKD_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


- Overlaps with DEGs 

```{r}
load(here::here("output/DEG-list.Rda"))


upf3b_cor = res_sig_dte$UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>%
  left_join(tx2gene, by = "tx_id") %>% 
    inner_join(DEGenes$UPF3B_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>%
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3d KD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
  
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3b_cor.png", plot = upf3b_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

upf3a_cor = res_sig_dte$UPF3A_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a KD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3a_cor.png", plot = upf3a_cor  , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)


upf3d_KD_cor = res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$DoubleKD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3 dKD DETs (LFC)") + ylab("Upf3 dKD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3dKD_cor.png", plot = upf3d_KD_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)


upf3a_OE_Upf3b_KD_DEGs_DSGs = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE Upf3b KD DETs (LFC)") + ylab("Upf3a OE Upf3b DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3a_Upf3b_kD_cor.png", plot = upf3a_OE_Upf3b_KD_DEGs_DSGs , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)



upf3b_KD_DEGs_DSGs = res_sig_dte$UPF3B_KD_Control%>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3B_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% 
  reshape2::melt() %>% 
  mutate(Comparison = c("Upf3b DETs"))

upf3a_KD_DEGs_DSGs = res_sig_dte$UPF3A_KD_Control%>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3a DETs"))


upf3d_KD_DEGs_DSGs= res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$DoubleKD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3d KD DETs"))

upf3a_OE_Upf3b_KD_DEGs_DSGs  = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3a OE Upf3b KD DETs"))

DEGs_DETs =rbind(upf3b_KD_DEGs_DSGs, upf3a_KD_DEGs_DSGs, upf3d_KD_DEGs_DSGs, upf3a_OE_Upf3b_KD_DEGs_DSGs ) %>%
   mutate(Comparison= factor(Comparison, levels = c(
     "Upf3a OE Upf3b KD DETs",
     "Upf3d KD DETs",
     "Upf3a DETs",
     "Upf3b DETs"))) %>%
     mutate(Var1 = factor(Var1, levels = c( "Discordant", "Concordance"))) %>%
 ggplot(aes(x = Comparison, y = value, fill = Var1)) +
  geom_bar(stat= "identity", alpha = 0.75, width= 0.65) + coord_flip() + scale_fill_manual(values = c( "#2E368F","#F15A2B")) +
  theme_bw() +
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=10, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(y= "Number of Overlapping transcripts",
                  fill = "Expression",
                  x= "")


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps-transcript-DEGs.svg", plot = DEGs_DETs,
       units = c("in"), height = 7, width = 9.26)
```


```{r}
 res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(res$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID, gene.FC  = logFC), by = "gene_id") %>% 
         with(cor.test(.$logFC, .$gene.FC))
```


```{r}


res_sig_dte = lapply(res_sig_dte, function(x){
  x %<>% 
    mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") 
})


go.obj <- newGeneOverlap(res_sig_dte$UPF3A_OE_UPF3B_KD_Control$gene_id, 
                         DEGenes$UPF3A_OE_UPF3B_KD_vs_Control$EnsID,
                         genome.size=length(res$UPF3B_vs_Control$EnsID))
  go.obj <- testGeneOverlap(go.obj)
 
  go.obj@pval
 go.obj@odds.ratio 
```



```{r}
## Save 
##dge.scaled.filtr
### res_sig_dte
```
