---
title: "Differential transcript expression analysis"
author: "Urwah Nawaz"
date: "2023-03-26"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: console
---

# Introduction


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```

```{r libraries}
source("code/libraries.R")
source("code/functions.R")
library(fishpond)
library(tximeta)
library(GeneOverlap)
library(ggside)
library(tidyquant)
library(stargazer)
library(ggfortify)
library(glue)
library(edgeR)
library(cowplot)
library(broom)
library(glmpca)
library(naniar)
library(bigPint)
```

## Set-up 

```{r tx-setup}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])

gdf = genes(EnsDb.Mmusculus.v79, return.type="DataFrame") %>% as.data.frame()
gene_stats = as.data.frame(gdf[,c("gene_id", "gene_biotype")])

tx2gene =full_join(tx2gene,gdf, by = "gene_id") %>%
  dplyr::select(-c("gene_biotype", "gene_seq_start",
                   "gene_seq_end",   "seq_name",
                   "seq_strand", "seq_coord_system",         "symbol",     "entrezid"))

```

```{r load-res}
load("output/limma-matrices.Rda")
load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")

```



```{r data-setup}
md.files = "data/LTK_Sample Metafile_V3.txt"
salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Salmon")
salmon = list.files(salmon.files, pattern = "transcripts$", full.names = TRUE)
names(salmon) <- paste(c(212:229))
md = read.table(md.files, header= TRUE) %>%  mutate(files = file.path(salmon, "quant.sf"))
comparisons = unique(md$Group)[-1]
colnames(md)[1] = c("names")
md$names = as.character(md$names)
md$Group <- factor(md$Group, 
                            levels=c("Control","UPF3B_KD", "UPF3A_KD", "UPF3A_OE", 
                                     "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
```

```{r}
group_cols <- hcl.colors(
  n = length(unique(md$Group)), 
  palette = "Zissou 1"
  ) %>%
  setNames(unique(md$Group))
```



Transcript-level counts were retrieved using the `catchSalmon()` function from the edgeR package. 
Once retrieved, the counts underwent rigourous QC. 

# Removal of outlier sample 

According to the quality control analysis, it was revealed that one of the samples was not clustering with its condition group. Investigation into GC content, transcript length and library size revealed no contribution of these factors. To ensure we are getting results that are not skewed we will remove the sample from the transcript level analysis. 


```{r include=FALSE}
catch <- catchSalmon(salmon)
colnames(catch$counts) =  paste(c(212:229))

cts.scaled = catch$counts/catch$annotation$Overdispersion

#md_filt = md %>% dplyr::filter(names != "223")


dge <- DGEList(counts = cts.scaled, 
               genes=catch$annotation, 
               samples = md,
               group = md$Group)

keep <- filterByExpr(dge)
dge.scaled.filtr <- dge[keep, , keep.lib.sizes = FALSE]
dge.scaled.filtr <- calcNormFactors(dge.scaled.filtr)

dim(dge.scaled.filtr$counts)

```

```{r fig.height = 7, fig.width=10, fig.cap="*Principal component analysis of transcript level data. PCA was performed on log2 transformed CPM after filtering lowly expressed genes. The PCA shows that samples are clustering close to their conditions based on PC1, however one of the samples of the UPF3A OE in UPF3B KD cell line (sample 223), seems to deviate from its condition group and the rest of the data, so needs to be further investigated*"}
 pca = cpm(dge.scaled.filtr, log = TRUE) %>% 
  t() %>% 
    prcomp(scale = TRUE) %>% 
  tidy() %>% 
  dplyr::rename(names = row) %>% 
  left_join(dge.scaled.filtr$samples, by = "names") %>%
   dplyr::filter(PC %in% 1:3) %>% 
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>% 
  ggplot(
    aes(PC1, PC2, colour = Group, size = lib.size/1e6)
  ) +
  geom_point() +
  geom_text_repel(aes(label = names), show.legend = FALSE) +
  scale_colour_manual(values = group_cols, 
                      labels = c("UPF3A_KD_UPF3B_KD" = "UPF3 dKD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  #scale_size_continuous(limits = c(5, 15), breaks = seq(5, 15, by = 5)) +
  theme_bw() + ggtitle("Transcript Level")



pca
  
```


```{r fig.height=5, fig.width=5}
 plotMDS(dge.scaled.filtr,col=as.numeric(dge.scaled.filtr$samples$group))
```



# Analysis

```{r include=FALSE}
catch <- catchSalmon(salmon[-12])
colnames(catch$counts) =  paste(c(212:222, 224:229))

cts.scaled = catch$counts/catch$annotation$Overdispersion

md_filt = md %>% dplyr::filter(names != "223")


dge <- DGEList(counts = cts.scaled, 
               genes=catch$annotation, 
               samples = md_filt,
               group = md_filt$Group)

keep <- filterByExpr(dge)
dge.scaled.filtr <- dge[keep, , keep.lib.sizes = FALSE]
dge.scaled.filtr <- calcNormFactors(dge.scaled.filtr)


cpm(dge.scaled.filtr, log = TRUE) %>% 
  t() %>% 
    prcomp(scale = TRUE) %>% 
  tidy() %>% 
  dplyr::rename(names = row) %>% 
  left_join(dge.scaled.filtr$samples, by = "names") %>%
   dplyr::filter(PC %in% 1:3) %>% 
  pivot_wider(names_from = "PC", names_prefix = "PC", values_from = "value") %>% 
  ggplot(
    aes(PC1, PC2, colour = Group, size = lib.size/1e6)
  ) +
  geom_point() +
  geom_text_repel(aes(label = names), show.legend = FALSE) +
  scale_colour_manual(values = group_cols, 
                      labels = c("UPF3A_KD_UPF3B_KD" = "UPF3 dKD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  #scale_size_continuous(limits = c(5, 15), breaks = seq(5, 15, by = 5)) +
  theme_bw() + ggtitle("Transcript Level")
  

```

```{r}
plotMDS(dge.scaled.filtr,col=as.numeric(dge.scaled.filtr$samples$group))
#legend("bottomleft", as.character(unique(dge.scaled.filtr$samples$group)), col=1:3, pch=20)

```

## Checking distribution of NIFs in all data 
```{r}
NMD_features = read.table("data/nif_output/Mus_musculus__nifs.tsv", sep = "\t", skip =1)

colnames(NMD_features)[1:4] = c("EnsID", "NIF_dEJ", "NIF_uORF", "NIF_long-3'utr")
NMD_features = NMD_features[,1:4]
tx_NMD = NMD_features %>% 
  mutate_at(.vars = "EnsID", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  dplyr::rename("tx_id" = "EnsID") %>% 
  right_join(dge.scaled.filtr$counts %>% 
               as.data.frame() %>%
               rownames_to_column("tx_id") %>%
               mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = ""),
             by="tx_id") 

dEJ = tx_NMD %>% 
  dplyr::filter(NIF_dEJ == "True") %>% 
  dplyr::select(tx_id)


uORF = tx_NMD %>% 
  dplyr::filter(NIF_uORF == "True") %>% 
  dplyr::select(tx_id)
UTR = tx_NMD %>% 
  dplyr::filter(`NIF_long-3'utr` == "True") %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = tx_NMD$tx_id)

UpSetR::upset(fromList(x), main.bar.color = c("#585A58", "#6d597a", "#eaac8b", "#fcb9b2", 
                                              "#585A58", "#585A58", "#585A58", "#585A58"), 
              sets.bar.color = c("#585A58", "#6d597a", "#eaac8b", "#fcb9b2"),
                  mainbar.y.label = "Overlap of Transcripts with NMD inducing features",
              sets.x.label = "Number of transcripts",
                text.scale = c(1.5,1.6,1.6, 1.6, 1.6), 
              point.size = 4.3)

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Main/NIF-transcriptome-2e.png"  ,  units = c("in"), height = 8, width = 14, dpi=1200)
```


## Differential expression analysis using edgeR

```{r fig.height=6, fig.width=6, fig.cap= "*Biological coeficcient of variation plot against the average abundance of each transcript. The plot shows the square-root estimates of the common, trended and tagwise NB dispersions.*"}
design <-  model.matrix(~0+Group, data = md_filt) %>% 
    set_colnames(gsub(pattern = "Group", replacement ="", x = colnames(.)))

dge.scaled.filtr <- estimateDisp(dge.scaled.filtr,design)

dge.scaled.filtr$common.dispersion

plotBCV(dge.scaled.filtr)
```

```{r fig.height=6, fig.width=6, fig.cap="*Quasi-likelihood dispersion aganist gene abundance. Estimates are shown for raw, tended and squeezed dispersions*"}
fit <- glmQLFit(dge.scaled.filtr,design)
plotQLDisp(fit)
```

```{r}
summary(fit$df.prior)
```


```{r}
contrasts <- makeContrasts(levels = colnames(design), 
                           UPF3B_KD_Control = UPF3B_KD - Control, 
                           UPF3A_KD_Control = UPF3A_KD - Control, 
                           UPF3A_OE_Control = UPF3A_OE - Control, 
                           UPF3A_KD_UPF3B_KD_Control = UPF3A_KD_UPF3B_KD - Control, 
                           UPF3A_OE_UPF3B_KD_Control = UPF3A_OE_UPF3B_KD - Control)


dgeLRT_obj = list()
res_dte = list()
metrics = list()
for (i in 1:ncol(contrasts)){
  temp = glmQLFTest(fit, contrast= contrasts[,i])
  tt = topTags(temp,n = Inf)
  tt %<>% as.data.frame() %>% rownames_to_column("tx_id") %>%
    left_join(decideTests(temp) %>%
                  as.data.frame() %>%
                  set_colnames(c("Results")) %>%
                  rownames_to_column("tx_id"))
  res_dte[[paste0(colnames(contrasts)[i])]] = tt
  dgeLRT_obj[[paste0(colnames(contrasts)[i])]] = temp
  
  setDT(tt)[]
  tt %<>% dplyr::rename("ID" = "tx_id")
  tt <- as.data.frame(tt)
  metrics[[paste0(colnames(contrasts)[i])]] = tt
}
```

### Checking normalization and QC

```{r}
scaled_matrix <- as.data.frame(t(apply(as.matrix(dge.scaled.filtr$counts), 1, scale)))
#scaled_matrix <- as.character(rownames(dge.scaled.filtr$counts))
colLength = length(scaled_matrix)
scaled_matrix<- scaled_matrix[,c(colLength, 1:(colLength-1))]
colnames(scaled_matrix) <- colnames(dge.scaled.filtr$counts)
nID <- which(is.nan(scaled_matrix[,2]))
scaled_matrix[nID,2:length(scaled_matrix)] <- 0
```

```{r}
scaled_matrix %<>% rownames_to_column("ID")

counts = log(dge.scaled.filtr$counts + 0.05)
counts  %<>% as.data.frame() %>%rownames_to_column("ID")
colnames(counts) =c("ID", "Control.1", "UPF3AKD.1", "UPF3AOE.1", "UPF3BKD.1", "UPF3AKDUPF3BKD.1", "UPF3AOEUPF3BKD.1",
                    "Control.2", "UPF3AKD.2", "UPF3AOE.2", "UPF3BKD.2", "UPF3AKDUPF3BKD.2",
                    "Control.3", "UPF3AKD.3", "UPF3AOE.3", "UPF3BKD.3", "UPF3AKDUPF3BKD.3", "UPF3AOEUPF3BKD.2")

names(metrics) = c("UPF3BKD_Control", "UPF3AKD_Control","UPF3AOE_Control", "UPF3AKDUPF3BKD_Control", "UPF3AOEUPF3BKD_Control")
ret = plotSM(counts[c(1:2, 5,8,11,13,16)], dataMetrics = metrics[1], saveFile = FALSE)
ret 
```

```{r}
upf3a_oe = plotSM(counts[c(1:2, 4,8,10,13,15)], dataMetrics = metrics[3], saveFile = FALSE)
upf3a_oe
```

### Distribution of results 
```{r}
plot_hist = function(df, comparison, legend = FALSE){
  ggplot(data= df, aes(x=PValue)) + geom_histogram(fill = "grey", color="black") + theme_bw() + xlab("Pvalue") + ylab("Frequency")
}
plot_fc = function(df){
  df %>% 
    dplyr::filter(FDR < 0.05) %>%  
    ggplot(aes(x=logFC)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white") +  
    geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept =c(log2(1.5), -log2(1.5)),color="red", linetype = "dashed")
}
```

### Fold change distributions 

```{r}
a_f = plot_fc(res_dte$UPF3B_KD_Control) 
b_f = plot_fc(res_dte$UPF3A_KD_Control)
c_f = plot_fc(res_dte$UPF3A_KD_UPF3B_KD_Control)
d_f = plot_fc(res_dte$UPF3A_OE_Control)
e_f = plot_fc(res_dte$UPF3A_OE_UPF3B_KD_Control)
```


```{r fig.height=10, fig.cap = "Distribution of p-values and log2Fold changes across conditions after running edgeR"}
a = plot_hist(res_dte$UPF3B_KD_Control) + ggtitle("(a) UPF3B KD vs Control")
b = plot_hist(res_dte$UPF3A_KD_Control) + ggtitle("(b) UPF3A KD vs Control")
c = plot_hist(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("(c) UPF3 dKD vs Control")
d =  plot_hist(res_dte$UPF3A_OE_Control) + ggtitle("(d) UPF3A OE vs Control")
e =  plot_hist(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("(e) UPF3A OE UPF3B KD vs Control")

#grid.arrange(a,b,c,d,e, ncol = 3, nrow=2)


grid.arrange(a, a_f,
             b,b_f, 
             c, c_f, 
             d, d_f, 
             e,e_f, ncol = 2, nrow=5)
```

```{r}
ma_plot = function(data){
  DEColours <- c("Downregulated" = "#14213d","Upregulated" = "#ff6700",  "Not Sig" = "#e6e8e6")
  temp = data %>% 
    mutate(diffExpressed = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",        
               ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))
  plot = temp %>% 
    ggplot(aes(y= logFC,
               x = logCPM,
               colour = as.factor(as.character(diffExpressed )))) +
    geom_point(alpha = 0.65, size = 3) +
    scale_colour_manual(values = DEColours) + theme_bw()+
    theme(legend.position = "top",
          axis.title.y = element_text(size = 12)) + geom_hline(yintercept = 0, 
                                                               linetype = "dashed", color = "grey") +
    # geom_ysidedensity(aes(x=after_stat(density),  
    #                          color = nif_label, fill = nif_label),  alpha=0.2, size = 1) +
    # scale_fill_manual(values = DEColours) + scale_size_manual(values = c("NIF+" = 3, 
    #                                                                      "NIF-" =2,
    #                                                                      "Not Sig"= 2)) +
    theme(
        axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
         strip.text.x = element_text(
        size = 12, face = "bold.italic", family = "serif"
        )) + labs(color = "Expression", 
                  fill = "",
                  size = "") + scale_alpha(guide = 'none')
  return(plot)
    
}


```


```{r fig.width=10, fig.height=10, fig.cap="MA plot of comprisons of KD/OE with Controls. Pink highlights significantly upregulated transcripts whereas blue highlights significantly downregulated transcripts"}
## Ma plot 

ma_1 = ma_plot(res_dte$UPF3B_KD_Control) + ggtitle("UPF3B KD vs Control")
ma_2 = ma_plot(res_dte$UPF3A_KD_Control) + ggtitle("UPF3A KD vs Control")
ma_3 = ma_plot(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("UPF3 dKD vs Control")
ma_4 = ma_plot(res_dte$UPF3A_OE_Control) + ggtitle("UPF3A OE vs Control")
ma_5 = ma_plot(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("UPF3A OE in UPF3B KD")



ggsave(plot=ma_1, filename = here::here("output/Thesis_plots/DETs/Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_2, filename = here::here("output/Thesis_plots/DETs/Upf3a_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_3, filename = here::here("output/Thesis_plots/DETs/Upf3_dKD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_4, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_5, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE_Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
```


```{r fig.height= 5, fig.width=5, fig.cap="*Ratio of up/down differentially expressed transcripts*"}
 ratios = data.frame()
 for (c in names(res_dte)){
   temp = table(res_dte[[c]]$Results) %>% 
     reshape2::melt() %>% 
     dplyr::filter(Var1 != 0) 
   ratio= temp$value[2]/temp$value[1]
   ratio = data.frame(Comparisons = c, 
            ratio = ratio)
   ratios = rbind(ratios, ratio)
   }

ratios$Comparisons = factor(ratios$Comparisons, levels = c("UPF3B_KD_Control", 
                                                           "UPF3A_KD_Control", 
                                                           "UPF3A_KD_UPF3B_KD_Control", 
                                                           "UPF3A_OE_Control", 
                                                           "UPF3A_OE_UPF3B_KD_Control"))

ratios_fig = ratios  %>%
  ggplot(aes(x =Comparisons, y = ratio)) + 
  geom_bar(stat = "identity",  colour = "white", width = .75, fill = "#e64055") + 
    theme_classic() + 
    theme(axis.text.x = element_text(angle = 70,  hjust=1, face = "bold"), 
          legend.position = "none", 
          legend.title= element_blank(), axis.title.y = element_text(size = 12), 
          axis.text.y = element_text(size = 14, face = "bold")) + 
    ylab("ratio up/down") + xlab("") + 
   scale_x_discrete(labels=c("UPF3A_KD_UPF3B_KD_Control" = "Upf3 dKD", "UPF3A_KD_Control" = "Upf3a KD", 
                                                            "UPF3B_KD_Control" = "Upf3b KD", "UPF3A_OE_Control" = "Upf3a OE", 
                                                            "UPF3A_OE_UPF3B_KD_Control" = "Upf3a OE Upf3b KD"))

# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/ratio_DET.svg", plot = ratios_fig, 
#        units = c("in"), height = 4, width = 4)
```


### Concordance and discordance of results 


```{r}
names(res_dte)
for_upf3b = names(res_dte)[-c(1)]

overlap_upf3b = list()
for (c in for_upf3b ){
  overlap = res_dte$UPF3B_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>% 
    inner_join(res_dte[[c]] %>% 
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_upf3b[[paste0(c, "_vs_Upf3b_KD")]] = total
  
}

for_upf3a = names(res_dte)[c(3:5)]
overlap_upf3a = list()
for (c in for_upf3a ){
  overlap =res_dte$UPF3A_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>%  
    inner_join(res_dte[[c]] %>%
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_upf3a[[paste0(c, "_vs_Upf3A_KD")]] = total
  
}


for_dkDupf3 = names(res_dte)[c(3,5)]
overlap_dkDupf3 = list()
for (c in for_dkDupf3 ){
  overlap = res_dte$UPF3A_KD_UPF3B_KD_Control %>% 
    dplyr::filter(FDR < 0.05) %>%  
    inner_join(res_dte[[c]] %>%
                 dplyr::filter(FDR < 0.05), by="tx_id") %>% 
      mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) 
  total = table(overlap$Label) %>% as.data.frame() 
  overlap_dkDupf3 [[paste0(c, "_vs_Upf3_dKD")]] = total
  
}



overlap_upf3b =do.call("rbind",overlap_upf3b) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 
  

overlap_upf3a =do.call("rbind",overlap_upf3a) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 


overlap_dkDupf3  =do.call("rbind",overlap_dkDupf3 ) %>% 
  rownames_to_column("Comparisons") %>% 
   mutate_at(.vars = "Comparisons", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") 


all_comparisons = rbind(overlap_upf3b,overlap_upf3a, overlap_dkDupf3) %>% 
  mutate(Var1 = factor(Var1, levels = c("Discordant", "Concordance" )), 
         Comparisons = factor(Comparisons, levels =
                                c("UPF3A_OE_UPF3B_KD_Control_vs_Upf3_dKD",
                                  "UPF3A_OE_Control_vs_Upf3_dKD",
                                  "UPF3A_OE_UPF3B_KD_Control_vs_Upf3A_KD",
                                  "UPF3A_OE_Control_vs_Upf3A_KD",
                                  "UPF3A_KD_UPF3B_KD_Control_vs_Upf3A_KD",      
                                  "UPF3A_OE_UPF3B_KD_Control_vs_Upf3b_KD",
                                  "UPF3A_OE_Control_vs_Upf3b_KD",
                                  "UPF3A_KD_UPF3B_KD_Control_vs_Upf3b_KD",
                                  "UPF3A_KD_Control_vs_Upf3b_KD"
                                                      ) )) %>%
  ggplot(aes(x = Comparisons, y = Freq, fill = Var1)) + 
  geom_bar(stat= "identity", alpha = 0.75, width= 0.65) + coord_flip() + scale_fill_manual(values = c( "#2E368F","#F15A2B")) +
  theme_bw() +
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=10, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(y= "Number of Overlapping transcripts",
                  fill = "Expression", 
                  x= "")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps-transcript.svg", plot = all_comparisons, 
       units = c("in"), height = 7, width = 9.26)
```

Significance of overlap and lists 
```{r}

library(GeneOverlap)

enrichment_DEGs_P = list()
enrichment_DEGs_OR = list()


test.list = list()
test.list$UPF3B_KD_Control = res_dte$UPF3B_KD_Control$tx_id[res_dte$UPF3B_KD_Contro$FDR < 0.05]
test.list$UPF3A_KD_Control =  res_dte$UPF3A_KD_Control$tx_id[res_dte$UPF3A_KD_Contro$FDR < 0.05]
test.list$UPF3A_KD_UPF3B_KD_Control =  res_dte$UPF3A_KD_UPF3B_KD_Control$tx_id[res_dte$UPF3A_KD_UPF3B_KD_Control$FDR < 0.05]
test.list$UPF3A_OE_UPF3B_KD_Control = res_dte$UPF3A_OE_UPF3B_KD_Control$tx_id[res_dte$UPF3A_OE_UPF3B_KD_Control$FDR < 0.05]
#test.list$UPF3A_OE_Control = res_dte$UPF3A_OE_Control$tx_id[res_dte$UPF3A_OE_Control$FDR < 0.05]
test.list$BG = res_dte$UPF3B_KD_Control$tx_id

for_upf3b = names(res_dte)[-c(1)]

for (c in for_upf3b ){
  
  go.obj <- newGeneOverlap(test.list$UPF3B_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3b_KD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3b_KD", "_", c)]] = go.obj@odds.ratio 
}
  
for_upf3a = names(res_dte)[c(3:5)]

for (c in for_upf3a ){
    go.obj <- newGeneOverlap(test.list$UPF3A_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3a_KD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3a_KD", "_", c)]] = go.obj@odds.ratio 
}


for_dkDupf3 = names(res_dte)[c(4:5)]
for (c in for_dkDupf3 ){
      go.obj <- newGeneOverlap(test.list$UPF3A_KD_UPF3B_KD_Control, 
                         test.list[[c]],
                         genome.size=length(test.list$BG))
  go.obj <- testGeneOverlap(go.obj)
  enrichment_DEGs_P[[paste0("Upf3_dKD", "_", c)]] = go.obj@pval
  enrichment_DEGs_OR[[paste0("Upf3_dKD", "_", c)]] = go.obj@odds.ratio 
  
}
    
odds_ratio = enrichment_DEGs_OR %>% do.call(rbind,.) %>% 
  set_colnames("OR") %>% 
  cbind(enrichment_DEGs_P %>% do.call(rbind,.) %>% set_colnames("pvalue")) %>% 
  as.data.frame() %>%
  rownames_to_column("comparisons") %>%
  mutate(comparsions = factor(comparisons, levels = c("Upf3b_KD_UPF3A_vs_Control", 
                                                      "Upf3b_KD_DoubleKD_vs_Control", 
                                                      "Upf3b_KD_UPF3A_OE_UPF3B_KD_vs_Control",
                                                      "Upf3a_KD_DoubleKD_vs_Control", 
                                                      "Upf3a_KD_UPF3A_OE_UPF3B_KD_vs_Control", 
                                                      "Upf3_dKD_UPF3A_OE_UPF3B_KD_vs_Control"))) %>%
  ggplot(aes(x=OR, y = comparisons)) + geom_bar(stat = "identity", fill = "#edafb8") +
  theme_bw() + geom_vline(xintercept = 1) + 
   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=12, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(x="Odds ratio")



#ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps_enrichment.svg", #plot = odds_ratio, 
 #      units = c("in"), height = 7, width = 9.26)

```

```{r}

res_sig_dte = lapply(res_dte, function(x){
  x %<>% 
    dplyr::filter(FDR < 0.05 ) %>% 
    as.data.frame()
})


upf3_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3_paralogs_cor.png", plot = upf3_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3d_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3dKD_Upf3b_cor.png", plot = upf3d_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3bKD_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a OE DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOE_Upf3b_cor.png", plot = upf3a_Upf3bKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3bKD_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3a OE Upf3b KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOEKD_Upf3b_cor.png", plot = upf3a_Upf3bKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a OE DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3aOE_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a OE Upf3b KD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aKD_Upf3aOEKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}
upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOE_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


```{r}
upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE Upf3b KD DETs (LFC)") + ylab("Upf3 dKD DETs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DETs/Upf3aOEKD_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

```


- Overlaps with DEGs 

```{r}
load("output/DEG-list.Rda")


upf3b_cor = res_sig_dte$UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>%
  left_join(tx2gene, by = "tx_id") %>% 
    inner_join(DEGenes$UPF3B_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>%
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3b KD DETs (LFC)") + ylab("Upf3d KD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
  
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3b_cor.png", plot = upf3b_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)

upf3a_cor = res_sig_dte$UPF3A_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a KD DETs (LFC)") + ylab("Upf3a KD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3a_cor.png", plot = upf3a_cor  , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)


upf3d_KD_cor = res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$DoubleKD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3 dKD DETs (LFC)") + ylab("Upf3 dKD DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3dKD_cor.png", plot = upf3d_KD_cor , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)


upf3a_OE_Upf3b_KD_DEGs_DSGs = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab("Upf3a OE Upf3b KD DETs (LFC)") + ylab("Upf3a OE Upf3b DEGs (LFC)") +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 13, family = "serif", color = "black"),
            axis.text.y = element_text(size = 13, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/overlaps/DEGs-DETs/Upf3a_Upf3b_kD_cor.png", plot = upf3a_OE_Upf3b_KD_DEGs_DSGs , 
       units = c("in"), height = 4.65, width = 4, dpi = 1200)



upf3b_KD_DEGs_DSGs = res_sig_dte$UPF3B_KD_Control%>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3B_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% 
  reshape2::melt() %>% 
  mutate(Comparison = c("Upf3b DETs"))

upf3a_KD_DEGs_DSGs = res_sig_dte$UPF3A_KD_Control%>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3a DETs"))


upf3d_KD_DEGs_DSGs= res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$DoubleKD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3d KD DETs"))

upf3a_OE_Upf3b_KD_DEGs_DSGs  = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  with(table(.$Label)) %>% reshape2::melt() %>% 
  mutate(Comparison = c("Upf3a OE Upf3b KD DETs"))

DEGs_DETs =rbind(upf3b_KD_DEGs_DSGs, upf3a_KD_DEGs_DSGs, upf3d_KD_DEGs_DSGs, upf3a_OE_Upf3b_KD_DEGs_DSGs ) %>%
   mutate(Comparison= factor(Comparison, levels = c(
     "Upf3a OE Upf3b KD DETs",
     "Upf3d KD DETs",
     "Upf3a DETs",
     "Upf3b DETs"))) %>%
     mutate(Var1 = factor(Var1, levels = c( "Discordant", "Concordance"))) %>%
 ggplot(aes(x = Comparison, y = value, fill = Var1)) +
  geom_bar(stat= "identity", alpha = 0.75, width= 0.65) + coord_flip() + scale_fill_manual(values = c( "#2E368F","#F15A2B")) +
  theme_bw() +
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=10, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(y= "Number of Overlapping transcripts",
                  fill = "Expression",
                  x= "")


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/all_overlaps-transcript-DEGs.svg", plot = DEGs_DETs,
       units = c("in"), height = 7, width = 9.26)
```


```{r}
 res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(res$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID, gene.FC  = logFC), by = "gene_id") %>% 
         with(cor.test(.$logFC, .$gene.FC))
```


```{r}


res_sig_dte = lapply(res_sig_dte, function(x){
  x %<>% 
    mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") 
})


go.obj <- newGeneOverlap(res_sig_dte$UPF3A_OE_UPF3B_KD_Control$gene_id, 
                         DEGenes$UPF3A_OE_UPF3B_KD_vs_Control$EnsID,
                         genome.size=length(res$UPF3B_vs_Control$EnsID))
  go.obj <- testGeneOverlap(go.obj)
 
  go.obj@pval
 go.obj@odds.ratio 
```

## DET statistics 
```{r}
NMD_features = read.table("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/data/nif_output/Mus_musculus__nifs.tsv", sep = "\t", skip =1)

colnames(NMD_features) = c("tx_id",	"NIF_dEJ", "NIF_uORF",	"NIF_long_3utr", "NIF|__|GC(cds[-10:termination_codon])",	"GC_5utr",	"GC_cds",	"GC_3utr",	"GC_exons",	"GC_introns",	"LENGTH_5utr", 	"LENGTH_cds",	"LENGTH_3utr",	"LENGTH_exons", "LENGTH_introns", "EXONS_count",	"dEJs|__|penultimate-exon>=55(nts)",	"uORFs|__|count",	"uORFs|__|gc-content",	"uORFs|__|lengths", "uORFs|__|max-length")

NMD_features$NIF_dEJ = ifelse(NMD_features$NIF_dEJ == "True",gsub("True", 1, NMD_features$NIF_dEJ), 0)
NMD_features$NIF_uORF = ifelse(NMD_features$NIF_uORF == "True",gsub("True", 1, NMD_features$NIF_uORF), 0)
NMD_features$NIF_long_3utr = ifelse(NMD_features$NIF_long_3utr == "True",gsub("True", 1, NMD_features$NIF_long_3utr), 0)
NMD_features %<>% mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")

NMD_features$NIF_dEJ[is.na(NMD_features$NIF_dEJ)] <- 0
NMD_features$NIF_uORF[is.na(NMD_features$NIF_uORF)] <- 0
NMD_features$NIF_long_3utr[is.na(NMD_features$NIF_long_3utr)] <- 0
```



```{r include=FALSE}
total = data.frame()
for (i in names(res_dte)){
    name = i
    #comparison = rep(name, 4) %>% as.data.frame()
    temp = res_dte[[i]]
    
    message("Merging DE results with NMD table for", paste(name))
    temp %<>%  as.data.frame() %>%
      mutate(tx_version = tx_id) %>%
      mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
      left_join(NMD_features, by = "tx_id")
    
    message("Summing all NMD values to determine number of NIFs per transcript for", paste0(name))
    temp$nif_feature = ifelse(rowSums(sapply(temp[, c(12:14)],
                       function(x) as.numeric(as.character(x)))) == 2 ,"Two",
                       ifelse(rowSums(sapply(temp[, c(12:14)],
                              function(x) as.numeric(as.character(x)))) == 3, "Three", 
                              ifelse(rowSums(sapply(temp[, c(12:14)],
                                                    function(x) as.numeric(as.character(x)))) == 1, "One", "None")))
    
    temp$nif_feature %<>%  replace_na("None")
    res_dte[[paste(name)]] = temp
    
    ## downregulation 
    message("Subsetting downregulated genes for ", paste0(name))
    down = temp %>%  as.data.frame() %>%  
        dplyr::filter(FDR < 0.05 & logFC < 0) 
    down = table(down$nif_feature)
    print(nrow(down))
    repetition = down %>% nrow() %>% as.integer()
    expression = rep(c("down"), repetition) %>% as.data.frame()
    comparison = rep(name, nrow(expression)) %>% as.data.frame()
    down = cbind(down, expression = expression)
    down = cbind(down, comparison = comparison)
    colnames(down) = c("NIF", "Value", "Expression", "Condition")
    
    ## upregulation 
    message("Subsetting upregulated genes for ", paste0(name))
    up = temp %>%  as.data.frame() %>%  
        dplyr::filter(FDR < 0.05 & logFC > 0 ) 
    up = table(up$nif_feature)
    print(dim(up))
    repetition = up %>% nrow() %>% as.integer()
    expression = rep(c("up"), repetition) %>% as.data.frame()
    comparison = rep(name, nrow(expression)) %>% as.data.frame()
    up = cbind(up, expression = expression)
    up = cbind(up, comparison = comparison)
    colnames(up) = c("NIF", "Value", "Expression", "Condition")
    
    ## bind everything together 
    message("Binding everything together")
    #together = rbind(up, down)
   
    #total = rbind(total,as.data.frame(up,down))
    total = rbind(total,up,down, stringsAsFactors = FALSE)
}
```

```{r fig.width=10}
total$NIF <- factor(total$NIF, levels = c("Three", "Two", "One", "None"))
up = total %>% 
  dplyr::filter(Expression == "up") %>% 
  ggplot(aes(x= Condition, y = Value, fill=NIF)) + geom_bar(stat = "identity", width = .5,color = "white") + coord_flip() +
  theme_bw() + ylab("Number of Upregulated Transcripts") + xlab("") + 
  scale_fill_manual(values = c("#FF8DD0","#E0155F", "#C40C71", "#370059")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"),
        legend.position = "bottom") + ylim(0, 4500) + 
  labs(fill = "Number of NMD inducible features (NIF)", labels = c("No NIF", "One NIF feature", "Two NIF features" , "Three NIF features"))

#ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/total_exp_up.svg", plot = up, 
 #      height = 6.31, width = 5.77, units = "in")


down = total %>% 
  dplyr::filter(Expression == "down") %>% 
  ggplot(aes(x= Condition, y = Value, fill=NIF)) + 
  geom_bar(stat = "identity", width = .5,color = "white") + coord_flip() +
  theme_bw() + ylab("Number of Downregulated Transcripts") + xlab("") + 
  scale_fill_manual(values = c("#FF8DD0","#E0155F", "#C40C71", "#370059")) +
  theme(axis.text.x = element_text(size = 10, face = "bold"), 
        legend.position = "none") + scale_y_reverse(limits=c(4500,0))  + 
  labs(fill = "Number of NMD inducible features (NIF)", labels = c("No NIF", "One NIF feature", "Two NIF features" , "Three NIF features"))



ggarrange(down, up, common.legend = TRUE)
#ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/total_exp_down.svg", plot = down, 
#       height = 6.31, width = 5.77, units = "in")
```


```{r}
up = total %>% 
  dplyr::filter(Expression == "up") %>% 
  ggplot(aes(x= Condition, y = Value, fill=NIF)) + geom_bar(position ="fill", stat = "identity",color = "white") + coord_flip() +
  theme_bw() + ylab("Proportion of Upregulated Transcripts") + xlab("") + 
  scale_fill_manual(values = c("#FF8DD0","#E0155F", "#C40C71", "#370059")) +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +   
  labs(fill = "Number of NMD inducible features (NIF)", 
       labels = c("No NIF", "One NIF feature", "Two NIF features" , "Three NIF features"))

down = total %>% 
  dplyr::filter(Expression == "down") %>% 
  ggplot(aes(x= Condition, y = Value, fill=NIF)) + 
  geom_bar(position="fill", stat="identity",color = "white") + coord_flip() +
  theme_bw() + ylab("Proportion of Downregulated Transcripts ") + xlab("") + 
  scale_fill_manual(values = c("#FF8DD0","#E0155F", "#C40C71", "#370059")) +
  theme(axis.text.x = element_text(size = 10, face = "bold")) + scale_y_reverse() +   
  labs(fill = "Number of NMD inducible features (NIF)", 
       labels = c("No NIF", "One NIF feature", "Two NIF features" , "Three NIF features"))


ggarrange(down, up,  common.legend = TRUE)
```





```{r}

res_sig = list()
res_sig$upf3b_kd = res_dte$UPF3B_KD_Control %>%  dplyr::filter(FDR < 0.05) %>%  left_join(tx2gene, by = "tx_id")
res_sig$upf3a_kd = res_dte$UPF3A_KD_Control %>%  dplyr::filter(FDR < 0.05) %>%  left_join(tx2gene, by = "tx_id")
res_sig$upf3_dkd = res_dte$UPF3A_KD_UPF3B_KD_Control %>%  dplyr::filter(FDR < 0.05) %>%  left_join(tx2gene, by = "tx_id")
res_sig$upf3a_oe = res_dte$UPF3A_OE_Control %>%  dplyr::filter(FDR < 0.05) %>%  left_join(tx2gene, by = "tx_id")
res_sig$upf3a_oe_upf3b_kd = res_dte$UPF3A_OE_UPF3B_KD_Control %>%  dplyr::filter(FDR < 0.05) %>%  left_join(tx2gene, by = "tx_id")
write.xlsx(res_sig, file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DET-analysis.xlsx")

```



```{r fig.width=12, fig.cap= "Number of NIF features in differentially expressed transcripts per comparison"}
nif_features = list()
for (i in names(res_dte)[-3]){
  name = i
  temp = res_dte[[i]]
  
  message("Counting NIFs in upregulated transcripts for ", name)
  sig_up = temp %>%  as.data.frame() %>%  
        dplyr::filter(FDR < 0.05, logFC > log2(1)) 
  utr = table(sig_up$NIF_long_3utr)[2]
  dej = table(sig_up$NIF_dEJ)[2]
  uorf = table(sig_up$NIF_uORF)[2]
  none = table(sig_up$nif_feature)[1]
  total_nif_up  = rbind(utr, dej, uorf, none) %>% 
    as.data.frame() %>% 
    rownames_to_column("NIF_feature") %>% 
    mutate(Condition = name, 
           Expression = "Up") %>% 
    dplyr::rename("Count" = "1")
  
  message("Counting NIFs in downregulated transcripts for ", name)
  
  sig_down = temp %>%  as.data.frame() %>%  
        dplyr::filter(FDR < 0.05, logFC < log2(1)) 
  utr = table(sig_down$NIF_long_3utr)[2]
  dej = table(sig_down$NIF_dEJ)[2]
  uorf = table(sig_down$NIF_uORF)[2]
   none = table(sig_down$nif_feature)[1]
  total_nif_down  = rbind(utr, dej, uorf, none) %>% 
    as.data.frame() %>% 
    rownames_to_column("NIF_feature") %>% 
    mutate(Condition = name, 
           Expression = "Down") %>% 
    dplyr::rename("Count" = "1")
  
  nif_features= rbind(nif_features, total_nif_up, total_nif_down, stringsAsFactors = FALSE)
  
}

OE_df = data.frame("NIF_feature" = "")

### For overexprssion - up
## 0 for long 3'UTR
## 0 for dEJ 
## 1 for uORF

### down 
## 3 long utr
## 0 dej
### 1 for uORF

OE_df = data.frame("NIF_feature" = rep(c("utr", "dej", "uorf"), 2), 
                   "Count" = c(0,0, 1,3,0, 1), 
                   "Condition" = rep(c("UPF3A_OE_Control")),
                   "Expression" = c(rep("Up",3), rep("Down", 3)))

nif_features= rbind(nif_features, OE_df, stringsAsFactors = FALSE)


condition.labs <- c("Upf3b KD", "Upf3a KD", "Upf3 dKD", "Upf3a OE", "Upf3a OE, Upf3b KD")
names(condition.labs) <- c("UPF3B_KD", "UPF3A_KD", "UPF3A_KD_UPF3B_KD",
                      "UPF3A_OE", "UPF3A_OE_UPF3B_KD") 
nif_number = nif_features %>%
  mutate(NIF_feature = gsub("utr", "Long 3'UTR", NIF_feature), 
         NIF_feature = gsub("uorf", "Upstream ORF", NIF_feature), 
         NIF_feature = gsub("dej", "dEJC", NIF_feature), 
         Condition = gsub("_Control", "", Condition)) %>%
  mutate(Condition = factor(Condition, levels = c("UPF3B_KD",
                                                  "UPF3A_KD", 
                                                  "UPF3A_KD_UPF3B_KD", 
                                                  "UPF3A_OE",
                                                  "UPF3A_OE_UPF3B_KD")), 
         Expression = factor(Expression, levels = c("Down", "Up")),
          NIF_feature = factor(NIF_feature, levels = c("none",
                                                             
                                                             "Upstream ORF",
                                                             "Long 3'UTR",
                                                       "dEJC"))) %>%
  
  ggplot(aes(x= NIF_feature, y = Count, fill = Expression)) + 
  geom_bar(stat = "identity", position = "stack", width = .5) + coord_flip() + 
  theme_bw() +
  facet_grid(~Condition,
             labeller = labeller(Condition = condition.labs))  +
  xlab("") + ylab("") + scale_fill_manual(values = c("Up"= "#b23a48",
                                                     "Down" = "grey")) + 
   
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
         strip.text.x = element_text(
        size = 12, face = "italic", family = "serif"
        ))

nif_number
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Main/Count_NIF_2f.png",
       plot = nif_number  ,  units = c("in"), height = 5, width = 15)

nif_prop =nif_features %>%
  mutate(NIF_feature = gsub("utr", "Long 3'UTR", NIF_feature), 
         NIF_feature = gsub("uorf", "Upstream ORF", NIF_feature), 
         NIF_feature = gsub("dej", "Downstream exon junction", NIF_feature), 
         Condition = gsub("_Control", "", Condition)) %>%
  mutate(Condition = factor(Condition, levels = c("UPF3B_KD",
                                                  "UPF3A_KD", 
                                                  "UPF3A_KD_UPF3B_KD", 
                                                  "UPF3A_OE",
                                                  "UPF3A_OE_UPF3B_KD")), 
         Expression = factor(Expression, levels = c("Down", "Up")),
          NIF_feature = factor(NIF_feature, levels = c("none",
                                                             
                                                             "uorf",
                                                             "Long 3'UTR",
                                                       "dej"))) %>%
  
  ggplot(aes(x= NIF_feature, y = Count, fill = Expression)) + 
  geom_bar(stat = "identity", position = "fill", width = .5) + coord_flip() + 
  theme_bw() +
  facet_grid(~Condition,
             labeller = labeller(Condition = condition.labs))  +
  xlab("") + ylab("") + scale_fill_manual(values = c("Up"= "#b23a48",
                                                     "Down" = "grey")) + 
   
     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
         strip.text.x = element_text(
        size = 12, face = "bold.italic", family = "serif"
        )) 

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Proprotion_NIF.svg",
       plot = nif_prop  ,  units = c("in"), height = 5, width = 14)
```

* Enrichment analysis of NIF+ features 

```{r}
# Calculating enrichment for this work 
### Ask Irina about adjusted pvalues 
enrichment_NIF_pvalue = list()
enrichment_NIF_OR = list()
for (c in names(res_dte)){
  test.list = list()
  test.list$BG = res_dte[[c]]$tx_id
  test.list$NIF = res_dte[[c]]$tx_id[res_dte[[c]]$nif_feature != "None"]
  test.list$sig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05]
  test.list$up = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC > 0]
  test.list$sig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC < 0]
  
  for (det in c("sig", "up", "down")) {
    go.obj <- newGeneOverlap(test.list[[det]],
                         test.list$NIF,
                         genome.size=length(test.list$BG))
    go.obj <- testGeneOverlap(go.obj)
    enrichment_NIF_pvalue[[paste0(c, "_", det)]] = go.obj@pval
    enrichment_NIF_OR[[paste0(c, "_", det)]] = go.obj@odds.ratio
    
  }
}

enrichment_NIF_pvalue = do.call(rbind, enrichment_NIF_pvalue) %>% as.data.frame()
colnames(enrichment_NIF_pvalue)[1] = c("pvalue")
enrichment_NIF_pvalue$padj = p.adjust(enrichment_NIF_pvalue$pvalue, method = "BH")
enrichment_NIF_OR  = do.call(rbind, enrichment_NIF_OR ) %>% as.data.frame()
colnames(enrichment_NIF_OR)[1] = c("OR")
NIF_enrichment = cbind(enrichment_NIF_pvalue, enrichment_NIF_OR)
```

```{r}
# NIF_enrichment %<>% rownames_to_column("condition")
# NIF_enrichment$Condition[grepl("UPF3B_KD_Control_", NIF_enrichment$condition)] = "UPF3B KD vs Controls"
# NIF_enrichment$Condition[grepl("UPF3A_KD_Control_", NIF_enrichment$condition)] = "UPF3A KD vs Controls"
# NIF_enrichment$Condition[grepl("UPF3A_OE_Control", NIF_enrichment$condition)] = "UPF3A OE vs Controls"
# NIF_enrichment$Condition[grepl("UPF3A_KD_UPF3B_KD_Control", NIF_enrichment$condition)] = "Double KD vs Controls"
# NIF_enrichment$Condition[grepl("UPF3A_OE_UPF3B_KD_Control", NIF_enrichment$condition)] = "UPF3A OE UPF3B KD vs Controls"
# NIF_enrichment$expression[grepl("_up", NIF_enrichment$condition)] = "Up"
# NIF_enrichment$expression[grepl("_down", NIF_enrichment$condition)] = "Down"
# NIF_enrichment$expression[grepl("_sig", NIF_enrichment$condition)] = "Sig"
# 
# NIF_enrichment$sig = ifelse(NIF_enrichment$pvalue < 0.05, "Sig", "Not Sig")
# 
# NIF_enrichment_plot = NIF_enrichment %>% 
#   mutate(Condition = factor(Condition, levels = c("UPF3B KD vs Controls", 
#                                                   "UPF3A KD vs Controls",
#                                                   "Double KD vs Controls", 
#                                                   "UPF3A OE vs Controls", 
#                                                   "UPF3A OE UPF3B KD vs Controls"))) %>%
#   ggplot(aes(y=condition, x=OR, color = expression)) +  
#    theme_bw()  +
#   geom_point(size = 6, aes(shape=sig)) + facet_grid(rows = vars(Condition),  scales = "free") +
#   theme(axis.text.y = element_blank()) + ylab("") + xlab("Odds ratio") + 
#   scale_color_manual(values=c("#2F124B", "#F1676B",  ))
# 
# 
# NIF_enrichment_plot
# 
# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/All_OR.svg",
#        plot = NIF_enrichment_plot ,  units = c("in"), height = 5.58, width = 4.69)

```



```{r}
enrichment_nif_pval = list()
enrichment_nif_OR = list()
nifs_list = c("NIF_uORF", "NIF_dEJ", "NIF_utr")

for (c in names(res_dte)){
  test.list = list()
  test.list$BG = res_dte[[c]]$tx_id
  test.list$NIF_uORF = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_uORF== 1 ]
  test.list$NIF_dEJ = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_dEJ== 1 ]
  test.list$NIF_utr = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_long_3utr== 1 ]
  test.list$sig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05]
  test.list$up = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC > 0]
  test.list$down = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC < 0]
  test.list$notsig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR > 0.05]
  for (nif in nifs_list){
     go.obj <- newGeneOverlap(test.list$notsig,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c, "_", nif, "_notsig")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c, "_", nif, "_notsig")]] = go.obj@odds.ratio
     go.obj <- newGeneOverlap(test.list$up,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c,"_",nif,  "_up")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c,"_",nif, "_up")]] = go.obj@odds.ratio
     go.obj <- newGeneOverlap(test.list$down,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c,"_",nif, "_down")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c,"_",nif, "_down")]] = go.obj@odds.ratio
    
  }
}


enrichment_nif_pval = do.call(rbind, enrichment_nif_pval) %>% as.data.frame()
colnames(enrichment_nif_pval)[1] = c("pvalue")
enrichment_nif_pval$padj = p.adjust(enrichment_nif_pval$pvalue, method = "BH")
enrichment_nif_OR  = do.call(rbind, enrichment_nif_OR) %>% as.data.frame()
colnames(enrichment_nif_OR)[1] = c("OR")
NIF_feature_enrichment = cbind(enrichment_nif_pval, enrichment_nif_OR)
```


```{r}
NIF_feature_enrichment %<>% rownames_to_column("condition")
NIF_feature_enrichment$Condition[grepl("UPF3B_KD_Control_", NIF_feature_enrichment$condition)] = "Upf3b KD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_KD_Control_", NIF_feature_enrichment$condition)] = "Upf3a KD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_OE_Control", NIF_feature_enrichment$condition)] = "Upf3a OE targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_KD_UPF3B_KD_Control", NIF_feature_enrichment$condition)] = "Upf3 dKD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_OE_UPF3B_KD_Control", NIF_feature_enrichment$condition)] = "Upf3a OE Upf3b KD targets"
NIF_feature_enrichment$expression[grepl("_up", NIF_feature_enrichment$condition)] = "Up"
NIF_feature_enrichment$expression[grepl("_down", NIF_feature_enrichment$condition)] = "Down"
NIF_feature_enrichment$expression[grepl("_notsig", NIF_feature_enrichment$condition)] = "All"

NIF_feature_enrichment$NIF[grepl("_utr_", NIF_feature_enrichment$condition)] = "Long 3'UTR"
NIF_feature_enrichment$NIF[grepl("_dEJ_", NIF_feature_enrichment$condition)] = "Downstream exon junction"
NIF_feature_enrichment$NIF[grepl("_uORF_", NIF_feature_enrichment$condition)] = "Upstream ORF"

NIF_feature_enrichment$sig = ifelse(NIF_feature_enrichment$pvalue < 0.05, "Significantly associated", "Not significantly associated")
```


```{r}
NIF_enrichment_F_plot = NIF_feature_enrichment %>% 
  mutate(Condition = factor(Condition, levels = c("Upf3b KD targets", 
                                                  "Upf3a KD targets",
                                                  "Upf3a OE targets", 
                                                  "Upf3 dKD targets", 
                                                  "Upf3a OE Upf3b KD targets"))) %>%
  ggplot(aes(y=Condition, x=OR, color = expression)) +  
   theme_bw()  + theme(axis.text.y = element_blank()) +
  geom_point(size = 6, aes(shape=sig)) + facet_grid(rows = vars(Condition), cols= vars(NIF), 
                                                    scales="free") +
  ylab("") + xlab("Odds ratio (OR)") + 
  scale_color_manual(values=c("All" = "grey",
                              "Up" = "#ee6c4d",
                              "Down" = "#293241")) +
  labs(shape = "",
       color = "Group") +
    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))


NIF_enrichment_F_plot

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIF_feature_OR.svg",
       plot = NIF_enrichment_F_plot ,  units = c("in"), height = 7.87, width =7.24)

```


```{r}
enrichment_nif_pval = list()
enrichment_nif_OR = list()
nifs_list = c("NIF_uORF", "NIF_dEJ", "NIF_utr")

for (c in names(res_dte)){
  test.list = list()
  test.list$BG = res_dte[[c]]$tx_id
  test.list$NIF_uORF = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_uORF== 1 ]
  test.list$NIF_dEJ = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_dEJ== 1 ]
  test.list$NIF_utr = res_dte[[c]]$tx_id[res_dte[[c]]$NIF_long_3utr== 1 & res_dte[[c]]$NIF_dEJ== 0 &  res_dte[[c]]$NIF_uORF== 0]
  test.list$sig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05]
  test.list$up = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC > 0]
  test.list$down = res_dte[[c]]$tx_id[res_dte[[c]]$FDR < 0.05 & res_dte[[c]]$logFC < 0]
  test.list$notsig = res_dte[[c]]$tx_id[res_dte[[c]]$FDR > 0.05]
  for (nif in nifs_list){
     go.obj <- newGeneOverlap(test.list$notsig,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c, "_", nif, "_notsig")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c, "_", nif, "_notsig")]] = go.obj@odds.ratio
     go.obj <- newGeneOverlap(test.list$up,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c,"_",nif,  "_up")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c,"_",nif, "_up")]] = go.obj@odds.ratio
     go.obj <- newGeneOverlap(test.list$down,
                         test.list[[nif]],
                         genome.size=length(test.list$BG))
     go.obj <- testGeneOverlap(go.obj)
     enrichment_nif_pval[[paste0(c,"_",nif, "_down")]] = go.obj@pval
     enrichment_nif_OR[[paste0(c,"_",nif, "_down")]] = go.obj@odds.ratio
    
  }
}


enrichment_nif_pval = do.call(rbind, enrichment_nif_pval) %>% as.data.frame()
colnames(enrichment_nif_pval)[1] = c("pvalue")
enrichment_nif_pval$padj = p.adjust(enrichment_nif_pval$pvalue, method = "BH")
enrichment_nif_OR  = do.call(rbind, enrichment_nif_OR) %>% as.data.frame()
colnames(enrichment_nif_OR)[1] = c("OR")
NIF_feature_enrichment = cbind(enrichment_nif_pval, enrichment_nif_OR)
```


```{r}
NIF_feature_enrichment %<>% rownames_to_column("condition")
NIF_feature_enrichment$Condition[grepl("UPF3B_KD_Control_", NIF_feature_enrichment$condition)] = "Upf3b KD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_KD_Control_", NIF_feature_enrichment$condition)] = "Upf3a KD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_OE_Control", NIF_feature_enrichment$condition)] = "Upf3a OE targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_KD_UPF3B_KD_Control", NIF_feature_enrichment$condition)] = "Upf3 dKD targets"
NIF_feature_enrichment$Condition[grepl("UPF3A_OE_UPF3B_KD_Control", NIF_feature_enrichment$condition)] = "Upf3a OE Upf3b KD targets"
NIF_feature_enrichment$expression[grepl("_up", NIF_feature_enrichment$condition)] = "Up"
NIF_feature_enrichment$expression[grepl("_down", NIF_feature_enrichment$condition)] = "Down"
NIF_feature_enrichment$expression[grepl("_notsig", NIF_feature_enrichment$condition)] = "All"

NIF_feature_enrichment$NIF[grepl("_utr_", NIF_feature_enrichment$condition)] = "Long 3'UTR"
NIF_feature_enrichment$NIF[grepl("_dEJ_", NIF_feature_enrichment$condition)] = "Downstream exon junction"
NIF_feature_enrichment$NIF[grepl("_uORF_", NIF_feature_enrichment$condition)] = "Upstream ORF"

NIF_feature_enrichment$sig = ifelse(NIF_feature_enrichment$pvalue < 0.05, "Significantly associated", "Not significantly associated")
```


```{r}
NIF_enrichment_F_plot = NIF_feature_enrichment %>% 
  mutate(Condition = factor(Condition, levels = c("Upf3b KD targets", 
                                                  "Upf3a KD targets",
                                                  "Upf3a OE targets", 
                                                  "Upf3 dKD targets", 
                                                  "Upf3a OE Upf3b KD targets"))) %>%
  ggplot(aes(y=Condition, x=OR, color = expression)) +  
   theme_bw()  + theme(axis.text.y = element_blank()) +
  geom_point(size = 6, aes(shape=sig)) + facet_grid(rows = vars(Condition), cols= vars(NIF), 
                                                    scales="free") +
  ylab("") + xlab("Odds ratio (OR)") + 
  scale_color_manual(values=c("All" = "grey",
                              "Up" = "#ee6c4d",
                              "Down" = "#293241")) +
  labs(shape = "",
       color = "Group") +
    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))


NIF_enrichment_F_plot


```



## Distribution of NIFs

```{r}
NIF_dist = function(res_matrix){
  temp = res_matrix
  dEJ_dist = temp %>% as.data.frame() %>%
    ggplot(aes(x=logFC)) +
    geom_density(data=dplyr::filter(temp,NIF_dEJ == 1, FDR < 0.05), fill="#FF2759", alpha = 0.5) +
    geom_density(data=dplyr::filter(temp, NIF_dEJ == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") + ggtitle("Density distribution of dEJ") + 
    xlim(-4,4)

  uORF_dist  = temp %>%  as.data.frame() %>%
    ggplot(aes(x=logFC)) +
    geom_density(data=dplyr::filter(temp,NIF_uORF == 1, FDR < 0.05),
                 fill="#FF2759", alpha = 0.5) +
    geom_density(data=dplyr::filter(temp, NIF_uORF == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") + ggtitle("Density distribution of uORF") + 
    xlim(-4,4)

  utr_dist = temp %>%  as.data.frame() %>%
    ggplot(aes(x=logFC)) + geom_density(data=dplyr::filter(temp, NIF_long_3utr == 1, FDR < 0.05), fill="#FF2759", alpha = 0.5) +
    geom_density(data=dplyr::filter(temp, NIF_long_3utr == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") +
    ggtitle("Density distribution Long 3' UTRs") + 
    xlim(-4,4)

  plot = ggarrange(dEJ_dist, uORF_dist, utr_dist, nrow=3)
  return(plot)
}



NIF_dist_hist = function(res_matrix){
  temp = res_matrix
  dEJ_dist = temp %>% as.data.frame() %>%
    ggplot(aes(x=logFC)) +
    geom_histogram(data=dplyr::filter(temp,NIF_dEJ == 1, FDR < 0.05), fill="#FF2759", alpha = 0.5) +
    geom_histogram(data=dplyr::filter(temp, NIF_dEJ == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") + ggtitle("Density distribution of dEJ")

  uORF_dist  = temp %>%  as.data.frame() %>%
    ggplot(aes(x=logFC)) +
    geom_histogram(data=dplyr::filter(temp,NIF_uORF == 1, FDR < 0.05),
                 fill="#FF2759", alpha = 0.5) +
    geom_histogram(data=dplyr::filter(temp, NIF_uORF == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") + ggtitle("Density distribution of uORF") 

  utr_dist = temp %>%  as.data.frame() %>%
    ggplot(aes(x=logFC)) + 
    geom_histogram(data=dplyr::filter(temp, NIF_long_3utr == 1, FDR < 0.05), fill="#FF2759", alpha = 0.5) +
    geom_histogram(data=dplyr::filter(temp, NIF_long_3utr == 0, FDR < 0.05), fill="lightgray", alpha = 0.2) +
    theme_bw() +
    ylab("Filtered Density") + xlab("Log2 Fold Change") + ggtitle("Density distribution Long 3' UTRs")

  plot = ggarrange(dEJ_dist, uORF_dist, utr_dist, nrow=3)
  return(plot)
}


```

* Distribution of any NIF + transcript
```{r}
nif_ma_plot = function(data){
  DEColours <- c("NIF-" = "#14213d","NIF+" = "#ff6700",  "Not Sig" = "#e6e8e6")
  temp = data %>% 
    mutate(nif_label = ifelse(nif_feature != "None" & FDR < 0.05, "NIF+",        
               ifelse(nif_feature == "None" & FDR < 0.05, "NIF-", "Not Sig")))
  plot = temp %>% 
    ggplot(aes(y= logFC,
               x = logCPM,
               colour = as.factor(as.character(nif_label)),
               size= nif_label)) +
    geom_point(alpha = 0.5) +
    scale_colour_manual(values = DEColours) + theme_bw()+
    theme(legend.position = "top",
          axis.title.y = element_text(size = 12)) + geom_hline(yintercept = 0, 
                                                               linetype = "dashed", color = "grey") +
    geom_ysidedensity(aes(x=after_stat(density),  
                             color = nif_label, fill = nif_label),  alpha=0.2, size = 1) +
    scale_fill_manual(values = DEColours) + scale_size_manual(values = c("NIF+" = 3, 
                                                                         "NIF-" =2,
                                                                         "Not Sig"= 2)) +
    theme(ggside.panel.scale.y = 0.2,
        axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
         strip.text.x = element_text(
        size = 12, face = "bold.italic", family = "serif"
        )) + labs(color = "Presence of NMD inducible feature", 
                  fill = "",
                  size = "") + scale_alpha(guide = 'none')
  return(plot)
    
}


ma_1 = nif_ma_plot(res_dte$UPF3B_KD_Control) + ggtitle("Upf3b KD DETs") +
  theme(legend.position = "none") +
  annotate("label", x =10, y = 4, label = "Upregulated=1860") +
  annotate("label", x =10, y = -4, label = "Downregulated =1328")

ma_2 = nif_ma_plot(res_dte$UPF3A_KD_Control) + ggtitle("Upf3a KD DETs") +
  annotate("label", x =10, y = 4, label = "Upregulated=1504") +
  annotate("label", x =10, y = -4, label = "Downregulated =816") 


ma_3 = nif_ma_plot(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("Upf3 dKD DETs") +
  annotate("label", x =10, y = 4, label = "Upregulated=4249") +
  annotate("label", x =10, y = -4, label = "Downregulated=3403") 

ma_4 = nif_ma_plot(res_dte$UPF3A_OE_Control) + ggtitle("Upf3a OE DETs") + 
    annotate("label", x =10, y = 4, label = "Upregulated=3") +
  annotate("label", x =10, y = -4, label = "Downregulated=8") 
ma_5 = nif_ma_plot(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("Upf3a OE Upf3b KD DETs") + 
   annotate("label", x =10, y = 4, label = "Upregulated=1246") +
  annotate("label", x =10, y = -4, label = "Downregulated=728") 



pdf(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/transcript_det_ma.pdf", height = 10, width = 18)
ggarrange(ma_1, ma_2, ma_3, 
            ma_4, ma_5, ncol=3, nrow=2, common.legend = TRUE)
dev.off()
```


```{r}
res_dte$UPF3A_OE_Control %>% as.data.frame %>%
    dplyr::filter(FDR < .05) %>% 
   with(., 
         table(sign.lfc=sign(logFC))) 
```

* Distribution of specific NIFs within Upf3b KD targets  

```{r fig.height=10, fig.width=8}
upf3b = NIF_dist(res_dte$UPF3B_KD_Control)

upf3b
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/UPF3B_NIF_dist.svg", 
       width = 7.08, height =6.89, plot = upf3b, units = "in")
```


* Distribution of specific NIFs within Upf3a KD targets 
```{r fig.height=10, fig.width=8}
upf3a = NIF_dist(res_dte$UPF3A_KD_Control) 
upf3a 
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/UPF3A_NIF_dist.svg", 
       width = 7.08, height =6.89, plot = upf3a, units = "in")
```

* Distribution of specific NIFs within Upf3 dKD targets 

```{r fig.height=10, fig.width=8}
dkd = NIF_dist(res_dte$UPF3A_KD_UPF3B_KD_Control) 
dkd 
```

* Distribution of specific NIFs within Upf3a OE targets 
```{r fig.height=10, fig.width=8}
upf3a_oe = NIF_dist(res_dte$UPF3A_OE_Control)
upf3a_oe 

```

* Distribution of specific NIFs within Upf3a OE in Upf3b KD targets 

```{r fig.height=10, fig.width=8}
upf3a_oe_b_kd =NIF_dist(res_dte$UPF3A_OE_UPF3B_KD_Control) 
upf3a_oe_b_kd
```


### Cumulative frequency of NIF

* 3'UTR length 

Transcripts were divided into three groups: 
 - Upregulated  (FDR < 0.05, log2FC > 0)
 - Downregulated (FDR < 0.05, log2FC < 0)
 - All (FDR > 0.05)


### UPF3B

* 3'UTR - length 

```{r}
library(ggstatsplot)

upf3b_A = res_dte$UPF3B_KD_Control %>%
  dplyr::mutate(sig = ifelse(FDR < 0.05, "DETs", "Not Sig")) %>%
  ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01)
) + ylim(0,15000) + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("DETs"= "#b23a48", "Not Sig" = "grey")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

upf3b_b = res_dte$UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>% 
    ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
) + ylim(0,15000) + theme_bw() + 
    scale_color_manual(values = c("Upregulated" = "#d7263d",
                               "Downregulated" = "#2e294e", 
                               "Not Sig" = "grey")) + 
       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + ylab("3'UTR length") + xlab("")


significance_testing = res_dte$UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) 
pairwise.t.test(significance_testing$LENGTH_3utr, significance_testing$sig, p.adjust.method = "bonferroni")

ggsave(upf3b_A , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3b_KD_utr_length.svg", height = 7, width = 8, units = "in")
ggsave(upf3b_b , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3b_KD_utr_length_DET.svg", height = 7, width = 8, units = "in")

```

```{r}
 upf3a_a = res_dte$UPF3A_KD_Control %>%
  dplyr::mutate(sig = ifelse(FDR < 0.05, "DETs", "Not Sig")) %>%
  ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
) +   ylim(0,15000) + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("DETs"= "#b23a48", "Not Sig" = "grey")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

upf3a_b  = res_dte$UPF3A_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>% 
    ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
  title = "Distribution of sepal length across Iris species"
) +  ylim(0,15000) + theme_bw() + 
    scale_color_manual(values = c("Upregulated" = "#d7263d",
                               "Downregulated" = "#2e294e", 
                               "Not Sig" = "grey")) + 
       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + ylab("3'UTR length") + xlab("")


significance_testing = res_dte$UPF3A_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))


pairwise.t.test(significance_testing$LENGTH_3utr, significance_testing$sig, p.adjust.method = "bonferroni")


ggsave(upf3a_a , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3a_KD_utr_length.svg", height = 7, width = 8, units = "in")
ggsave(upf3a_b , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3a_KD_utr_length_DET.svg", height = 7, width = 8, units = "in")
```




```{r}
 upf3dKD_a = res_dte$UPF3A_KD_UPF3B_KD_Control %>%
  dplyr::mutate(sig = ifelse(FDR < 0.05, "DETs", "Not Sig")) %>%
  ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
) +   ylim(0,15000) + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("DETs"= "#b23a48", "Not Sig" = "grey")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

 upf3dKD_b =res_dte$UPF3A_KD_UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>% 
    ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
  title = "Distribution of sepal length across Iris species"
) +  ylim(0,15000) + theme_bw() + 
    scale_color_manual(values = c("Upregulated" = "#d7263d",
                               "Downregulated" = "#2e294e", 
                               "Not Sig" = "grey")) + 
       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + ylab("3'UTR length") + xlab("")

 
 significance_testing = res_dte$UPF3A_KD_UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))


pairwise.t.test(significance_testing$LENGTH_3utr, significance_testing$sig, p.adjust.method = "bonferroni")



ggsave(upf3dKD_a , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3_dKD_utr_length.svg", height = 7, width = 8, units = "in")
ggsave(upf3dKD_b , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3_dKD_utr_length_DET.svg", height = 7, width = 8, units = "in")

```



```{r}
 upf3aOE_a = res_dte$UPF3A_OE_Control %>%
  dplyr::mutate(sig = ifelse(FDR < 0.05, "DETs", "Not Sig")) %>%
  ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
) +   ylim(0,15000) + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("DETs"= "#b23a48", "Not Sig" = "grey")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

 upf3aOE_b =res_dte$UPF3A_OE_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>% 
    ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
  title = "Distribution of sepal length across Iris species"
) +  ylim(0,15000) + theme_bw() + 
    scale_color_manual(values = c("Upregulated" = "#d7263d",
                               "Downregulated" = "#2e294e", 
                               "Not Sig" = "grey")) + 
       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + ylab("3'UTR length") + xlab("")


```

```{r}
upf3aOE_upf3bkD_a = res_dte$UPF3A_OE_UPF3B_KD_Control %>%
  dplyr::mutate(sig = ifelse(FDR < 0.05, "DETs", "Not Sig")) %>%
  ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
) +   ylim(0,15000) + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("DETs"= "#b23a48", "Not Sig" = "grey")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

upf3aOE_upf3bkD_b = res_dte$UPF3A_OE_UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>% 
    ggbetweenstats(
  x     = sig,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01),
  title = "Distribution of sepal length across Iris species"
) +  ylim(0,15000) + theme_bw() + 
    scale_color_manual(values = c("Upregulated" = "#d7263d",
                               "Downregulated" = "#2e294e", 
                               "Not Sig" = "grey")) + 
       theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + ylab("3'UTR length") + xlab("")


 significance_testing = res_dte$UPF3A_OE_UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))

pairwise.t.test(significance_testing$LENGTH_3utr, significance_testing$sig, p.adjust.method = "bonferroni")


ggsave(upf3aOE_upf3bkD_a , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3a_OE_upf3b_KD_utr_length.svg", height = 7, width = 8, units = "in")
ggsave(upf3aOE_upf3bkD_b  , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/NIFs/upf3a_OE_upf3b_KD_utr_length_DET.svg", height = 7, width = 8, units = "in")

```



```{r}
upf3b_utr = res_dte$UPF3B_KD_Control %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>%
  mutate(normalized_3utr = LENGTH_3utr/Length ) %>%
  ggplot(aes(x=normalized_3utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Normalized 3'UTR length")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/upf3b_3utr.svg", plot = upf3b_utr,
       units = c("in"), height = 5.83, width = 7.52)
```



```{r}


```



```{r}
plot.ecdf.3utr.normalized = function(df) {
  plot = df %>% 
    mutate(sig = ifelse(logFC > 0 & FDR < 0.05, 
                        "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05,
                             "Downregulaated", "Not Sig"))) %>%
  mutate(normalized_3utr = LENGTH_3utr/Length ) %>%
  ggplot(aes(x=normalized_3utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Normalized 3'UTR length")
  return(plot)
}

plot.ecdf.3utr = function(df, nif) {
  plot = df %>% 
    mutate(sig = ifelse(logFC > 0 & FDR < 0.05, 
                        "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05,
                             "Downregulaated", "Not Sig"))) %>%
  mutate(normalized_3utr = LENGTH_3utr/Length ) %>%
  ggplot(aes(x=LENGTH_3utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("3'UTR length")
  return(plot)
}
```




```{r}
upf3b_utr = plot.ecdf.3utr(res_dte$UPF3B_KD_Control)
upf3a_utr = plot.ecdf.3utr(res_dte$UPF3A_KD_Control)
upf3dKD_utr = plot.ecdf.3utr(res_dte$UPF3A_KD_UPF3B_KD_Control)
upf3aOE_utr = plot.ecdf.3utr(res_dte$UPF3A_OE_Control)
upf3aOEbKD_utr = plot.ecdf.3utr(res_dte$UPF3A_OE_UPF3B_KD_Control)

grid.arrange(upf3b_utr, upf3a_utr, upf3dKD_utr, 
            upf3aOE_utr, upf3aOEbKD_utr, ncol=3, nrow=2)
```


```{r}
upf3b_utr = plot.ecdf.3utr.normalized(res_dte$UPF3B_KD_Control)
upf3a_utr = plot.ecdf.3utr.normalized(res_dte$UPF3A_KD_Control)
upf3dKD_utr = plot.ecdf.3utr.normalized(res_dte$UPF3A_KD_UPF3B_KD_Control)
upf3aOE_utr = plot.ecdf.3utr.normalized(res_dte$UPF3A_OE_Control)
upf3aOEbKD_utr = plot.ecdf.3utr.normalized(res_dte$UPF3A_OE_UPF3B_KD_Control)

grid.arrange(upf3b_utr, upf3a_utr, upf3dKD_utr, 
            upf3aOE_utr, upf3aOEbKD_utr, ncol=3, nrow=2)
```

* GC content in 3'UTR

```{r}
plot.ecdf.3GC = function(df) {
  plot = df %>% 
    mutate(sig = ifelse(logFC > 0 & FDR < 0.05, 
                        "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05,
                             "Downregulaated", "Not Sig"))) %>%
  ggplot(aes(x=GC_3utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("GC content of 3'UTR")
  return(plot)
}
```


```{r}
upf3b_utr = plot.ecdf.3GC(res_dte$UPF3B_KD_Control)
upf3a_utr = plot.ecdf.3GC(res_dte$UPF3A_KD_Control)
upf3dKD_utr = plot.ecdf.3GC(res_dte$UPF3A_KD_UPF3B_KD_Control)
upf3aOE_utr = plot.ecdf.3GC(res_dte$UPF3A_OE_Control)
upf3aOEbKD_utr = plot.ecdf.3GC(res_dte$UPF3A_OE_UPF3B_KD_Control)

grid.arrange(upf3b_utr, upf3a_utr, upf3dKD_utr, 
            upf3aOE_utr, upf3aOEbKD_utr, ncol=3, nrow=2)
```


* 5' UTR length
```{r}

plot.ecdf.5utr.normalized = function(df) {
  plot = df %>% 
    mutate(sig = ifelse(logFC > 0 & FDR < 0.05, 
                        "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05,
                             "Downregulaated", "Not Sig"))) %>%
  mutate(normalized_5utr = LENGTH_5utr/Length ) %>%
  ggplot(aes(x=normalized_5utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Normalized 3'UTR length")
  return(plot)
}

plot.ecdf.5utr = function(df) {
  plot = df %>% 
    mutate(sig = ifelse(logFC > 0 & FDR < 0.05, 
                        "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05,
                             "Downregulaated", "Not Sig"))) %>%
  ggplot(aes(x=LENGTH_5utr, color = sig)) + stat_ecdf(geom = "point", size=0.8) +
  scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("3'UTR length")
  return(plot)
}
```

```{r}
upf3b_utr = plot.ecdf.5utr(res_dte$UPF3B_KD_Control)
upf3a_utr = plot.ecdf.5utr(res_dte$UPF3A_KD_Control)
upf3dKD_utr = plot.ecdf.5utr(res_dte$UPF3A_KD_UPF3B_KD_Control)
upf3aOE_utr = plot.ecdf.5utr(res_dte$UPF3A_OE_Control)
upf3aOEbKD_utr = plot.ecdf.5utr(res_dte$UPF3A_OE_UPF3B_KD_Control)

grid.arrange(upf3b_utr, upf3a_utr, upf3dKD_utr, 
            upf3aOE_utr, upf3aOEbKD_utr, ncol=3, nrow=2)
```


```{r}
upf3b_utr = plot.ecdf.5utr.normalized(res_dte$UPF3B_KD_Control)
upf3a_utr = plot.ecdf.5utr.normalized(res_dte$UPF3A_KD_Control)
upf3dKD_utr = plot.ecdf.5utr.normalized(res_dte$UPF3A_KD_UPF3B_KD_Control)
upf3aOE_utr = plot.ecdf.5utr.normalized(res_dte$UPF3A_OE_Control)
upf3aOEbKD_utr = plot.ecdf.5utr.normalized(res_dte$UPF3A_OE_UPF3B_KD_Control)

grid.arrange(upf3b_utr, upf3a_utr, upf3dKD_utr, 
            upf3aOE_utr, upf3aOEbKD_utr, ncol=3, nrow=2)
```


* uORFs 

```{r}
uorfs_upf3a = res_dte$UPF3A_KD_Control %>%
  mutate(count_uORF = `uORFs|__|count`) %>%
  mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",
                                     between(count_uORF, 1, 2) ~"1-2",
                                     between(count_uORF, 3,4 ) ~ "3-4",
                                     count_uORF > 5 ~ ">5")) %>%
  drop_na(uORF_count_bins) %>%
  mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",
                                                  "1-2",
                                                  "3-4",
                                                  ">5"))) %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%
  ggplot(aes(x=uORF_count_bins, fill = sig)) + geom_bar(stat="count", position = "fill", color = "white") +
  theme_bw() + scale_fill_manual(values = c("#2F124B", "#FD9675", "#EA2A5F"))# + 
  theme(legend.position = "none")

uorfs_upf3a 
```




<!-- # UPF3B transcripts  -->

<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>% -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--       DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- ## UPF3B transcripts with NMD features  -->

<!-- ```{r fig.cap= "*NIF in upregulated transcripts in UPF3B*"} -->
<!-- sig_res= res_dte$UPF3B_KD_Control  %>%  -->
<!--   dplyr::filter(FDR < 0.05, abs(logFC) > log2(1)) -->


<!-- dEJ = sig_res %>%  -->
<!--   dplyr::filter(NIF_dEJ == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- uORF = sig_res %>%  -->
<!--   dplyr::filter(NIF_uORF == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- UTR = sig_res %>%  -->
<!--   dplyr::filter(NIF_long_3utr == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- x = list( "dEJ" = dEJ$tx_id,  -->
<!--           "uORF" = uORF$tx_id,  -->
<!--           "Long 3'UTR" = UTR$tx_id,  -->
<!--           "All transcript" = sig_res$tx_id) -->

<!-- upset(fromList(x)) -->

<!-- ``` -->


<!-- ### Characteristics of 3'UTR -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05, NIF_long_3utr == 1) %>%  -->
<!--   ggplot(aes(x=GC_3utr, y =logFC)) + geom_point() -->

<!-- ``` -->

<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   mutate(Expression = ifelse(logFC > 1, "Up", "Down")) %>%  -->
<!--   ggplot(aes(x=Expression, y =LENGTH_3utr)) + geom_boxplot() -->


<!-- ``` -->
<!-- ## Comparison of DEGs with DETs for UPF3B KD  -->

<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(dplyr::rename(DEGenes$UPF3B_vs_Control, "gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   ggscatter(y="logFC", x="Coef.UPF3B_vs_Control",  cor.coef = TRUE, add = "reg.line", color ="NIF_long_3utr", -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray"))  +  -->
<!--   xlab("UPF3B KD (gene level logFC)") + ylab("UPF3B KD (transcript level logFC)") -->


<!-- ``` -->

<!-- ### Genes with NIFs -->
<!-- ```{r} -->
<!-- genes_nif = res_dte$UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(dplyr::rename(DEGenes$UPF3B_vs_Control, "gene_id" = "ensembl_gene_id"), by = "gene_id") -->


<!-- dim(genes_nif) -->

<!-- length(unique(genes_nif$gene.y)) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- NIF_dist_hist(genes_nif) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--    inner_join(tx2gene, by = "tx_id") %>% -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3B_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id") %>%  -->
<!--               dplyr::select(contains("UPF3B_vs_Control"), gene_id), by = "gene_id") %>%  -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->

<!-- # UPF3A transcripts  -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_Control %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>% -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--       DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- ## UPF3A transcripts with NMD features  -->

<!-- ```{r fig.cap= "*NIF in upregulated transcripts in UPF3B*"} -->
<!-- sig_res= res_dte$UPF3A_KD_Control  %>%  -->
<!--   dplyr::filter(FDR < 0.05, abs(logFC) > log2(1)) -->


<!-- dEJ = sig_res %>%  -->
<!--   dplyr::filter(NIF_dEJ == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- uORF = sig_res %>%  -->
<!--   dplyr::filter(NIF_uORF == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- UTR = sig_res %>%  -->
<!--   dplyr::filter(NIF_long_3utr == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- x = list( "dEJ" = dEJ$tx_id,  -->
<!--           "uORF" = uORF$tx_id,  -->
<!--           "Long 3'UTR" = UTR$tx_id,  -->
<!--           "All transcript" = sig_res$tx_id) -->

<!-- upset(fromList(x)) -->

<!-- ``` -->


<!-- ### Comparison of DEGs with DETs -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(dplyr::rename(DEGenes$UPF3A_vs_Control, "gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   ggscatter(y="logFC", x="Coef.UPF3A_vs_Control",  cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray"))  +  -->
<!--   xlab("UPF3A KD (gene level logFC)") + ylab("UPF3A KD (transcript level logFC)") -->


<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--    inner_join(tx2gene, by = "tx_id") %>% -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3A_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id") %>%  -->
<!--               dplyr::select(contains("UPF3A_vs_Control"), gene_id), by = "gene_id") %>%  -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->



<!-- # UPF3 dKD -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_UPF3B_KD_Control %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>% -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--       DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->


<!-- ### Comparison of DEGs with DETs -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(dplyr::rename(DEGenes$DoubleKD_vs_Control, "gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   ggscatter(y="logFC", x="Coef.DoubleKD_vs_Control",  cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--   xlab("UPF3 dKD (gene level logFC)") + ylab("UPF3 dKD (transcript level logFC)") -->

<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--    inner_join(tx2gene, by = "tx_id") %>% -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$DoubleKD_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id") %>%  -->
<!--               dplyr::select(contains("DoubleKD_vs_Control"), gene_id), by = "gene_id") %>%  -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->




<!-- ## Transcripts with NIFs -->

<!-- ```{r fig.cap= "*NIF in upregulated transcripts in UPF3B*"} -->
<!-- sig_res= res_dte$UPF3A_KD_UPF3B_KD_Control  %>%  -->
<!--   dplyr::filter(FDR < 0.05, abs(logFC) > log2(1)) -->


<!-- dEJ = sig_res %>%  -->
<!--   dplyr::filter(NIF_dEJ == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- uORF = sig_res %>%  -->
<!--   dplyr::filter(NIF_uORF == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- UTR = sig_res %>%  -->
<!--   dplyr::filter(NIF_long_3utr == 1) %>%  -->
<!--   dplyr::select(tx_id) -->

<!-- x = list( "dEJ" = dEJ$tx_id,  -->
<!--           "uORF" = uORF$tx_id,  -->
<!--           "Long 3'UTR" = UTR$tx_id,  -->
<!--           "All transcript" = sig_res$tx_id) -->

<!-- upset(fromList(x)) -->

<!-- ``` -->

<!-- # UPF3A OE  -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_OE_Control %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>% -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--       DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_OE_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--    inner_join(tx2gene, by = "tx_id") %>% -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3A_OE_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id") %>%  -->
<!--               dplyr::select(contains("UPF3A_OE_vs_Control"), gene_id), by = "gene_id") %>%  -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->


<!-- ``` -->

<!-- # UPF3A OE in UPF3B KD  -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_OE_UPF3B_KD_Control %>%  -->
<!--   as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>% -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--       DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->

<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_OE_UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(dplyr::rename(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control, "gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   ggscatter(y="logFC", x="Coef.UPF3A_OE_UPF3B_KD_vs_Control",  cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--   xlab("UPF3A OE in UPF3B KD (gene level logFC)") + ylab("UPF3A OE in UPF3B KD (transcript level logFC)") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- res_dte$UPF3A_OE_UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::select(tx_id, Overdispersion, logFC,  -->
<!--                 logCPM, PValue, FDR, NIF_dEJ, NIF_uORF, LENGTH_3utr,  -->
<!--                 `NIF|__|GC(cds[-10:termination_codon])`) %>%  -->
<!--    inner_join(tx2gene, by = "tx_id") %>% -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id") %>%  -->
<!--               dplyr::select(contains("UPF3A_OE_UPF3B_KD_vs_Control"), gene_id), by = "gene_id") %>%  -->
<!--      DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--             options = list(scrollX = TRUE,  -->
<!--                            scrollCollapse = TRUE,  -->
<!--                            pageLength = 5, autoWidth = TRUE,  -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->


<!-- ``` -->


<!-- # Are DET shared between UPF3B KD and UPF3A KD?  -->

<!-- ```{r} -->
<!-- degs_b= res_dte$UPF3B_KD_Control %>% as.data.frame %>% -->
<!--   dplyr::filter(FDR < .05) %>%  -->
<!--   with(.,  -->
<!--        table(sign.lfc=sign(logFC))) %>% as.data.frame() -->
<!-- ``` -->

<!-- In UPF3B KD, there are 1860 upregulated genes aand 1328 downregulated genes  -->

<!-- ```{r} -->

<!-- res_dte$UPF3A_KD_Control %>% as.data.frame %>% -->
<!--   dplyr::filter(FDR < .05) %>%  -->
<!--   with(.,  -->
<!--        table(sign.lfc=sign(logFC))) -->

<!-- ``` -->

<!-- In UPF3A KD, there are 1504 upregulated genes and 816 downregulated genes  -->


<!-- ```{r} -->
<!-- x = list( "UPF3B KD" = res_dte$UPF3B_KD_Control$tx_id[res_dte$UPF3B_KD_Control$FDR < 0.05],   -->
<!--     "UPF3A KD" = res_dte$UPF3A_KD_Control$tx_id[res_dte$UPF3A_KD_Control$FDR < 0.05],  -->
<!--     "Double KD" = res_dte$UPF3A_KD_UPF3B_KD_Control$tx_id[res_dte$UPF3A_KD_UPF3B_KD_Contro$FDR < 0.05], -->
<!--     "UPF3A OE" = res_dte$UPF3A_OE_Control$tx_id[res_dte$UPF3A_OE_Control$FDR < 0.05],  -->
<!--     "UPF3A OE UPF3B KD" = res_dte$UPF3A_OE_UPF3B_KD_Control$tx_id[res_dte$UPF3A_OE_UPF3B_KD_Control$FDR < 0.05]) -->

<!-- ggvenn(x[c(1,2)]) -->
<!-- ``` -->

<!-- ```{r fig.height=8, fig.width=15} -->
<!-- grid.arrange((res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B) %>%  -->
<!--   inner_join(dplyr::filter(as.data.frame(res_dte$UPF3A_KD_Control), FDR < 0.05), by = "tx_id")  %>%  -->
<!--    ggscatter(., x = "logFC.UPF3B", y = "logFC", color = "NIF_uORF", -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  +  -->
<!--   geom_xsidedensity(aes(у = after_stat(density), fill = NIF_uORF), alpha = 0.5,  -->
<!--                     size = 1,position = "stack") + geom_ysidedensity(aes(x=after_stat(density),  -->
<!--                                                                          fill = NIF_uORF),  -->
<!--                                                                      alpha=0.5, size = 1) + -->
<!--     xlab("UPF3B KD (log2FC)") + ylab("UPF3A KD (log2FC)")), -->

<!-- (res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B) %>%  -->
<!--   inner_join(dplyr::filter(as.data.frame(res_dte$UPF3A_KD_Control), FDR < 0.05), by = "tx_id")  %>%  -->
<!--    ggscatter(., x = "logFC.UPF3B", y = "logFC", color = "NIF_dEJ", -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  +  -->
<!--   geom_xsidedensity(aes(у = after_stat(density), fill = NIF_dEJ), alpha = 0.5,  -->
<!--                     size = 1,position = "stack") + geom_ysidedensity(aes(x=after_stat(density),  -->
<!--                                                                          fill = NIF_dEJ),  -->
<!--                                                                      alpha=0.5, size = 1)  -->
<!--  + xlab("UPF3B KD (log2FC)")+ ylab("UPF3A KD (log2FC)")),  -->


<!-- (res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B) %>%  -->
<!--   inner_join(dplyr::filter(as.data.frame(res_dte$UPF3A_KD_Control), FDR < 0.05), by = "tx_id")  %>%  -->
<!--    ggscatter(., x = "logFC.UPF3B", y = "logFC", color = "NIF_long_3utr", -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  +  -->
<!--   geom_xsidedensity(aes(у = after_stat(density), fill = NIF_long_3utr), alpha = 0.5,  -->
<!--                     size = 1,position = "stack") + geom_ysidedensity(aes(x=after_stat(density),  -->
<!--                                                                          fill = NIF_long_3utr),  -->
<!--                                                                      alpha=0.5, size = 1) + -->
<!--    xlab("UPF3B KD (log2FC)") + ylab("UPF3A KD (log2FC)") ),  -->
<!-- ncol=3 -->

<!-- ) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>%  -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B) %>%  -->
<!--   left_join(dplyr::filter(as.data.frame(res_dte$UPF3A_KD_Control), Results != 0), by = "tx_id")  %>%  -->
<!--    ggscatter(., x = "logFC.UPF3B", y = "logFC", color= "nif_feature",  -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()    -->
<!-- ``` -->


<!-- ```{r} -->
<!-- res_dte$UPF3A_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>%  -->
<!--   dplyr::rename("logFC.UPF3A" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3A) %>%  -->
<!--   left_join(dplyr::filter(as.data.frame(res_dte$UPF3B_KD_Control), Results == 0), by = "tx_id")  %>%  -->
<!--    ggscatter(., x = "logFC.UPF3A", y = "logFC", -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()   -->

<!-- ``` -->

<!-- ### What is the distribution between transcripts in UPF3B KD and UPF3A OE?  -->

<!-- ```{r} -->
<!-- ggvenn(x[c(1, 4)]) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(Results != 0) %>%  -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B) %>%  -->
<!--   left_join(as.data.frame(res_dte$UPF3A_OE_Control, by = "tx_id")) %>%  -->
<!--    ggscatter(., x = "logFC.UPF3B", y = "logFC", color = "nif_feature", -->
<!--               cor.coef = TRUE, add = "reg.line",  -->
<!--               conf.int = TRUE, add.params = list(color = "blue", -->
<!--                                                  fill = "lightgray")) +  -->
<!--     theme_bw()  + xlab("UPF3B KD") + ylab("UPF3A OE") +  -->
<!--   geom_xsidedensity(aes(у = after_stat(density), fill = nif_feature), alpha = 0.5,  -->
<!--                     size = 1,position = "stack") + geom_ysidedensity(aes(x=after_stat(density),  -->
<!--                                                                          fill = nif_feature),  -->
<!--                                                                      alpha=0.5, size = 1) + -->
<!--   scale_color_tq()  + scale_fill_tq() + geom_hline(yintercept = 0 ) -->
<!-- ``` -->



<!-- ### Transcripts arising from a single gene  -->
<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>% as.data.frame() %>%  -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>% -->
<!--   dplyr::rename("logFC.UPF3B" = "logFC") %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B, gene_id) %>%  -->
<!--   inner_join(res_dte$UPF3A_KD_Control %>%  -->
<!--                dplyr::filter(FDR < 0.05) %>%  -->
<!--                inner_join(tx2gene, by = "tx_id"),  -->
<!--              by = "gene_id", relationship = "many-to-many")  -->



<!-- overlap_single_genes = overlap[duplicated(overlap$gene_id) |  -->
<!--                                  duplicated(overlap$gene_id, fromLast=TRUE),] %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>%  -->
<!--   dplyr::select(tx_id, logFC.UPF3B, logFC, gene_id, gene) %>%  -->
<!--   ggscatter(., x = "logFC.UPF3B", y = "logFC", label = "gene", -->
<!--               cor.coef = TRUE,  -->
<!--               conf.int = TRUE) + -->
<!--   geom_vline(xintercept = 0) + geom_hline(yintercept = 0) -->


<!-- my_gg = overlap_single_genes + geom_point_interactive(aes(tooltip = gene, data_id = gene), -->
<!--                                             size = 1, hover_nearest = TRUE) -->

<!-- #overlap[duplicated(overlap$gene_id),] -->


<!-- ``` -->


<!-- ### Overlapping transcripts that were dysregulated at gene level -->

<!-- ```{r} -->
<!-- UPF3B_trans_overlap = res_dte$UPF3B_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3B_vs_Control %>%  -->
<!--                dplyr::filter(Res.UPF3B_vs_Control !=0, Res.UPF3A_vs_Control !=0) %>% -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by= "gene_id") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- UPF3A_trans_overlap = res_dte$UPF3A_KD_Control %>% -->
<!--   dplyr::filter(FDR < 0.05) %>%  -->
<!--   inner_join(tx2gene, by = "tx_id") %>%  -->
<!--   mutate(gene = mapIds(org.Mm.eg.db, keys=gene_id,  -->
<!--                        column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% -->
<!--   inner_join(DEGenes$UPF3A_vs_Control %>%  -->
<!--                dplyr::filter(Res.UPF3B_vs_Control !=0, Res.UPF3A_vs_Control !=0) %>% -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by= "gene_id") -->


<!-- write.csv(UPF3B_trans_overlap,  -->
<!--           file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/UPF3B_trans_overlap.csv") -->

<!-- write.csv(UPF3A_trans_overlap,  -->
<!--           file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/UPF3A_trans_overlap.csv") -->

<!-- ``` -->

<!-- ```{r} -->
<!-- ggvenn(data = list("UPF3A" = UPF3A_trans_overlap$gene_id,  -->
<!--                    "UPF3B" = UPF3B_trans_overlap$gene_id)) -->

<!-- overlap_list =calculate.overlap(x=list("UPF3A" = UPF3A_trans_overlap$gene_id,  -->
<!--                    "UPF3B" = UPF3B_trans_overlap$gene_id)) -->

<!-- subset(b, !(y %in% a$x)) -->
<!-- overlap_list$a1[!overlap_list$a1 %in% overlap_list$a3]  -->

<!-- overlap_list_transcript = calculate.overlap(x=list("UPF3A" = UPF3A_trans_overlap$tx_id,  -->
<!--                    "UPF3B" = UPF3B_trans_overlap$tx_id)) -->

<!-- overlap_list_transcript$a2[!overlap_list_transcript$a2 %in% overlap_list_transcript$a3]  -->

<!-- ggvenn(data = list("UPF3A" = UPF3A_trans_overlap$tx_id,  -->
<!--                    "UPF3B" = UPF3B_trans_overlap$tx_id)) -->

<!-- ``` -->


<!-- ## Genes and transcripts correlation  -->

<!-- ```{r fig.cap = " All transcripts with upregulated genes and their biotypes"} -->
<!-- ## Upregulated genes with DETs that may be protein coding  -->

<!-- ## UPF3B   -->
<!-- res_dte$UPF3B_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   dplyr::filter(nif_feature != "None") %>%  -->
<!--   left_join(DEGenes$UPF3B_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n)) %>%  -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n)) + -->
<!--     geom_bar( stat = "identity") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v75)", y = "Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black"))+ -->
<!-- theme(strip.text = element_text(color = "white", face = "bold")) -->
<!-- ``` -->


<!-- ```{r fig.cap = " All transcripts with upregulated genes and their biotypes - genes only"} -->
<!-- ## Upregulated genes with DETs that may be protein coding  -->

<!-- ## UPF3B   -->
<!-- res_dte$UPF3B_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   dplyr::filter(nif_feature != "None") %>%  -->
<!--   left_join(DEGenes$UPF3B_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--   distinct(gene_id, .keep_all=TRUE) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n)) %>%  -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n)) + -->
<!--     geom_bar( stat = "identity") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v75)", y = "Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black"))+ -->
<!-- theme(strip.text = element_text(color = "white", face = "bold")) -->
<!-- ``` -->

<!-- ### UPF3A  -->

<!-- ```{r fig.cap = " All transcripts with upregulated genes and their biotypes"} -->
<!-- ## Upregulated genes with DETs that may be protein coding  -->

<!-- ## UPF3B   -->
<!-- res_dte$UPF3A_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   dplyr::filter(nif_feature != "None") %>%  -->
<!--   left_join(DEGenes$UPF3A_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n)) %>%  -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n)) + -->
<!--     geom_bar( stat = "identity") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v75)", y = "Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black"))+ -->
<!-- theme(strip.text = element_text(color = "white", face = "bold")) -->
<!-- ``` -->


<!-- ```{r fig.cap = " All transcripts with upregulated genes and their biotypes - genes only"} -->
<!-- ## Upregulated genes with DETs that may be protein coding  -->

<!-- ## UPF3B   -->
<!-- res_dte$UPF3A_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   dplyr::filter(nif_feature != "None") %>%  -->
<!--   left_join(DEGenes$UPF3A_vs_Control %>%  -->
<!--               dplyr::rename("gene_id" = "ensembl_gene_id"), by = "gene_id") %>%  -->
<!--   dplyr::filter(Res.UPF3B_vs_Control == 1) %>% -->
<!--   distinct(gene_id, .keep_all=TRUE) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n)) %>%  -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n)) + -->
<!--     geom_bar( stat = "identity") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v75)", y = "Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black"))+ -->
<!-- theme(strip.text = element_text(color = "white", face = "bold")) -->
<!-- ``` -->


<!-- * Majority of transcripts are protein coding  -->
<!-- * Not looking at transcripts that are pseudogenes  -->
<!-- * Gene ontology on the protein coding genes - with transcripts  -->
<!-- ** All conditions - transcript and gene level comparison  -->

<!-- ** Capacity to do genes with isoforms with different NIFs - try in UPF3B  -->

<!-- ** Is NMD a cell-type specific thing? --? Maybe UPF1 work with brain coexp networks  -->


<!-- Any transcript that is expression - 40k  -->
<!-- transcripts that have multiple transcripts and have a look at the CPM - Just do it in UPF3B - average of the three samples -->

<!-- ## Data export  -->


<!-- ```{r} -->
<!-- save(res_dte, file=here::here("output/DET-list.Rda")) -->
<!-- ``` -->



<!-- ```{r} -->
<!--   temp = res_dte$UPF3B_KD_Control -->

<!--   message("Counting NIFs in upregulated transcripts for ", name) -->
<!--   sig_up = temp %>%  as.data.frame() %>%   -->
<!--         dplyr::filter(FDR < 0.05, logFC > log2(1))  -->
<!--   utr = table(sig_up$NIF_long_3utr)[2] -->
<!--   dej = table(sig_up$NIF_dEJ)[2] -->
<!--   uorf = table(sig_up$NIF_uORF)[2] -->
<!--   none = table(sig_up$nif_feature)[1] -->
<!--   total_nif_up  = rbind(utr, dej, uorf, none) %>%  -->
<!--     as.data.frame() %>%  -->
<!--     rownames_to_column("NIF_feature") %>%  -->
<!--     mutate(Condition = name,  -->
<!--            Expression = "Up") %>%  -->
<!--     dplyr::rename("Count" = "1") -->

<!--   message("Counting NIFs in downregulated transcripts for ", name) -->

<!--   sig_down = temp %>%  as.data.frame() %>%   -->
<!--         dplyr::filter(FDR < 0.05, logFC < log2(1))  -->
<!--   utr = table(sig_down$NIF_long_3utr)[2] -->
<!--   dej = table(sig_down$NIF_dEJ)[2] -->
<!--   uorf = table(sig_down$NIF_uORF)[2] -->

<!--   total_nif_down  = rbind(utr, dej, uorf) %>%  -->
<!--     as.data.frame() %>%  -->
<!--     rownames_to_column("NIF_feature") %>%  -->
<!--     mutate(Condition = name,  -->
<!--            Expression = "Down") %>%  -->
<!--     dplyr::rename("Count" = "1") -->

<!--   nif_features= rbind(nif_features, total_nif_up, total_nif_down, stringsAsFactors = FALSE) -->

<!-- ``` -->



<!-- # Plots to finalise  -->




<!-- # Transcript types - count -->

<!-- ```{r fig.cap = " All transcripts with upregulated genes and their biotypes - genes only"} -->
<!-- transcript_types = res_dte$UPF3A_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--   dplyr::filter(FDR < 0.05) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n),  -->
<!--            label = "Upf3a targets") %>% -->
<!--   rbind(res_dte$UPF3B_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--   dplyr::filter(FDR < 0.05) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n), -->
<!--            label = "Upf3b targets")) %>%  -->
<!--   dplyr::filter(n > 40) %>% -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n, fill = label)) + -->
<!--     geom_bar( stat = "identity", alpha = 0.6, color = "black") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v79)", y = "Transcript Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black")) + -->
<!-- theme(strip.text = element_text(color = "white", face = "bold"),  -->
<!--       legend.position = "none") + -->
<!--   scale_fill_manual(values= c("#EA2A5F", "#2F124B")) + -->
<!-- geom_text(aes(label=n), hjust=-.5, size=3.5) + ylim(0, 2800) +  -->
<!--   facet_grid(~label) -->

<!-- transcript_types_up = res_dte$UPF3A_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--   dplyr::filter(FDR < 0.05, logFC > 0) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n),  -->
<!--            label = "Upf3a targets") %>% -->
<!--   rbind(res_dte$UPF3B_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--   dplyr::filter(FDR < 0.05, logFC > 0) %>% -->
<!--   dplyr::count(tx_biotype) %>% -->
<!--     mutate(sum=sum(n),frac = n / sum(n), -->
<!--            label = "Upf3b targets")) %>%  -->
<!--   dplyr::filter(n > 40) %>% -->
<!--   ggplot(aes(x = as.factor(tx_biotype), y = n, fill = label)) + -->
<!--     geom_bar( stat = "identity", alpha = 0.6, color = "black") + -->
<!--     coord_flip() + -->
<!--     theme_bw() + -->
<!--     theme(axis.text.x = element_text(angle = -90)) + -->
<!--     labs(x = "Transcript biotype (Ensembl v79)", y = "Transcript Count") + -->
<!-- theme(panel.spacing = unit(1, "lines")) + -->
<!-- theme(strip.background = element_rect(fill = "black")) + -->
<!-- theme(strip.text = element_text(color = "white", face = "bold"),  -->
<!--       legend.position = "none") + -->
<!--   scale_fill_manual(values= c("#EA2A5F", "#2F124B")) + -->
<!-- geom_text(aes(label=n), hjust=-.5, size=3.5) + ylim(0, 2800) +  -->
<!--   facet_grid(~label) -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/transcript_types_all.svg", plot = transcript_types,  -->
<!--        units = c("in"), height = 5.85, width = 10.29) -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/transcript_types_up.svg", plot = transcript_types_up,  -->
<!--        units = c("in"), height = 5.85, width = 10.29) -->
<!-- ``` -->


<!-- # Transcript types cumulative frequency  -->

<!-- ## UPF3A -->
<!-- ```{r} -->
<!-- cf_upf3a = res_dte$UPF3A_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(CF_types = ifelse(tx_biotype == "retained_intron", "Retained Intron",  -->
<!--                            ifelse(tx_biotype == "protein_coding", "Protein Coding",  -->
<!--                                   ifelse(tx_biotype == "processed_transcript", "Processed Transcript",  -->
<!--                                          ifelse(tx_biotype == "nonsense_mediated_decay", "NMD", "All other types"))))) %>%  -->
<!--   drop_na(CF_types) %>% -->
<!--   dplyr::filter(FDR < 0.05) %>% -->
<!--   ggplot(aes(x=logFC, color = CF_types)) + stat_ecdf(geom = "step", size=1) + -->

<!--   scale_color_manual(values = c("#585A58","#5800FF" ,"#F5EB13", "#EE3224", "#22B58C")) +  -->
<!--   theme_bw() +  -->
<!--   ylab("Cumulative Frequency") + xlab("Log FC") +  -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +  -->
<!--   ggtitle("Transcript type")  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/cf_upf3a.svg", plot = cf_upf3a,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->

<!-- ``` -->

<!-- ## UPF3B -->
<!-- ```{r} -->
<!-- cf_upf3b = res_dte$UPF3B_KD_Control %>%  -->
<!--   left_join(tx2gene) %>%  -->
<!--   mutate(CF_types = ifelse(tx_biotype == "retained_intron", "Retained Intron",  -->
<!--                            ifelse(tx_biotype == "protein_coding", "Protein Coding",  -->
<!--                                   ifelse(tx_biotype == "processed_transcript", "Processed Transcript",  -->
<!--                                          ifelse(tx_biotype == "nonsense_mediated_decay", "NMD", "All other types"))))) %>%  -->
<!--   drop_na(CF_types) %>% -->
<!--   dplyr::filter(FDR < 0.05) %>% -->
<!--   ggplot(aes(x=logFC, color = CF_types)) + stat_ecdf(geom = "step", size=1) + -->
<!--   scale_color_manual(values = c("#585A58","#5800FF" ,"#F5EB13", "#EE3224", "#22B58C"))+  -->
<!--   theme_bw() +  -->
<!--   ylab("Cumulative Frequency") + xlab("Log FC") +  -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +  -->
<!--   ggtitle("Transcript type")  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/cf_upf3b.svg", plot = cf_upf3b,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->

<!-- ``` -->



<!-- ## Transcript length  -->

<!-- ```{r} -->
<!-- res_dte$UPF3B_KD_Control %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%  -->
<!--   mutate(normalized_3utr = LENGTH_3utr/Length ) %>% -->
<!--   ggplot(aes(x=Length, color = sig)) + stat_ecdf(geom = "point", size=0.8) + -->
<!--   scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +  -->
<!--   ylab("Cumulative Frequency") + xlab("") + theme(legend.position = "none") +  -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") + xlim(0, 25000) +  -->
<!--   ggtitle("Transcript length") -->
<!-- ``` -->




<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/upf3b_5utr.svg", plot = upf3b_5utr,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->
<!-- ``` -->




<!-- * dEJ and logFC -->

<!-- ```{r} -->
<!--  res_dte$UPF3B_KD_Control %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--     mutate_if(is.numeric, ~replace_na(., 0)) %>% -->
<!--   drop_na(NIF_dEJ) %>% -->
<!--   ggplot(aes(x=logFC, color = NIF_dEJ)) + stat_ecdf(geom = "step", size=0.8) + -->
<!--   scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +  -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +  -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +  -->
<!--   ggtitle("LogfC and dEJ") -->

<!-- res_dte$UPF3A_KD_Control %>%  -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>% -->
<!--     mutate_if(is.numeric, ~replace_na(., 0)) %>% -->
<!--   drop_na(NIF_dEJ) %>% -->
<!--   ggplot(aes(x=logFC, color = NIF_dEJ)) + stat_ecdf(geom = "step", size=0.8) + -->
<!--   scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +  -->
<!--   ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +  -->
<!--   geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +  -->
<!--   ggtitle("LogfC and dEJ") -->



<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/upf3a_3utr.svg", plot= upf3a_3utr,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->
<!-- ``` -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/upf3a_5utr.svg", plot= upf3a_5utr,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->

<!-- ``` -->

<!-- #### GC content in the 3'UTR region  -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/upf3a_3utr_gc.svg", plot= upf3a_3utr_gc,  -->
<!--        units = c("in"), height = 5.83, width = 7.52) -->
<!-- ``` -->

<!-- ## Bins of uORFs  -->

<!-- ```{r} -->
<!-- uorfs_upf3a = res_dte$UPF3A_KD_Control %>% -->
<!--   mutate(count_uORF = `uORFs|__|count`) %>%  -->
<!--   mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",  -->
<!--                                      between(count_uORF, 1, 2) ~"1-2",  -->
<!--                                      between(count_uORF, 3,4 ) ~ "3-4",  -->
<!--                                      count_uORF > 5 ~ ">5")) %>%  -->
<!--   drop_na(uORF_count_bins) %>% -->
<!--   mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",  -->
<!--                                                   "1-2", -->
<!--                                                   "3-4",  -->
<!--                                                   ">5"))) %>% -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%  -->
<!--   ggplot(aes(x=uORF_count_bins, fill = sig)) + geom_bar(stat="count", position = "dodge", color = "white") +  -->
<!--   theme_bw() + scale_fill_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) +  -->
<!--   coord_cartesian(ylim=c(0,1200)) + theme(legend.position = "none") -->


<!-- total = res_dte$UPF3A_KD_Control %>% -->
<!--   mutate(count_uORF = `uORFs|__|count`) %>%  -->
<!--   mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",  -->
<!--                                      between(count_uORF, 1, 2) ~"1-2",  -->
<!--                                      between(count_uORF, 3,4 ) ~ "3-4",  -->
<!--                                      count_uORF > 5 ~ ">5"),  -->
<!--          sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%  -->
<!--   drop_na(uORF_count_bins) %>% -->
<!--   mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",  -->
<!--                                                   "1-2", -->
<!--                                                   "3-4",  -->
<!--                                                   ">5"))) %>% -->
<!--   ggplot(aes(x=uORF_count_bins, fill = sig)) +  -->
<!--   geom_bar(stat="count", position = "dodge", color = "white") +  -->
<!--   theme_bw() + scale_fill_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) +  -->
<!--   coord_cartesian(ylim=c(4500,16000)) + theme(legend.position = "none") -->

<!-- ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/NIF-analysis/uORFs_upf3a.svg",  -->
<!--        units = "in",  height = 5.83, width = 7.52, plot = uorfs_upf3a ) -->


<!-- ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/NIF-analysis/uORFs_upf3a_total.svg",  -->
<!--        units = "in",  height = 3, width = 7.52, plot = total) -->
<!-- ``` -->

<!-- ### Upf3b  -->

<!-- ```{r} -->
<!-- uorfs_upf3b = res_dte$UPF3B_KD_Control %>% -->
<!--   mutate(count_uORF = `uORFs|__|count`) %>%  -->
<!--   mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",  -->
<!--                                      between(count_uORF, 1, 2) ~"1-2",  -->
<!--                                      between(count_uORF, 3,4 ) ~ "3-4",  -->
<!--                                      count_uORF > 5 ~ ">5")) %>%  -->
<!--   drop_na(uORF_count_bins) %>% -->
<!--   mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",  -->
<!--                                                   "1-2", -->
<!--                                                   "3-4",  -->
<!--                                                   ">5"))) %>% -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>%  -->
<!--   ggplot(aes(x=uORF_count_bins, fill = sig)) + geom_bar(stat="count", position = "dodge", color = "white") +  -->
<!--   theme_bw() + scale_fill_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) +  -->
<!--   coord_cartesian(ylim=c(0,1200)) + theme(legend.position = "none") -->


<!-- total = res_dte$UPF3B_KD_Control %>% -->
<!--   mutate(count_uORF = `uORFs|__|count`) %>%  -->
<!--   mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",  -->
<!--                                      between(count_uORF, 1, 2) ~"1-2",  -->
<!--                                      between(count_uORF, 3,4 ) ~ "3-4",  -->
<!--                                      count_uORF > 5 ~ ">5")) %>%  -->
<!--   drop_na(uORF_count_bins) %>% -->
<!--   mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",  -->
<!--                                                   "1-2", -->
<!--                                                   "3-4",  -->
<!--                                                   ">5"))) %>% -->
<!--   mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",  -->
<!--                       ifelse(logFC < 0 & FDR < 0.05, "Downregulaated", "Not Sig"))) %>%  -->
<!--   ggplot(aes(x=uORF_count_bins, fill = sig)) +  -->
<!--   geom_bar(stat="count", position = "dodge", color = "white") +  -->
<!--   theme_bw() + scale_fill_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) +  -->
<!--   coord_cartesian(ylim=c(4500,16000)) + theme(legend.position = "none") -->

<!-- ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/NIF-analysis/uORFs_upf3b.svg",  -->
<!--        units = "in",  height = 5.83, width = 7.52, plot = uorfs_upf3b) -->


<!-- ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/NIF-analysis/uORFs_upf3b_total.svg",  -->
<!--        units = "in",  height = 3, width = 7.52, plot = total) -->
<!-- ``` -->

# Differences between targets of Upf3a and Upf3b KDs

In this section I will perform a cumulative frequency analysis of the transcript features of Upf3a KD targets and Upf3b KD targets to better show the differences in impact. Subsequently, I will perform the KS statistic to see the significance of the difference. 

- 3'UTR 

One of the most prominant features in this analysis was the long 3'UTR. I want to see how the fold changes differ of the target transcripts containing the long 3'UTR, as well as the length and the GC content. 

* Fold change
```{r UTR-FC}
utr_FC = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(NIF_long_3utr == 1 & FDR < 0.05) %>% 
  dplyr::select(logFC) %>%
  mutate(Targets = "Upf3b KD") %>% 
  rbind(res_dte$UPF3A_KD_Control %>%
          dplyr::filter(NIF_long_3utr == 1 & FDR < 0.05) %>% 
          dplyr::select(logFC) %>% 
          mutate(Targets = "Upf3a KD")) %>% 
  ggplot(aes(x=logFC, color = Targets)) + stat_ecdf(geom = "point", size=0.2) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Log fold change of target transcripts with long 3'UTR") + 
  theme(legend.position = "none")

ggsave(plot = utr_FC, filename = here::here("output/Transcript/Thesis_figures/NIF_features/3UTR_FC.svg"), 
       height=8, width=8)

upf3b_fc = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(NIF_long_3utr == 1 & FDR < 0.05) %>% 
  dplyr::select(logFC)

upf3a_fc =  res_dte$UPF3A_KD_Control %>%
  dplyr::filter(NIF_long_3utr == 1 & FDR < 0.05) %>% 
  dplyr::select(logFC)

ks.test(upf3b_fc$logFC, upf3a_fc$logFC)
```

* Length 
```{r}
utr_length =res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(LENGTH_3utr) %>%
  mutate(Targets = "Upf3b KD") %>% 
  rbind(res_dte$UPF3A_KD_Control %>%
          dplyr::filter(FDR < 0.05) %>% 
          dplyr::select(LENGTH_3utr) %>% 
          mutate(Targets = "Upf3a KD")) %>% 
  ggplot(aes(x=LENGTH_3utr, color = Targets)) + stat_ecdf(geom = "point", size=0.2) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("3'UTR length") + xlim(0,7500) +
    theme(legend.position = "none")

ggsave(plot = utr_length, filename = here::here("output/Transcript/Thesis_figures/NIF_features/3UTR_length.svg"), 
       height=8, width=8)

upf3b_length = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(LENGTH_3utr)

upf3a_length =  res_dte$UPF3A_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(LENGTH_3utr)

ks.test(upf3b_length$LENGTH_3utr, upf3a_length$LENGTH_3utr)
```

* GC content 

```{r}
utr_GC= res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(GC_3utr) %>%
  mutate(Targets = "Upf3b KD") %>% 
  rbind(res_dte$UPF3A_KD_Control %>%
          dplyr::filter(FDR < 0.05) %>% 
          dplyr::select(GC_3utr) %>% 
          mutate(Targets = "Upf3a KD")) %>% 
  ggplot(aes(x=GC_3utr, color = Targets)) + stat_ecdf(geom = "point", size=0.2) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("3'UTR GC content") +
    theme(legend.position = "none")

ggsave(plot = utr_GC, filename = here::here("output/Transcript/Thesis_figures/NIF_features/3UTR_GC.svg"), 
       height=8, width=8)

upf3b_GC = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(GC_3utr)

upf3a_GC =  res_dte$UPF3A_KD_Control %>%
  dplyr::filter(FDR < 0.05) %>% 
  dplyr::select(GC_3utr)

ks.test(upf3b_GC$GC_3utr, upf3a_GC$GC_3utr)


```

- Downstream exon junction 

```{r}
dEJ_FC = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_dEJ == 1) %>% 
  dplyr::select(logFC) %>%
  mutate(Targets = "Upf3b KD") %>% 
  rbind(res_dte$UPF3A_KD_Control %>%
          dplyr::filter(FDR < 0.05 & NIF_dEJ ==1) %>% 
          dplyr::select(logFC) %>% 
          mutate(Targets = "Upf3a KD")) %>% 
  ggplot(aes(x=logFC, color = Targets)) + stat_ecdf(geom = "point", size=0.5) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Log fold change of target transcripts with downstream exon junction") + 
  xlab("Log2 fold change") +
    theme(legend.position = "none")


ggsave(plot = dEJ_FC, filename = here::here("output/Transcript/Thesis_figures/NIF_features/dEJ_FC.svg"), 
       height=8, width=8)
upf3b_dEJ = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_dEJ ==1) %>% 
  dplyr::select(logFC)

upf3a_dEJ =  res_dte$UPF3A_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_dEJ ==1) %>% 
  dplyr::select(logFC)

ks.test(upf3b_dEJ$logFC, upf3a_dEJ$logFC)

```

- upstream ORF 

```{r}
uORF_FC = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_uORF == 1) %>% 
  dplyr::select(logFC) %>%
  mutate(Targets = "Upf3b KD") %>% 
  rbind(res_dte$UPF3A_KD_Control %>%
          dplyr::filter(FDR < 0.05 & NIF_uORF ==1) %>% 
          dplyr::select(logFC) %>% 
          mutate(Targets = "Upf3a KD")) %>% 
  ggplot(aes(x=logFC, color = Targets)) + stat_ecdf(geom = "point", size=0.5) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Log fold change of target transcripts with uORF") + 
  xlab("Log2 fold change")



ggsave(plot = uORF_FC, filename = here::here("output/Transcript/Thesis_figures/NIF_features/uORF_FC.svg"), 
       height=8, width=8)
upf3b_uORF = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_uORF ==1) %>% 
  dplyr::select(logFC)

upf3a_uORF =  res_dte$UPF3A_KD_Control %>%
  dplyr::filter(FDR < 0.05 & NIF_uORF ==1) %>% 
  dplyr::select(logFC)

ks.test(upf3b_uORF$logFC, upf3a_uORF$logFC)
```

- Number of uORFs per comparison
```{r}
res_dte$UPF3A_KD_Control %>%
  mutate(count_uORF = `uORFs|__|count`) %>%
  mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",
                                     between(count_uORF, 1, 2) ~"1-2",
                                     between(count_uORF, 3,4 ) ~ "3-4",
                                     count_uORF > 5 ~ ">5")) %>%
  drop_na(uORF_count_bins) %>%
  mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",
                                                  "1-2",
                                                  "3-4",
                                                  ">5"))) %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%
 
  mutate(Targets = "Upf3a targets") %>%
  dplyr::select(sig, uORF_count_bins, Targets) %>%
  rbind(res_dte$UPF3B_KD_Control %>%
   mutate(count_uORF = `uORFs|__|count`) %>%
   mutate(uORF_count_bins = case_when(count_uORF == 0 ~ "0",
                                       between(count_uORF, 1, 2) ~"1-2",
                                       between(count_uORF, 3,4 ) ~ "3-4",
                                      count_uORF > 5 ~ ">5")) %>%
    drop_na(uORF_count_bins) %>%
    mutate(uORF_count_bins = factor(uORF_count_bins, levels = c("0",
                                                  "1-2",
                                                  "3-4",
                                                  ">5"))) %>%
  mutate(sig = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",
                      ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig"))) %>%
  mutate(sig = factor(sig, levels = c("Upregulated", "Downregulated", "Not Sig" ))) %>%
    as.data.frame() %>%
    mutate(Targets = "Upf3b targets") %>%
  dplyr::select(sig, uORF_count_bins, Targets)) %>%
  # dplyr::filter(sig != "Not Sig") %>%
  ggplot(aes(x=uORF_count_bins, fill = sig, color = "white")) + geom_bar(stat="count", position = "fill", color = "white") +
  theme_bw() + scale_fill_manual(values = c("Upregulated"= "#7D81BE", "Downregulated"= "#F1676B",
                                            "Not sig" ="grey")) + facet_wrap(~Targets)


```

- Previously annotated transcripts

Using the ensembl transcript annotations, I annotated the biotypes of transcripts. For the purpose of this, transcripts labelled with NMD and intron retention for grouped into one category

```{r}
 upf3b_tx_types = res_dte$UPF3B_KD_Control %>%  
 left_join(tx2gene) %>% 
  dplyr::filter(FDR < 0.05) %>%  
  mutate(CF_types = ifelse(tx_biotype == "retained_intron" |tx_biotype ==  "nonsense_mediated_decay", "NMD target", 
                            ifelse(tx_biotype == "protein_coding", "Protein Coding", 
                                 ifelse(tx_biotype == "processed_transcript", "Processed Transcript", "All other types"))))

upf3a_tx_types = res_dte$UPF3A_KD_Control %>%  
 left_join(tx2gene) %>% 
  dplyr::filter(FDR < 0.05) %>%  
  mutate(CF_types = ifelse(tx_biotype == "retained_intron" | tx_biotype == "nonsense_mediated_decay", "NMD target", 
                            ifelse(tx_biotype == "protein_coding", "Protein Coding", 
                                 ifelse(tx_biotype == "processed_transcript", "Processed Transcript", "All other types"))))

upf3b_tx_types %>% 
  dplyr::filter(CF_types == "NMD target") %>% 
  dplyr::select(logFC) %>% 
  mutate(Target = "Upf3b") %>%
  rbind(upf3a_tx_types %>% 
          dplyr::filter(CF_types== "NMD target") %>% 
          dplyr::select(logFC) %>% 
          mutate(Target = "Upf3a")) %>% 
  ggplot(aes(x=logFC, color = Target)) + stat_ecdf(geom = "point", size=0.5) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Log fold change of target transcripts with uORF") + 
  xlab("Log2 fold change") +
    theme(legend.position = "none")

upf3b_NMD =upf3b_tx_types %>% 
  dplyr::filter(CF_types == "NMD target") %>% 
  dplyr::select(logFC)

upf3a_NMD =upf3a_tx_types %>% 
  dplyr::filter(CF_types == "NMD target") %>% 
  dplyr::select(logFC)

ks.test(upf3b_NMD$logFC, upf3a_NMD$logFC)
```

# Variable importance

```{r}
library(party)


upf3b_tx = res_dte$UPF3B_KD_Control %>%
  dplyr::filter(FDR < 0.05, logFC > 0)

upf3b_rf = upf3b_tx %>% dplyr::select(tx_id,logFC,  GC_3utr, GC_5utr,  GC_cds, LENGTH_3utr,
                           LENGTH_5utr, LENGTH_cds, "uORFs|__|count", Length,
                           `dEJs|__|penultimate-exon>=55(nts)`) %>%
  column_to_rownames("tx_id")



cf1 <- cforest(logFC ~ . , data=upf3b_rf, control=cforest_unbiased(mtry=2,ntree=50))

imp_upf3b = varimp(cf1) %>% reshape2::melt() %>%
  rownames_to_column("Variable")


imp_upf3b %>%
  ggplot(aes(x=value, y = Variable)) + geom_bar(stat ="identity")


### Upf3a

upf3a_tx = res_dte$UPF3A_KD_Control %>%
  dplyr::filter(FDR < 0.05, logFC >0)


upf3a_rf = upf3a_tx %>% dplyr::select(tx_id,logFC,  GC_3utr, GC_5utr,  GC_cds, LENGTH_3utr,
                                      LENGTH_5utr, LENGTH_cds, "uORFs|__|count", Length,
                                      `dEJs|__|penultimate-exon>=55(nts)`) %>%
  column_to_rownames("tx_id")


cf1 <- cforest(logFC ~ . , data=upf3a_rf, control=cforest_unbiased(mtry=2,ntree=50))
cf1
imp_upf3a = varimp(cf1) %>% reshape2::melt() %>%
  rownames_to_column("Variable")


imp_upf3a %>%
  ggplot(aes(x=value, y = Variable)) + geom_bar(stat ="identity")




```
