---
title: "Differential-transcript-usage"
author: "unawaz1996"
date: "2023-03-22"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```


# Background 

# Set-up 


# Background

- What is differential transcript usage 

- Why is isoform switching and differential transcript usage important to use in NMD? 

- What are some NMD features we should be looking out for - signs of escape from NMD 

Identify and compare the cellular and molecular roles Upf3a through gain and loss of function approaches in mouse L-cells 

 
# Analysis

```{r}
source("code/libraries.R")
```


<!-- # ```{r} -->
<!-- # salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Salmon") -->
<!-- # md.files = "data/LTK_Sample Metafile_V3.txt" -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # myComparison <- data.frame( -->
<!-- #   condition_1 = c("Control", "Control", "UPF3B_KD", "Control", "Control"), -->
<!-- #   condition_2 = c("UPF3A_KD","UPF3B_KD", "UPF3A_KD", "UPF3A_KD_UPF3B_KD", "UPF3A_OE")) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame") -->
<!-- # tx2gene = as.data.frame(txdf[,c("tx_id","gene_id")]) -->
<!-- # tab <- table(tx2gene$gene_id) -->
<!-- # tx2gene$ntx <- tab[match(tx2gene$gene_id, names(tab))] -->
<!-- #  -->
<!-- #  -->
<!-- # salmon = list.files(salmon.files, pattern = "transcripts$", full.names = TRUE) -->
<!-- # md = read.table(md.files, header= TRUE) -->
<!-- # md$Group = factor(md$Group) -->
<!-- # rownames(md) = md$Sample -->
<!-- # colnames(md)[1] = c("sample") -->
<!-- #  -->
<!-- # md %<>% mutate(path = salmon) -->
<!-- # all_files = file.path(salmon, "quant.sf") -->
<!-- # names(all_files) <- paste(c(212:229)) -->
<!-- #  -->
<!-- # all_files = all_files[-c(6,12,18)] -->
<!-- #  -->
<!-- # md.files = "data/LTK_Sample Metafile_V3.txt" -->
<!-- # design = read.table(md.files, header= TRUE) -->
<!-- # design %<>% dplyr::select(Sample, Group) -->
<!-- # colnames(design) = c("sampleID", "condition") -->
<!-- # design$sampleID = paste(design$sampleID, "L4_transcripts", sep = "_") -->
<!-- # design$sampleID = as.character(design$sampleID) -->
<!-- # design=design[-c(6,12, 18),] -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # salmonQuant <- importIsoformExpression( -->
<!-- #     parentDir = salmon.files,  -->
<!-- #     addIsofomIdAsColumn = TRUE -->
<!-- # ) -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- #  -->
<!-- # aSwitchList <- importRdata( -->
<!-- #   isoformCountMatrix   = salmonQuant$counts[,-c(7, 13, 19)], -->
<!-- #   isoformRepExpression = salmonQuant$abundanc[,-c(7, 13, 19)], -->
<!-- #   designMatrix         = design, -->
<!-- #   addAnnotatedORFs = TRUE, -->
<!-- #   comparisonsToMake = myComparison, -->
<!-- #   isoformExonAnnoation = "/home/neuro/Documents/NMD_analysis/Analysis/Genome/Mus_musculus.GRCm39.107.gtf.gz", -->
<!-- #   isoformNtFasta       = "/home/neuro/Documents/NMD_analysis/Analysis/Genome/for_salmon/Mus_musculus.GRCm39.cdna.all.fa", -->
<!-- #   fixStringTieAnnotationProblem = TRUE, -->
<!-- #   showProgress = FALSE -->
<!-- # ) -->
<!-- # ``` -->
<!-- #  -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # aSwitchListFiltered = preFilter( -->
<!-- #   switchAnalyzeRlist = aSwitchList, -->
<!-- #   geneExpressionCutoff = .5, -->
<!-- #   isoformExpressionCutoff = 0, -->
<!-- #   removeSingleIsoformGenes = TRUE) -->
<!-- #  -->
<!-- # ``` -->
<!-- #  -->
<!-- # ```{r} -->
<!-- # ASwitchListAnalyzed = isoformSwitchTestDEXSeq( -->
<!-- #   switchAnalyzeRlist = aSwitchListFiltered, -->
<!-- #   reduceToSwitchingGenes=TRUE, -->
<!-- #   showProgress = TRUE) -->
<!-- #  -->
<!-- # ``` -->

```{r}
#new = ASwitchListAnalyzed

load("output/ISAR/new_res/ISAR_results.Rda")
summary(ASwitchListAnalyzed)

```


<!-- ```{r} -->
<!-- exampleSwitchListAnalyzed <- extractSequence( -->
<!--     ASwitchListAnalyzed, -->
<!--     pathToOutput = 'output/ISAR/new_res', -->
<!--     writeToFile=TRUE # to avoid output when running this example data -->
<!-- ) -->

<!-- ``` -->

<!-- # ```{r} -->
<!-- # ASwitchListAnalyzed = analyzeCPC2( -->
<!-- #   switchAnalyzeRlist = ASwitchListAnalyzed,  -->
<!-- #   pathToCPC2resultFile = "output/ISAR/new_res/result_cpc2.txt",  -->
<!-- #   removeNoncodinORFs   = FALSE) -->
<!-- #  -->
<!-- # ASwitchListAnalyzed <- analyzeSignalP( -->
<!-- #   switchAnalyzeRlist       = ASwitchListAnalyzed, -->
<!-- #   pathToSignalPresultFile  = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/ISAR/new_res/prediction_results.txt") -->
<!-- #  -->
<!-- # ASwitchListAnalyzed <- analyzeIUPred2A( -->
<!-- #   switchAnalyzeRlist = ASwitchListAnalyzed,  -->
<!-- #   pathToIUPred2AresultFile = "output/ISAR/new_res/isoformSwitchAnalyzeR_isoform_AA.result" -->
<!-- # ) -->
<!-- #  -->
<!-- # ASwitchListAnalyzed <- analyzeAlternativeSplicing( -->
<!-- #   switchAnalyzeRlist = ASwitchListAnalyzed, -->
<!-- #   quiet=TRUE) -->
<!-- #  -->
<!-- # ``` -->

```{r}
# ASwitchListAnalyzed <- analyzeIUPred2A(
#   switchAnalyzeRlist = ASwitchListAnalyzed,  
#   pathToIUPred2AresultFile = "output/ISAR/new_res/isoformSwitchAnalyzeR_isoform_AA.result")



# ASwitchListAnalyzed <- analyzeSignalP(
#    switchAnalyzeRlist       = ASwitchListAnalyzed,
#    pathToSignalPresultFile  = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/ISAR/new_res/SIgnalIP/prediction_results.txt")
```


<!-- # -->
<!-- # ASwitchListAnalyzed <- analyzeSignalP( -->
<!-- #   switchAnalyzeRlist       = ASwitchListAnalyzed, -->
<!-- #   pathToSignalPresultFile  = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/ISAR/new_res/prediction_results.txt") -->
```{r}
consequencesOfInterest <- c('intron_retention',
                            'NMD_status', 'IDR_identified', 
                            'IDR_type')
ASwitchListAnalyzed <- analyzeSwitchConsequences(
    ASwitchListAnalyzed,
    consequencesToAnalyze = consequencesOfInterest, 
    dIFcutoff = 0.1, # very high cutoff for fast runtimes - you should use the default (0.1)
    showProgress=FALSE
)

```

# Results 
Results for each analysis and comparison. Columns indicate as following: 


- `IF1`: Isoform fraction (IF) ratio in Controls calculated by dividing the isoform expression with the expression of the parent gene
- `IF2`: IF ration in Condition calculated by dividing the isoform expression with the expression of the parent gene 
- `dIF`: The difference in IF values (IF2 - IF1)
- `isoform_switch_q_value`: Isoform level False Discovery Rate (FDR) corrected P-values 



* All data

```{r}
ASwitchListAnalyzed$isoformFeatures %>% 
  #dplyr::filter(condition_2 == "UPF3B_KD" & isoform_switch_q_value < 0.05 & abs(dIF) > 0.1 ) %>% 
  #dplyr::select(-c("condition_1", "condition_2", "iso_ref", "gene_ref"), -contains("gene_value"), 
   #             -contains("gene_stderr_"), -contains("iso_value"), -contains("iso_stderr")) %>%
  dplyr::select("gene_name", everything()) %>% 
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


* UPF3B KD 

```{r}
ASwitchListAnalyzed$isoformFeatures %>% 
  dplyr::filter(condition_2 == "UPF3B_KD" & isoform_switch_q_value < 0.05 & abs(dIF) > 0.1 ) %>% 
  dplyr::select(-c("condition_1", "condition_2", "iso_ref", "gene_ref"), -contains("gene_value"), 
                -contains("gene_stderr_"), -contains("iso_value"), -contains("iso_stderr")) %>%
  dplyr::select("gene_name", everything()) %>% 
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```





* UPF3A KD 

```{r}
ASwitchListAnalyzed$isoformFeatures %>% 
  dplyr::filter(condition_2 == "UPF3A_KD" & isoform_switch_q_value < 0.05 & abs(dIF) > 0.1 ) %>% 
  dplyr::select(-c("condition_1", "condition_2", "iso_ref", "gene_ref"), -contains("gene_value"), 
                -contains("gene_stderr_"), -contains("iso_value"), -contains("iso_stderr")) %>%
  dplyr::select("gene_name", everything()) %>% 
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


* Double KD 

```{r}
ASwitchListAnalyzed$isoformFeatures %>% 
  dplyr::filter(condition_2 == "UPF3A_KD_UPF3B_KD" & isoform_switch_q_value < 0.05 & abs(dIF) > 0.1 ) %>% 
  dplyr::select(-c("condition_1", "condition_2", "iso_ref", "gene_ref"), -contains("gene_value"), 
                -contains("gene_stderr_"), -contains("iso_value"), -contains("iso_stderr")) %>%
  dplyr::select("gene_name", everything()) %>% 
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```




# Overview of results 
After performing the analysis and annotating the transcripts we now want to see a global overview of the transcript usage. 
```{r}
#load("../output/ISAR/switchList.Rda")
```


```{r}
 extractSwitchSummary(
     ASwitchListAnalyzed,
     filterForConsequences = FALSE) %>% pander(style="rmarkdown", 
                                               caption= "Summary of switches across all comparisons. Results show number of significant isoforms. switches and gene ")
```




```{r}
 extractSwitchSummary(
     ASwitchListAnalyzed,
     filterForConsequences = TRUE) %>% pander(style="rmarkdown", 
                                               caption= "Summary of switches across all comparisons with functional consequences. Results show number of significant isoforms. switches and gene ")
```


```{r}
extractConsequenceSummary(
    ASwitchListAnalyzed,
    consequencesToAnalyze='all',
    plotGenes = FALSE,           # enables analysis of genes (instead of isoforms)
    asFractionTotal = FALSE      # enables analysis of fraction of significant features
)
```


```{r fig.width=10}
extractConsequenceEnrichment(
    ASwitchListAnalyzed,
    consequencesToAnalyze='all',
    analysisOppositeConsequence = TRUE,
    returnResult = FALSE)
```




```{r fig.height=12, fig.width=15}
extractSplicingEnrichmentComparison(
    ASwitchListAnalyzed,
    splicingToAnalyze = c('A3','A5','ATSS','ATTS'), # the splicing highlighted above
    returnResult = FALSE # if TRUE returns a data.frame with the results
)
```


<!-- # ```{r} -->
<!-- # ggplot(data=ASwitchListAnalyzed$isoformFeatures, aes(x=dIF, y=-log10(isoform_switch_q_value))) + -->
<!-- #      geom_point(aes( color=PTC & isoform_switch_q_value < 0.05 & abs(dIF) > 0.1 ), # default cutoff -->
<!-- #         size=0.5) + -->
<!-- #     geom_point(data=filter(., is.na(PTC)), aes(color="lightgray"), alpha = 0.5) +  -->
<!-- # 	  geom_point(data=filter(dIF_all, PTC == FALSE), aes(color="royalblue"), alpha = 0.5) + -->
<!-- # 	  geom_point(data=filter(dIF_all, PTC == TRUE), aes(color="red"), alpha = 0.5) + -->
<!-- #     geom_hline(yintercept = -log10(0.05), linetype='dashed') + # default cutoff -->
<!-- #     geom_vline(xintercept = c(-0.1, 0.1), linetype='dashed') + # default cutoff -->
<!-- #     facet_wrap( ~ condition_2) + -->
<!-- #     #facet_grid(condition_1 ~ condition_2) + # alternative to facet_wrap if you have overlapping conditions -->
<!-- #     scale_color_manual('PTC', values = c('black','red')) + -->
<!-- #     labs(x='dIF', y='-Log10 ( Isoform Switch Q Value )') + -->
<!-- #     theme_bw() -->
<!-- #  -->
<!-- # ``` -->

### Switch plots of specific genes 

```{r}
switchingIso <- extractTopSwitches( 
    ASwitchListAnalyzed, 
    filterForConsequences = TRUE, 
    n = NA,                  # n=NA: all features are returned
    extractGenes = FALSE,    # when FALSE isoforms are returned
    sortByQvals = TRUE
)
```

```{r}
switchingIso %>% 
  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```


<!-- # ```{r} -->
<!-- # switchPlot( -->
<!-- #     exampleSwitchListAnalyzed, -->
<!-- #     gene='Cdkl2', -->
<!-- #     condition1 = 'Control', -->
<!-- #     condition2 = 'UPF3A_KD', -->
<!-- #     localTheme = theme_bw(base_size = 13) # making text sightly larger for vignette -->
<!-- # ) -->
<!-- #  -->
<!-- # ``` -->



## Data for each analysis



# Exporting data 

```{r}
save(ASwitchListAnalyzed, file=here::here("output/ISAR/new_res/ISAR_results.Rda"))
```

