---
title: "Enichment-analysis-fgsea"
author: "unawaz1996"
date: "2023-03-16"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```

# Fast gene set enrichment analysis 
```{r load libraries}
source("code/libraries.R")
source("code/functions.R")
```

```{r}
theme_set(theme_bw())
panderOptions("table.split.table", Inf)
panderOptions("table.style", "rmarkdown")
panderOptions("big.mark", ",")
```

```{r load data}
load("output/limma-matrices.Rda")
#load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")
```

Fast Gene Set Enrichment Analysis implements a fast version of the gsea algorithm. As a result, more permutations are made to get more fine grained p-values. 


`fgsea` results can be interpreted as following: 

- `ES`: Enrichment score which is calculated based on the rank
- `NES`: Normalized Enrichment Score


# Gene sets

- Hallmark Gene sets 

```{r}
library(msigdbr)
hm = msigdbr(species = "mouse", category = "H") %>% 
  distinct(gs_name, ensembl_gene, .keep_all = TRUE)
```

- C2 genesets

Using the MSigDB genesets as above, we will conduct a fgsea analysis on the sets of interest including KEGG, Wikipathways and Reactome. In our analysis, we will be using 2,465 gene sets in total.  

```{r}
c2 =  msigdbr(species = "mouse", category = c("C2")) %>% 
  dplyr::filter(gs_subcat == "CP:REACTOME") %>% 
  distinct(gs_name, ensembl_gene, .keep_all = TRUE)

```


```{r}
ah <- AnnotationHub() %>%
    subset(species == "Mus musculus") %>%
    subset(rdataclass == "EnsDb") %>% 
  subset(genome == "GRCm39")
ensDb <- ah[["AH104895"]]
grTrans <- transcripts(ensDb)
trLengths <- exonsBy(ensDb, "tx") %>%
    width() %>%
    vapply(sum, integer(1))
mcols(grTrans)$length <- trLengths[names(grTrans)]

genesGR = genes(ensDb)
```

```{r}
id2Name <- structure(
  genesGR$gene_name,
  names = genesGR$gene_id
) %>% 
  .[!duplicated(names(.))]
```

### Hallmark gene sets 

```{r include = FALSE}
fgsea_res = run_fgsea(v = v,  contrasts = contrasts, design = design, idx = hm )
```


```{r}
sigRes = list()

sigRes[["UPF3B_KD_vs_Control"]] = fgsea_res$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) 
  

sigRes[["UPF3A_KD_vs_Control"]] = fgsea_res$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

sigRes[["UPF3A_OE_vs_Control"]] = fgsea_res$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

sigRes[["DoubleKD_vs_Control"]] = fgsea_res$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

sigRes[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)
```

```{r}
sigRes$UPF3B_KD_vs_Control %>% 
  separate_rows(fgsea.leadingEdge_genes, sep = ";") %>% 
  mutate(leadingEdge = gsub(" ", "", fgsea.leadingEdge_genes)) %>% 
  dplyr::rename("SYMBOL" = "fgsea.leadingEdge_genes") %>% 
  inner_join(DEGenes$UPF3B_vs_Control$gene %>%
              as.data.frame() %>% 
              set_colnames("SYMBOL"), by = "SYMBOL")

```

```{r}
sigRes_summary = lapply(sigRes, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES, leSize, fgsea.size) })



sig_heatmap_mh = sigRes_summary %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>% 
  mutate(comparison=gsub("\\.[0-9]*$","", comparison)) %>% as.data.frame()
```



```{r}
x = list("Upf3b" = sigRes$UPF3B_KD_vs_Control$fgsea.Geneset, 
         "Upf3a" = sigRes$UPF3A_KD_vs_Control$fgsea.Geneset)
ggvenn(x)
```



```{r}
genesets = calculate.overlap(x)
```

What are the genes (that are also DEG) in each category?

```{r}
in_common_upf3b = sigRes$UPF3B_KD_vs_Control %>%  
  dplyr::filter(fgsea.Geneset %in% genesets$a3) %>%
  separate_rows(fgsea.leadingEdge_genes, sep=";") %>%
    dplyr::rename("SYMBOL" = "fgsea.leadingEdge_genes") %>% 
  dplyr::mutate(Comparison = "Upf3b KD") %>% as.data.frame() 



in_common_upf3a = sigRes$UPF3A_KD_vs_Control %>%  
  dplyr::filter(fgsea.Geneset %in% genesets$a3) %>%
  separate_rows(fgsea.leadingEdge_genes, sep=";") %>%
    dplyr::rename("SYMBOL" = "fgsea.leadingEdge_genes") %>% 
  dplyr::mutate(Comparison = "Upf3a KD")
```

```{r}
in_common_upf3a %<>%
  dplyr::filter(fgsea.Geneset %in% c("HALLMARK_MTORC1_SIGNALING")) %>% as.data.frame()

in_common_upf3b %<>%
  dplyr::filter(fgsea.Geneset %in% c("HALLMARK_MTORC1_SIGNALING")) %>% as.data.frame()
```

```{r}
x = list("Upf3b" = in_common_upf3b$SYMBOL, 
         "Upf3a" = in_common_upf3a$SYMBOL)
ggvenn(x)
```

```{r}
in_common_upf3a$SYMBOL
```

## Reactome datasets 

```{r include = FALSE}
fgsea_res_c2 = run_fgsea(v = v,  contrasts = contrasts, design = design, idx = c2)
```


```{r}
sigres_c2 = list()

sigres_c2[["UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

sigres_c2[["UPF3A_KD_vs_Control"]] = fgsea_res_c2$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

sigres_c2[["UPF3A_OE_vs_Control"]] = fgsea_res_c2$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

sigres_c2[["DoubleKD_vs_Control"]] = fgsea_res_c2$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

sigres_c2[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  )
```

```{r}
sigRes_all = list()

sigRes_all[["UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>%
  rbind(sigRes$UPF3B_KD_vs_Control)

sigRes_all[["UPF3A_KD_vs_Control"]] = fgsea_res_c2$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_KD_vs_Control)

sigRes_all[["UPF3A_OE_vs_Control"]] = fgsea_res_c2$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>%
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_OE_vs_Control)

sigRes_all[["DoubleKD_vs_Control"]] = fgsea_res_c2$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>%
  rbind(sigRes$DoubleKD_vs_Control)

sigRes_all[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_OE_UPF3B_KD_vs_Control)
```


```{r}
sigRes_all = list()

sigRes_all[["UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>%
  rbind(sigRes$UPF3B_KD_vs_Control)

sigRes_all[["UPF3A_KD_vs_Control"]] = fgsea_res_c2$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_KD_vs_Control)

sigRes_all[["UPF3A_OE_vs_Control"]] = fgsea_res_c2$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>%
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_OE_vs_Control)

sigRes_all[["DoubleKD_vs_Control"]] = fgsea_res_c2$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>%
  rbind(sigRes$DoubleKD_vs_Control)

sigRes_all[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
    dplyr::filter(padj < 0.05) %>% 
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) %>% 
  rbind(sigRes$UPF3A_OE_UPF3B_KD_vs_Control)
```


```{r}
#write.xlsx(sigRes_all, file= "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/fgsea-output.xlsx")
```



```{r}
sigRes_summary_c2 = lapply(sigres_c2, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES, leSize, fgsea.size) })



sig_heatmap = sigRes_summary_c2 %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>%
  mutate(comparison=gsub("\\.[0-9]*$","", comparison))

all_heatmap =rbind(sig_heatmap_mh, sig_heatmap)
```



## Upf3a and Upf3b
```{r}
x = list("Upf3b" = sigRes_all$UPF3B_KD_vs_Control$fgsea.Geneset, 
         "Upf3a" = sigRes_all$UPF3A_KD_vs_Control$fgsea.Geneset)
ggvenn(x)
```



```{r}
x = list("Upf3b" = sigRes_all$UPF3B_KD_vs_Control$fgsea.Geneset, 
         "Upf3a" = sigRes_all$UPF3A_KD_vs_Control$fgsea.Geneset)
ggvenn(x)
```



```{r}
genesets = calculate.overlap(x)
```

<!-- What are the genes (that are also DEG) in each category? -->

<!-- ```{r} -->
<!-- in_common_upf3b = sigRes_all$UPF3B_KD_vs_Control %>%   -->
<!--   #dplyr::filter(fgsea.Geneset %in% genesets$a3) %>% -->
<!--   separate_rows(fgsea.leadingEdge_genes, sep=";") %>% -->
<!--     dplyr::rename("SYMBOL" = "fgsea.leadingEdge_genes") %>%  -->
<!--   as.data.frame() #%>% -->
<!-- #  inner_join(DEGenes$UPF3B_vs_Control$gene %>% -->
<!--  #              as.data.frame() %>% -->
<!--   #             set_colnames("SYMBOL"), by = "SYMBOL") -->



<!-- in_common_upf3a = sigRes_c2$UPF3A_KD_vs_Control %>%   -->
<!--   dplyr::filter(fgsea.Geneset %in% genesets$a3) %>% -->
<!--   separate_rows(fgsea.leadingEdge, sep=";") %>% -->
<!--     dplyr::rename("SYMBOL" = "fgsea.leadingEdge") %>% -->
<!--   as.data.frame() #%>% -->
<!--  # right_join(DEGenes$UPF3A_vs_Control$gene %>% -->
<!--   #             as.data.frame() %>% -->
<!--    #            set_colnames("SYMBOL"), by = "SYMBOL") -->

<!-- ``` -->


```{r}
x = list("Upf3b" = in_common_upf3b$SYMBOL, 
         "Upf3a" = in_common_upf3a$SYMBOL)
ggvenn(x)
```


Next I wanted to categorize my genesets 

```{r}
categories_reactome = read_excel(here::here("data/genesets/genesets-categories.xls")) %>% 
  as.data.frame() %>% 
  dplyr::select(-thesis_category)

all_heatmap %<>%
  dplyr::rename(Geneset = "fgsea.Geneset") %>% 
  left_join(categories_reactome)

all_heatmap$Category[4] <- c("Extracellular matrix")
```




<!-- ## Summary of results from all comparisons  -->


<!-- ```{r} -->
<!-- enrich_GSEA = list("UPF3B KD" = sigRes$UPF3B_KD_vs_Control$fgsea.Geneset,  -->
<!--                 "UPF3A KD" = sigRes$UPF3A_KD_vs_Control$fgsea.Geneset,  -->
<!--                 "Double KD" = sigRes$DoubleKD_vs_Control$fgsea.Geneset,  -->
<!--                 "UPF3A OE" = sigRes$UPF3A_OE_vs_Control$fgsea.Geneset,  -->
<!--                 "UPF3A OE, UPF3B KD" = sigRes$UPF3A_OE_UPF3B_KD_vs_Control$fgsea.Geneset) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- allRes_summary = lapply(allRes, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES, fgsea.padj) }) -->

<!-- allRes_summary = allRes_summary %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>%  -->
<!--   mutate(comparison=gsub("\\.[0-9]*$","", comparison)) #%>%  -->
<!--   xtabs(fgsea.NES ~ comparison + fgsea.Geneset,.) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- rg <- max(abs(allRes_summary$fgsea.NES)); -->

<!-- allRes_summary %>%  -->
<!--   head(10) %>% -->
<!-- ggplot(aes(fgsea.Geneset, forcats::fct_rev(comparison),  -->
<!--            fill = fgsea.NES, size = -log(fgsea.padj))) + -->
<!--   geom_point(shape = 21, stroke = 1) + -->
<!--   geom_hline(yintercept = seq(.5, 4.5, 1), size = .2) + -->
<!--   scale_x_discrete(position = "top") + -->
<!--   scale_radius(range = c(1, 15)) + -->
<!--   scale_fill_gradient(low = "#2F124B", high = "#EA2A5F", -->
<!--                       breaks = seq(-rg, rg, length.out = 100)) + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom",  -->
<!--         panel.grid.major = element_blank(), -->
<!--         legend.text = element_text(size = 8), -->
<!--         legend.title = element_text(size = 8)) + -->
<!--   guides(size = guide_legend(override.aes = list(fill = NA, color = "black", stroke = .25),  -->
<!--                              label.position = "bottom", -->
<!--                              title.position = "right",  -->
<!--                              order = 1), -->
<!--          fill = guide_colorbar(ticks.colour = NA, title.position = "top", order = 2)) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- library(ggplot2) -->

<!-- set.seed(42) -->

<!-- d <- data.frame( -->
<!--   x = rep(paste("Team", LETTERS[1:8]), 4), -->
<!--   y = rep(paste("Task", 1:4), each = 8), -->
<!--   value = runif(32) -->
<!-- ) -->

<!-- ggplot(d, aes(x, forcats::fct_rev(y), fill = value, size = value)) + -->
<!--   geom_point(shape = 21, stroke = 0) + -->
<!--   geom_hline(yintercept = seq(.5, 4.5, 1), size = .2) + -->
<!--   scale_x_discrete(position = "top") + -->
<!--   scale_radius(range = c(1, 15)) + -->
<!--   scale_fill_gradient(low = "orange", high = "blue", breaks = c(0, .5, 1), labels = c("Great", "OK", "Bad"), limits = c(0, 1)) + -->
<!--   theme_minimal() + -->
<!--   theme(legend.position = "bottom",  -->
<!--         panel.grid.major = element_blank(), -->
<!--         legend.text = element_text(size = 8), -->
<!--         legend.title = element_text(size = 8)) + -->
<!--   guides(size = guide_legend(override.aes = list(fill = NA, color = "black", stroke = .25),  -->
<!--                              label.position = "bottom", -->
<!--                              title.position = "right",  -->
<!--                              order = 1), -->
<!--          fill = guide_colorbar(ticks.colour = NA, title.position = "top", order = 2)) + -->
<!--   labs(size = "Area = Time Spent", fill = "Score:", x = NULL, y = NULL) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- sigRes_summary = lapply(sigRes, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES, fgsea.padj) }) -->

<!-- sig_heatmap = sigRes_summary %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>%  -->
<!--   mutate(comparison=gsub("\\.[0-9]*$","", comparison)) %>%  -->
<!--   xtabs(fgsea.NES ~ comparison + fgsea.Geneset,.) -->




<!-- sig_heatmap %<>% as.data.frame.matrix(sig_heatmap, row.names = c("Double KD", "UPF3A KD",  -->
<!--                                                                  "UPF3A OE, UPF3B KD", "UPF3A OE", -->
<!--                                                                  "UPF3B KD")) -->



<!-- rg <- sig_heatmap %>% max(abs(.)); -->

<!-- # class(sig_heatmap) -->
<!-- # #[c(5, 2, 1, 4,3),] -->
<!-- # WP_heatmap = sig_heatmap[c(5, 2, 1, 4,3),] %>%  -->
<!-- #   dplyr::select(contains("WP_")) %>% -->
<!-- #     pheatmap(, -->
<!-- #         scale="none",  -->
<!-- #         cluster_cols = FALSE, -->
<!-- #         cluster_rows = FALSE, -->
<!-- #         breaks = seq(-rg, rg, length.out = 100), -->
<!-- #         clustering_distance_rows = "euclidean",  -->
<!-- #         border_color = "white",  -->
<!-- #         treeheight_row = 0, -->
<!-- #         fontsize = 6, -->
<!-- #         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!-- #         labels_row = row.names(.),  -->
<!-- #         cellwidth = 15,  -->
<!-- #         cellheight = 15 -->
<!-- #     ) -->
<!-- # ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/WP_pathway_fgsea.svg", width=18.80, height =7.79, units = "in", plot= WP_heatmap) -->

<!-- sig_heatmap[c(5, 2, 1, 4,3),] %>%  -->
<!--   as.data.frame() %>% -->
<!--   rownames_to_column("conditions") %>%  -->
<!--   dplyr::filter("UPF3A KD" != 0  & "UPF3B KD" > 0) %>% -->
<!--   column_to_rownames("conditions") %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.) -->
<!--     ) -->

<!-- HM_heatmap = sig_heatmap[c(5, 2, 1, 4,3),] %>%  -->
<!--   dplyr::select(contains("HALLMARK_")) %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 15,  -->
<!--         cellheight = 15 -->
<!--     ) -->
<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/KEGG_pathway_fgsea_July.svg", width=18.80, height =7.79, units = "in", plot= KEGG_heatmap) -->

<!-- ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/HM_pathway_fgsea_July.svg", width=18.80, height =7.79, units = "in", plot= HM_heatmap) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- allRes_summary = lapply(allRes, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES) }) -->


<!-- all_heatmap = allRes_summary %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>%  -->
<!--   mutate(comparison=gsub("\\.[0-9]*$","", comparison)) %>%  -->
<!--   xtabs(fgsea.NES ~ comparison + fgsea.Geneset,.) -->


<!-- all_heatmap  %<>% as.data.frame.matrix(all_heatmap , row.names = c("Double KD", "UPF3A KD",  -->
<!--                                                                  "UPF3A OE, UPF3B KD", "UPF3A OE", -->
<!--                                                                  "UPF3B KD")) -->


<!-- all_heatmap[c(5, 2, 1, 4,3),] %>%  -->
<!--   dplyr::select(contains("HALLMARK_")) %>% -->
<!--     pheatmap(, -->
<!--         scale="none",  -->
<!--         cluster_cols = FALSE, -->
<!--         cluster_rows = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 15,  -->
<!--         cellheight = 15 -->
<!--     ) -->


<!-- ``` -->




<!-- ```{r} -->
<!-- bf_genes = read_excel("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Upf3b-high-confident-genes.xlsx") %>%  -->
<!--   as.data.frame() #%>% -->
<!--  # dplyr::filter(logFC.x > 0) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- sigRes_all$UPF3B_KD_vs_Control %>%   -->
<!--   #dplyr::filter(fgsea.Geneset %in% genesets$a3) %>% -->
<!--   separate_rows(fgsea.leadingEdge_gesnes, sep=";") %>% -->
<!--     dplyr::rename("gene.x" = "fgsea.leadingEdge_genes") %>%  -->
<!--   as.data.frame() %>%  -->
<!--   inner_join(bf_genes, by = "gene.x") -->
<!-- ``` -->


<!-- # Hallmark sets  -->

<!-- ```{r message = FALSE, include = FALSE} -->
<!-- fgsea_resH = run_fgsea(v = v,  contrasts = contrasts, design = design, idx = hm) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- sigRes_H = list() -->

<!-- sigRes_H[["UPF3B_KD_vs_Control"]] = fgsea_resH$UPF3B_vs_Control %>% as.data.frame %>% -->
<!--     dplyr::filter(fgsea.padj < 0.05)  -->

<!-- sigRes_H[["UPF3A_KD_vs_Control"]] = fgsea_resH$UPF3A_vs_Control %>% as.data.frame %>% -->
<!--     dplyr::filter(fgsea.padj < 0.05)  -->

<!-- sigRes_H[["UPF3A_OE_vs_Control"]] = fgsea_resH$UPF3A_OE_vs_Control %>% as.data.frame %>% -->
<!--     dplyr::filter(fgsea.padj < 0.05)  -->

<!-- sigRes_H[["DoubleKD_vs_Control"]] = fgsea_resH$DoubleKD_vs_Control %>% as.data.frame %>% -->
<!--     dplyr::filter(fgsea.padj < 0.05)  -->

<!-- sigRes_H[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_resH$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>% -->
<!--     dplyr::filter(fgsea.padj < 0.05)  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- x = list("Upf3b" = sigRes_H$UPF3B_KD_vs_Control$fgsea.Geneset,  -->
<!--          "Upf3a" = sigRes_H$UPF3A_KD_vs_Control$fgsea.Geneset) -->
<!-- ggvenn(x) -->

<!-- ``` -->




<!-- ```{r fig.height=10, fig.width=10} -->
<!-- sigRes_H_summary = lapply(sigRes_H, function(x) { x %>% dplyr::select(fgsea.Geneset, fgsea.NES) }) -->



<!-- sig_heatmap = sigRes_H_summary %>% do.call(rbind, .) %>% rownames_to_column("comparison") %>%  -->
<!--   mutate(comparison=gsub("\\.[0-9]*$","", comparison)) %>%  -->
<!--   xtabs(fgsea.NES ~ comparison + fgsea.Geneset,.) -->

<!-- rg <- sig_heatmap %>% max(abs(.)); -->


<!-- hallmark_heatmap = sig_heatmap[c(5, 2, 1, 4,3),]  %>%  -->
<!--     pheatmap(, -->
<!--         scale="none", -->
<!--         cluster_rows = FALSE,  -->
<!--         cluster_cols = FALSE, -->
<!--         breaks = seq(-rg, rg, length.out = 100), -->
<!--         clustering_distance_rows = "euclidean",  -->
<!--         border_color = "white",  -->
<!--         treeheight_row = 0, -->
<!--         fontsize = 6, -->
<!--         color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100), -->
<!--         labels_row = row.names(.),  -->
<!--         cellwidth = 15, -->
<!--         cellheight=15) -->

<!-- ggsave(filename="/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/fgsea_gsea_H.svg",  -->
<!--        plot = hallmark_heatmap, width =9.23, height = 7.79, units = "in") -->

<!-- #%>% -->
<!--  # pheatmap(border_color = "white", cellheight =15, cellwidth = 15, scale = "none") -->


<!-- ``` -->



## Creating bubble heatmaps 

```{r}
heatmap_1 =all_heatmap  |> 
  mutate(geneRatio = leSize/fgsea.size,
         Geneset = tolower(gsub("REACTOME_", "", Geneset))) |>
  mutate(Geneset = gsub("_", " ", Geneset)) |>
  dplyr::filter(Category %in%  c("Cell Cycle",
                                 "Gene expression (transcription)",
                                 "Metabolism of RNAs", 
                                 "Extracellular matrix",
                                 "Cellular responses to stimuli")) |>
  mutate(Category = factor(Category, levels =  c("Cell Cycle",
                                 "Gene expression (transcription)",
                                 "Metabolism of RNAs", 
                                 "Extracellular matrix",
                                 "Cellular responses to stimuli")),
         comparison = factor(comparison, levels = c("UPF3B_KD_vs_Control",
                                                    "UPF3A_KD_vs_Control",
                                                    "DoubleKD_vs_Control",
                                                    "UPF3A_OE_vs_Control",
                                                    "UPF3A_OE_UPF3B_KD_vs_Control"))) |>
  ggplot(aes(x = comparison, y = fct_reorder(Geneset, as.integer(Category)))) +
  geom_point(
    aes(fill = fgsea.NES, size = leSize), shape = 21
  ) +
    theme_bw() +
  theme(
    legend.position = 'top',
    text = element_text(color = 'grey40')
  )  +
  scale_fill_gradient2(low="#7CCCBE",mid="#E8C1A0", high = "#F4593E") +
  #  scale_color_manual(values = "black") +
  labs(fill = "Normalized enrichment score") +
    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black", angle = 90),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top")
heatmap_1 
```

```{r}
heatmap_2 = all_heatmap  |> 
  mutate(geneRatio = leSize/fgsea.size,
         Geneset = tolower(gsub("REACTOME_", "", Geneset))) |>
  mutate(Geneset = gsub("_", " ", Geneset)) |>
  dplyr::filter(Category %in%  c( "Signal Transduction",
                                 "Developmental Biology",
                                 "Metabolism of proteins",
                                 "Metabolism of Proteins",
                                 "Chromatin organization",
                                 "DNA Replication",
                                 "DNA Repair",
                                 "Disease",
                                 "Immune",
                                 "Metabolism",
                                 "Programmed Cell Death",
                                 "Autophagy"))  |>
  mutate(Category = factor(Category, levels =  c("Signal Transduction",
                                 "Developmental Biology",
                                 "Metabolism of proteins",
                                 "Metabolism of Proteins",
                                 "Chromatin organization",
                                 "DNA Replication",
                                 "DNA Repair",
                                 "Disease",
                                 "Immune",
                                 "Metabolism",
                                 "Programmed Cell Death",
                                 "Autophagy")),
         comparison = factor(comparison, levels = c("UPF3B_KD_vs_Control",
                                                    "UPF3A_KD_vs_Control",
                                                    "DoubleKD_vs_Control",
                                                    "UPF3A_OE_vs_Control",
                                                    "UPF3A_OE_UPF3B_KD_vs_Control"))) |>
  ggplot(aes(x = comparison, y =fct_reorder(Geneset, as.integer(Category)))) +
  geom_point(
    aes(fill = fgsea.NES, size = leSize), shape = 21
  ) +
    theme_bw() +
  theme(
    legend.position = 'top',
    text = element_text(color = 'grey40')
  )  +
  scale_fill_gradient2(low="#7CCCBE",mid="#E8C1A0", high = "#F4593E") +
  #  scale_color_manual(values = "black") +
  labs(fill = "Normalized enrichment score") +
    theme(#axis.text.x = element_text(size = 12, family = "serif", color = "black", angle = 90),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top")

heatmap_2
```


```{r}
#save(all_heatmap, file= "output/fgsea-heatmap-data.Rda")
```


```{r}
upf3a_KD_apoptosis = fgsea_res$UPF3A_vs_Control %>% as.data.frame 


upf3a_KD_apoptosis_react = fgsea_res_c2$UPF3A_vs_Control %>% as.data.frame
```

```{r fig.height =6, fig.width=10}
apoptosis_pathways = upf3a_KD_apoptosis %>%
  mutate(Geneset = c("Hallmark")) %>%
  rbind(upf3a_KD_apoptosis_react %>%
          mutate(Geneset = c("Reactome"))) %>%
  dplyr::filter(fgsea.Geneset %in% c("HALLMARK_APOPTOSIS", "HALLMARK_P53_PATHWAY", 
                                      "REACTOME_APOPTOSIS", "REACTOME_PROGRAMMED_CELL_DEATH", 
                                      "REACTOME_INTRINSIC_PATHWAY_FOR_APOPTOSIS")) %>%
  dplyr::mutate(
    sig_label = dplyr::case_when(
      fgsea.padj < 0.001 ~ "***",
      fgsea.padj < 0.01  ~ "**",
      fgsea.padj < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    bar_colour = ifelse(fgsea.NES > 0, "positive", "negative"),
    fgsea.Geneset = fgsea.Geneset %>%
      gsub("^HALLMARK_", "", .) %>%
      gsub("^REACTOME_", "", .) %>%
      gsub("_", " ", .) %>%
      stringr::str_to_title()
  ) %>%
  ggplot(aes(x = fgsea.Geneset, y = fgsea.NES, fill = bar_colour)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "black") +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(
    label = sig_label,
    y = ifelse(fgsea.NES > 0, fgsea.NES + 0.05, fgsea.NES - 0.05)
  ), size = 4) +
  scale_fill_manual(values = c("positive" = "red", "negative" = "blue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12, colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"), 
        legend.position = "none") +
  facet_grid(~Geneset, scales = "free_x", space = "free_x") +
  labs(y = "Normalised enrichment score", x = "")


#ggsave(apoptosis_pathways, filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/Apoptosis-pathways.png", 
#,       height = 5, width = 9, units = "in")


library(ragg)
agg_png("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/Apoptosis-pathways.png",
         height = 5, width = 9, units = "in", res = 300)
print(apoptosis_pathways)
dev.off()

```


```{r}
ApopRes = list()

ApopRes[["UPF3B_KD_vs_Control"]] = fgsea_res$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj) 
  

ApopRes[["UPF3A_KD_vs_Control"]] = fgsea_res$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

ApopRes[["UPF3A_OE_vs_Control"]] = fgsea_res$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

ApopRes[["DoubleKD_vs_Control"]] = fgsea_res$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)

ApopRes[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) %>% 
  dplyr::arrange(fgsea.padj)
```


```{r}
AposRes_c2 = list()

AposRes_c2[["UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3B_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

AposRes_c2[["UPF3A_KD_vs_Control"]] = fgsea_res_c2$UPF3A_vs_Control %>% as.data.frame %>%
  mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

AposRes_c2[["UPF3A_OE_vs_Control"]] = fgsea_res_c2$UPF3A_OE_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

AposRes_c2[["DoubleKD_vs_Control"]] = fgsea_res_c2$DoubleKD_vs_Control %>% as.data.frame %>%
 mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  ) 

AposRes_c2[["UPF3A_OE_UPF3B_KD_vs_Control"]] = fgsea_res_c2$UPF3A_OE_UPF3B_KD_vs_Control %>% as.data.frame %>%
mutate(padj = p.adjust(fgsea.pval, "bonf")) %>%
  mutate(
    leSize = vapply(fgsea.leadingEdge, length, integer(1)),
   fgsea.leadingEdge_genes = vapply(
     fgsea.leadingEdge,
     function(x) paste(id2Name[x], collapse = "; "),
     character(1)
   ) 
  )



apop_all =ApopRes %>% 
  do.call("rbind", .) %>%
  mutate(Geneset = c("Hallmark")) %>%
  rbind(AposRes_c2 %>% 
           do.call("rbind", .) %>% 
          mutate(Geneset = c("Reactome"))) %>%

dplyr::filter(fgsea.Geneset %in% c("HALLMARK_APOPTOSIS", "HALLMARK_P53_PATHWAY", 
                                      "REACTOME_APOPTOSIS", "REACTOME_PROGRAMMED_CELL_DEATH", 
                                      "REACTOME_INTRINSIC_PATHWAY_FOR_APOPTOSIS")) %>% 
  rownames_to_column("Comparsion") %>% 
  mutate(Comparison = sub("\\.\\d+$", "", Comparsion)) %>%
    mutate(Comparison = Comparison %>%
    sub("_vs_Control", " vs Control", .) %>%
    sub("_KD", " KD", .) %>%
    gsub("_", " ", .) %>%
    tools::toTitleCase() %>%
    gsub("\\bKd\\b", "KD", .)) %>%
mutate(Comparison = factor(Comparison, levels = c("UPF3B KD vs Control", 
                                                   "UPF3A KD vs Control", 
                                                   "DoubleKD vs Control", 
                                                   "UPF3A OE vs Control", 
                                                   "UPF3A OE UPF3B KD vs Control")))%>%
    dplyr::mutate(
    sig_label = dplyr::case_when(
      fgsea.padj < 0.001 ~ "***",
      fgsea.padj < 0.01  ~ "**",
      fgsea.padj < 0.05  ~ "*",
      TRUE ~ "ns"
    ),
    bar_colour = ifelse(fgsea.NES > 0, "positive", "negative"),
    fgsea.Geneset = fgsea.Geneset %>%
      gsub("^HALLMARK_", "", .) %>%
      gsub("^REACTOME_", "", .) %>%
      gsub("_", " ", .) %>%
      stringr::str_to_title()
  ) %>%
  ggplot(aes(x = fgsea.Geneset, y = fgsea.NES, fill = bar_colour)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "black") +
  geom_bar(stat = "identity", width = 0.5) +
  geom_text(aes(
    label = sig_label,
    y = ifelse(fgsea.NES > 0, fgsea.NES + 0.05, fgsea.NES - 0.05)
  ), size = 4) +
  scale_fill_manual(values = c("positive" = "red", "negative" = "blue")) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12, colour = "black"),
        axis.text.y = element_text(size = 12, colour = "black"), 
        legend.position = "none") +
  facet_grid(Comparison ~Geneset , scales = "free_x", space = "free_x") +
  labs(y = "Normalised enrichment score", x = "")


library(ragg)
agg_png("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/Apoptosis-pathways_all.png",
         height = 12, width = 9, units = "in", res = 300)

print(apop_all)

dev.off()
```



```{r}
#ggsave(heatmap_1, filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/fgsea-plots-1.svg",# 
  #     height = 14, width = 9, units = "in")
```

```{r}
#ggsave(heatmap_2, filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/fgsea-plots-2.svg", 
 #      height = 14, width = 9, units = "in")
```
