---
title: "Enichment-analysis-goseq"
author: "unawaz1996"
date: "2023-03-16"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options: 
    chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = FALSE,
	warning = FALSE,
	message = FALSE
)
```

```{r load libraries}
source("code/libraries.R")
source("code/functions.R")
```

```{r load data}
#load("output/limma-matrices.Rda")
#load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")
```


In this workbook, we are testing for enrichment within discrete sets of DE genes as defined in the DEG analysis. 

In order to do this, we will be using `goseq`, an R package that performs enrichment analysis whilst taking length bias into account. It does so by calculating a Probability Weighting Function (PWF), which gives the probability that a gene will be differentially expressed based on length alone. 

The PWD is calculated by fitting a monotonic spline to the binary data series of differential expression (1 = DE, 0= Not DE) as a function of gene length. 
The PWF is used to weight the chance of selecting each gene when forming a null distribution for GO category membership. The fact that the PWF is calculated directly from the dataset under consideration makes this approach robust, only correcting for the length bias present in the data.

Results can be interpreted as following: 

- `overrpresented p-value`: Over representation in this analysis means that there are more DE genes in categort than expected. P-value relates to the probability of observing this number of DE genes in category by chance
- `numDEinCat`:  number of DE genes in category
- `Expected`: Expected number of genes in category
- `adjP` : Bonferroni adjustment of over represented p-value
- `FDR`: FDR adjustment of over represented p-value

# Databases used for testing 
Data was sourced using the `misigdbr` package. 

`# Enrichment in the DE Gene Set 



For the analysis of gene-sets from GO database, gene-sets were restricted to those with 3 or more steps back to the ontology root terms. 


The first step of analysis using goseq, regardless of the gene-set, is estimation of the probability weight function (PWF) which quantifies the probability of a gene being considered as DE based on a single covariate.


## GO terms 

For gene ontology analysis, we will be using the GO summaries method. Essentially, this analysis involves creating a graph for each ontology term, and removing the node `all`, as this redundant. For each GO term, we get the ontology it belongs to, the shortest path back to the root note, the longest path to the root node and whether it GO term is a terminal node. 

```{r}
load("/home/neuro/Documents/NMD_analysis/Analysis/Genome/mouse_length.Rda")
comparisons = c("UPF3B_vs_Control", "UPF3A_vs_Control", 
                "UPF3A_OE_vs_Control", "DoubleKD_vs_Control", "UPF3A_OE_UPF3B_KD_vs_Control")
```


```{r fig.caption= "UPF3B vs Controls"}
gene_ont= list()
genes = as.integer(res$UPF3B_vs_Control$EnsID %in% DEGenes$UPF3B_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = res$UPF3B_vs_Control$EnsID
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")
gene_ont$Upf3b_KD = get_ontologies(genes = genes)
```

```{r}
gene_ont$Upf3b_KD$Genes = sapply(gene_ont$Upf3b_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$Upf3b_KD$Genes = vapply(gene_ont$Upf3b_KD$Genes, paste, collapse = ", ", character(1L))

gene_ont$Upf3b_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```




```{r fig.caption= "UPF3A vs Controls"}
genes = as.integer(res$UPF3A_vs_Control$EnsID %in% DEGenes$UPF3A_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = res$UPF3A_vs_Control$EnsID 
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")
gene_ont$Upf3a_KD = get_ontologies(genes = genes)
```

```{r}
gene_ont$Upf3a_KD$Genes = sapply(gene_ont$Upf3a_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$Upf3a_KD$Genes = vapply(gene_ont$Upf3a_KD$Genes, paste, collapse = ", ", character(1L))

gene_ont$Upf3a_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```



```{r fig.caption= "Double KD vs Controls"}
genes = as.integer(res$DoubleKD_vs_Control$EnsID %in% DEGenes$DoubleKD_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = res$DoubleKD_vs_Control$EnsID 
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")
gene_ont$Double_KD = get_ontologies(genes = genes)
```


```{r}
gene_ont$Double_KD$Genes = sapply(gene_ont$Double_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$Double_KD$Genes = vapply(gene_ont$Double_KD$Genes, paste, collapse = ", ", character(1L))

gene_ont$Double_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```



```{r fig.caption= "UPF3A OE UPF3B KD  vs Controls"}
genes = as.integer(res$UPF3A_OE_UPF3B_KD_vs_Control$EnsID %in% DEGenes$UPF3A_OE_UPF3B_KD_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = res$UPF3A_OE_UPF3B_KD_vs_Control$EnsID
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")

gene_ont$UPF3A_OE_UPF3B_KD = get_ontologies(genes = genes)
```


```{r}
gene_ont$UPF3A_OE_UPF3B_KD$Genes = sapply(gene_ont$UPF3A_OE_UPF3B_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$UPF3A_OE_UPF3B_KD$Genes = vapply(gene_ont$UPF3A_OE_UPF3B_KD$Genes, paste, collapse = ", ", character(1L))

gene_ont$UPF3A_OE_UPF3B_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```



```{r}
write.xlsx(gene_ont, file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/GOseq-results.xlsx")
```


### Overlap 


```{r}
library(ggvenn)
library(VennDiagram)
x = list("Upf3b" = gene_ont$Upf3b_KD$id,
         "upf3a" = gene_ont$Upf3a_KD$id,
         "upf3 dKD" = gene_ont$Double_KD$id, 
         "upf3a OE upf3b KD" = gene_ont$UPF3A_OE_UPF3B_KD$id)

ggvenn(x[c(1,2)])


overlap = calculate.overlap(x[c(1,2)])
overlap$a3
```

<!-- # ```{r fig.height=10, fig.width=10, fig.caption= "Chord plot showing association of thedifferentially expressed genes found in the comparison of  UPF3B KDs to Controls with GO terms. The plot also shows the log fold change of the differentially expressed genes from the DE analysis where red indicates up-regulated and blue denotes downregulation "} -->
<!-- # terms = gene_ontologies$UPF3B_vs_Control %>%  -->
<!-- #     dplyr::rename(ID = id, Category = ontology, adj_pval = adjP) -->
<!-- # genes = DEGenes$UPF3B_vs_Control %>%  -->
<!-- #     dplyr::select(gene, contains("UPF3B_vs_Control")) %>%  -->
<!-- #     dplyr::rename(ID = gene, logFC = Coef.UPF3B_vs_Control) -->
<!-- # goplot_data = circle_dat(terms, genes) -->
<!-- #  -->
<!-- # processes = unique(goplot_data$term) -->
<!-- #  -->
<!-- # genes_of_interest = goplot_data %>%  -->
<!-- #   dplyr::select(genes, logFC) %>% distinct(genes, logFC, .keep_all = TRUE) #%>%  -->
<!-- #   #with(structure(logFC, names = genes)) -->
<!-- #  -->
<!-- # chord= chord_dat(goplot_data, genes_of_interest, processes) -->
<!-- # GOChord(chord, space = 0.02, gene.order = 'logFC', gene.space = 0.25, gene.size = 5, -->
<!-- #         border.size = 0.25) -->
<!-- #  -->
<!-- # ``` -->



```{r}
gene_ont$Upf3b_KD %<>% 
  mutate(comparison = c("Upf3b KD"))

gene_ont$Upf3a_KD %<>% 
  mutate(comparison = c("Upf3a KD"))

gene_ont$Double_KD %<>% 
  mutate(comparison = c("Upf3 dKD"))

gene_ont$UPF3A_OE_UPF3B_KD %<>% 
  mutate(comparison = c("Upf3a OE - Upf3b dKD"))
```

```{r fig.height=10, fig.width=10, fig.caption= "Chord plot showing association of thedifferentially expressed genes found in the comparison of  UPF3A KDs to Controls with GO terms. The plot also shows the log fold change of the differentially expressed genes from the DE analysis where red indicates up-regulated and blue denotes downregulation"}

terms = gene_ont$Upf3a_KD %>%
    dplyr::rename(ID = id, Category = ontology, adj_pval = adjP)
genes = DEGenes$UPF3A_vs_Control %>% 
    dplyr::rename(ID = gene)
goplot_data = circle_dat(terms, genes)

processes = unique(goplot_data$term)

genes_of_interest = goplot_data %>%
  dplyr::select(genes, logFC) %>% distinct(genes, logFC, .keep_all = TRUE) #%>%
  #with(structure(logFC, names = genes))

chord= chord_dat(goplot_data, genes_of_interest, processes)
GOChord(chord, space = 0.02, gene.order = 'logFC', gene.space = 0.25, gene.size = 5,
        border.size = 0.25)

```




<!-- ```{r fig.height=10, fig.width=10, fig.caption= "Chord plot showing association of the differentially expressed genes found in the comparison of  UPF3A OE, UPF3B KDs to Controls with GO terms. The plot also shows the log fold change of the differentially expressed genes from the DE analysis where red indicates up-regulated and blue denotes downregulation"} -->
<!-- terms = gene_ontologies$UPF3A_OE_UPF3B_KD_vs_Control %>% -->
<!--     dplyr::rename(ID = id, Category = ontology, adj_pval = adjP) -->
<!-- genes = DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% -->
<!--     dplyr::select(gene, contains("UPF3A_OE_UPF3B_KD_vs_Control")) %>% -->
<!--     dplyr::rename(ID = gene, logFC = Coef.UPF3A_OE_UPF3B_KD_vs_Control) -->
<!-- goplot_data = circle_dat(terms, genes) -->

<!-- processes = unique(goplot_data$term) -->
<!-- s -->
<!-- genes_of_interest = goplot_data %>% -->
<!--   dplyr::select(genes, logFC) %>% distinct(genes, logFC, .keep_all = TRUE) #%>% -->
<!--   #with(structure(logFC, names = genes)) -->

<!-- chord= chord_dat(goplot_data, genes_of_interest, processes) -->
<!-- GOChord(chord, space = 0.02, gene.order = 'logFC', gene.space = 0.25, gene.size = 5, -->
<!--         border.size = 0.25) -->

<!-- ``` -->

### Summary plot of gene ontology terms 

```{r fig.height=10, fig.width=8, fig.caption= "Summary bubbleplot"}
go_plot = do.call(rbind, gene_ont) %>%
  mutate(ontology = factor(ontology, levels = c("BP", "MF", "CC")), 
         comparison = factor(comparison, levels = c("Upf3b KD", "Upf3a KD", 
                                                    "Upf3 dKD", "Upf3a OE - Upf3b dKD"))) %>%
  mutate(term = str_to_title(term)) %>%
    ggplot(aes(x = -log10(over_represented_pvalue), y = fct_reorder(term, as.integer(ontology)), size = numDEInCat, fill = ontology)) + 
        geom_point(alpha=0.85, shape=21, color = "white") + 
   geom_segment( aes(y=term, yend=term, x=0, xend=(-log10(over_represented_pvalue)-0.15)), linewidth=0.3)   + theme_bw() + 
  theme(axis.text.y = element_text( hjust=1, colour = "black")) + 
  labs(fill = "Gene Ontology") + facet_grid(~comparison, scales ="free_y") + 
      theme(axis.text.x = element_text(size = 10, family = "serif", color = "black"),
              axis.text.y = element_text(size = 10, family = "serif"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top", 
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line = element_line(colour = "black")) + geom_vline(xintercept = 1, linetype="dotdash", color = "grey") +
  scale_fill_manual(values =c("BP" = "#EF4242",
                              "MF" = "#5E4A87", 
                              "CC" = "#50A89D")) +
  labs(xlab = bquote(-log["10"]("Overrepresented P-value")), 
       ylab= " ", 
       size = "DEGs")
go_plot    

#-log10(adjP)
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/GO-DEGs.png", plot = go_plot,  units = c("in"), height = 7.5, width = 13.5, dpi = 1200)

```



```{r}
do.call(rbind, gene_ont) %>% 
  separate_rows(Genes, sep=",") %>%
  dplyr::filter(comparison == "Upf3b KD" | comparison == "Upf3a KD") %>% 
  dplyr::filter(Genes %in% cor_genes$gene.y) %>% 
  as.data.frame()

```

### All GO terms 

```{r fig.caption= "UPF3B vs Controls"}
gene_ont_all = list()
analysis = limma_results %>% dplyr::select(ensembl_gene_id, contains("UPF3B_vs_Control")) %>% 
  set_colnames(gsub(paste0(".", "UPF3B_vs_Control"), "", colnames(.)))
genes = as.integer(limma_results$ensembl_gene_id %in% DEGenes$UPF3B_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = analysis$ensembl_gene_id
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")
gene_ont_all$Upf3b_KD = get_ontologies(genes = genes)
```

```{r}
gene_ont_all$Upf3b_KD$Genes = sapply(gene_ont_all$Upf3b_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont_all$Upf3b_KD$Genes = vapply(gene_ont_all$Upf3b_KD$Genes, paste, collapse = ", ", character(1L))

gene_ont_all$Upf3b_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```




```{r fig.caption= "UPF3A vs Controls"}
analysis = limma_results %>% dplyr::select(ensembl_gene_id, contains("UPF3A_vs_Control")) %>% 
  set_colnames( gsub(paste0(".", "UPF3A_vs_Control"), "", colnames(.)))
genes = as.integer(limma_results$ensembl_gene_id %in% DEGenes$UPF3A_vs_Control$EnsID)
not_na = !is.na(genes)
names(genes) = analysis$ensembl_gene_id
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")
gene_ont_all$Upf3a_KD = get_ontologies_all(genes = genes)
```

```{r}
gene_ont_all$Upf3a_KD$Genes = sapply(gene_ont_all$Upf3a_KD$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont_all$Upf3a_KD$Genes = vapply(gene_ont$Upf3a_KD_all$Genes, paste, collapse = ", ", character(1L))

gene_ont_all$Upf3a_KD %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in UPF3B KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


```{r}
gene_ont_all$Upf3b_KD %<>% 
  mutate(comparison = c("Upf3b KD"))

gene_ont_all$Upf3a_KD %<>% 
  mutate(comparison = c("Upf3a KD"))

gene_ont_all$Double_KD %<>% 
  mutate(comparison = c("Upf3 dKD"))

gene_ont_all$UPF3A_OE_UPF3B_KD %<>% 
  mutate(comparison = c("Upf3a OE - Upf3b dKD"))
```

```{r fig.height=10, fig.width=8, fig.caption= "Summary bubbleplot"}
do.call(rbind, gene_ont) %>%
  mutate(ontology = factor(ontology, levels = c("BP", "MF", "CC")), 
         comparison = factor(comparison, levels = c("Upf3b KD", "Upf3a KD", 
                                                    "Upf3 dKD", "Upf3a OE - Upf3b dKD"))) %>% 
    ggplot(aes(x = comparison, y = fct_reorder(term, as.integer(ontology)), size = numDEInCat)) + 
     labs(x = 'Over-represented Pvalue', y = "")+
        geom_point(alpha=0.7, shape=21, color="black") + 
        scale_fill_manual(values = c("#EA2A5F", "#FCC98A", "#2F124B"))  + theme_bw() + theme(axis.text.y = element_text( hjust=1, face = "bold")) +       labs(fill = "Gene Ontology") #+ facet_grid(rows = vars(comparison), scales ="free_y")
    


```
#### Considering upregulated genes only 


```{r}
gene_ontologie_up = list()
for (c in comparisons){
    analysis = limma_results %>% dplyr::select(ensembl_gene_id, contains(c)) 
    colnames(analysis) = gsub(paste0(".", c), "", colnames(analysis))
    genes = as.integer(analysis$Res == 1)
    not_na = !is.na(genes)
    names(genes) = analysis$ensembl_gene_id
    genes = genes[not_na]
    gene_ont = get_ontologies(genes = genes)
    gene_ont %<>% as.data.frame() %>% 
        mutate(Comparison = c(c))
    gene_ontologie_up[[paste0(c)]] = gene_ont
    }

```

```{r fig.height=10, fig.width=8, fig.caption= "Summary bubbleplot"}
do.call(rbind, gene_ontologie_up[-3]) %>%
    ggplot(aes(x = over_represented_pvalue, y = term, fill = ontology, size = numDEInCat)) + 
     labs(x = 'Over-represented Pvalue', y = "")+
        geom_point(alpha=0.7, shape=21, color="black") + 
        scale_fill_manual(values = c("#EA2A5F", "#FCC98A", "#2F124B"))  + theme_bw() + theme(axis.text.y = element_text( hjust=1, face = "bold")) +       labs(fill = "Gene Ontology") + facet_grid(rows = vars(Comparison), scales ="free_y")
    


```


# Shared transcripts between the KDs 

```{r}
analysis = limma_results
genes = as.integer(analysis$Res.UPF3B_vs_Control != 0 & analysis$Res.UPF3A_vs_Control !=0)
not_na = !is.na(genes)
names(genes) = analysis$ensembl_gene_id
genes = genes[not_na]
gene_ont = get_ontologies(genes = genes)
```

```{r}
gene_ont$Genes = sapply(gene_ont$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$Genes = vapply(gene_ont$Genes, paste, collapse = ", ", character(1L))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in the 144 shared genes between UFP3A KD and UPF3B KD",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, 
                      shortest_path, longest_path, terminal_node,
                      numDEInCat, numInCat, Expected, adjP)) %>%
  dplyr::select(id, term,ontology, everything()) %>%
  dplyr::rename("GO ID" = id, "Term" = term,
                "Over-represented p-value" = "over_represented_pvalue", 
                "FDR adj p-value" = "FDR", "Genes in category" = "Genes") %>% 
  stargazer(title="Gene ontoloty terms for 144 overlapping genes",  summary=FALSE, rownames = FALSE, out = "output/DEG/Thesis_tables/GO/overlapping_DEGs.tex")
```


```{r fig.height=10, fig.width=10, fig.caption= "Chord plot showing association of thedifferentially expressed genes found in the comparison of  UPF3B KDs to Controls with GO terms. The plot also shows the log fold change of the differentially expressed genes from the DE analysis where red indicates up-regulated and blue denotes downregulation "}
terms = gene_ont %>% 
    dplyr::rename(ID = id, Category = ontology, adj_pval = adjP)
genes = DEGenes$UPF3A_vs_Control %>% 
    dplyr::select(gene, contains("UPF3A_vs_Control")) %>% 
    dplyr::rename(ID = gene, logFC = Coef.UPF3A_vs_Control)
goplot_data = circle_dat(terms, genes)

processes = unique(goplot_data$term)

genes_of_interest = goplot_data %>% 
  dplyr::select(genes, logFC) %>% distinct(genes, logFC, .keep_all = TRUE) #%>% 
  #with(structure(logFC, names = genes))

chord= chord_dat(goplot_data, genes_of_interest, processes)
chord_plot = GOChord(chord, space = .06, gene.order = 'logFC', gene.space = 0.25, gene.size = 4,
        border.size = 0.25)
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/corr_144_GO_upf3a.svg", plot = chord_plot, 
       units = c("in"), width =5.78, height = 6.45)
```

```{r}
GOHeat(chord, nlfc = 1)

```




### UPF3B specific genes 

```{r}
analysis = limma_results
genes = as.integer(analysis$Res.UPF3B_vs_Control != 0 & analysis$Res.UPF3A_vs_Control ==0)
not_na = !is.na(genes)
names(genes) = analysis$ensembl_gene_id
genes = genes[not_na]
gene_ont = get_ontologies(genes = genes)
```

```{r}
gene_ont$Genes = sapply(gene_ont$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })

gene_ont$Genes = vapply(gene_ont$Genes, paste, collapse = ", ", character(1L))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in the 144 shared genes between UFP3A KD and UPF3B KD",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, 
                      shortest_path, longest_path, terminal_node,
                      numDEInCat, numInCat, Expected, adjP)) %>%
  dplyr::select(id, term,ontology, everything()) %>%
  dplyr::rename("GO ID" = id, "Term" = term,
                "Over-represented p-value" = "over_represented_pvalue", 
                "FDR adj p-value" = "FDR", "Genes in category" = "Genes") %>% 
  stargazer(title="Gene ontoloty terms for Upf3b",  summary=FALSE, rownames = FALSE, out = "output/DEG/Thesis_tables/GO/Upf3b_DEGs.tex")
```


```{r fig.height=10, fig.width=10, fig.caption= "Chord plot showing association of thedifferentially expressed genes found in the comparison of  UPF3B KDs to Controls with GO terms. The plot also shows the log fold change of the differentially expressed genes from the DE analysis where red indicates up-regulated and blue denotes downregulation "}
terms = gene_ont %>% 
    dplyr::rename(ID = id, Category = ontology, adj_pval = adjP)
genes = DEGenes$UPF3B_vs_Control %>% 
    dplyr::select(gene, contains("UPF3A_vs_Control")) %>% 
    dplyr::rename(ID = gene, logFC = Coef.UPF3A_vs_Control)
goplot_data = circle_dat(terms, genes)

processes = unique(goplot_data$term)

genes_of_interest = goplot_data %>% 
  dplyr::select(genes, logFC) %>% distinct(genes, logFC, .keep_all = TRUE) #%>% 
  #with(structure(logFC, names = genes))

chord= chord_dat(goplot_data, genes_of_interest, processes)
chord_plot = GOChord(chord, space = .06, gene.order = 'logFC', gene.space = 0.25, gene.size = 4,
        border.size = 0.25)
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/DEG/Thesis_figures/corr_144_GO_upf3a.svg", plot = chord_plot, 
       units = c("in"), width =5.78, height = 6.45)
```

```{r}
GOHeat(chord, nlfc = 1)

chord
```


### UPF3A specific terms 

```{r}
analysis = limma_results
genes = as.integer(analysis$Res.UPF3B_vs_Control == 0 & analysis$Res.UPF3A_vs_Control !=0)
not_na = !is.na(genes)
names(genes) = analysis$ensembl_gene_id
genes = genes[not_na]
gene_ont = get_ontologies(genes = genes)
```

```{r}
gene_ont$Genes = sapply(gene_ont$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })


gene_ont$Genes = vapply(gene_ont$Genes, paste, collapse = ", ", character(1L))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in the 144 shared genes between UFP3A KD and UPF3B KD",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

gene_ont %>% 
    dplyr::select(-c( contains("Sig"), EnsemblID, 
                      shortest_path, longest_path, terminal_node,
                      numDEInCat, numInCat, Expected, adjP)) %>%
  dplyr::select(id, term,ontology, everything()) %>%
  dplyr::rename("GO ID" = id, "Term" = term,
                "Over-represented p-value" = "over_represented_pvalue", 
                "FDR adj p-value" = "FDR", "Genes in category" = "Genes") %>% 
  stargazer(title="Gene ontoloty terms for Upf3b",  summary=FALSE, rownames = FALSE, out = "output/DEG/Thesis_tables/GO/Upf3a_DEGs.tex")
```

