---
title: "Publication-Supplementary-Figures"
author: "urwahnawaz"
date: "2026-03-05"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: console
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	message = FALSE
)
```


# Introduction

```{r load}
source("code/plots.R") 
source("code/functions.R")
source("code/libraries.R")
library(biomaRt)
library(nVennR)
library(eulerr)
library(ggsci)
library(janitor)
library(ggside)
library(ggstatsplot)
library(stargazer)
library(coin)
library(GeneOverlap)
```

```{r}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])
```


```{r}
ma_plot = function(data){
  DEColours <- c("Downregulated" = "#14213d","Upregulated" = "#ff6700",  "Not Sig" = "#e6e8e6")
  temp = data %>% 
    mutate(diffExpressed = ifelse(logFC > 0 & FDR < 0.05, "Upregulated",        
               ifelse(logFC < 0 & FDR < 0.05, "Downregulated", "Not Sig")))
  plot = temp %>% 
    ggplot(aes(y= logFC,
               x = logCPM,
               colour = as.factor(as.character(diffExpressed )))) +
    geom_point(alpha = 0.65, size = 3) +
    scale_colour_manual(values = DEColours) + theme_bw()+
    theme(legend.position = "top",
          axis.title.y = element_text(size = 12)) + geom_hline(yintercept = 0, 
                                                               linetype = "dashed", color = "grey") +
    # geom_ysidedensity(aes(x=after_stat(density),  
    #                          color = nif_label, fill = nif_label),  alpha=0.2, size = 1) +
    # scale_fill_manual(values = DEColours) + scale_size_manual(values = c("NIF+" = 3, 
    #                                                                      "NIF-" =2,
    #                                                                      "Not Sig"= 2)) +
    theme(
        axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.text.y = element_text(size = 12, family = "serif", color = "black"), 
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
         strip.text.x = element_text(
        size = 12, face = "bold.italic", family = "serif"
        )) + labs(color = "Expression", 
                  fill = "",
                  size = "") + scale_alpha(guide = 'none')
  return(plot)
    
}
```

# Introduction 

```{r}
load("output/DEG-list.Rda")
load("output/Stability/differential-stability.Rda")
load(file=here::here("output/DET-list.Rda"))
```

## Supplementary Figure 1

### Supplementary Figure 1c

### Supplementary Figure 1d

## Supplementary Figure 2

### Supplementary Figure 2a

```{r}
res_sig_dte = lapply(res_dte, function(x){
  x %<>% 
    dplyr::filter(FDR < 0.05 ) %>% 
    as.data.frame()
})

res_sig_dte$UPF3B_KD_Control %>% 
  with(table(sign(.$logFC)))

res_sig_dte$UPF3A_KD_Control %>% 
  with(table(sign(.$logFC)))

res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  with(table(sign(.$logFC)))

res_sig_dte$UPF3A_OE_Control %>% 
  with(table(sign(.$logFC)))
```


```{r fig.width=10, fig.height=10, fig.cap="MA plot of comprisons of KD/OE with Controls. Pink highlights significantly upregulated transcripts whereas blue highlights significantly downregulated transcripts"}
## Ma plot 

ma_1 = ma_plot(res_dte$UPF3B_KD_Control) + ggtitle("UPF3B KD vs Control")
ma_2 = ma_plot(res_dte$UPF3A_KD_Control) + ggtitle("UPF3A KD vs Control")
ma_3 = ma_plot(res_dte$UPF3A_KD_UPF3B_KD_Control) + ggtitle("UPF3 dKD vs Control")
ma_4 = ma_plot(res_dte$UPF3A_OE_Control) + ggtitle("UPF3A OE vs Control")
ma_5 = ma_plot(res_dte$UPF3A_OE_UPF3B_KD_Control) + ggtitle("UPF3A OE in UPF3B KD")



ggsave(plot=ma_1, filename = here::here("output/Thesis_plots/DETs/Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_2, filename = here::here("output/Thesis_plots/DETs/Upf3a_KD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_3, filename = here::here("output/Thesis_plots/DETs/Upf3_dKD.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_4, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE.png"), height =4, width =4, 
       dpi = 1200)
ggsave(plot=ma_5, filename = here::here("output/Thesis_plots/DETs/Upf3a_OE_Upf3b_KD.png"), height =4, width =4, 
       dpi = 1200)
```



### Supplementary Figure 2b

## Supplementary Figure 3

### Supplementary Figure 3a

```{r}
res_sig_dte = lapply(res_dte, function(x){
  x %<>% 
    dplyr::filter(FDR < 0.05 ) %>% 
    as.data.frame()
})

upf3_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
             cor.coeff.args = list(size = 5, family = "serif",  label.y = 6), 
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab(expression(italic(Upf3b) ~ "KD DETs (LFC)")) +
ylab(expression(italic(Upf3a) ~ "KD DETs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

upf3_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3_paralogs_cor.png", plot = upf3_cor , 
       units = c("in"), height = 4, width = 4, dpi = 1200)
```

```{r}

upf3d_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
            cor.coeff.args = list(size = 5, family = "serif", label.y = 7),
              size=5, alpha =0.75, color = "Label",
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab(expression(italic(Upf3b) ~ "KD DETs (LFC)")) + 
 ylab(expression(italic(Upf3) ~ "dKD DETs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
upf3d_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3dKD_Upf3b_cor.png", plot = upf3d_cor , 
       units = c("in"), height = 4, width = 4, dpi = 1200)

```

```{r}

upf3b_Upf3aOE_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif", label.y = -0.05),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab(expression(italic(Upf3b) ~ "KD DETs (LFC)")) + ylab(expression(italic(Upf3a) ~ "OE DETs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
upf3b_Upf3aOE_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aOE_Upf3b_cor.png", plot = upf3b_Upf3aOE_cor, 
       units = c("in"), height = 4, width = 4, dpi = 1200)

```

```{r}

upf3b_Upf3aOEKD_cor = res_sig_dte$UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
             cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab(expression(italic(Upf3b) ~ "KD DETs (LFC)")) + ylab(expression(atop(italic(Upf3a) ~ "OE",
                     italic(Upf3b) ~ "KD DETs (LFC)"))) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  

upf3b_Upf3aOEKD_cor


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aOEKD_Upf3b_cor.png", plot = upf3b_Upf3aOEKD_cor, 
       units = c("in"), height = 4, width = 4.45, dpi = 1200)

```


```{r}

upf3a_Upf3dKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
             cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() + xlab(expression(italic(Upf3a) ~ "KD DETs (LFC)"))+ 
   ylab(expression(italic(Upf3) ~ "dKD DETs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
upf3a_Upf3dKD_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aKD_Upf3dKD_cor.png", plot = upf3a_Upf3dKD_cor, 
       units = c("in"), height = 4, width = 4, dpi = 1200)

```

```{r}

upf3a_Upf3aOE_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() +
  xlab(expression(italic(Upf3a) ~ "KD DETs (LFC)"))+ 
   ylab(expression(italic(Upf3a) ~ "OE DETs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aKD_Upf3aOE_cor.png", plot = upf3a_Upf3aOE_cor, 
       units = c("in"), height = 4, width = 4, dpi = 1200)

```



```{r}

Upf3aKD_Upf3aOEKD_cor = res_sig_dte$UPF3A_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif", label.y = 7),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
  xlab(expression(italic(Upf3a) ~ "KD DETs (LFC)")) + ylab(expression(atop(italic(Upf3a) ~ "OE",
                     italic(Upf3b) ~ "KD DETs (LFC)"))) +
    theme_bw() + 
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aKD_Upf3aOEKD_cor.png", plot = Upf3aKD_Upf3aOEKD_cor, 
       units = c("in"), height = 4, width = 4.45, dpi = 1200)

```


```{r}
upf3aOE_Upf3dKD_cor = res_sig_dte$UPF3A_OE_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join(res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    xlab(expression(italic(Upf3a) ~ "OE DETs (LFC)")) +   
  ylab(expression(italic(Upf3) ~ "dKD DETs (LFC)")) +
    theme_bw() +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output//Figures/Supplementary/Figure-3/DETs/Upf3aOE_Upf3dKD_cor.png", plot = upf3aOE_Upf3dKD_cor, 
       units = c("in"), height = 4.2, width = 4.2, dpi = 1200)

```


```{r}
upf3aOEKD_Upf3dKD_cor = res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC) %>% 
  inner_join( res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
  dplyr::select(tx_id, logFC), by = "tx_id") %>% 
  mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif", label.y = 7),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    xlab(expression(italic(Upf3) ~ "dKD DETs (LFC)")) + ylab(expression(atop(italic(Upf3a) ~ "OE",
                     italic(Upf3b) ~ "KD DETs (LFC)"))) +
    theme_bw() + 
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#3c1642", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
  
upf3aOEKD_Upf3dKD_cor 

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DETs/Upf3aOEKD_Upf3dKD_cor.png", plot = upf3aOEKD_Upf3dKD_cor, 
       units = c("in"), height = 4, width = 4.45, dpi = 1200)

```

```{r}
ggarrange(upf3_cor,upf3d_cor, upf3b_Upf3aOE_cor,
  upf3b_Upf3aOEKD_cor,upf3a_Upf3dKD_cor, upf3a_Upf3aOE_cor,
  Upf3aKD_Upf3aOEKD_cor, upf3aOE_Upf3dKD_cor, upf3aOEKD_Upf3dKD_cor, ncol =3, nrow=3 )
```


### Supplementary Figure 3b


```{r}

upf3b_cor = res_sig_dte$UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>%
  left_join(tx2gene, by = "tx_id") %>% 
    inner_join(DEGenes$UPF3B_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>%
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
             cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
   xlab(expression(italic(Upf3b) ~ "KD DETs (LFC)"))+ 
   ylab(expression(italic(Upf3b) ~ "KD DEGs (LFC)")) +
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  theme_bw() +
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
upf3b_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DET-DEGs/Upf3b_cor.png", plot = upf3b_cor , 
       units = c("in"), height = 4, width = 4.1, dpi = 1200)
```

```{r}
upf3a_cor = res_sig_dte$UPF3A_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
                cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
     xlab(expression(italic(Upf3a) ~ "KD DETs (LFC)"))+ 
   ylab(expression(italic(Upf3a) ~ "KD DEGs (LFC)")) + 
    theme_bw() + 
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) + 
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

upf3a_cor

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DET-DEGs/Upf3a_cor.png", plot = upf3a_cor  , 
       units = c("in"), height = 4, width = 4.1, dpi = 1200)
```

```{r}
upf3d_KD_cor = res_sig_dte$UPF3A_KD_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$DoubleKD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
    ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
              cor.coeff.args = list(size = 5, family = "serif"),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
    theme_bw() +   xlab(expression(italic(Upf3) ~ "dKD DETs (LFC)"))+ 
   ylab(expression(italic(Upf3) ~ "dKD DEGs (LFC)")) + 
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DET-DEGs/Upf3dKD_cor.png", plot = upf3d_KD_cor , 
       units = c("in"), height = 4, width = 4.1, dpi = 1200)
```

```{r}
upf3a_OE_Upf3b_KD_DEGs_DSGs = res_sig_dte$UPF3A_OE_UPF3B_KD_Control %>% 
   mutate_at(.vars = "tx_id", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  left_join(tx2gene, by = "tx_id") %>%
    inner_join(DEGenes$UPF3A_OE_UPF3B_KD_vs_Control %>% 
                 dplyr::rename(gene_id = EnsID), by = "gene_id") %>% 
          mutate(Label = ifelse(logFC.x > 0 & logFC.y >0 | 
                            logFC.x < 0 & logFC.y <0, "Concordance", "Discordant")) %>% 
  ggscatter(., x= "logFC.x", y="logFC.y", 
              cor.coef = TRUE, add = "reg.line", 
              size=5, alpha =0.75, color = "Label",
            cor.coeff.args = list(size = 5, family = "serif", label.y = 5),
              conf.int = TRUE, add.params = list(color = "blue",
                                                 fill = "lightgray")) + 
  xlab(expression(atop(italic(Upf3a) ~ "OE",
                     italic(Upf3b) ~ "KD DETs (LFC)"))) + 
  ylab(expression(atop(italic(Upf3a) ~ "OE",
                     italic(Upf3b) ~ "KD DEGs (LFC)"))) +
    theme_bw() + 
  geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  scale_color_manual(values = c("#c9184a", "#BAB8B5")) +  
  theme(axis.text.x = element_text(size = 18, family = "serif", color = "black"),
            axis.text.y = element_text(size = 18, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Figure-3/DET-DEGs/Upf3a_Upf3b_kD_cor.png", plot = upf3a_OE_Upf3b_KD_DEGs_DSGs , 
       units = c("in"), height = 4.45, width = 4.45, dpi = 1200)
```


## Supplementary Figure 5 

### Supplementary Figure 5a

```{r}
upf3a_res = z[[1]]
upf3a_res %>%
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3a_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
  drop_na("results") %>%
    ggplot(aes(x = log2FoldChange,
               y = -log10(padj),
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
        xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value"))) +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) +
  ylim(0,7.5) +theme(axis.text.x = element_text(size = 25, family = "serif", color = "black"),
              axis.text.y = element_text(size = 25, family = "serif", color = "black"),
        axis.title = element_text(size=25, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
         legend.text = element_text(family = "serif", size = 15), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "") +   xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value")))
volc
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-stability-Supp-5a.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)
```

### Supplementary Figure 5b

```{r}
upf3a_res =  z[[1]] %>% as.data.frame() %>%
  dplyr::filter(padj < 0.05) %>%
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsemblID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsemblID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "Upf3a KD DEG" = DEGenes$UPF3A_vs_Control$EnsID,
    "Upf3a KD DSGs" = upf3a_res$EnsemblID)


ggvenn(x[c(1,2)], fill_color = c("#ef476f", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 10,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)

 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-stability_overlap_fig3b.png",  units = c("in"), height = 6, width = 6, dpi =1200)

```

```{r}
x = calculate.overlap(x)
```

```{r}
go.obj <- newGeneOverlap(DEGenes$UPF3A_vs_Control$EnsID,
                         upf3a_res$EnsemblID,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```

### Supplementary Figure 5c 


```{r}
overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  inner_join(upf3a_res %>% 
               dplyr::rename(EnsID = EnsemblID), by="EnsID") %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature) ,"NIF-", "NIF+"))

overlap_DSA =overlapping_upf3a %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
             cor.coeff.args = list(size = 6, family = "serif"),
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  ylab("Upf3a DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 14, family = "serif", color = "black"),
            axis.text.y = element_text(size = 14, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif", size = 12),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .45) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) +
labs(  y = expression(italic(Upf3a) ~ "KD DSG (" * log[2] * "FC)"),
  x = expression(italic(Upf3a) ~ "KD DEG (" * log[2] * "FC)"))

overlap_DSA

 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-stability-overlap_5c.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 8, dpi =1200)

```

### Supplementary Figure 5d

```{r}

non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>%
    dplyr::rename("gene_id" = "EnsID") %>%
  left_join(res_dte$UPF3A_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
  mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))

nif_overlap = overlapping_upf3a %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  dplyr::select(nif) %>%
   mutate(group = "DEGs/DSGs\nOverlapping") %>%
  rbind(non_overlapping_upf3a %>%
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping\nDEGs")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = ""#,
 # ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + theme_bw() +
    labs(y="Prpoportion of transcripts") +
  theme(axis.text = element_text(color ="black", size =15),
        axis.title = element_text(color ="black", size =15),
         legend.box.background = element_rect(color = "black"))

nif_overlap


 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-bf-NIF-proportion_fig5d.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```

### Supplementary Figure 5e

```{r}
upf3dKD_res = z[[3]]
upf3dKD_res %>%
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3dKD_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
  drop_na(results) %>%
    ggplot(aes(x = log2FoldChange,
               y = -log10(padj),
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
        xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value"))) +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) +
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 25, family = "serif", color = "black"),
              axis.text.y = element_text(size = 25, family = "serif", color = "black"),
        axis.title = element_text(size=25, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "")
volc

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3dKD_stability_5e.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)


```

### Supplementary Figure 5f

* Correlation between DSE and DEGs

```{r}
upf3dKD_res =  z[[3]] %>% as.data.frame() %>%
  dplyr::filter(padj < 0.05) %>%
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsemblID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsemblID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "Upf3 dKD DEG" = DEGenes$DoubleKD_vs_Control$EnsID,
    "Upf3 dKD DSG" = upf3dKD_res$EnsemblID)


ggvenn(x, fill_color = c("#7175b3", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 10,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)


 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3dKD-stability_overlap_fig3b.png",  units = c("in"), height = 6, width = 6, dpi =1200)
```

```{r}
x = calculate.overlap(x)
```

```{r}
go.obj <- newGeneOverlap(DEGenes$DoubleKD_vs_Control$EnsID,
                         upf3dKD_res$EnsemblID,
                         genome.size=length(res$DoubleKD_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```

### Supplementary Figure 5g


```{r}
overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>%
  inner_join(upf3dKD_res %>% dplyr::rename(EnsID =EnsemblID ), by="EnsID") %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" |is.na(nif_feature),"NIF-", "NIF+"))

overlap_DSA = overlapping_upf3dKD %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
             cor.coeff.args = list(size = 7, family = "serif"), 
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif", size =12),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .45) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) +
  labs(  y = expression(italic(Upf3) ~ "dKD DSG (" * log[2] * "FC)"),
  x = expression(italic(Upf3) ~ "dKD DEG (" * log[2] * "FC)"))


overlap_DSA

 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3dKD-stability-overlap.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)



```

```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
  mutate(nif= ifelse(nif_feature == "None","NIF-", "NIF+"))
```


```{r}
nif_overlap =overlapping_upf3dKD %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  dplyr::select(nif) %>%
   mutate(group = "DEGs/DSGs\nOverlapping") %>%
  rbind(non_overlapping_upf3dKD  %>%
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping\nDEGs")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = ""
#  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + theme_bw() +
    labs(y="Prpoportion of transcripts") +
  theme(axis.text = element_text(color ="black", size =15),
        axis.title = element_text(color ="black", size =15),
         legend.box.background = element_rect(color = "black"))

nif_overlap


 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3d-bf-NIF-proportion_fig5d.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)




# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```


