---
title: "RNA-stability"
author: "unawaz1996"
date: "2023-05-05"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  
  chunk_output_type: console
---

# Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	message = FALSE
)
```

```{r libraries}

library(eisaR)
source("code/libraries.R")
source("/home/neuro/Documents/NMD_analysis/other_repos/DiffRAC/DiffRAC.R")
source("code/functions.R")
library(coin)
library(gplots)
library(stargazer)
library(janitor)
library(ggside)
library(ggstatsplot)
library(GeneOverlap)
```

```{r data-load}
load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")
load("output/DET-list.Rda")
```


```{r mouse-length}
load("/home/neuro/Documents/NMD_analysis/Analysis/Genome/mouse_length.Rda")
```

```{r functions}

get_ontologies_stb = function(genes,tree,pathway){

    ## Overlap the DE genes with the genes in the length data vector
    select_genes <- as.vector(names(lengthData)%in%names(genes))
    new_lengthData <- lengthData[select_genes]
    select_genes <- as.vector(names(genes) %in% names(new_lengthData))
    genes = genes[select_genes]

    ## Calculate a Probability Weighting Function
    ## This files a monotonic spline
    pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)


    if (missing(pathway)){
        message("Using GO from GOseq")

        Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE)
        goterms <- Go.wall$category
        goList <- getGeneLists(pwf, goterms, "mm39", "ensGene")
        Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]])
        Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L))
        goRes <- Go.wall %>%
            dplyr::select(-under_represented_pvalue) %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            left_join(goSummaries)
        nGenes <- c(BG = sum(genes == 0), DE = sum(genes))
        goFinal <- goRes %>%
            dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > tree) %>%
            arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>%
            arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(Sig_adjP, numDEInCat > Expected)


    } else {
        Go.wall=goseq(pwf, gene2cat = pathway, use_genes_without_cat = TRUE)
        goFinal = Go.wall %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(Sig_adjP) %>%
            dplyr::arrange(Sig_adjP)

    }

    return(goFinal)

}
get_ontologies = function(genes, pathway){

    ## Overlap the DE genes with the genes in the length data vector
    select_genes <- as.vector(names(lengthData)%in%names(genes))
    new_lengthData <- lengthData[select_genes]
    select_genes <- as.vector(names(genes) %in% names(new_lengthData))
    genes = genes[select_genes]

    ## Calculate a Probability Weighting Function
    ## This files a monotonic spline
    pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)


    if (missing(pathway)){
        message("Using GO from GOseq")

        Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE)
        goterms <- Go.wall$category
        goList <- getGeneLists(pwf, goterms, "mm39", "ensGene")
        Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]])
        Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L))
        goRes <- Go.wall %>%
            dplyr::select(-under_represented_pvalue) %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            left_join(goSummaries)
        nGenes <- c(BG = sum(genes == 0), DE = sum(genes))
        goFinal <- goRes %>%
            dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > 3) %>%
            arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>%
            arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = over_represented_pvalue < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(Sig_adjP, numDEInCat > Expected)


    } else {
        Go.wall=goseq(pwf, gene2cat = pathway, use_genes_without_cat = TRUE)
        goFinal = Go.wall %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = over_represented_pvalue < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(Sig_adjP) %>%
            dplyr::arrange(Sig_adjP)

    }

    return(goFinal)

}

plot_fc = function(df){
  df %>% 
    dplyr::filter(padk < 0.05) %>%  
    ggplot(aes(x=log2FoldChange)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white") +  
    geom_density(alpha=.2, fill="#FF6666") + geom_vline(xintercept =c(log2(1.5), -log2(1.5)),color="red", linetype = "dashed")
}
```

```{r}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])
```

```{r}
NMD_features = read.table("data/nif_output/Mus_musculus__nifs.tsv", sep = "\t", skip =1)

colnames(NMD_features)[1:4] = c("EnsID", "NIF_dEJ", "NIF_uORF", "NIF_long_3utr")
NMD_features = NMD_features[,1:4]
NMD_features$NIF_dEJ = ifelse(NMD_features$NIF_dEJ == "True",gsub("True", 1, NMD_features$NIF_dEJ), 0)
NMD_features$NIF_uORF = ifelse(NMD_features$NIF_uORF == "True",gsub("True", 1, NMD_features$NIF_uORF), 0)
NMD_features$NIF_long_3utr = ifelse(NMD_features$NIF_long_3utr == "True",gsub("True", 1, NMD_features$NIF_long_3utr), 0)
NMD_features %<>% mutate_at(.vars = "EnsID", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")

NMD_features$NIF_dEJ[is.na(NMD_features$NIF_dEJ)] <- 0
NMD_features$NIF_uORF[is.na(NMD_features$NIF_uORF)] <- 0
NMD_features$NIF_long_3utr[is.na(NMD_features$NIF_long_3utr)] <- 0
```

# RNA-stability analysis 

It has been suggested that, while exonic read counts in RNA-seq data correspond to steady-state mRNA abundance, changes in the abundance of intronic reads can be used to estimate the change in transcription rate. Through this concept, a change in exonic reads without a corresponding change in intronic reads is diagnostic of differential RNA stability, while concurrent changes in both exonic and intronic reads suggest altered transcription. DiffRAC is a framework that converts unspliced/spliced relationships into a generalized linear model whose parameters can then be inferred from sequencing count data.  


## Differential stability analysis 

```{r}
md = read.table("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/data/LTK_Sample Metafile_V3.txt", header = TRUE, sep="\t")
md$Group = factor(md$Group, levels = c("Control", "UPF3A_KD", "UPF3A_OE", 
                                       "UPF3B_KD", "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
sampleTable= read.csv("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/data/data2.txt", header=TRUE, sep = ",")
inputFolder="/home/neuro/Documents/NMD_analysis/Analysis/Results/EISA"

de_exonic  = DESeqDataSetFromHTSeqCount( # omits special rows from htseq-count
		    sampleTable = sampleTable[sampleTable$ReadType=="exonic",],
		    directory = inputFolder,
		    design = ~ 1 # required parameter
		)

countsEx= counts(de_exonic)

de_intron = DESeqDataSetFromHTSeqCount( # omits special rows from htseq-count
		    sampleTable = sampleTable[sampleTable$ReadType=="intronic",],
		    directory = inputFolder,
		    design = ~ 1 # required parameter
		)
countsIn = counts(de_intron)
```

```{r}
md$Samples = colnames(countsEx)
md %<>% dplyr::select(Samples, Group)

rownames(md) = md$Samples

md$Group = factor(md$Group, levels = c("Control", "UPF3A_KD", "UPF3A_OE", 
                                       "UPF3B_KD", "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
```

```{r}
countsEx = countsEx[rownames(countsEx) %in% rownames(countsIn),]
countsIn = countsIn[rownames(countsIn) %in% rownames(countsEx),]
```

```{r include=FALSE}
diffrac_res <- DiffRAC( ~Group,
         md,
         countsEx,countsIn,
         "sample",
         optimizeBias = T)
```


### Results 

```{r}
resultsTable = data.frame()
names = c("GroupUPF3A_KD.Ratio","GroupUPF3B_KD.Ratio",
          "GroupUPF3A_KD_UPF3B_KD.Ratio","GroupUPF3A_OE.Ratio",
          "GroupUPF3A_OE_UPF3B_KD.Ratio")

for (analysis in names){
  temp =  as.data.frame(results(diffrac_res$dds,name = analysis))
  stability_total = temp %>%
    dplyr::filter(padj < 0.05) %>%
  with(., 
       table(sign.lfc=sign(log2FoldChange))) %>%
    as.data.frame() %>%
    t() %>%
    as.data.frame() %>% 
    row_to_names(row_number = 1) %>%
    mutate(Group = analysis)
  resultsTable = rbind(resultsTable, stability_total)
}

resultsTable %<>% 
  dplyr::rename("Destabilized" = "-1", "Stabilized" = "1") %>% 
  rownames_to_column("random") %>% 
  dplyr::select(Group, everything(), -c(random)) %>% 
  mutate(Group = ifelse(Group == 'GroupUPF3A_KD.Ratio',"Upf3a KD targets",
                          ifelse(Group == 'GroupUPF3B_KD.Ratio',"Upf3b KD targets", 
                                 ifelse(Group == 'GroupUPF3A_KD_UPF3B_KD.Ratio',"Upf3 dKD targets",
                                        ifelse(Group == 'GroupUPF3A_OE.Ratio',"Upf3a OE targets", "Upf3a OE Upf3b KD targets" )))))



resultsTable %>%
  DT::datatable(caption = "Differential stability estimates")
```


```{r}
z <- lapply(names, function(x) {
       tmp <- results(diffrac_res$dds, name = x) %>% 
         as.data.frame() 
       tmp %<>% mutate(group = as.character(x))
       tmp %<>% rownames_to_column("EnsemblID")
       #names(tmp) <- paste(x, names(tmp), sep = ".")
       return(tmp)})

alltables <- do.call(rbind, z)
```


```{r fig.height=10, fig.width=10, fig.cap="Distribution of RNA-stability estimates across comparisons"}
DEColours <- c("-1" = "#4657A6","1" = "#C32137",  "0" = "#D3D5E3")

stability_dist = alltables %>% 
  mutate(results = sign(log2FoldChange)) %>%
  dplyr::filter(padj < 0.05) %>%
  mutate(group=factor(group, levels = c("GroupUPF3B_KD.Ratio",
                                        "GroupUPF3A_KD.Ratio", 
                                        "GroupUPF3A_KD_UPF3B_KD.Ratio", 
                                        "GroupUPF3A_OE.Ratio",
                                        "GroupUPF3A_OE_UPF3B_KD.Ratio"))) %>%
  ggplot(aes(x=log2FoldChange, fill=as.factor(results))) +
    geom_histogram(alpha = 0.65) + theme_bw() +
  scale_fill_manual(values=DEColours ) +
  theme(legend.position="none") + xlim(-2, 2) +
  facet_wrap(~group) +
theme(panel.spacing = unit(1, "lines")) +
theme(strip.text = element_text(color = "white", face = "bold"))


stability_dist
# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Stability/Thesis_figures/stability_dist.svg", 
#       width = 8.99, height=7.5, units = "in", plot = stability_dist)
```


## Upf3b KD stability estimates and overlap with differential expression analyses 

```{r fig.width=10, fig.height=15, fig.cap="Upf3b KD DSA volcano plot"}
upf3b_res = z[[2]]

DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3b_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
    ggplot(aes(x = log2FoldChange, 
               y = -log10(padj), 
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("Upf3b KD differentially stabilised genes") + 
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) + 
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
              axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "") 

volc
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-stability.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)


```


* Correlation between DSE and DEGs

```{r fig.width=8, fig.height=8, fig.cap="Overlap of DEG and DSA analysis from differential stability estimates"}
upf3b_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3B_KD.Ratio")) %>% 
  dplyr::filter(padj < 0.05) %>% 
  rownames_to_column("EnsID") %>% 
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"), 
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"),
           type =  mapIds(org.Mm.eg.db, keys=EnsID,  column="GENETYPE",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


upf3b_res %>% with(., table(type))
x= list( "Upf3b KD DEGs" = DEGenes$UPF3B_vs_Control$EnsID, 
    "Upf3b KD DSGs" = upf3b_res$EnsID)

ggvenn(x[c(1,2)])
x = calculate.overlap(x)

ggvenn(x, fill_color = c("#2F124B", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 4.5, 
       stroke_size = 0.45, fill_alpha = 0.8, 
       stroke_color = "white", show_percentage = FALSE)
```


```{r}
  go.obj <- newGeneOverlap(DEGenes$UPF3B_vs_Control$EnsID, 
                         upf3b_res$EnsID,
                         genome.size=length(res$UPF3B_vs_Control$EnsID))
  go.obj <- testGeneOverlap(go.obj)


```

```{r fig.width=7, fig.height=8, fig.cap="Cumulative changes of LFC of gene expression results from genes which overlap with DSA versus genes which are non-overlapping"}

overlapping_upf3b  = DEGenes$UPF3B_vs_Control %>% 
  inner_join(upf3b_res, by="EnsID") %>% 
mutate(Status = "Overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3b = DEGenes$UPF3B_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
  mutate(Status = "Non-overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

rbind(overlapping_upf3b, non_overlapping_upf3b) %>% 
  ggplot(aes(x=logFC, color = Status)) + stat_ecdf(geom = "step", size=0.8) +
  #scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + 
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("")

ks.test(overlapping_upf3b$logFC, non_overlapping_upf3b$logFC)
```

<!-- # ```{r} -->
<!-- # upf3b_res %>%  -->
<!-- #   dplyr::select(EnsID, log2FoldChange) %>%  -->
<!-- #   inner_join(DEGenes$UPF3B_vs_Control, by="EnsID") %>%  -->
<!-- #   mutate(Label = ifelse(logFC > 0 & log2FoldChange   >0 |  -->
<!-- #                             logFC < 0 & log2FoldChange   <0, "Concordance", "Discordant")) %>% -->
<!-- #  # mutate(Res.UPF3B_vs_Control = as.character(Res.UPF3B_vs_Control)) %>% -->
<!-- #   tter(.,x="logFC", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line", -->
<!-- #               conf.int = TRUE,color = "Label", add.params = list(color = "blue", -->
<!-- #                                                  fill = "lightgray"),size=5, alpha =0.6) +  -->
<!-- #     theme_bw() + xlab("Upf3b KD DEG (log2FC)") + -->
<!-- #   ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +  -->
<!-- #     geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!-- #   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!-- #   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #         axis.title = element_text(size=15, family = "serif"),  -->
<!-- #          legend.box.background = element_rect(color = "black"), -->
<!-- #           legend.text = element_text(family = "serif"),  -->
<!-- #           legend.title = element_text( family = "serif"), -->
<!-- #            plot.title = element_text(family = "serif", size =20),          -->
<!-- #         panel.grid.major.x = element_blank(), -->
<!-- #         panel.grid.major.y = element_line( size=.1 ), -->
<!-- #               legend.position = "top", -->
<!-- #         strip.text.y = element_text( -->
<!-- #         size = 10, face = "bold.italic", family = "serif" -->
<!-- #         ), -->
<!-- #          strip.text.x = element_text( -->
<!-- #         size = 10, family = "serif" -->
<!-- #         )) + labs(color = "Expression") -->
<!-- # ``` -->


```{r fig.width=10, fig.height=10, fig.cap="Correlation analysis of DEG and DSA for Upf3b KD"}
overlapping_upf3b = DEGenes$UPF3B_vs_Control %>% 
  inner_join(upf3b_res, by="EnsID") %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
    left_join(res_dte$UPF3B_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))

overlap_DSA = overlapping_upf3b %>%  
  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3b KD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DSE (log2FC)")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-stability-overlap.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)
```            

```{r}
transcript_overlap = overlapping_upf3b %>%  
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3b KD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DET (log2FC)")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)
```

```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3b = DEGenes$UPF3B_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
  left_join(res_dte$UPF3B_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>% 
  mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))


table(overlapping_upf3b$nif) %>% 
  as.data.frame() %>% 
  mutate(group = "Overlap") %>% 
  rbind(table(non_overlapping_upf3b$nif) %>% 
          as.data.frame() %>% 
          mutate(group = "Non-overlapping"))
```

```{r}
nif_overlap = overlapping_upf3b %>% 
  distinct(gene_id, .keep_all = TRUE) %>%  
  dplyr::select(nif) %>% 
   mutate(group = "Overlap") %>% 
  rbind(non_overlapping_upf3b %>% 
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = "",
  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey"))

nif_overlap 


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```


```{r}
dEJ = non_overlapping_upf3b %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = non_overlapping_upf3b %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = non_overlapping_upf3b %>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = non_overlapping_upf3b$tx_id)
UpSetR::upset(fromList(x))


dEJ = overlapping_upf3b %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = overlapping_upf3b %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = overlapping_upf3b %>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = overlapping_upf3b$tx_id)
UpSetR::upset(fromList(x))

```


<!-- # ```{r} -->
<!-- #  overlapping %>%   -->
<!-- #   ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line", -->
<!-- #               conf.int = TRUE, color = "nif", add.params = list(color = "blue", -->
<!-- #                                                  fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +  -->
<!-- #     theme_bw() + xlab("Upf3b KD DEG (log2FC)") + -->
<!-- #   #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +  -->
<!-- #     geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!-- #   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!-- #   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #         axis.title = element_text(size=15, family = "serif"),  -->
<!-- #          legend.box.background = element_rect(color = "black"), -->
<!-- #           legend.text = element_text(family = "serif"),  -->
<!-- #           legend.title = element_text( family = "serif"), -->
<!-- #            plot.title = element_text(family = "serif", size =20),          -->
<!-- #         panel.grid.major.x = element_blank(), -->
<!-- #         panel.grid.major.y = element_line( size=.1 ), -->
<!-- #               legend.position = "top", -->
<!-- #         strip.text.y = element_text( -->
<!-- #         size = 10, face = "bold.italic", family = "serif" -->
<!-- #         ), -->
<!-- #          strip.text.x = element_text( -->
<!-- #         size = 10, family = "serif" -->
<!-- #         )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") + -->
<!-- #    theme(ggside.panel.scale = .3) + -->
<!-- #   scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) + -->
<!-- #   scale_color_manual(values = c("NIF+" = "#b23a48", -->
<!-- #                                 "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48", -->
<!-- #                                 "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DSE (log2FC)") -->
<!-- # ``` -->



## Upf3a KD stability estimates and overlap with differential expression analyses 

```{r fig.width=10, fig.height=15, fig.cap="Upf3a KD DSA volcano plot"}
upf3a_res = z[[1]]
upf3a_res %>% 
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3a_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
    ggplot(aes(x = log2FoldChange, 
               y = -log10(padj), 
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("Upf3a KD differentially stabilised genes") + 
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) + 
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
              axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "")
volc
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a-stability.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)

```


* Correlation between DSE and DEGs

```{r}
upf3a_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3A_KD.Ratio")) %>% 
  dplyr::filter(padj < 0.05) %>% 
  rownames_to_column("EnsID") %>% 
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"), 
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "DEG" = DEGenes$UPF3A_vs_Control$EnsID, 
    "Differential stability" = upf3a_res$EnsID)

ggvenn(x[c(1,2)])

ggvenn(x, fill_color = c("#ef476f", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 4.5, 
       stroke_size = 0.45, fill_alpha = 0.8, 
       stroke_color = "white", show_percentage = FALSE)

```

```{r}
x = calculate.overlap(x)
```

```{r}
  go.obj <- newGeneOverlap(DEGenes$UPF3A_vs_Control$EnsID, 
                         upf3a_res$EnsID,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
  go.obj <- testGeneOverlap(go.obj)


```

```{r}
overlapping_upf3a = DEGenes$UPF3A_vs_Control %>% 
  inner_join(upf3a_res, by="EnsID") %>% 
mutate(Status = "Overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
  mutate(Status = "Non-overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

rbind(overlapping_upf3a, non_overlapping_upf3a) %>% 
  ggplot(aes(x=logFC, color = Status)) + stat_ecdf(geom = "step", size=0.8) +
  #scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + 
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("")

ks.test(overlapping_upf3a$logFC, non_overlapping_upf3a$logFC)
```



```{r}
#res_sig = res[res$ensembl_gene_id %in% DEGenes$UPF3B_vs_Control$ensembl_gene_id,]

upf3a_res %>% 
  dplyr::select(EnsID, log2FoldChange) %>% 
  inner_join(DEGenes$UPF3A_vs_Control, by="EnsID") %>% 
  mutate(Label = ifelse(logFC > 0 & log2FoldChange   >0 | 
                            logFC < 0 & log2FoldChange   <0, "Concordance", "Discordant")) %>%
 # mutate(Res.UPF3B_vs_Control = as.character(Res.UPF3B_vs_Control)) %>%
  ggscatter(.,x="logFC", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE,color = "Label", add.params = list(color = "blue",
                                                 fill = "lightgray"),size=5, alpha =0.6) + 
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  ylab("Upf3a DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression")
```


```{r}
overlapping_upf3a = DEGenes$UPF3A_vs_Control %>% 
  inner_join(upf3a_res, by="EnsID") %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
    left_join(res_dte$UPF3A_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature) ,"NIF-", "NIF+"))

 overlap_DSA =overlapping_upf3a %>%  
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  ylab("Upf3a DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .4) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DSE (log2FC)")

overlap_DSA
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a-stability-overlap.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)


```


```{r}
transcript_overlap= overlapping_upf3a %>%  
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .4) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3a KD DET (log2FC)")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)
```


```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
    inner_join(res_dte$UPF3A_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>% 
  mutate(nif= ifelse(nif_feature == "None","NIF-", "NIF+"))

dEJ = non_overlapping_upf3a %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = non_overlapping_upf3a %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = non_overlapping_upf3a %>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = non_overlapping_upf3a$tx_id)
UpSetR::upset(fromList(x))


dEJ = overlapping_upf3a %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = overlapping_upf3a %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = overlapping_upf3a %>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = overlapping_upf3a$tx_id)
UpSetR::upset(fromList(x))

```


```{r}
table(overlapping_upf3a$nif) %>% 
  as.data.frame() %>% 
  mutate(group = "Overlap") %>% 
  rbind(table(non_overlapping_upf3a$nif) %>% 
          as.data.frame() %>% 
          mutate(group = "Non-overlapping")) %>% 
  ggplot(aes(x=group, y=Freq, fill = Var1)) +   
  geom_col(position = "fill", width = 0.5) +
  scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + coord_flip() + 
  theme_bw() + labs(x="", y="Proportion of transcripts", 
                    fill = "NMD inducible feature") +
    theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

non_overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
  left_join(res_dte$UPF3A_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>% 
  mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))

```

```{r}
nif_overlap = overlapping_upf3a %>% 
  distinct(gene_id, .keep_all = TRUE) %>%  
  dplyr::select(nif) %>% 
   mutate(group = "Overlap") %>% 
  rbind(non_overlapping_upf3a %>% 
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = "",
  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey"))

nif_overlap 


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```


### Identifying bona fide candidates 

To further differentiate transcriptional from post-transcriptional changes of gene expression and identify bone fide NMD targets in mouse L-cells, DEGs were overlapped with the differential mRNA stability estimates. If a gene was upregulated and stabilized or downregulated and destabilization, it would be considered a strong probable  indicator of post-transcriptional gene expression change.  On the other hand, discordance between the stability estimates and gene expression fold changes would be indicative change at the transcriptional level.

```{r}
bf_candidates = list()
```

```{r}
bf_up = overlapping_upf3b %>% 
  dplyr::filter(logFC.x > 0) %>% 
   dplyr::filter(nif == "NIF+") %>% 
   mutate(mgis = mapIds(org.Mm.eg.db, gene_id, "MGI","ENSEMBL"))
  
heat_map = bf_up  %>%
  dplyr::select( gene_id, NIF_dEJ,NIF_uORF, NIF_long_3utr, GC_3utr, LENGTH_3utr) 
heat_map$NIF_dEJ = as.numeric(heat_map$NIF_dEJ)
heat_map$NIF_uORF = as.numeric(heat_map$NIF_uORF)
heat_map$NIF_long_3utr = as.numeric(heat_map$NIF_long_3utr)
bf_candidates$upf3b = heat_map %>% 
  dplyr::group_by( gene_id) %>%
  dplyr::summarise(across(.cols = starts_with("NIF"),.fns = sum)) %>% 
  column_to_rownames("gene_id") %>% 
  mutate(Stabilized = 1, 
         Upregulated = 1) %>% 
  dplyr::select(Upregulated, Stabilized, everything()) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) 
```

<!-- # ```{r} -->
<!-- # x= list( "bona fide" = bf_up$gene_id,  -->
<!-- #     "upf3a DEGs" = DEGenes$UPF3A_vs_Control$EnsID) -->
<!-- #  -->
<!-- # ggvenn(x[c(1,2)]) -->
<!-- #  -->
<!-- # bf_up %>%  -->
<!-- #   inner_join(DEGenes$UPF3A_vs_Control %>%  -->
<!-- #                dplyr::rename("gene_id" = "EnsID"), by = "gene_id") %>%  -->
<!-- #   dplyr::select(logFC.x, logFC) -->
<!-- #  -->
<!-- # ``` -->

```{r}
bf_up  %>% 
  dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id",
                "Adjusted P-value (Gene)" = "adj.P.Val", 
                "Log2FoldChange (Gene)" ="logFC.x",
                "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR",
                 "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ", 
                "NIF_uORF","GC_3utr","LENGTH_3utr") %>% 
  mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"), 
         NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>%
   DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
                 caption = "Upregulated and stabilized targets of Upf3b. Targets were selected if they were differentially stabilized with gene expression results upregulation  Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.", 
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


```

<!-- * Downregulated -->

<!-- ```{r} -->
<!-- upf3b_bf_targets_down = overlapping %>% -->
<!--   dplyr::filter(logFC.x < -log2(1.5), nif_feature != "None") -->

<!-- ``` -->


<!-- ```{r} -->
<!-- upf3b_bf_targets_down %>% -->
<!--   dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id", -->
<!--                 "Adjusted P-value (Gene)" = "adj.P.Val", -->
<!--                 "Log2FoldChange (Gene)" ="logFC.x", -->
<!--                 "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR", -->
<!--                  "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ", -->
<!--                 "NIF_uORF","GC_3utr","LENGTH_3utr") %>% -->
<!--   mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"), -->
<!--          NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>% -->
<!--    DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                  caption = "Downregulated and destabilized targets of Upf3b. Targets were selected if they were differentially destabilized with gene expression results downregulation. Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.", -->
<!--             options = list(scrollX = TRUE, -->
<!--                            scrollCollapse = TRUE, -->
<!--                            pageLength = 5, autoWidth = TRUE, -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->


<!-- ``` -->


```{r}
upf3a_up = overlapping_upf3a %>% 
  dplyr::filter(logFC.x > 0) %>% 
   dplyr::filter(nif == "NIF+") 
```

```{r}
upf3a_up  %>% 
  dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id",
                "Adjusted P-value (Gene)" = "adj.P.Val", 
                "Log2FoldChange (Gene)" ="logFC.x",
                "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR",
                 "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ", 
                "NIF_uORF","GC_3utr","LENGTH_3utr") %>% 
  mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"), 
         NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>%
   DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
                 caption = "Upregulated and stabilized targets of Upf3a. Targets were selected if they were differentially stabilized with gene expression results upregulation  Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.", 
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


``` 

```{r}
heat_map = upf3a_up  %>%
  dplyr::select( gene_id, NIF_dEJ,NIF_uORF, NIF_long_3utr, GC_3utr, LENGTH_3utr) 
heat_map$NIF_dEJ = as.numeric(heat_map$NIF_dEJ)
heat_map$NIF_uORF = as.numeric(heat_map$NIF_uORF)
heat_map$NIF_long_3utr = as.numeric(heat_map$NIF_long_3utr)
bf_candidates$upf3a = heat_map %>% 
  dplyr::group_by(gene_id) %>%
  dplyr::summarise(across(.cols = starts_with("NIF"),.fns = sum)) %>% 
  column_to_rownames("gene_id") %>% 
  mutate(Stabilized = 1, 
         Upregulated = 1) %>% 
  dplyr::select(Upregulated, Stabilized, everything()) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) 

upf3a_candidates = bf_candidates$upf3a %>% t() %>%
    pheatmap(cellwidth = 10, cellheight = 10,
           	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
           cluster_rows = FALSE, 
           cluster_cols = FALSE)

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-bf-candidates-features.svg"), 
    height = 5, width =15, plot =  upf3a_candidates)
```


```{r}
upf3a_down = overlapping_upf3a %>% 
  dplyr::filter(logFC.x < 0) %>% 
   dplyr::filter(nif == "NIF+") 
```


### Overlap between Upf3a and Upf3b KD bona fide candidates 

```{r}
x= list( "Upf3b" = rownames(bf_candidates$upf3b), 
    "Upf3a" = rownames(bf_candidates$upf3a))
ggvenn(x)
```

```{r}
x = calculate.overlap(x)
upf3b_only = x$a1[!x$a1 %in% x$a3]
```

```{r}
upf3a_only = x$a2[!x$a2 %in% x$a3]

```

```{r}
overlapping_bf = x$a3
overlapping_GE = lapply(res, function(x){
  y = x %>% 
    dplyr::filter(EnsID %in% overlapping_bf ) %>%
    dplyr::select(EnsID,logFC, gene) %>% 
    dplyr::arrange(gene)
  })

ol_bf_exp = do.call(cbind, overlapping_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.gene") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))

rg <- ol_bf_exp %>%  max(abs(.))


ol_exp_heatmap =ol_bf_exp %>%
  t() %>% 
    pheatmap(,
        scale="none", 
        cluster_cols = TRUE,
        cluster_rows= FALSE,
        breaks = seq(-3, 3, length.out = 100),
        clustering_distance_rows = "euclidean", 
        border_color = "white", 
        treeheight_row = 0,
        fontsize = 6,
        color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100),
        labels_row = row.names(.), 
        cellwidth = 10, cellheight = 10 
    )

ggsave(filename  = here::here("output/Thesis_plots/overlap-bf-exp.svg"), 
    height = 5, width =15, plot =  ol_exp_heatmap)

```



```{r}
upf3b_GE = lapply(res, function(x){
  y = x %>% 
    dplyr::filter(EnsID %in% upf3b_only) %>%
    dplyr::select(EnsID,gene,logFC) %>% 
    dplyr::arrange(gene)
    
    
})

upf3b_bf_exp = do.call(cbind, upf3b_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.gene") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))



upf3b_exp_heatmap =upf3b_bf_exp %>%
  t() %>% 
    pheatmap(,
        scale="none", 
        cluster_cols = TRUE,
        cluster_rows= FALSE,
        breaks = seq(-3, 3, length.out = 100),
        clustering_distance_rows = "euclidean", 
        border_color = "white", 
        treeheight_row = 0,
        fontsize = 6,
        color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100),
        labels_row = row.names(.), 
        cellwidth = 10, cellheight = 10 
    )

ggsave(filename  = here::here("output/Thesis_plots/Upf3b-bf-exp.svg"), 
    height = 5, width =15, plot =  upf3b_exp_heatmap)
```


```{r}
upf3a_GE = lapply(res, function(x){
  y = x %>% 
    dplyr::filter(EnsID %in% upf3a_only) %>%
    dplyr::select(EnsID,gene,logFC) %>% 
    dplyr::arrange(gene)
    
    
})

upf3a_bf_exp = do.call(cbind, upf3a_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.gene") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))

upf3a_exp_heatmap =upf3a_bf_exp %>%
  t() %>% 
    pheatmap(,
        scale="none", 
        cluster_cols = TRUE,
        cluster_rows= FALSE,
        breaks = seq(-3, 3, length.out = 100),
        clustering_distance_rows = "euclidean", 
        border_color = "white", 
        treeheight_row = 0,
        fontsize = 6,
        color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100),
        labels_row = row.names(.), 
        cellwidth = 10, cellheight = 10 
    )

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-bf-exp.svg"), 
    height = 5, width =15, plot =  upf3a_exp_heatmap)
```


## For feature heatmap 
```{r}
upf3b_bf_feature = do.call(cbind, upf3b_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.EnsID") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))
upf3b_order <- dist(upf3b_bf_exp, method="euclidean")
upf3b_order <- hclust(upf3b_order )
upf3b_order = upf3b_order$labels[c(upf3b_order$order)]
upf3b_order %<>% as.data.frame() %>% 
  set_colnames("genes")
```

```{r}
upf3b_candidates = bf_candidates$upf3b %>%
  rownames_to_column("EnsID") %>%
  dplyr::filter(EnsID %in% upf3b_only)  %>% 
  left_join(upf3b_GE$UPF3B_vs_Control %>%
              dplyr::select(EnsID, gene), by = "EnsID") %>%
  dplyr::select(-EnsID) 

upf3b_candidates =  upf3b_candidates [ order(match(upf3b_candidates$gene, upf3b_order$genes)), ]

upf3b_feature_map = upf3b_candidates  %>%  as.data.frame() %>%
    rownames_to_column("x") %>%
   column_to_rownames("gene") %>%
    dplyr::select(-x) %>%
   t() %>%
     pheatmap(cellwidth = 10, cellheight = 10,
            	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
            cluster_rows = FALSE, 
            cluster_cols = FALSE, 
           fontsize = 6)

ggsave(filename  = here::here("output/Thesis_plots/Upf3b-bf-features.svg"), 
    height = 5, width =15, plot =  upf3b_feature_map)
```

```{r}
upf3a_bf_feature = do.call(cbind, upf3a_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.EnsID") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))
upf3a_order <- dist(upf3a_bf_exp, method="euclidean")
upf3a_order <- hclust(upf3a_order )
upf3a_order = upf3a_order$labels[c(upf3a_order$order)]
upf3a_order %<>% as.data.frame() %>% 
  set_colnames("genes")
```

```{r}
upf3a_candidates = bf_candidates$upf3a %>%
  rownames_to_column("EnsID") %>%
  dplyr::filter(EnsID %in% upf3a_only)  %>% 
  left_join(upf3a_GE$UPF3A_vs_Control %>%
              dplyr::select(EnsID, gene), by = "EnsID") %>%
  dplyr::select(-EnsID) 

upf3a_candidates =  upf3a_candidates [ order(match(upf3a_candidates$gene, upf3a_order$genes)), ]

upf3a_feature_map = upf3a_candidates  %>%  as.data.frame() %>%
    rownames_to_column("x") %>%
   column_to_rownames("gene") %>%
    dplyr::select(-x) %>%
   t() %>%
     pheatmap(cellwidth = 10, cellheight = 10,
            	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
            cluster_rows = FALSE, 
            cluster_cols = FALSE,
           fontsize = 6)

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-bf-features.svg"), 
    height = 5, width =15, plot =  upf3a_feature_map)
```

```{r}
ol_bf_feature = do.call(cbind, overlapping_GE) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.EnsID") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))
ol_order <- dist(ol_bf_feature, method="euclidean")
ol_order <- hclust(ol_order)
ol_order = ol_order$labels[c(ol_order$order)]
ol_order %<>% as.data.frame() %>% 
  set_colnames("genes")
```

```{r}
ol_candidates = bf_candidates$upf3a %>%
  rownames_to_column("EnsID") %>%
  dplyr::filter(EnsID %in% overlapping_bf)  %>% 
  left_join(overlapping_GE$UPF3B_vs_Control %>%
              dplyr::select(EnsID, gene), by = "EnsID") %>%
  dplyr::select(-EnsID) 

ol_candidates =  ol_candidates [ order(match(ol_candidates$gene, ol_order$genes)), ]

ol_feature_map = ol_candidates  %>%  as.data.frame() %>%
    rownames_to_column("x") %>%
   column_to_rownames("gene") %>%
    dplyr::select(-x) %>%
   t() %>%
     pheatmap(cellwidth = 10, cellheight = 10,
            	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
            cluster_rows = FALSE, 
            cluster_cols = FALSE, 
           fontsize = 6)

ggsave(filename  = here::here("output/Thesis_plots/overlapping-bf-features.svg"), 
    height = 5, width =15, plot =  ol_feature_map)
```


### Cumulative frequency 

```{r}
freq_upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) %>% 
  mutate(Condition = "Upf3b KD") %>% 
  rbind(DEGenes$UPF3A_vs_Control %>% 
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) %>% 
  mutate(Condition = "Upf3a KD ")) %>% 
  ggplot(aes(x=logFC, color = Condition)) + stat_ecdf(geom = "step", size=1) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Upf3b bona fide genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") + 
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))  + 
   annotate("point", x = 0.78, y = 0, colour = "#2F124B") +
  annotate("point", x = 0.760, y = 0, colour = "#EA2A5F") +
  annotate("text", x=2.5, y = 0.19, label = "D = 0.28571, p-value = 0.2639", fontface =7)

bf_upf3b_upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) 

bf_upf3b_upf3a = DEGenes$UPF3A_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) 
ks.test(bf_upf3b_upf3b$logFC, bf_upf3b_upf3a$logFC)

median(bf_upf3b_upf3b$logFC)
median(bf_upf3b_upf3a$logFC)

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b_bf_cumulative.svg", plot =freq_upf3b  , 
       units = c("in"), height = 5.65, width = 5)
```


```{r}
freq_upf3a = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) %>% 
  mutate(Condition = "Upf3b KD") %>% 
  rbind(DEGenes$UPF3A_vs_Control %>% 
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) %>% 
  mutate(Condition = "Upf3a KD ")) %>% 
  ggplot(aes(x=logFC, color = Condition)) + stat_ecdf(geom = "step", size=1) +
  scale_color_manual(values = c("#EA2A5F", "#2F124B")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Upf3a bona fide genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") + 
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))  + 
   annotate("point", x = 0.89, y = 0, colour = "#2F124B") +
  annotate("point", x =1.15, y = 0, colour = "#EA2A5F") +
  annotate("text", x=2.5, y = 0.19, label = "D = 0.31143, p-value = 0.08512", fontface =7)

bf_upf3a_upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) 

bf_upf3a_upf3a = DEGenes$UPF3A_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) 
ks.test(bf_upf3a_upf3b$logFC, bf_upf3a_upf3a$logFC)

median(bf_upf3a_upf3b$logFC)
median(bf_upf3a_upf3a$logFC)

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a_bf_cumulative.svg", plot =freq_upf3a  , 
       units = c("in"), height = 5.65, width = 5)
```


### Upf3a downregulated 

```{r}


upf3a_down_GE = lapply(res, function(x){
  y = x %>% 
    dplyr::filter(EnsID %in% upf3a_down$gene_id ) %>%
    dplyr::select(EnsID,logFC, gene) %>% 
    dplyr::arrange(gene)
  })


upf3a_down_exp = do.call(cbind, upf3a_down_GE ) %>% 
  dplyr::rename("GeneID" = "UPF3B_vs_Control.gene") %>% 
  column_to_rownames("GeneID") %>%
  dplyr::select(contains("logFC"))

upf3a_exp_heatmap =upf3a_down_exp %>%
  t() %>% 
    pheatmap(,
        scale="none", 
        cluster_cols = TRUE,
        cluster_rows= FALSE,
        breaks = seq(-3, 3, length.out = 100),
        clustering_distance_rows = "euclidean", 
        border_color = "white", 
        treeheight_row = 0,
        fontsize = 6,
        color = colorRampPalette(c("#2F124B", "#6E74B4","#F9F9F9", "#FD9675", "#EA2A5F"))(100),
        labels_row = row.names(.), 
        cellwidth = 10, cellheight = 10 
    )

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-down-exp.svg"), 
    height = 5, width =15, plot =  upf3a_exp_heatmap)
```


```{r}
heat_map = upf3a_down  %>%
  dplyr::select( gene_id, NIF_dEJ,NIF_uORF, NIF_long_3utr, GC_3utr, LENGTH_3utr) 
heat_map$NIF_dEJ = as.numeric(heat_map$NIF_dEJ)
heat_map$NIF_uORF = as.numeric(heat_map$NIF_uORF)
heat_map$NIF_long_3utr = as.numeric(heat_map$NIF_long_3utr)
bf_candidates$upf3a_down = heat_map %>% 
  dplyr::group_by( gene_id) %>%
  dplyr::summarise(across(.cols = starts_with("NIF"),.fns = sum)) %>% 
  column_to_rownames("gene_id") %>% 
  mutate(Stabilized = 1, 
         Upregulated = 1) %>% 
  dplyr::select(Upregulated, Stabilized, everything()) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) 


upf3a_bf_heatmap = bf_candidates$upf3a_down  %>%  as.data.frame() %>%
  rownames_to_column("gene") %>%
  mutate(gene_id = mapIds(org.Mm.eg.db, keys=gene,
                       column="SYMBOL",keytype="ENSEMBL", multiVals="first")) %>% 
   column_to_rownames("gene_id") %>%
    dplyr::select(-gene) %>%
   t() %>%
     pheatmap(cellwidth = 10, cellheight = 10,
            	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
            cluster_rows = FALSE, 
            cluster_cols = FALSE, 
           fontsize = 6)

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-down-bf.svg"), 
    height = 5, width =15, plot =  upf3a_bf_heatmap)
```


## Upf3 dKD


```{r}
upf3dKD_res = z[[3]]
upf3dKD_res %>% 
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3dKD_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
    ggplot(aes(x = log2FoldChange, 
               y = -log10(padj), 
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("Upf3 dKD differentially stabilised genes") + 
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) + 
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
              axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "")
volc
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-stability.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)


```

* Correlation between DSE and DEGs

```{r}
upf3dKD_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3A_KD_UPF3B_KD.Ratio")) %>% 
  dplyr::filter(padj < 0.05) %>% 
  rownames_to_column("EnsID") %>% 
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"), 
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "DEG" = DEGenes$DoubleKD_vs_Control$EnsID, 
    "Differential stability" = upf3dKD_res$EnsID)


ggvenn(x, fill_color = c("#7175b3", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 4.5, 
       stroke_size = 0.45, fill_alpha = 0.8, 
       stroke_color = "white", show_percentage = FALSE)

```

```{r}
x = calculate.overlap(x)
```

```{r}
  go.obj <- newGeneOverlap(DEGenes$DoubleKD_vs_Control$EnsID, 
                         upf3dKD_res$EnsID,
                         genome.size=length(res$DoubleKD_vs_Control$EnsID))
  go.obj <- testGeneOverlap(go.obj)


```

## Cumulative changes 


```{r}
overlapping = DEGenes$DoubleKD_vs_Control %>% 
  inner_join(upf3dKD_res, by="EnsID") %>% 
mutate(Status = "Overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping = DEGenes$DoubleKD_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
  mutate(Status = "Non-overlap") %>% 
  dplyr::select(EnsID, logFC, Status)

rbind(overlapping, non_overlapping) %>% 
  ggplot(aes(x=logFC, color = Status)) + stat_ecdf(geom = "step", size=0.8) +
  #scale_color_manual(values = c("#2F124B", "#FD9675", "#EA2A5F")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") + 
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("")

ks.test(overlapping$logFC, non_overlapping$logFC)
```



<!-- # ```{r} -->
<!-- # #res_sig = res[res$ensembl_gene_id %in% DEGenes$UPF3B_vs_Control$ensembl_gene_id,] -->
<!-- #  -->
<!-- # upf3dKD_res %>%  -->
<!-- #   dplyr::select(EnsID, log2FoldChange) %>%  -->
<!-- #   inner_join(DEGenes$DoubleKD_vs_Control, by="EnsID") %>%  -->
<!-- #   mutate(Label = ifelse(logFC > 0 & log2FoldChange   >0 |  -->
<!-- #                             logFC < 0 & log2FoldChange   <0, "Concordance", "Discordant")) %>% -->
<!-- #   dplyr::filter(Label == "Discordant") %>% -->
<!-- #  # mutate(Res.UPF3B_vs_Control = as.character(Res.UPF3B_vs_Control)) %>% -->
<!-- #   ggscatter(.,x="logFC", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line", -->
<!-- #               conf.int = TRUE,color = "Label", add.params = list(color = "blue", -->
<!-- #                                                  fill = "lightgray"),size=5, alpha =0.6) +  -->
<!-- #     theme_bw() + xlab("Upf3a KD DEG (log2FC)") + -->
<!-- #   ylab("Upf3a DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +  -->
<!-- #     geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") + -->
<!-- #   geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") + -->
<!-- #   theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #         axis.title = element_text(size=15, family = "serif"),  -->
<!-- #          legend.box.background = element_rect(color = "black"), -->
<!-- #           legend.text = element_text(family = "serif"),  -->
<!-- #           legend.title = element_text( family = "serif"), -->
<!-- #            plot.title = element_text(family = "serif", size =20),          -->
<!-- #         panel.grid.major.x = element_blank(), -->
<!-- #         panel.grid.major.y = element_line( size=.1 ), -->
<!-- #               legend.position = "top", -->
<!-- #         strip.text.y = element_text( -->
<!-- #         size = 10, face = "bold.italic", family = "serif" -->
<!-- #         ), -->
<!-- #          strip.text.x = element_text( -->
<!-- #         size = 10, family = "serif" -->
<!-- #         )) + labs(color = "Expression") -->
<!-- # ``` -->


```{r}
overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>% 
  inner_join(upf3dKD_res, by="EnsID") %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" |is.na(nif_feature),"NIF-", "NIF+"))

overlap_DSA = overlapping_upf3dKD %>%  
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3 dKD DEG (log2FC)") +
  ylab("Upf3 dKD DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DSE (log2FC)")

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-stability-overlap.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)


overlap_DSA
```            


```{r}
transcript_overlap= overlapping_upf3dKD %>%  
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) + 
    theme_bw() + xlab("Upf3 dKD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) + 
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3 dKD DET (log2FC)")

transcript_overlap
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)
```


```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>% 
    dplyr::rename("gene_id" = "EnsID") %>% 
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>% 
               left_join(tx2gene) %>% 
               dplyr::filter(FDR < 0.05),by = "gene_id") %>% 
  mutate(nif= ifelse(nif_feature == "None","NIF-", "NIF+"))

dEJ = non_overlapping_upf3dKD %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = non_overlapping_upf3dKD %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = non_overlapping_upf3dKD %>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = non_overlapping_upf3dKD$tx_id)
UpSetR::upset(fromList(x))


dEJ = overlapping_upf3dKD %>% 
  dplyr::filter(NIF_dEJ == 1) %>% 
  dplyr::select(tx_id)

uORF = overlapping_upf3dKD %>% 
  dplyr::filter(NIF_uORF == 1) %>% 
  dplyr::select(tx_id)
UTR = overlapping_upf3dKD%>% 
  dplyr::filter(NIF_long_3utr == 1) %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "Long 3'UTR" = UTR$tx_id, 
          "All transcripts" = overlapping_upf3dKD$tx_id)
UpSetR::upset(fromList(x))

```


<!-- # ```{r} -->
<!-- # table(overlapping_upf3dKD$nif) %>%  -->
<!-- #   as.data.frame() %>%  -->
<!-- #   mutate(group = "Overlap") %>%  -->
<!-- #   rbind(table(non_overlapping$nif) %>%  -->
<!-- #           as.data.frame() %>%  -->
<!-- #           mutate(group = "Non-overlapping")) %>%  -->
<!-- #   ggplot(aes(x=group, y=Freq, fill = Var1)) +    -->
<!-- #   geom_col(position = "fill", width = 0.5) + -->
<!-- #   scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + coord_flip() +  -->
<!-- #   theme_bw() + labs(x="", y="Proportion of transcripts",  -->
<!-- #                     fill = "NMD inducible feature") + -->
<!-- #     theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #             axis.text.y = element_text(size = 12, family = "serif", color = "black"), -->
<!-- #         axis.title = element_text(size=15, family = "serif"),  -->
<!-- #          legend.box.background = element_rect(color = "black"), -->
<!-- #           legend.text = element_text(family = "serif"),  -->
<!-- #           legend.title = element_text( family = "serif"), -->
<!-- #            plot.title = element_text(family = "serif", size =20),          -->
<!-- #         panel.grid.major.x = element_blank(), -->
<!-- #         panel.grid.major.y = element_line( size=.1 ), -->
<!-- #               legend.position = "top", -->
<!-- #         strip.text.y = element_text( -->
<!-- #         size = 10, face = "bold.italic", family = "serif" -->
<!-- #         ), -->
<!-- #          strip.text.x = element_text( -->
<!-- #         size = 10, family = "serif" -->
<!-- #         )) -->
<!-- #  -->
<!-- #  -->
<!-- # ``` -->

```{r}
nif_overlap =overlapping_upf3dKD %>% 
  distinct(gene_id, .keep_all = TRUE) %>%  
  dplyr::select(nif) %>% 
   mutate(group = "Overlap") %>% 
  rbind(non_overlapping_upf3dKD  %>% 
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = "",
  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey"))

nif_overlap 


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```

```{r}
upf3dKD_up = overlapping_upf3dKD %>% 
  dplyr::filter(logFC.x > 0) %>% 
   dplyr::filter(nif == "NIF+") 
```

```{r}
table(upf3dKD_up$NIF_dEJ)
table(upf3dKD_up$NIF_uORF)
table(upf3dKD_up$NIF_long_3utr)

dKD_features = data.frame("Feature" = c("dEJC", "uORF", "Long 3'UTR", "dEJC Overlap with Upf3b", "dEJC overlap with Upf3a"), 
                              "Count" = c(table(upf3dKD_up$NIF_dEJ)[2],
                                          table(upf3dKD_up$NIF_uORF)[2],
                                          table(upf3dKD_up$NIF_long_3utr)[2], 
                                          9, 1)) %>% 
  mutate(Feature = factor(Feature, levels=c("dEJC", "uORF", "Long 3'UTR", "dEJC Overlap with Upf3b", "dEJC overlap with Upf3a"))) %>%
  ggplot(aes(x=Feature, y = Count, fill=Feature)) + geom_bar(stat = "identity", width = .55) + theme_classic() +
  ylab("Number of transcripts") + xlab("") + geom_vline(xintercept = 3.5, linetype = "dashed", colour = "black", size = 1.5) + 
   scale_fill_manual(values =c("#6d597a", "#eaac8b", "#fcb9b2", "#585A58","#585A58")) + theme(axis.text.x = element_text(size = 12, family = "serif", color = "black", angle = 75),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-bf-NIF-features.svg", plot = dKD_features,  units = c("in"), height = 6, width = 5.5)
```

```{r}
upf3dKD_dej= upf3dKD_up %>% 
  dplyr::filter(NIF_dEJ == 1) 
upf3b_dej = bf_up %>% 
                 dplyr::filter(NIF_dEJ == 1)

upf3a_dej = upf3a_up %>% 
                 dplyr::filter(NIF_dEJ == 1)
 x = list("upf3b" = upf3b_dej$tx_id, 
          "upf3 dKD" = upf3dKD_dej$tx_id, 
          "upf3a" = upf3a_dej$tx_id)
 
 
 ggvenn(x)

```

```{r}
heat_map = upf3dKD_up %>%
  dplyr::select(gene.y, NIF_dEJ,NIF_uORF, NIF_long_3utr, GC_3utr, LENGTH_3utr) 
heat_map$NIF_dEJ = as.numeric(heat_map$NIF_dEJ)
heat_map$NIF_uORF = as.numeric(heat_map$NIF_uORF)
heat_map$NIF_long_3utr = as.numeric(heat_map$NIF_long_3utr)
bf_candidates$upf3dKD_up = heat_map %>% 
  dplyr::group_by(gene.y) %>%
  dplyr::summarise(across(.cols = starts_with("NIF"),.fns = sum)) %>% 
  column_to_rownames("gene.y") %>% 
  mutate(Stabilized = 1, 
         Upregulated = 1) %>% 
  dplyr::select(Upregulated, Stabilized, everything()) %>% 
  mutate_if(is.numeric, ~1 * (. != 0)) 

upf3dKD_candidates = bf_candidates$upf3dKD_up %>% t() %>%
    pheatmap(cellwidth = 10, cellheight = 10,
           	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
           cluster_rows = FALSE, 
           cluster_cols = FALSE)

ggsave(filename  = here::here("output/Thesis_plots/Upf3dKD-bf-candidates-features.svg"), 
    height = 5, width =15, plot =  upf3dKD_candidates)
```

```{r}
nif_overlap = overlapping_upf3dKD %>% 
  distinct(gene_id, .keep_all = TRUE) %>%  
  dplyr::select(nif) %>% 
   mutate(group = "Overlap") %>% 
  rbind(non_overlapping_upf3dKD %>% 
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = "",
  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey"))


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-bf-NIF-proportion.svg", plot = nif_overlap ,  units = c("in"), height = 7, width = 7)
```

<!-- # ```{r} -->
<!-- # x= list("Upf3b KD" = upf3b_candidates,  -->
<!-- #         "Upf3a KD" = upf3a_cadidates, -->
<!-- #         "Upf3 dKD" = upf3dKD_candidates) -->
<!-- # ggvenn(x) -->
<!-- # ``` -->


<!-- # ```{r} -->
<!-- # genes =DEGenes$UPF3B_vs_Control %>%  -->
<!-- #      mutate(mgis = mapIds(org.Mm.eg.db, EnsID, "MGI","ENSEMBL")) -->
<!-- #    -->
<!-- # mgi_ids = unique(genes$mgis) %>%  -->
<!-- #   as.data.frame() -->
<!-- #  -->
<!-- # write.table(mgi_ids, "Upf3b-all-DEGs-genes.tsv", sep = "\t") -->
<!-- # ``` -->



```{r}
x= list( "Upf3b KD" = rownames(bf_candidates$upf3b), 
    "Upf3a KD" = rownames(bf_candidates$upf3a),
    "Upf3 dKD" = unique(upf3dKD_up$gene_id))

ggvenn(x, fill_color = c("#2F124B", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 4.5, 
       stroke_size = 0.45, fill_alpha = 0.8, 
       stroke_color = "white")

```

```{r}
x = calculate.overlap(x)
```



```{r}
upf3d_KD_only = x$a7 %>% 
  as.data.frame() %>% 
  set_colnames("gene_id") %>%
  left_join(upf3dKD_up %>% 
              dplyr::select(gene_id, gene.x), by = "gene_id")
```

```{r}
upf3a_candidates = bf_candidates$upf3dKD_up %>%
  rownames_to_column("EnsID") %>%
  dplyr::filter(EnsID %in% x$a7)  %>% 
  left_join(upf3a_GE$UPF3A_vs_Control %>%
              dplyr::select(EnsID, gene), by = "EnsID") %>%
  dplyr::select(-EnsID) 

upf3a_candidates =  upf3a_candidates [ order(match(upf3a_candidates$gene, upf3a_order$genes)), ]

upf3a_feature_map =  bf_candidates$upf3dKD_up %>% 
  as.data.frame() %>%
  rownames_to_column("gene") %>%
  dplyr::filter(gene %in% upf3d_KD_only$gene.x) %>%
  column_to_rownames("gene") %>%
   t() %>%
     pheatmap(cellwidth = 10, cellheight = 10,
            	color = (c("#fffdf7", "#b23a48")), 
           border_color = "white", 
            cluster_rows = FALSE, 
            cluster_cols = FALSE,
           fontsize = 6)

ggsave(filename  = here::here("output/Thesis_plots/Upf3a-bf-features.svg"), 
    height = 5, width =15, plot =  upf3a_feature_map)
```

```{r}
freq_upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) %>% 
  mutate(Condition = "Upf3b KD") %>% 
  rbind(res$UPF3A_vs_Control %>% 
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) %>% 
  mutate(Condition = "Upf3a KD")) %>% 
  rbind(res$DoubleKD_vs_Control %>% 
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) %>% 
  mutate(Condition = "Upf3 dKD")) %>% 
  ggplot(aes(x=logFC, color = Condition)) + stat_ecdf(geom = "step", size=1) +
  scale_color_manual(values = c("Upf3a KD" ="#EA2A5F", "Upf3b KD"="#2F124B","Upf3 dKD" ="grey")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Upf3b bona fide genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") + 
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))  + 
   annotate("point", x = 0.78, y = 0, colour = "#2F124B") +
  annotate("point", x = 0.23, y = 0, colour = "#EA2A5F") +
  annotate("point", x = 0.98, y = 0, colour = "grey")+
  annotate("text", x=2.5, y = 0.19, label = "Upf3a KD genes: D = 0.8, p < 2.2e-16", fontface =7, color="#EA2A5F") + 
  annotate("text", x=2.5, y = 0.15, label = "Upf3 dKD genes: D = 0.3, p = 3.4e-3", fontface =7, color="grey")

bf_upf3b_upf3b = DEGenes$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) 

bf_upf3b_upf3a = res$UPF3A_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) 
ks.test(bf_upf3b_upf3b$logFC, bf_upf3b_upf3a$logFC)

bf_upf3b_upf3d = res$DoubleKD_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3b)) 
ks.test(bf_upf3b_upf3b$logFC, bf_upf3b_upf3a$logFC)
ks.test(bf_upf3b_upf3b$logFC, bf_upf3b_upf3d$logFC)

median(bf_upf3b_upf3b$logFC)
median(bf_upf3b_upf3a$logFC)
median(bf_upf3b_upf3d$logFC)

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b_bf_cumulative.svg", plot =freq_upf3b  , 
       units = c("in"), height = 5.65, width = 5)
```


```{r}
freq_upf3a = res$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) %>% 
  mutate(Condition = "Upf3b KD") %>% 
  rbind(DEGenes$UPF3A_vs_Control %>% 
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) %>% 
  mutate(Condition = "Upf3a KD")) %>% 
  rbind(res$DoubleKD_vs_Control %>%
          dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) %>% 
          mutate(Condition = "Upf3 dKD")) %>%
  ggplot(aes(x=logFC, color = Condition)) + stat_ecdf(geom = "step", size=1) +
  scale_color_manual(values = c("Upf3a KD" ="#EA2A5F", "Upf3b KD"="#2F124B","Upf3 dKD" ="grey")) + theme_bw() +
  ylab("Cumulative Frequency") + xlab("") +# theme(legend.position = "none") +
  geom_hline(yintercept = c(1, 0),  linetype = "dashed", color = "grey") +
  ggtitle("Upf3b bona fide genes") + theme(legend.position = "none") + xlab("Log 2 Fold Change") + 
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),         
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))  + 
   annotate("point", x = 0.59, y = 0, colour = "#2F124B") +
  annotate("point", x =1.15, y = 0, colour = "#EA2A5F") +
  annotate("point", x = 1.38, y = 0, colour = "grey")+
  annotate("text", x=2.5, y = 0.19, label = "Upf3b KD genes: D = 0.51, p = 1.73e-05", fontface =7, color="#2F124B") + 
  annotate("text", x=2.5, y = 0.15, label = "Upf3 dKD genes: D = 0.20, p = 3.05e-01", fontface =7, color="grey")


bf_upf3a_upf3b = res$UPF3B_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) 

bf_upf3a_upf3a = DEGenes$UPF3A_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a))

bf_upf3a_upf3dkd = res$DoubleKD_vs_Control %>% 
  dplyr::filter(EnsID %in% rownames(bf_candidates$upf3a)) 
ks.test(bf_upf3a_upf3a$logFC, bf_upf3a_upf3b$logFC)
ks.test(bf_upf3a_upf3a$logFC, bf_upf3a_upf3dkd$logFC)

median(bf_upf3a_upf3b$logFC)
median(bf_upf3a_upf3a$logFC)
median(bf_upf3a_upf3dkd$logFC)

ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a_bf_cumulative.svg", plot =freq_upf3a  , 
       units = c("in"), height = 5.65, width = 5)
```


#### Export tables 


```{r}
overlapping_lists = list()

overlapping_lists$upf3dKD = overlapping_upf3dKD

overlapping_lists$upf3a = overlapping_upf3a

overlapping_lists$upf3b = overlapping_upf3b 

write.xlsx(overlapping_lists, file= here::here("output/Overlap-stability-DEG.xlsx"))
```

```{r}

bona_fide = list()
bona_fide$upf3dKD = upf3dKD_up
bona_fide$upf3a = upf3a_up
bona_fide$upf3b = bf_up

write.xlsx(bona_fide , file= here::here("output/Bona-fide-genes.xlsx"))

```

### Functional relevance 

- Upf3b 
```{r}

analysis =  res$UPF3B_vs_Control
genes = as.integer(analysis$EnsID %in% rownames(bf_candidates$upf3b))
not_na = !is.na(genes)
names(genes) = analysis$EnsID
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")

    
Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE)
goterms <- Go.wall$category
goList <- getGeneLists(pwf, goterms, "mm39", "ensGene")
Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]])
Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L))
goRes <- Go.wall %>%
            dplyr::select(-under_represented_pvalue) %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            left_join(goSummaries)
nGenes <- c(BG = sum(genes == 0), DE = sum(genes))
goFinal <- goRes %>%
            dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > 3) %>%
            arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>%
            arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(over_represented_pvalue < 0.05, numDEInCat > Expected)
  
```


```{r}

goFinal$Genes = sapply(goFinal$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })
goFinal$Genes = vapply(goFinal$Genes , paste, collapse = ", ", character(1L))


goFinal %>% 
    dplyr::select(-c(contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in Upf1 KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


```{r}

analysis =  res$UPF3A_vs_Control
genes = as.integer(analysis$EnsID %in% rownames(bf_candidates$upf3a))
not_na = !is.na(genes)
names(genes) = analysis$EnsID
genes = genes[not_na]
    
select_genes <- as.vector(names(lengthData)%in%names(genes))
new_lengthData <- lengthData[select_genes]
select_genes <- as.vector(names(genes) %in% names(new_lengthData))
genes = genes[select_genes]
pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE)
plotPWF(pwf, main = "Bias from gene length")

    
Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE)
goterms <- Go.wall$category
goList <- getGeneLists(pwf, goterms, "mm39", "ensGene")
Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]])
Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L))
goRes <- Go.wall %>%
            dplyr::select(-under_represented_pvalue) %>%
            dplyr::filter(numDEInCat > 0) %>%
            dplyr::rename(id = category) %>%
            left_join(goSummaries)
nGenes <- c(BG = sum(genes == 0), DE = sum(genes))
goFinal <- goRes %>%
            dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > 3) %>%
            arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>%
            arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>%
            mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"),
                   Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>%
            dplyr::filter(numDEInCat > Expected)
  
```


```{r}

goFinal$Genes = sapply(goFinal$EnsemblID, function(x){
            tmp = strsplit(as.character(x), ",") %>% unlist()
            tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim)
            tmp = tmp[[1]]
            tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% 
                unname
            return(tmp)
        })
goFinal$Genes = vapply(goFinal$Genes , paste, collapse = ", ", character(1L))


goFinal %>% 
    dplyr::select(-c(contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>%
 DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
               caption ="Significant gene ontology terms in Upf1 KD compared to controls",
               style = "auto", width = NULL, height = NULL, elementId = NULL,
            options = list(scrollX = TRUE, 
                           scrollCollapse = TRUE, 
                           pageLength = 5, autoWidth = TRUE, 
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))
```


