---
title: "RNA-stability"
author: "unawaz1996"
date: "2023-05-05"
output:
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    autodep = TRUE,
    eval = TRUE,
	echo = TRUE,
	warning = FALSE,
	message = FALSE
)
```

```{r libraries}

library(eisaR)
source("code/libraries.R")
source("/home/neuro/Documents/NMD_analysis/other_repos/DiffRAC/DiffRAC.R")
source("code/functions.R")
library(coin)
library(gplots)
library(stargazer)
library(janitor)
library(ggside)
library(ggstatsplot)
library(GeneOverlap)
```

```{r}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])
```

```{r data-load}
load("output/DEG-limma-results.Rda")
load("output/DEG-list.Rda")
load("output/DET-list.Rda")
```

```{r}
NMD_features = read.table("data/nif_output/Mus_musculus__nifs.tsv", sep = "\t", skip =1)

colnames(NMD_features)[1:4] = c("EnsID", "NIF_dEJ", "NIF_uORF", "NIF_long_3utr")
NMD_features = NMD_features[,1:4]
NMD_features$NIF_dEJ = ifelse(NMD_features$NIF_dEJ == "True",gsub("True", 1, NMD_features$NIF_dEJ), 0)
NMD_features$NIF_uORF = ifelse(NMD_features$NIF_uORF == "True",gsub("True", 1, NMD_features$NIF_uORF), 0)
NMD_features$NIF_long_3utr = ifelse(NMD_features$NIF_long_3utr == "True",gsub("True", 1, NMD_features$NIF_long_3utr), 0)
NMD_features %<>% mutate_at(.vars = "EnsID", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "")

NMD_features$NIF_dEJ[is.na(NMD_features$NIF_dEJ)] <- 0
NMD_features$NIF_uORF[is.na(NMD_features$NIF_uORF)] <- 0
NMD_features$NIF_long_3utr[is.na(NMD_features$NIF_long_3utr)] <- 0
```


# Introduction

It has been suggested that while exonic read counts in RNA-seq data correspond to steady-state mRNA abundance, changes in the abundance of intronic reads can be used to estimate the change in transcription rate. Through this concept, a change in exonic reads without a corresponding change in intronic reads is diagnostic of differential RNA stability, while concurrent changes in both exonic and intronic reads suggest altered transcription. DiffRAC is a framework that converts unspliced/spliced relationships into a generalized linear model whose parameters can then be inferred from sequencing count data.  

## Differential stability analysis 

```{r}
md = read.table("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/data/LTK_Sample Metafile_V3.txt", header = TRUE, sep="\t")
md$Group = factor(md$Group, levels = c("Control", "UPF3A_KD", "UPF3A_OE", 
                                       "UPF3B_KD", "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
sampleTable= read.csv("/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/data/data2.txt", header=TRUE, sep = ",")
inputFolder="/home/neuro/Documents/NMD_analysis/Analysis/Results/EISA"

de_exonic  = DESeqDataSetFromHTSeqCount( # omits special rows from htseq-count
		    sampleTable = sampleTable[sampleTable$ReadType=="exonic",],
		    directory = inputFolder,
		    design = ~ 1 # required parameter
		)

countsEx= counts(de_exonic)

de_intron = DESeqDataSetFromHTSeqCount( # omits special rows from htseq-count
		    sampleTable = sampleTable[sampleTable$ReadType=="intronic",],
		    directory = inputFolder,
		    design = ~ 1 # required parameter
		)
countsIn = counts(de_intron)
```

```{r}
md$Samples = colnames(countsEx)
md %<>% dplyr::select(Samples, Group)

rownames(md) = md$Samples

md$Group = factor(md$Group, levels = c("Control", "UPF3A_KD", "UPF3A_OE", 
                                       "UPF3B_KD", "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))
```


```{r}
countsEx = countsEx[rownames(countsEx) %in% rownames(countsIn),]
countsIn = countsIn[rownames(countsIn) %in% rownames(countsEx),]
```

```{r include=FALSE}
diffrac_res <- DiffRAC( ~Group,
         md,
         countsEx,countsIn,
         "sample",
         optimizeBias = T)
```




### Overal results 

```{r}
resultsTable = data.frame()
names = c("GroupUPF3A_KD.Ratio","GroupUPF3B_KD.Ratio",
          "GroupUPF3A_KD_UPF3B_KD.Ratio","GroupUPF3A_OE.Ratio",
          "GroupUPF3A_OE_UPF3B_KD.Ratio")

for (analysis in names){
  temp =  as.data.frame(results(diffrac_res$dds,name = analysis))
  stability_total = temp %>%
    dplyr::filter(padj < 0.05) %>%
  with(., 
       table(sign.lfc=sign(log2FoldChange))) %>%
    as.data.frame() %>%
    t() %>%
    as.data.frame() %>% 
    row_to_names(row_number = 1) %>%
    mutate(Group = analysis)
  resultsTable = rbind(resultsTable, stability_total)
}

resultsTable %<>% 
  dplyr::rename("Destabilized" = "-1", "Stabilized" = "1") %>% 
  rownames_to_column("random") %>% 
  dplyr::select(Group, everything(), -c(random)) %>% 
  mutate(Group = ifelse(Group == 'GroupUPF3A_KD.Ratio',"Upf3a KD targets",
                          ifelse(Group == 'GroupUPF3B_KD.Ratio',"Upf3b KD targets", 
                                 ifelse(Group == 'GroupUPF3A_KD_UPF3B_KD.Ratio',"Upf3 dKD targets",
                                        ifelse(Group == 'GroupUPF3A_OE.Ratio',"Upf3a OE targets", "Upf3a OE Upf3b KD targets" )))))



resultsTable %>%
  DT::datatable(caption = "Differential mRNA stability estimates")
```


```{r}
z <- lapply(names, function(x) {
       tmp <- results(diffrac_res$dds, name = x) %>% 
         as.data.frame() 
       tmp %<>% mutate(group = as.character(x))
       tmp %<>% rownames_to_column("EnsemblID")
       #names(tmp) <- paste(x, names(tmp), sep = ".")
       return(tmp)})
save(z,file = here::here("output/Stability/differential-stability.Rda"))

alltables <- do.call(rbind, z)
``` 


```{r fig.height=10, fig.width=10, fig.cap="Distribution of RNA-stability estimates across comparisons"}
DEColours <- c("-1" = "#4657A6","1" = "#C32137",  "0" = "#D3D5E3")

stability_dist = alltables %>% 
  mutate(results = sign(log2FoldChange)) %>%
  dplyr::filter(padj < 0.05) %>%
  mutate(group=factor(group, levels = c("GroupUPF3B_KD.Ratio",
                                        "GroupUPF3A_KD.Ratio", 
                                        "GroupUPF3A_KD_UPF3B_KD.Ratio", 
                                        "GroupUPF3A_OE.Ratio",
                                        "GroupUPF3A_OE_UPF3B_KD.Ratio"))) %>%
  ggplot(aes(x=log2FoldChange, fill=as.factor(results))) +
    geom_histogram(alpha = 0.65) + theme_bw() +
  scale_fill_manual(values=DEColours ) +
  theme(legend.position="none") + xlim(-2, 2) +
  facet_wrap(~group) +
theme(panel.spacing = unit(1, "lines"))


stability_dist
 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Stability/Thesis_figures/stability_dist.png", 
       width = 8.99, height=7.5, units = "in", plot = stability_dist)
```

### Upf3b KD NMD genes 


```{r fig.width=10, fig.height=15, fig.cap="Upf3b KD DSA volcano plot"}
upf3b_res = z[[2]]

DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3b_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
  drop_na("results") %>%
    ggplot(aes(x = log2FoldChange, 
               y = -log10(padj), 
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value"))) +
  
  #  labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
  #  ggtitle("Upf3b KD differentially stabilised genes") + 
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) + 
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 20, family = "serif", color = "black"),
              axis.text.y = element_text(size = 20, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"), 
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif", size = 25), 
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "") 

volc
```




* Correlation between DSE and DEGs

- Overlap

```{r fig.width=8, fig.height=8, fig.cap="Overlap of DEG and DSA analysis from differential stability estimates"}
upf3b_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3B_KD.Ratio")) %>%
  dplyr::filter(padj < 0.05) %>%
  rownames_to_column("EnsID") %>%
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first"),
           type =  mapIds(org.Mm.eg.db, keys=EnsID,  column="GENETYPE",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "Upf3b KD DEGs" = DEGenes$UPF3B_vs_Control$EnsID,
    "Upf3b KD DSGs" = upf3b_res$EnsID)


ggvenn(x[c(1,2)], fill_color = c("#2F124B", "#801D2E"),
       set_name_color = "black", text_color = "white", text_size = 4.5,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)
```


```{r}
x = calculate.overlap(x)

go.obj <- newGeneOverlap(DEGenes$UPF3B_vs_Control$EnsID,
                         upf3b_res$EnsID,
                         genome.size=length(res$UPF3B_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```

- Correlation

```{r fig.width=10, fig.height=10, fig.cap="Correlation analysis of DEG and DSA for Upf3b KD"}
overlapping_upf3b = DEGenes$UPF3B_vs_Control %>%
  inner_join(upf3b_res, by="EnsID") %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))

overlap_DSA = overlapping_upf3b %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif",
            cor.coeff.args = list(size = 6, family = "serif"),
            add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
            axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=20, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(
  y = expression(italic(Upf3b) ~ "KD DSG (" * log[2] * "FC)"),
  x = expression(italic(Upf3b) ~ "KD DEG (" * log[2] * "FC)"))
overlap_DSA


 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Main/Upf3b-stability-overlap_3c.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)
```


```{r}
transcript_overlap = overlapping_upf3b %>%
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3b KD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DET (log2FC)")

# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3b-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)
```

```{r}
ggarrange(overlap_DSA, transcript_overlap)
```

- Transcripts with NIF features

```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3b = DEGenes$UPF3B_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>%
    dplyr::rename("gene_id" = "EnsID") %>%
  left_join(res_dte$UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
  mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))


table(overlapping_upf3b$nif) %>%
  as.data.frame() %>%
  mutate(group = "Overlap") %>%
  rbind(table(non_overlapping_upf3b$nif) %>%
          as.data.frame() %>%
          mutate(group = "Non-overlapping"))
```

```{r}
nif_overlap = overlapping_upf3b %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  dplyr::select(nif) %>%
   mutate(group = "DEGs/DSGs\nOverlapping") %>%
  rbind(non_overlapping_upf3b %>%
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping\nDEGs")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = ""#,
  #ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + theme_bw() +
  labs(y="Prpoportion of transcripts") +
  theme(axis.text = element_text(color ="black", size =12),
         legend.box.background = element_rect(color = "black"))

nif_overlap


# overlapping_upf3b %>%
#   distinct(gene_id, .keep_all = TRUE) %>%
#   dplyr::select(nif) %>%
#    mutate(group = "Overlap") %>%
#   rbind(non_overlapping_upf3b %>%
#           distinct(gene_id, .keep_all = TRUE) %>%
#          dplyr::select(nif) %>%
#           mutate(group = "Non-overlapping"))

 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Main/Upf3b-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 4.5, dpi =1200)
```


### Upf3a KD NMD genes


```{r fig.width=10, fig.height=15, fig.cap="Upf3a KD DSA volcano plot"}
upf3a_res = z[[1]]
upf3a_res %>%
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3a_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
  drop_na("results") %>%
    ggplot(aes(x = log2FoldChange,
               y = -log10(padj),
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("Upf3a KD differentially stabilised genes") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) +
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
              axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "") +   xlab(bquote(log["2"](FC))) +
  ylab(bquote(-log["10"]("adjusted P-value")))
volc
ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-stability-Supp-5a.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)

```


* Correlation between DSE and DEGs

```{r}
upf3a_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3A_KD.Ratio")) %>%
  dplyr::filter(padj < 0.05) %>%
  rownames_to_column("EnsID") %>%
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "DEG" = DEGenes$UPF3A_vs_Control$EnsID,
    "Differential stability" = upf3a_res$EnsID)


ggvenn(x[c(1,2)], fill_color = c("#ef476f", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 10,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)

```

```{r}
x = calculate.overlap(x)
```

```{r}
go.obj <- newGeneOverlap(DEGenes$UPF3A_vs_Control$EnsID,
                         upf3a_res$EnsID,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```


```{r}
overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  inner_join(upf3a_res, by="EnsID") %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature) ,"NIF-", "NIF+"))

overlap_DSA =overlapping_upf3a %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
             cor.coeff.args = list(size = 6, family = "serif"),
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  ylab("Upf3a DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .4) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) +
labs(  y = expression(italic(Upf3a) ~ "KD DSG (" * log[2] * "FC)"),
  x = expression(italic(Upf3a) ~ "KD DEG (" * log[2] * "FC)"))

overlap_DSA

 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-stability-overlap_5c.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)

```


```{r}
transcript_overlap= overlapping_upf3a %>%
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3a KD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .4) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3a KD DET (log2FC)")
# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3a-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)



```


```{r}
ggarrange(overlap_DSA, transcript_overlap)
```


```{r}

non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3a = DEGenes$UPF3A_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>%
    dplyr::rename("gene_id" = "EnsID") %>%
  left_join(res_dte$UPF3A_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
  mutate(nif= ifelse(nif_feature == "None" | is.na(nif_feature),"NIF-", "NIF+"))

nif_overlap = overlapping_upf3a %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  dplyr::select(nif) %>%
   mutate(group = "DEGs/DSGs\nOverlapping") %>%
  rbind(non_overlapping_upf3a %>%
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping\nDEGs")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = ""#,
 # ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey")) + theme_bw() +
    labs(y="Prpoportion of transcripts") +
  theme(axis.text = element_text(color ="black", size =12),
         legend.box.background = element_rect(color = "black"))

nif_overlap


 ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Figures/Supplementary/Upf3a-bf-NIF-proportion_fig5d.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```

### Upf3 dKD NMD genes




```{r}
upf3dKD_res = z[[3]]
upf3dKD_res %>%
  dplyr::filter(padj < 0.05) %>%
  with(., table(sign(log2FoldChange)))


DEColours <- c("Destabilized" = "#2F124B","Stabilised" = "#ee6c4d",  "Not significant" = "#D3D5E3", "NA" = "#D3D5E3")
volc = upf3dKD_res %>%
   mutate(results = ifelse(log2FoldChange> 0 & padj < 0.05, "Stabilised",
                           ifelse(padj < 0.05 & log2FoldChange < 0, "Destabilized", "Not significant"))) %>%
    ggplot(aes(x = log2FoldChange,
               y = -log10(padj),
               colour = as.factor(as.character(results )))) +
    geom_point(alpha = 0.8, size =3) +
    scale_colour_manual(values = DEColours) + theme_classic() +
    geom_hline(yintercept = -log10(0.05), color = "grey60", size = 0.5, lty = "dashed") +
    labs(x = "log2 Fold Change", y = "-log10 adj p-value") +
    ggtitle("Upf3 dKD differentially stabilised genes") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + xlim(-3, 3) +
  ylim(0,7.5) +
        theme(axis.text.x = element_text(size = 15, family = "serif", color = "black"),
              axis.text.y = element_text(size = 15, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =15),
        legend.position = "top"
        )  + labs(color = "")
volc
# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-stability.png", plot = volc,  units = c("in"), height = 6, width = 6, dpi =1200)


```


* Correlation between DSE and DEGs

```{r}
upf3dKD_res =  as.data.frame(results(diffrac_res$dds,name = "GroupUPF3A_KD_UPF3B_KD.Ratio")) %>%
  dplyr::filter(padj < 0.05) %>%
  rownames_to_column("EnsID") %>%
    mutate(gene = mapIds(org.Mm.eg.db, keys=EnsID,  column="SYMBOL",keytype="ENSEMBL", multiVals="first"),
           entrez = mapIds(org.Mm.eg.db, keys=EnsID,  column="ENTREZID",keytype="ENSEMBL", multiVals="first")) %>% drop_na(entrez)


x= list( "DEG" = DEGenes$DoubleKD_vs_Control$EnsID,
    "Differential stability" = upf3dKD_res$EnsID)


ggvenn(x, fill_color = c("#7175b3", "#801D2E", "#6b3b30"),
       set_name_color = "black", text_color = "white", text_size = 4.5,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)

```

```{r}
x = calculate.overlap(x)
```

```{r}
go.obj <- newGeneOverlap(DEGenes$DoubleKD_vs_Control$EnsID,
                         upf3dKD_res$EnsID,
                         genome.size=length(res$DoubleKD_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)


```

```{r}
overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>%
  inner_join(upf3dKD_res, by="EnsID") %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
   mutate(nif= ifelse(nif_feature == "None" |is.na(nif_feature),"NIF-", "NIF+"))

overlap_DSA = overlapping_upf3dKD %>%
  ggscatter(.,x="logFC.x", y ="log2FoldChange", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3 dKD DEG (log2FC)") +
  ylab("Upf3 dKD DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3b KD DSE (log2FC)")

# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-stability-overlap.png", plot = overlap_DSA,  units = c("in"), height = 7, width = 7, dpi =1200)



```


```{r}
transcript_overlap= overlapping_upf3dKD %>%
#  distinct(gene_id, .keep_all = TRUE) %>%
  ggscatter(.,x="logFC.x", y ="logFC.y", cor.coef = TRUE, add = "reg.line",
              conf.int = TRUE, color = "nif", add.params = list(color = "blue",
                                                 fill = "lightgray"), alpha =0.8, shape = "nif", size =3) +
    theme_bw() + xlab("Upf3 dKD DEG (log2FC)") +
  #ylab("Upf3b DSG (log2FC)") +   scale_color_manual(values = c("#F15A2B", "#2E368F")) +
    geom_hline(yintercept = 0, size = 0.5,lty = "dashed", color = "grey60") +
  geom_vline(xintercept = 0, size = 0.5, lty = "dashed", color = "grey60") + theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
            axis.text.y = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "top",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        )) + labs(color = "Expression") + geom_xsideboxplot(aes(y = nif, fill= nif), orientation = "y") +
   theme(ggside.panel.scale = .3) +
  scale_xsidey_discrete()  + geom_ysidedensity(aes(x = after_stat(density), fill = nif), alpha = 0.5) +
  scale_color_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) +  scale_fill_manual(values = c("NIF+" = "#b23a48",
                                "NIF-" = "grey")) + scale_shape_manual(values = c(19,15)) + labs(y= "Upf3 dKD DET (log2FC)")

ggarrange(overlap_DSA,transcript_overlap)
# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-transcript-overlap.svg", plot = transcript_overlap,  units = c("in"), height = 7, width = 7)
```

```{r}
non_intersect = x$a1[!x$a1 %in% x$a3]
non_overlapping_upf3dKD = DEGenes$DoubleKD_vs_Control %>%
  dplyr::filter(EnsID %in% non_intersect) %>%
    dplyr::rename("gene_id" = "EnsID") %>%
    left_join(res_dte$UPF3A_KD_UPF3B_KD_Control %>%
               left_join(tx2gene) %>%
               dplyr::filter(FDR < 0.05),by = "gene_id") %>%
  mutate(nif= ifelse(nif_feature == "None","NIF-", "NIF+"))
```


```{r}
nif_overlap =overlapping_upf3dKD %>%
  distinct(gene_id, .keep_all = TRUE) %>%
  dplyr::select(nif) %>%
   mutate(group = "Overlap") %>%
  rbind(non_overlapping_upf3dKD  %>%
          distinct(gene_id, .keep_all = TRUE) %>%
         dplyr::select(nif) %>%
          mutate(group = "Non-overlapping")) %>%

  mutate(nif = factor(nif, levels = c("NIF+", "NIF-"))) %>%
ggbarstats(
  x                = nif,
  y                = group,

xlab = "",
legend.title     = "",
  ggplot.component = list(ggplot2::scale_x_discrete(guide = ggplot2::guide_axis(n.dodge = 2))),
) + scale_fill_manual(values = c("NIF+" = "#b23a48","NIF-" = "grey"))

nif_overlap


# ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/Upf3dKD-bf-NIF-proportion.png", plot = nif_overlap ,  units = c("in"), height = 6, width = 6, dpi =1200)
```

### List of DD-NMD genes across analyses

```{r}
DD_NMD_candidates = list()

DD_NMD_candidates$Upf3b = overlapping_upf3b %>%
  dplyr::filter(logFC.x > 0) %>%
   dplyr::filter(nif == "NIF+")


DD_NMD_candidates$Upf3a = overlapping_upf3a %>%
  dplyr::filter(logFC.x > 0) %>%
   dplyr::filter(nif == "NIF+")

DD_NMD_candidates$Upf3dKD = overlapping_upf3dKD %>%
  dplyr::filter(logFC.x > 0) %>%
   dplyr::filter(nif == "NIF+")
```



- Upf3b DD-NMD genes

```{r}
DD_NMD_candidates$Upf3b   %>%
  dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id",
                "Adjusted P-value (Gene)" = "adj.P.Val",
                "Log2FoldChange (Gene)" ="logFC.x",
                "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR",
                 "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ",
                "NIF_uORF","GC_3utr","LENGTH_3utr") %>%
  mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"),
         NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>%
   DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
                 caption = "Upregulated and stabilized targets of Upf3b. Targets were selected if they were differentially stabilized with gene expression results upregulation  Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.",
            options = list(scrollX = TRUE,
                           scrollCollapse = TRUE,
                           pageLength = 5, autoWidth = TRUE,
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))

```

- Upf3a

```{r}
DD_NMD_candidates$Upf3a  %>%
  dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id",
                "Adjusted P-value (Gene)" = "adj.P.Val",
                "Log2FoldChange (Gene)" ="logFC.x",
                "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR",
                 "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ",
                "NIF_uORF","GC_3utr","LENGTH_3utr") %>%
  mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"),
         NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>%
   DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
                 caption = "Upregulated and stabilized targets of Upf3a. Targets were selected if they were differentially stabilized with gene expression results upregulation  Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.",
            options = list(scrollX = TRUE,
                           scrollCollapse = TRUE,
                           pageLength = 5, autoWidth = TRUE,
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


```



- Upf3 dKD

```{r}
DD_NMD_candidates$Upf3dKD  %>%
  dplyr::select("Gene" = "gene.y", "Gene ID" = "gene_id",
                "Adjusted P-value (Gene)" = "adj.P.Val",
                "Log2FoldChange (Gene)" ="logFC.x",
                "Transcript ID" = "tx_id", "Adjusted P-value (Transcript)" = "FDR",
                 "Log2FoldChange (Transcript)" = "logFC.y", "NIF_dEJ",
                "NIF_uORF","GC_3utr","LENGTH_3utr") %>%
  mutate(NIF_uORF = ifelse(NIF_uORF == 1, "Yes", "No"),
         NIF_dEJ = ifelse(NIF_dEJ  == 1, "Yes", "No")) %>%
   DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'),
                 caption = "Upregulated and stabilized targets of Upf3 dKD. Targets were selected if they were differentially stabilized with gene expression results upregulation  Targets were further filtered based on if they had a corresponding differentially expressed transcript with an NIF.",
            options = list(scrollX = TRUE,
                           scrollCollapse = TRUE,
                           pageLength = 5, autoWidth = TRUE,
                           dom = 'Blfrtip',
                           buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'),
                           lengthMenu = list(c(10,25,50,-1),
                                             c(10,25,50,"All"))))


```

### Overlap of targets


```{r}
x= list( "Upf3b KD" = DD_NMD_candidates$Upf3b$gene_id,
    "Upf3a KD" =DD_NMD_candidates$Upf3a$gene_id,
    "Upf3 dKD" = DD_NMD_candidates$Upf3dKD$gene_id)

ggvenn(x, fill_color = c("#1c174a", "#ef476f", "#7175b3"),
       set_name_color = "black", text_color = "white", text_size = 5,
       stroke_size = 0.45, fill_alpha = 0.8,
       stroke_color = "white", show_percentage = FALSE)
```

```{r}
overlaps = calculate.overlap(x)
```

```{r}
go.obj <- newGeneOverlap(DD_NMD_candidates$Upf3b$gene_id,
                         DD_NMD_candidates$Upf3a$gene_id,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))

```

```{r}
go.obj <- newGeneOverlap(DD_NMD_candidates$Upf3dKD$gene_id,
                         DD_NMD_candidates$Upf3a$gene_id,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```

```{r}
go.obj <- newGeneOverlap(DD_NMD_candidates$Upf3dKD$gene_id,
                         DD_NMD_candidates$Upf3b$gene_id,
                         genome.size=length(res$UPF3A_vs_Control$EnsID))
go.obj <- testGeneOverlap(go.obj)

message(paste("Odds ratio=", go.obj@odds.ratio))
message(paste("p-value=", go.obj@pval))
```

### NMD inducible features in each of the DD-NMD genes


- Upf3b


```{r}
tx_upf3b = DD_NMD_candidates$Upf3b  %>%
  dplyr::select(NIF_dEJ,NIF_uORF, NIF_long_3utr)


tx_upf3b$NIF_dEJ = as.numeric(tx_upf3b$NIF_dEJ)
tx_upf3b$NIF_uORF = as.numeric(tx_upf3b$NIF_uORF)
tx_upf3b$NIF_long_3utr = as.numeric(tx_upf3b$NIF_long_3utr)

colSums(tx_upf3b)
```


- Upf3a
```{r}
tx_upf3a = DD_NMD_candidates$Upf3a  %>%
  dplyr::select(NIF_dEJ,NIF_uORF, NIF_long_3utr)


tx_upf3a$NIF_dEJ = as.numeric(tx_upf3a$NIF_dEJ)
tx_upf3a$NIF_uORF = as.numeric(tx_upf3a$NIF_uORF)
tx_upf3a$NIF_long_3utr = as.numeric(tx_upf3a$NIF_long_3utr)

colSums(tx_upf3a)
```

- Upf3 dKD

```{r}
tx_upf3dkd = DD_NMD_candidates$Upf3dKD  %>%
  dplyr::select(gene_id, NIF_dEJ,NIF_uORF, NIF_long_3utr) %>%
  dplyr::filter(gene_id %in% overlaps$a7)


tx_upf3dkd$NIF_dEJ = as.numeric(tx_upf3dkd$NIF_dEJ)
tx_upf3dkd$NIF_uORF = as.numeric(tx_upf3dkd$NIF_uORF)
tx_upf3dkd$NIF_long_3utr = as.numeric(tx_upf3dkd$NIF_long_3utr)

colSums(tx_upf3dkd[,-1])

tx_upf3dkd %>%
  dplyr::filter(NIF_dEJ == 1)
```

```{r}
 DD_NMD_candidates$Upf3dKD %>%  drop_na(LENGTH_3utr) %>% with(median(LENGTH_3utr))
 DD_NMD_candidates$Upf3b %>%  drop_na(LENGTH_3utr) %>% with(median(LENGTH_3utr))
DD_NMD_candidates$Upf3a %>%  drop_na(LENGTH_3utr) %>% with(median(LENGTH_3utr))

t.test(DD_NMD_candidates$Upf3b$LENGTH_3utr, DD_NMD_candidates$Upf3a$LENGTH_3utr)
```

```{r}
library(ggstatsplot)
UTR = lapply(DD_NMD_candidates, function(x) {
  x %<>%
    dplyr::select(LENGTH_3utr)
})


 
utr_length = do.call(rbind, UTR) %>%
  rownames_to_column("Condition") %>%
   mutate_at(.vars = "Condition", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>%
    ggbetweenstats(
  x     = Condition,
  y     = LENGTH_3utr,
  p.adjust.method  = "bonferroni",
  ggsignif.args    = list(textsize = 4, tip_length = 0.01)
)  + theme_bw() + ylab("3'UTR length") + xlab("") +
  scale_color_manual(values = c("Upf3b"= "#1c174a","Upf3a" = "#ef476f",
                                "Upf3dKD" = "#7175b3")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))


ggsave(utr_length , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/DD-NMD/DD-NMD-UTR.svg", height = 6, width = 6, units = "in")
```

```{r}
uorf = lapply(DD_NMD_candidates, function(x){
  x%<>%
    dplyr::select(NIF_uORF) %>%
    drop_na() %>%
    mutate(NIF_uORF = as.numeric(NIF_uORF)) %>%
    colSums()
})

uorf$Upf3dKD = DD_NMD_candidates$Upf3dKD %>%
  dplyr::filter(gene_id %in% overlaps$a7) %>%
    dplyr::select(NIF_uORF) %>%
    drop_na() %>%
    mutate(NIF_uORF = as.numeric(NIF_uORF)) %>%
    colSums()




utr_length = do.call(rbind, uorf) %>%
  as.data.frame() %>%
  rownames_to_column("Condition") %>%
  ggplot(aes(x=Condition, y = NIF_uORF, fill = Condition)) + geom_bar(stat= "identity", width =0.5) +
  theme_bw() + ylab("# Transcripts with uORFs") + xlab("") +
  scale_fill_manual(values = c("Upf3b"= "#1c174a","Upf3a" = "#ef476f",
                                "Upf3dKD" = "#7175b3")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))


ggsave(utr_length , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/DD-NMD/DD-NMD-uORF.png", height = 6, width = 6, units = "in")
```


```{r}
dEJ = lapply(DD_NMD_candidates, function(x){
  x%<>%
    dplyr::select(NIF_dEJ) %>%
    drop_na() %>%
    mutate(NIF_dEJ = as.numeric(NIF_dEJ)) %>%
    colSums()
})

dEJ$Upf3dKD = DD_NMD_candidates$Upf3dKD %>%
  dplyr::filter(gene_id %in% overlaps$a7) %>%
    dplyr::select(NIF_dEJ) %>%
    drop_na() %>%
    mutate(NIF_dEJ = as.numeric(NIF_dEJ)) %>%
    colSums()


#t.test(DD_NMD_candidates$Upf3b$NIF_dEJ, DD_NMD_candidates$Upf3a$NIF_dEJ)

utr_length = do.call(rbind, dEJ) %>%
  as.data.frame() %>%
  rownames_to_column("Condition") %>%
  ggplot(aes(x=Condition, y = NIF_dEJ, fill = Condition)) + geom_bar(stat= "identity", width =0.5) +
  theme_bw() + ylab("# Transcripts with dEJs") + xlab("") +
  scale_fill_manual(values = c("Upf3b"= "#1c174a","Upf3a" = "#ef476f",
                                "Upf3dKD" = "#7175b3")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))

utr_length 
ggsave(utr_length , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/DD-NMD/DD-NMD-dEJ.png", height = 6, width = 6, units = "in")
```



```{r}
utr  = lapply(DD_NMD_candidates, function(x){
  x%<>%
    dplyr::select(NIF_long_3utr) %>%
    drop_na() %>%
    mutate(NIF_long_3utr = as.numeric(NIF_long_3utr)) %>%
    colSums()
})

utr$Upf3dKD = DD_NMD_candidates$Upf3dKD %>%
  dplyr::filter(gene_id %in% overlaps$a7) %>%
    dplyr::select(NIF_long_3utr) %>%
    drop_na() %>%
    mutate(NIF_long_3utr = as.numeric(NIF_long_3utr)) %>%
    colSums()


#t.test(DD_NMD_candidates$Upf3b$NIF_dEJ, DD_NMD_candidates$Upf3a$NIF_dEJ)

utr_length = do.call(rbind, utr) %>%
  as.data.frame() %>%
  rownames_to_column("Condition") %>%
  ggplot(aes(x=Condition, y = NIF_long_3utr, fill = Condition)) + geom_bar(stat= "identity", width =0.5) +
  theme_bw() + ylab("# Transcripts with long 3'UTRs") + xlab("") +
  scale_fill_manual(values = c("Upf3b"= "#1c174a","Upf3a" = "#ef476f",
                                "Upf3dKD" = "#7175b3")) +
      theme(axis.text.x = element_text(size = 12, family = "serif", color = "black"),
        axis.title = element_text(size=15, family = "serif"),
         legend.box.background = element_rect(color = "black"),
          legend.text = element_text(family = "serif"),
          legend.title = element_text( family = "serif"),
           plot.title = element_text(family = "serif", size =20),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line( size=.1 ),
              legend.position = "none",
        strip.text.y = element_text(
        size = 10, face = "bold.italic", family = "serif"
        ),
         strip.text.x = element_text(
        size = 10, family = "serif"
        ))
utr_length

ggsave(utr_length , filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Thesis_plots/DD-NMD/DD-NMD-UTR-presence.png", height = 6, width = 6, units = "in")
```


### Heatmaps of targets

```{r}

heat_map = DD_NMD_candidates$Upf3dKD   %>%
  dplyr::filter(gene_id %in% overlaps$a7) %>%
  dplyr::select(gene.x, NIF_dEJ,NIF_uORF, NIF_long_3utr, GC_3utr, LENGTH_3utr)
heat_map$NIF_dEJ = as.numeric(heat_map$NIF_dEJ)
heat_map$NIF_uORF = as.numeric(heat_map$NIF_uORF)
heat_map$NIF_long_3utr = as.numeric(heat_map$NIF_long_3utr)
heat_map  = heat_map %>%
  dplyr::group_by(gene.x) %>%
  dplyr::summarise(across(.cols = starts_with("NIF"),.fns = sum)) %>%
  column_to_rownames("gene.x") %>%
  mutate(Stabilized = 1,
         Upregulated = 1) %>%
  dplyr::select(Upregulated, Stabilized, everything()) %>%
  mutate_if(is.numeric, ~1 * (. != 0))


upf3dKD_candidates = heat_map%>% head(60) %>% t() %>%
    pheatmap(cellwidth = 10, cellheight = 10,
           	color = (c("#fffdf7", "#b23a48")),
           border_color = "white",
           cluster_rows = FALSE,
           cluster_cols = FALSE)

#ggsave(filename  = here::here("output/Thesis_plots/Upf3d-KD-candidates-features-1.svg"),
 #   height = 5, width =15, plot =  upf3dKD_candidates)
```

```{r}
upf3dKD_candidates_1 = heat_map%>% tail(47) %>% t() %>%
    pheatmap(cellwidth = 10, cellheight = 10,
           	color = (c("#fffdf7", "#b23a48")),
           border_color = "white",
           cluster_rows = FALSE,
           cluster_cols = FALSE)

#ggsave(filename  = here::here("output/Thesis_plots/Upf3d-KD-candidates-features-2.svg"),
 #   height = 5, width =15, plot =  upf3dKD_candidates_1)
```

<!-- ## Gene ontology enrichment analysis -->

<!-- ```{r mouse-length} -->
<!-- load("/home/neuro/Documents/NMD_analysis/Analysis/Genome/mouse_length.Rda") -->
<!-- ``` -->

<!-- - Upf3b -->

<!-- ```{r} -->
<!-- analysis =  res$UPF3B_vs_Control -->
<!-- genes = as.integer(analysis$EnsID %in% DD_NMD_candidates$Upf3b$gene_id) -->
<!-- not_na = !is.na(genes) -->
<!-- names(genes) = analysis$EnsID -->
<!-- genes = genes[not_na] -->

<!-- select_genes <- as.vector(names(lengthData)%in%names(genes)) -->
<!-- new_lengthData <- lengthData[select_genes] -->
<!-- select_genes <- as.vector(names(genes) %in% names(new_lengthData)) -->
<!-- genes = genes[select_genes] -->
<!-- pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE) -->
<!-- plotPWF(pwf, main = "Bias from gene length") -->


<!-- Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE) -->
<!-- goterms <- Go.wall$category -->
<!-- goList <- getGeneLists(pwf, goterms, "mm39", "ensGene") -->
<!-- Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]]) -->
<!-- Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L)) -->
<!-- goRes <- Go.wall %>% -->
<!--             dplyr::select(-under_represented_pvalue) %>% -->
<!--             dplyr::filter(numDEInCat > 0) %>% -->
<!--             dplyr::rename(id = category) %>% -->
<!--             left_join(goSummaries) -->
<!-- nGenes <- c(BG = sum(genes == 0), DE = sum(genes)) -->
<!-- goFinal <- goRes %>% -->
<!--             dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > 3) %>% -->
<!--             arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>% -->
<!--             arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>% -->
<!--             mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"), -->
<!--                    Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>% -->
<!--             dplyr::filter(over_represented_pvalue < 0.05,  numDEInCat > Expected) -->

<!-- ``` -->


<!-- ```{r} -->

<!-- goFinal$Genes = sapply(goFinal$EnsemblID, function(x){ -->
<!--             tmp = strsplit(as.character(x), ",") %>% unlist() -->
<!--             tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim) -->
<!--             tmp = tmp[[1]] -->
<!--             tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% -->
<!--                 unname -->
<!--             return(tmp) -->
<!--         }) -->
<!-- goFinal$Genes = vapply(goFinal$Genes , paste, collapse = ", ", character(1L)) -->


<!-- goFinal %>% -->
<!--     dplyr::select(-c(contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>% -->
<!--  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                caption ="Significant gene ontology terms in Upf1 KD compared to controls", -->
<!--                style = "auto", width = NULL, height = NULL, elementId = NULL, -->
<!--             options = list(scrollX = TRUE, -->
<!--                            scrollCollapse = TRUE, -->
<!--                            pageLength = 5, autoWidth = TRUE, -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->

<!-- - Upf3a -->

<!-- ```{r} -->
<!-- analysis =  res$UPF3A_vs_Control -->
<!-- genes = as.integer(analysis$EnsID %in% DD_NMD_candidates$Upf3a$gene_id) -->
<!-- not_na = !is.na(genes) -->
<!-- names(genes) = analysis$EnsID -->
<!-- genes = genes[not_na] -->

<!-- select_genes <- as.vector(names(lengthData)%in%names(genes)) -->
<!-- new_lengthData <- lengthData[select_genes] -->
<!-- select_genes <- as.vector(names(genes) %in% names(new_lengthData)) -->
<!-- genes = genes[select_genes] -->
<!-- pwf=nullp(genes,bias.data=new_lengthData, plot.fit = FALSE) -->
<!-- plotPWF(pwf, main = "Bias from gene length") -->


<!-- Go.wall=goseq(pwf, "mm39", "ensGene", use_genes_without_cat = TRUE) -->
<!-- goterms <- Go.wall$category -->
<!-- goList <- getGeneLists(pwf, goterms, "mm39", "ensGene") -->
<!-- Go.wall$EnsemblID <- sapply(Go.wall$category, function(x) goList[[x]]) -->
<!-- Go.wall$EnsemblID <- vapply(Go.wall$EnsemblID, paste, collapse = ", ", character(1L)) -->
<!-- goRes <- Go.wall %>% -->
<!--             dplyr::select(-under_represented_pvalue) %>% -->
<!--             dplyr::filter(numDEInCat > 0) %>% -->
<!--             dplyr::rename(id = category) %>% -->
<!--             left_join(goSummaries) -->
<!-- nGenes <- c(BG = sum(genes == 0), DE = sum(genes)) -->
<!-- goFinal <- goRes %>% -->
<!--             dplyr::filter(numDEInCat > 2) %>% dplyr::filter(shortest_path > 3) %>% -->
<!--             arrange(desc(shortest_path), EnsemblID) %>% distinct(EnsemblID, .keep_all = TRUE) %>% -->
<!--             arrange(over_represented_pvalue) %>% mutate(Expected = round(nGenes[["DE"]] * numInCat / nGenes[["BG"]], 0)) %>% -->
<!--             mutate(adjP = p.adjust(over_represented_pvalue, "bonferroni"),FDR = p.adjust(over_represented_pvalue, "fdr"), -->
<!--                    Sig_adjP = adjP < 0.05, Sig_FDR = FDR < 0.05) %>% -->
<!--             dplyr::filter(over_represented_pvalue < 0.05, numDEInCat > Expected) -->

<!-- ``` -->


<!-- ```{r} -->

<!-- goFinal$Genes = sapply(goFinal$EnsemblID, function(x){ -->
<!--             tmp = strsplit(as.character(x), ",") %>% unlist() -->
<!--             tmp %<>% as.data.frame() %>% mutate_if(is.character, str_trim) -->
<!--             tmp = tmp[[1]] -->
<!--             tmp = mapIds(org.Mm.eg.db, keys=tmp,  column="SYMBOL",keytype="ENSEMBL", multiVals="first") %>% -->
<!--                 unname -->
<!--             return(tmp) -->
<!--         }) -->
<!-- goFinal$Genes = vapply(goFinal$Genes , paste, collapse = ", ", character(1L)) -->


<!-- goFinal %>% -->
<!--     dplyr::select(-c(contains("Sig"), EnsemblID, shortest_path, longest_path, terminal_node)) %>% -->
<!--  DT::datatable(filter = 'top', extensions = c('Buttons', 'FixedColumns'), -->
<!--                caption ="Significant gene ontology terms in Upf1 KD compared to controls", -->
<!--                style = "auto", width = NULL, height = NULL, elementId = NULL, -->
<!--             options = list(scrollX = TRUE, -->
<!--                            scrollCollapse = TRUE, -->
<!--                            pageLength = 5, autoWidth = TRUE, -->
<!--                            dom = 'Blfrtip', -->
<!--                            buttons = c('copy', 'csv', 'excel', 'pdf', 'colvis'), -->
<!--                            lengthMenu = list(c(10,25,50,-1), -->
<!--                                              c(10,25,50,"All")))) -->
<!-- ``` -->
