---
title: "transcript-preprocessing"
author: "unawaz1996"
date: "2023-03-24"
output: 
  html_notebook: default
  workflowr::wflow_html:
    code_folding: hide
editor_options:
  chunk_output_type: console
---

# Transcript level exploration of data 

## Filtering the data 
Over here I will go through various preprocessing steps for filtering transcripts 

```{r}
source("code/libraries.R")
source("code/functions.R")
library(fishpond)
library(tximeta)
library(corrplot)
```

```{r}
ah <- AnnotationHub() %>%
    subset(species == "Mus musculus") %>%
    subset(rdataclass == "EnsDb") %>% 
  subset(genome == "GRCm39")
ensDb <- ah[["AH104895"]]

```

```{r}
## Create manual transcriptome annotation

txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])


indexDir = file.path("/home/neuro/Documents/NMD_analysis/Analysis/Genome/for_salmon/April_2023/decoy_awaresalmon_index/")
fasta = file.path("/home/neuro/Documents/NMD_analysis/Analysis/Genome/for_salmon/Mus_musculus.GRCm39.cdna.all.fa")
gtf = file.path("/home/neuro/Documents/NMD_analysis/Analysis/Genome/Mus_musculus.GRCm39.107.gtf.gz")
tmp <- tempdir()
jsonFile <- file.path(tmp, paste0(basename(indexDir), ".json"))

makeLinkedTxome(indexDir=indexDir,
                source="Ensembl", organism="Mus musculus",
                release="107", genome="GRCm39.107",
                fasta=fasta, gtf=gtf,
                jsonFile=jsonFile)

```



```{r}
md.files = "data/LTK_Sample Metafile_V3.txt"
salmon.files = ("/home/neuro/Documents/NMD_analysis/Analysis/Results/Quantification/Salmon_decoy_aware/GRm39")
salmon = list.files(salmon.files, pattern = "transcripts$", full.names = TRUE)
names(salmon) <- paste(c(212:229))
md = read.table(md.files, header= TRUE) %>%  mutate(files = file.path(salmon, "quant.sf"))
comparisons = unique(md$Group)[-1]
colnames(md)[1] = c("names")
md$names = as.character(md$names)

md$Group <- factor(md$Group, 
                            levels=c("Control","UPF3B_KD", "UPF3A_KD", "UPF3A_OE", 
                                     "UPF3A_KD_UPF3B_KD", "UPF3A_OE_UPF3B_KD"))

se <- tximeta(md)


#txi<- tximport(all_files, type="salmon", txOut=TRUE,
 #              countsFromAbundance="scaledTPM", tx2gene = tx2gene, ignoreTxVersion = TRUE, ignoreAfterBar = TRUE)
```

# Comparison of data distributions before and after filtering of low expressed transcripts and genes 
Before and after filtering 

```{r}
txdf = transcripts(EnsDb.Mmusculus.v79, return.type="DataFrame")
tx2gene = as.data.frame(txdf[,c("tx_id","gene_id", "tx_biotype")])

## Get rid of versions id from transcript id 
# txi[c(1,2,4)] = lapply(txi[c(-3, -5)], function(x){
#   x %>% as.data.frame() %>%
#     rownames_to_column("tx") %>% 
#     mutate_at(.vars = "tx", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
#     column_to_rownames("tx")
# })
```


```{r}
## Filtering threshold
keep = (rowSums(se@assays@data$abundance >= .5) >= 3)

before =  se@assays@data$abundance %>% as.data.frame() %>%
    mutate(rowMeans = rowMeans(.)) %>% 
  rownames_to_column("tx_id") %>%
  left_join(tx2gene, by = "tx_id") %>%
  group_by(gene_id) %>% mutate(ntx = n()) %>% 
dplyr::select(gene_id, tx_id, tx_biotype, rowMeans, ntx) 

after = se@assays@data$abundance[keep,] %>% as.data.frame() %>%
    mutate(rowMeans = rowMeans(.)) %>% 
  rownames_to_column("tx_id") %>%
  left_join(tx2gene, by = "tx_id") %>%
  group_by(gene_id) %>% mutate(ntx = n()) %>% 
dplyr::select(gene_id, tx_id, tx_biotype, rowMeans, ntx) 


df <- gdata::combine(before, after) %>% dplyr::rename(Filter=source)

dplyr::group_by(df,Filter) %>% dplyr::summarise(median(rowMeans),sd(rowMeans))


ntxInfo  <- df %>% dplyr::group_by(Filter) %>%
    dplyr::distinct(gene_id,.keep_all = TRUE)  
biotypeCount <- df %>% dplyr::group_by(Filter) %>%
    dplyr::count(tx_biotype) %>%
    mutate(sum=sum(n),frac = n / sum(n)) 

```

```{r fig.height=6, fig.width=6, fig.cap= "*Log10 of the mean CPMs (counts per million) over all samples before and after filtering out low-expressed transcripts and genes*"}
distribution = ggplot(df, aes(x = log10(rowMeans), col = Filter)) +
    geom_histogram(aes(y = ..density.., fill = Filter), alpha = 0.4, bins = 100) +
    geom_density(size = 1.1) +
    scale_x_continuous(name = "log10(mean TPM)", limits = c(-8,10)) +
    theme_bw() +
    scale_color_manual(values = c("#FCC98A", "#DF1861"))+
    scale_fill_manual(values = c("#FCC98A", "#DF1861"))
distribution
ggsave(plot = dist, file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/transcript-distribution.svg", 
       height = 4.29, width = 8.02, units = "in")
```

```{r fig.cap="*Violin plots showing the distribution of the number of transcripts per gene (in logarithmic scale). Violin width is scaled by the total number of observations while jittered points represent actual observations.*"}
biocount = ggplot(subset(biotypeCount, frac >= 0.001), aes(x = as.factor(tx_biotype), y = frac, fill = Filter)) +
    geom_bar( stat = "identity") +
    coord_flip() +
    facet_wrap(.~Filter) +
    theme_bw() +
    theme(axis.text.x = element_text(angle = -90)) +
    labs(x = "Transcript biotype (Ensembl v75)", y = "Frequency") +
        scale_color_manual(values = c("#FCC98A", "#DF1861"))+
    scale_fill_manual(values = c("#FCC98A", "#DF1861"))+  
theme(panel.spacing = unit(1, "lines")) +
theme(strip.background = element_rect(fill = "black"))+
theme(strip.text = element_text(color = "white", face = "bold"))
ggsave(plot = biocount, file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/transcript-composition.svg", 
       height = 4.05, width = 7.74, units = "in")
```


```{r}
ggplot(ntxInfo, aes(x = Filter, y = log10(ntx))) +
    geom_violin(scale = "count", aes(fill = Filter)) +
    geom_jitter(width = 0.1, size = 0.3, alpha = 0.01) +
    theme_bw() +
    scale_y_continuous(name = "log10(# transcripts per gene)") + 
    labs(x = "", title = "b") +
        scale_color_manual(values = c("deepskyblue3", "darkblue"))+
    scale_fill_manual(values = c("deepskyblue3", "darkblue"))   
```
 


```{r}
log10(se@assays@data$abundance + 0.01) %>%
    melt() %>%
    dplyr::filter(is.finite(value)) %>% 
    ggplot(aes(x=value, color = as.character(Var2))) +
    geom_density() +
    ggtitle("Before filtering") +
    labs(x = "logTPM", y = "Proportion of Genes") + 
    theme_bw() +
    theme(legend.position = "none")

log10(se@assays@data$abundance[keep,] + 0.01) %>%
    melt() %>%
    dplyr::filter(is.finite(value)) %>% 
    ggplot(aes(x=value, color = as.character(Var2))) +
    geom_density() +
    ggtitle("Before filtering") +
    labs(x = "logTPM", y = "Proportion of Genes") + 
    theme_bw() +
    theme(legend.position = "none")
```

### Overlap of transcripts with genes from other analysis 

```{r}
tx_genes =  tximport(all_files, type="salmon", txOut=FALSE,
                      countsFromAbundance="scaledTPM", tx2gene = tx2gene, ignoreTxVersion = TRUE, ignoreAfterBar = TRUE)
keep.genes = (rowSums(tx_genes$abundance >= .5 ) >= 3)
tx_genes = tx_genes$abundance[keep.genes,]

tx_for_comp = txi$abundance[keep,] %>% as.data.frame %>% 
    rownames_to_column("tx_id") %>% 
  mutate(ens_gene = mapIds(org.Mm.eg.db, keys=.$tx_id,
                             column="ENSEMBL",keytype="ENSEMBLTRANS",
                         multiVals="first")) 




## Roughly 2754 genes have matching transcripts 

x = list( "genes" = rownames(tx_genes), 
    "tx" = tx_for_comp$ens_gene)

ggvenn(x[c(1,2)])

```

```{r}
tx_for_comp = tx_for_comp[tx_for_comp$ens_gene %in% rownames(tx_genes),]
length(unique(tx_for_comp$ens_gene))

tx_for_comp %<>% left_join(tx2gene, by = "tx_id")

biotypeCount_genes <- tx_for_comp  %>% melt() %>% 
    dplyr::count(tx_biotype) %>%
    mutate(sum=sum(n),frac = n / sum(n)) 


ggplot(subset(biotypeCount_genes, frac >= 0.001), aes(x = as.factor(tx_biotype), y = frac)) +
    geom_bar( stat = "identity") +
    coord_flip() +
    theme_bw() +
    theme(axis.text.x = element_text(angle = -90)) +
    labs(x = "Transcript biotype (Ensembl v75)", y = "Frequency", title = "c") +
theme(panel.spacing = unit(1, "lines")) +
theme(strip.background = element_rect(fill = "black"))+
theme(strip.text = element_text(color = "white", face = "bold"))


```

### Specific to NMD features
```{r}

filt_genes = se@assays@data$abundance[keep,] %>% as.data.frame() %>% 
  rownames_to_column("tx_id")
NMD_features = read.table("data/nif_output/Mus_musculus__nifs.tsv", sep = "\t", skip =1)

colnames(NMD_features)[1:4] = c("EnsID", "NIF_dEJ", "NIF_uORF", "NIF_long-3'utr")
NMD_features = NMD_features[,1:4]
tx_NMD = NMD_features %>% 
  mutate_at(.vars = "EnsID", .funs = gsub, pattern = "\\.[0-9]*$", replacement = "") %>% 
  dplyr::rename("tx_id" = "EnsID") %>% 
  right_join(filt_genes, by="tx_id") 

dEJ = tx_NMD %>% 
  dplyr::filter(NIF_dEJ == "True") %>% 
  dplyr::select(tx_id)

uORF = tx_NMD %>% 
  dplyr::filter(NIF_uORF == "True") %>% 
  dplyr::select(tx_id)
UTR = tx_NMD %>% 
  dplyr::filter(`NIF_long-3'utr` == "True") %>% 
  dplyr::select(tx_id)

x = list( "dEJ" = dEJ$tx_id, 
          "uORF" = uORF$tx_id, 
          "UTR" = UTR$tx_id, 
          "All transcript" = tx_NMD$tx_id)

plot = upset(fromList(x))

ggsave(file = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/total-transcript-composision.svg", 
       height =5.77 ,width=8.10, units = "in", plot = plot)
```

## Principal component analysis
```{r}
pca = log2(se@assays@data$abundance[keep,]  + 10^-6 ) %>%
    t() %>% 
    prcomp(scale = TRUE)


pcaVars <- percent_format(0.1)(summary(pca)$importance["Proportion of Variance",]) 

md$Group = factor(md$Group)


pca$x[,1:2] %>% 
    as.data.frame()  %>%
    cbind(md) %>%
    as.tibble() %>%
    ggplot(aes(PC1, PC2, shape = Condition_UPF3B, label = Condition_UPF3B, color=Group)) +
    geom_point(size = 11.5) +
    theme_bw() + scale_shape_manual(values = c(0,16)) + labs(shape="UPF3B status", colour= "Samples") +
    theme(axis.title.y = element_text(size = 12), axis.title.x = element_text(size = 12), 
          axis.text = element_text(size = 10)) +
    xlab(paste0("Principal Component 1 (28.1%)")) + ylab("Principal Component 2 (16.6%)") +
    scale_color_manual(values = c("#FCC98A",  "#2F124B", "#6E74B4",  "#EA2A5F", "#E12F8F", "#FD9675"),
                      labels = c("UPF3A_KD_UPF3B_KD" = "Double KD", "UPF3A_KD" = "UPF3A KD", 
                                                                   "UPF3B_KD" = "UPF3B KD", "UPF3A_OE" = "UPF3A OE", 
                                                                   "UPF3A_OE_UPF3B_KD" = "UPF3A OE UPF3B KD")) +
  labs(x = paste0("PC1 (", pcaVars[["PC1"]], ")"),
    y = paste0("PC2 (", pcaVars[["PC2"]], ")"))


ggsave(filename = "/home/neuro/Documents/NMD_analysis/Analysis/NMD-analysis/output/Transcript/Thesis_figures/Transcript_PCA.svg", plot = transcript_PCA, units = c("in"), height = 5.60, width = 6.65)
```


## PCA with respect to transcript length and GC content 


```{r}
rawFqc <- list.files(
    path = "data/fastqc/Raw",
    pattern = "zip",
    full.names = TRUE
    ) %>%
    FastqcDataList()

ah <- AnnotationHub() %>%
    subset(species == "Mus musculus") %>%
    subset(rdataclass == "EnsDb") %>% 
  subset(genome == "GRCm39")
ensDb <- ah[["AH104895"]]
grTrans <- transcripts(ensDb)
trLengths <- exonsBy(ensDb, "tx") %>%
    width() %>%
    vapply(sum, integer(1))
mcols(grTrans)$length <- trLengths[names(grTrans)]

gc <- getModule(rawFqc, "Per_sequence_GC") 

gcDev <- gc %>%
  mutate(names = gsub("\\_.*", "", Filename)) %>% 
    left_join(md, by = "names") %>%
    group_by(names) %>%
    mutate(Freq = Count / sum(Count)) %>%
    left_join(
        getGC(gcTheoretical, "Mmusculus", "Trans")) %>%
    dplyr::rename(actual = Mmusculus) %>%
    mutate(res = Freq - actual) %>%
    summarise(ss = sum(res^2), n = n()) %>%
    ungroup() %>%
    mutate(sd = sqrt(ss / (n - 1)))

rawGC <- gc %>% 
    group_by(Filename) %>% 
    mutate(Freq = Count / sum(Count)) %>%
    dplyr::filter(GC_Content > 70) %>% 
    summarise(Freq = sum(Freq)) %>% 
    arrange(desc(Freq)) %>%
     mutate(names = gsub("\\_.*", "", Filename)) %>% 
    left_join(md, by= "names") 


gcInfo <- se@assays@data$abundance[keep,] %>%
    as.data.frame() %>%
    rownames_to_column("tx_id")  %>%
    as_tibble() %>%
    pivot_longer(
        cols =  starts_with("2"), 
        names_to = "sampleName", 
        values_to = "counts"
    ) %>%
    dplyr::filter(
        counts > 0
    )  %>%
    left_join(
        mcols(grTrans) %>% as.data.frame()
    ) %>% 
   dplyr::select(
        ends_with("id"), sampleName, counts, gc_content, length
    ) %>%
    split(f = .$sampleName) %>%
    lapply(function(x){
        DataFrame(
            gc = Rle(x$gc_content/100, x$counts),
            logLen = Rle(log10(x$length), x$counts)
        )
    }
    ) 


gcSummary <- gcInfo %>%
    vapply(function(x){
        c(mean(x$gc), sd(x$gc), mean(x$logLen), sd(x$logLen))
        }, numeric(4)
        ) %>%
    t() %>%
    set_colnames(
        c("mn_gc", "sd_gc", "mn_logLen", "sd_logLen")
    ) %>%
    as.data.frame() %>%
    rownames_to_column("names") %>%
    as_tibble() %>%
    left_join(dplyr::filter(md)) #%>%
    #dplyr::select(starts_with("sample"), contains("_"))


```


```{r}

pca$x %>%
    as.data.frame() %>%
  rownames_to_column("names") %>%
    left_join(gcSummary) %>%
    as_tibble() %>%
     left_join(
        dplyr::filter(rawGC)
    ) %>% 
  left_join(gcDev) %>%
    dplyr::select(
        PC1, PC2, PC3, 
        Mean_GC = mn_gc, 
        Mean_Length = mn_logLen, 
        Initial_GC70 = Freq,
        SD = sd,
        Group
    ) %>% 
    mutate(Group = as.numeric(as.factor(Group))) %>%
    cor() %>% 
    corrplot(
        type = "lower", 
        diag = FALSE, 
        addCoef.col = 1)




```
